### 2020-12-08 write/manuscript.tex
initial commit

\documentclass[a4paper,11pt]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{natbib}

\date{Draft \today}
\title{Towards a Universal Veering Profile for Turbulent Ekman Flow at arbitrary Reynolds number \\
  \normalsize LES and DNS of Turbulent Ekman Flow \\}

\begin{document}

\maketitle

\begin{abstract}
  An analytical formulation for the profiles of stream- and span-wise velocity in turbulent Ekman flow
  is developed based on asymptotic theory as well as data from direct numerical simulation and large-eddy simulation.
  %
  The profile of the stream-wise component follows the classical viscous, logarithmic and wake scaling.
  %
  The span-wise component poses a conceptual challenge to the channel-flow analogy
  in the context of asymptotic mixing, and it exhibits a mixed scaling in the surface layer, but follows
  outer scaling for most of the outer layer.
  %
  In both profiles, the wake region can be described by smooth error-function transition from a logarithmic layer
  to the constant outer velocity.

\end{abstract}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\section{Introduction}

The exact veering of the wind away from its geostrophic direction is due to direct and indirect (due to turbulence) frictional effects,
and it plays a role for various practical problems: 
%
It sets the amount of cross-isobaric volume flow and is thus a key factor in the life-cycle of large-scale synoptic systems. This mechanism is commonly
referred to as 'Ekman pumping' referring to the perturbation of the geostrophic equilibrium that causes the veering as it was first described
and solved for analytically by \cite{Ekman:1905tl}.
%
On the mesoscale, shear in the surface layer is known to be crucial for the initiation and sustainment of strong convective events (REFERENCE?).
% 
And locally, directional sheer of the wind in the upper part of the surface layer may cause a yaw for large wind power generation devices (REFERENCE?).
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\section{Formulation}
%
\paragraph{Physical Problem.} We consider incompressible, turbulent Ekman flow, i.e. the turbulent flow over a flat rotating plate, as
a physical model for the truly neutral atmospheric boundary layer (ABL).
%
That is, the f-plane approximation is used confining effects of rotation to the horizontal components of
velocity and neglecting dynamical effects due to latitudinal changes of the rate of rotation. 
%
%%PAR
%
\paragraph{Notation.} The dimensional velocity vector of the numerical simulations is $\underline{U} = (U_1,U_2,U_3) = (U,V,W)$ over the coordinate system
$Oxyz$, where an approximate alignment with the surface shear stress is achieved.
%
For analysis of the results, we use two coordinate systems that are
(i)  exactly aligned with the surface shear stress and labelled by an upper index $\alpha$ as in $\underline{U}^\alpha$ for the velocity vector, and
(ii) the coordinate system aligned with the free-atmosphere geostrophic wind labelled by an upper index $G$ as in $\underline{U}^G$.
%
We denote by $u_\star$ the modulus surface friction velocity and let $Z_\star=1/u_\star$ for brevity.

\section{Towards a Universal Velocity Profile for the Turbulent Ekman Layer}
\subsection{Drag-Law} 
The drag law for Ekman flow, including prediction of surface veering, has received considerable attention in the past.
\begin{itemize}
\item \cite{Rossby:1935wa} first approach;
\item \cite{Tennekes:1973jh, Blackadar:1968ew} asymptotic similarity;
\item \cite{Spalart:1989p2436, Spalart:2008p2432, Spalart:2009p2466} seminal LES studies in search for a universal log-law;
\item \cite{Ansorge:2014hf, Ansorge:2019di} re-evaluation of the constants with geophysical focus;
\item \cite{Hogstrom:1988wg, Hgstrm:1996kt} (re-)consideration of observational data.
  %

\end{itemize}
\begin{figure}
  \centerline{\includegraphics[width=0.7\textwidth]{ustar_alpha.pdf}}
  \caption{Variation of $Z_\star$ and surface veering with Reynolds number according to
    the theory by Spalart et al. (1989) and as estimated from DNS data}
  \label{fig:drag_law}
\end{figure} 
%
Attempts were also undertaken to obtain profiles of the wind speed: 
\cite{Gryning:2007dy} present an extension of the wind-speed profile beyond the surface layer using a neutral referece profile
and a stability correction;
\cite{Kelly:2010ev}, based on a probabilistic representation of stratification, develop a model for the long-term mean windspeed
in the atmospheric boundary layer and compare this with observation at different sites; 
\cite{Kelly:2016bs} demonstrate the effect of such improved model for wind-energy applications.
%
These investigations merely focus on the wind speed and thus do not regard the actual vector, i.e. the veering of
the wind with height is not described and there is little knowledge on the profile of the span-wise velocity component
and the precise shape of the hodograph in the limit of a truly neutral Ekman boundary layer.
%
While this is indeed an academic problem, the Ekman boundary layer often serves as limit for stratified boundary layers
and thus not only needs to be understood but also well-described as reference for higher-order approaches that take into
account possible effects of stratification, roughness or other physical complications of the geophysical system.
%
\subsection{Stream-wise Velocity} 
For the stream-wise velocity profile (that in non-rotating flows due to the geometry is always aligned with the surface shear stress),
well-established theories exist for various regimes according to their distance from the wall and the relative role of
viscosity, turbulence and interaction with the outer region of the flow with the logarithmic law for the mean velocity as a central anchor point:
\begin{subequations}
  \begin{itemize}
  \item In immediate vicinity to the surface, local turbulent mixing cannot occur for the no-slip/no-penetration boundary condition,
    and the mean velocity is described by a viscous profile of the form
    \begin{equation}
      (U,W)^{\alpha+}  = (z^+,0)  
    \end{equation}
    where the direction of 
    in absence of roughness elements are marginal linear regime / viscous sub-layer \cite{Foken:2002gv,Foken:1978bc} 
  \item buffer layer transition from linear regime to log-law  
  \item log-layer (K\'arm\'an, Prandtl, Jimenez, Marusic)
  \item outer log-layer () 
  \item wake region
  \item outer region 
  \end{itemize}
\end{subequations} 
%
\begin{figure}
  \includegraphics[width=\textwidth]{u_profile.pdf}
  \caption{Shear-aligned profiles of velocity components $U^{\alpha+}$ in inner (left) and outer (right) units.} 
\end{figure}

Existing theories applied to the boundary layer:
\begin{itemize}
\item \cite{Etling:2002kx, Emeis:2018vq}  (Sec. 21.10; Eq. 21.48) -- ad-hoc matching at Prandtl-layer height; requires $\alpha$ 
\item \cite{Emeis:2007cy} (Sec. 3; Eq. 3.1-3.19) -- 1D profile; considering constant (?) veering 
\item \cite{Gryning:2007dy} (Mean wind profiles)
\end{itemize}

For non-rotating, non-external flows theoretical considerations by Townsend (cite to attached-eddy hypothesis)
have recently been confirmed from both wind-tunnel  experiments and numerical simulation \cite{Ng:2011bk,Marusic:2013hf}
implying physical models for the second moments (Reynolds stresses) of the flow.
%
\subsection{Span-wise velocity} 
%
The background rotation and associated veering of the surface wind implies a non-zero profile for the span-wise velocity that
challenges the conventional assumptions related to the so-called channel-flow analogy:
%
While the universal profiles in vicinity of the wall implies that the profile be zero or at least small with respect to the stream-wise
component, the veering requires a value of $V_{top}=U_G\sin\alpha$ in the free stream (and thus also at the top of the boundary layer if we assume
that substantial velocity gradients are confined to the turbulent part of the flow). 
%
This is commonly shown in terms of velocity hodographs aligned with the outer, geostrophic flow (cf. Fig.~\ref{fig:hodograph}) and normalized by
the geostrophic wind. 
%
Thus, the drag-law already implies three important properties of these hodographs,
(i) the boundary conditions at the surface,
(ii) the boundary condition at the top, and
(iii) the inclination of the hodograph at the origin by the surface veering:
\begin{subequations} 
\begin{eqnarray} 
 \underline{U}^{G}(z=0) &=& 0,\\ 
 \underline{U}^{G}(z\mapsto\infty) &=& \underline{G}^G = \left(\begin{array}{c}1\\0\end{array}\right) \\
   \left.\frac{\partial V^G}{\partial U^G}\right|_{z=0} &=& \sin\alpha
\end{eqnarray} 
\end{subequations} 
%
Outer scaling of the velocity profile further implies that the velocity deficity of the spanwise component $V^\alpha-G^\alpha$
be a universal function of the outer height $z^-$, i.e.  
\begin{eqnarray}
  V^\alpha-G^\alpha = f^\alpha_V(z^-).  
\end{eqnarray}
%
Interpreting the shear-aligned spanwise velocity deficit $f_{V}^{\alpha}(z^-)$ as 
a signature of outer rotation may lead one to conclude that this universal function
is appropriate down to the surface--irrespective of the Reynolds number.
%
This cannot be the case for the variation of $f_{V}^{\alpha}(z^-)$ across the boundary layer (i.e. between $0<z^-<1$)
must fulfill the drag law and boundary conditions;
at the same time, the weak dependence of $\alpha$ and $u_\star$ on $Re$ implies this range depends on $Re$.
%
%%%PAR
%
\paragraph{Outer Layer.} In the outer region of the flow (for $z^-\mapsto 1$), $f_V(z^-)$, should govern the
spanwise velocity profile which indeed is supported by our DNS data (Fig.\ref{fig:outer_w}):
%
The deficit is close to zero for $z^-\gtrsim 1$, and can be approximated by a logarithmic profile in the
range $0.15\lesssim z^- \lesssim 0.4$.
%
The slope of this profile is best fit by $\kappa_V=0.836 (1 + 150 / Re_D) \kappa$, where the correction term for low Reynolds number
may be omitted for $Re_D \gg 1000$.
%
Based on Fig.~\ref{fig:outer_w}, we pick the offset as
a constant fraction (0.906) of $V^\alpha_G=Z_\star \sin\alpha$, i.e
\begin{subequations} 
\begin{align}
  V_\text{log}(z^-)  = \frac{1}{\kappa_V} \log\left(\frac{z^-}{0.4}\right) + 0.906V_G^{\alpha}. 
\end{align} 
%
%%%PAR
% 
The eventually converged simulations at intermediate $Re_D$ (where the DNS could be run over several inertial periods)
exhibit a super-geostrophic maximum (positiv velocity deficit) in the span-wise component.
%
This maximum is not present or less pronounced for the simulations at higher $Re$ where closer analysis reveals that the
velocity profiles in the outer layer are still subject to a small though finite adaptation process on the inertial time
scales $2\pi/f$ due to the relatively short time span for which these cases could be simulated.
%
While there is no physical model for this wake region that leads to an analytical velocity profile formulation in the span-wise direction,
an error-function transition from the outer logarithmic law to the boundary condition provides a reasonable fit:
%
\begin{align}
  V_\text{outer}(z^-) = (1-w_V)V_\text{log}(z^-) + w_V V_G^\alpha
\end{align} 
with
\begin{align} 
w_V = \frac{\mathrm{erf}\left[\xi_\text{outer} \log(z/z_\text{ref})\right]  + 1}{2}
\end{align}
where $\xi_\text{outer}$ is a transition scale that defines the width of the wake region and $z_\text{ref}$ defines the height.
Based on the DNS data, we choose $\xi_\text{outer}=2$ and let $z_\text{ref}$ such that $V_\text{log}(z_\text{ref}) = 0.98V_G^\alpha$.
There is, however, a degree of ambiguity in the exact choice of these parameters due to the imperfect convergence of the
DNS data to the statistical equilibrium.
%
\end{subequations} 
%
%%% PAR
%
\paragraph{Inner Layer.} It is important to note that the logarithmic layer observed here for the spanwise velocity does not
conincide with the log-region of the streamwise velocity but is rather a 'continuation' through the outer layer.
%
This is consistent with the recent finding of a second log-region in the outer region of Ekman flow
(... et al. (incl. Sullivan), JAS 2019).
%
Below this region, the gradients in span-wise velocity are rather small and the span-wise velocity
monotonically approaches its surface boundary condition $V(z=0)=0$.
%
While the stream-wise velocity follows a universal inner scaling that has acquired its universal, $Re$-independent shape for $Re_D> \mathcal{O}\left(10^3\right)$, the span-wise component that defines how the velocity vector veers when the surface is approached,
does not collapse in inner units, and there is, most importantly no sign of convergence even at the highest Reynolds numbers for
which simulations where carried out.
%
Even though the simplest assumption $V=0$ is reasonable for the lower part of the surface surface layer ($z^-<10^{-3}$), it does
not appropriately capture the profile in the rest of the surface layer:

First, $V=0$  implies a discontinuity in the velocity profile at $z^-=0.1$, where the outer scaling found above 
yields a finite value at geophysical $Re$, i.e. there is non-zero veering in the upper part of the
surface layer--as is well-known also from field observation. 
%
Second, the layer around $z^-=0.1$ is crucial to obtain the characteristic and well-established shape of the hodographs as the layer
where $V$ sets in marks the 'maximum' of $V^-$ vs. $U^-$.
%
This illustrates that the higher-order (in terms of the channel-flow analogy where $V=0$) correction follows a mixed scaling when the surface is approached which has been demonstrated for higher-order terms in other types of flow \citep{Mellado:2016hv}. 
%
\par
%
The scale for the magnitude of the span-wise velocity component is $u_\star\sin\alpha$. 
%
Based on our DNS data, we suggest that the Reynolds number scaling of this velocity-magnitude scale is captured
by $Re_\tau^{-1/2}$ which is indeed known from the generalization of higher-order statistics, such as turbulent fluxes
in the inner layer \citep{Marusic:2013hf} that also follow a mixed scaling in the inner layer.
%
We then parameterize the spanwise velocity at 10 wall units as anchor point in the inner layer:
\begin{subequations}
\begin{align}
  V_{10}\equiv V(z^+=10)=750 \frac{u_\star\sin\alpha}{\sqrt{Re_\tau}}.
\end{align} 
%
This leaves us with three fixed points of the velocity profile in the inner layer, namely
(i) the boundary condition $V_0=0$,
(ii) $V_{10}$ at $z^+=10$, and 
(iii) the lower end of the logarithmic profile at $z^-=0.1$ where the latter two are semi-empirically estimated
from DNS data. 
%
In absence of well-established scaling considerations for the span-wise velocity, the choice of profile fits joining
these three points is indeed arbitrary, but we can resort to the DNS data for an empirical approach and find
that a square-root profile fits $V(z^+)$ in the surface layer.
%
A linear profile for $V$ is employed in the viscous sub-layer.
%
Based on the physical extent of the viscous sub-layer in Ekman flow around five wall units \citep{Foken:2002gv, Ansorge:2019di},
we choose $z^+=5$ to transition from one to the other and note that $V$ is already very small at this height.
%
The span-wise velocity profile in the surface layer is then estimated as 
\begin{align}
  \left.V(z^+)\right|_{\text{inner}} = \left\{\begin{array}{c c l}
  a_1 z^+&;& z^+ \le 5\\ 
  b_1+b_2 \sqrt{z^+}&;& 5<z^+<Re_\tau/10 
  \label{eqn:w_profile_inner}
  \end{array} \right, 
\end{align}
with $b_1$ and $b_2$ estimated such that
\begin{eqnarray}
  \begin{array}{r c l} 
    \left.V(z^+=10)\right|_\text{inner}&=&V_{10}\\
    \left.V(z^+=Re_\tau/10)\right|_\text{inner}&=&V_{\text{outer}}(0.1)\\
  \end{array}
  \Rightarrow \left\{ \begin{array}{l}
    b_2=\dfrac{V_\text{outer}(0.1) - V_{10}}{\sqrt{Re_\tau/10}-\sqrt{10}}  \\[1.5em]
    b_1= V_{10} - \sqrt{10}b_2
  \end{array}\right. 
\end{eqnarray}
We then estimate $\alpha$ from the matching condition at $z^+=5$, i.e.
\begin{align}
  5a = b_1+\sqrt{5}b_2 \Rightarrow a = \frac15 \left[V_{10} + (\sqrt{5}- \sqrt{10})\left(\frac{V_\text{outer}-V_{10}}{\sqrt{Re_\tau/10}-\sqrt{10}}\right) \right].
  \label{eqn:alpha_coefficient} 
\end{align}
%
%%%PAR
%
\paragraph{Matching region.} While the profile composed of $V_\text{inner}(z^+\le 0.1 Re_\tau)$, $V_\text{outer}(z^->0.1)$ is continuous,
it is not smooth at $z^-=0.1$, i.e. at the transition from power-law ($V\propto \sqrt{z^+}$) to logarithmic scaling.
%
To alleviate this issue, we use a second-order polynomial for transition from the inner to the outer layer in the range
$z_\text{low}<z<z_\text{up}$ such that
\begin{align} 
  V_\text{trans}(z^-) &=& V_\text{inner}(z^+_\text{low}) + \Delta V \left(a z_{arg} + b (z_{arg})^2\right)
\end{align}
with $\Delta V &=& V_\text{outer}(z^-_\text{up})-V_\text{inner}(z^+_\text{low})$
and  $z_{arg} = (z - z_\text{low})/(z_\text{up}-z_\text{low})$.
It is  $a+b=1$ for $V_\text{trans}(z^-_\text{up}) = V_\text{outer}(z^-_\text{up})$, and we constrain $a$ by
\begin{align}
  \left.\frac{\partial V_\text{trans}}{\partial z^-}\right|_{z=z_\text{low}} = \left.\frac{\partial V_\text{inner}}{\partial z^-}\right|_{z=z_\text{low}},
\end{align}
where we find that $z^-_\text{low}=0.06$ and $z^-_\text{up}=0.13$ yield satisfactory agreement with DNS data.

%
\begin{figure}
  \includegraphics[width=1.0\textwidth]{w_profile.pdf}
  \caption{Profiles of shear-aligned span-wise velocity $(W^\alpha)^+$ versus inner and outer height.
  Dashed lines show DNS data, thick, opaque lines are from the semi-emprical theory developed above. }
  \label{fig:outer_w} 
\end{figure}
\end{subequations}

\begin{figure}
  \phantom{AAA}\textbf{(a)\hspace{0.3\textwidth}(b)\hspace{0.3\textwidth}(c)}\\
  \includegraphics[width=\textwidth]{ekman_profiles_data.pdf}
  \caption{ \textbf{(a)} Velocity deficity, \textbf{(b)} velocity profile in shear-aligned hodographs and \textbf{(c)} hodograph in geostrophy-aligned coordinates.
    Thick, solid lines show theory, dashed lines data from DNS. }
  \label{fig:hodograph} 
\end{figure}

\section{Comparison with other theories}

\textbf{Compare} to Emeis (2002) and Gryning (2007); highlight explicit knowledge on veering-profile $\rightarrow$ directional sheer;
%
\par%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
Implications for \textbf{K-theory} (we now can consider that shear and stress are not necessarily perfectly aligned).
$\rightarrow$ can we do something to infer a K-profile from these theoretical considerations? 
%
\begin{figure}
  \includegraphics[width=\textwidth]{mag_dir.pdf} 
  \caption{Total velocity and veering (in degrees) vs inner and outer height.
    Dashed lines show DNS data, thick lines are from semi-empirical theory.} 
\end{figure} 
%
\section{Conclusions}
%
Applications:
\begin{itemize}
\item reference-shear for neutral profile approaches(systematic!) $\rightarrow$ wind engineering! 
\item initial condition for LES/DNS to eliminate/minimize inertial oscillation in Benchmark simulations
\item 
\bibliographystyle{plainnat}
\bibliography{ansorge.bib}
\end{document}

### 2020-12-08 write/stress_rotation.tex
initial commit

\documentclass[a4paper,12pt]{article}

\usepackage{amsmath} 
\usepackage{amssymb}

\title{Rotation of Reynolds stress tensor [for shear-alignment]} 
\begin{document} 
Let $u$, $v$ and $w$ (and the corresponding Reynolds stress tensor) 
be defined on a grid $Ox \times Oy \times Oz$ where the unit vector 
$\hat{e}_x$ is turned by $\theta$ with respect to the [negative] 
direction of surface shear stress \footnote{Here, negative means that we 
consider the grid aligned with the positive vector of the velocity gradient 
at the surface}
We require the velocities on the grid $Ox_\star \times Oy \times Oz_\star$ 
that is aligned with the surface shear stress, i.e. 
\begin{align}  
  \left(\begin{array}{c}
    \hat{e}_{x_\star}\\
    \hat{e}_y\\
    \hat{e}_{z_\star}
  \end{array} \right) = \left(\begin{array}{ccc}
    \cos\theta& 0 &-\sin\theta \\ 
    0 & 1 & 0 \\
    \sin\theta & 0 & \cos\theta 
  \end{array} \right)\left(\begin{array}{c} 
    \hat{e}_x \\ \hat{e}_y \\ \hat{e}_z
   \end{array}\right). 
\end{align} 
This implies 
\begin{eqnarray*} 
  u_\theta&=&u\cos\theta  +w\sin\theta \\ 
  w_\theta&=&-u\sin\theta +w\cos\theta. 
\end{eqnarray*}  
Let $c_\theta\equiv \cos\theta$, $s_\theta\equiv \sin_\theta$ and $m_\theta=\sin\theta\cos\theta$.
The corresponding Reynolds stress tensor is 
\begin{align} 
  \left(\begin{array}{ccc} 
    u_\theta u_\theta & u_\theta v_\theta & u_\theta w_\theta \\ 
    u_\theta v_\theta & v_\theta v_\theta & v_\theta w_\theta \\ 
    u_\theta w_\theta & v_\theta w_\theta & w_\theta w_\theta 
  \end{array} \right) =
  \left(\begin{array}{ccc} 
    (u c_\theta + w s_\theta )^2&
    uvc_\theta +wvs_\theta&
    (uc_\theta +ws_\theta)(-us_\theta +wc_\theta)\\
    & vv & (-us_\theta+wc_\theta)v\\ 
    && (-us_\theta +w c_\theta)^2 \end{array}\right)\\
  = \left(\begin{array}{ccc}  
    u^2c_\theta^2 +w^2 s_\theta^2 + 2uw m_\theta&
    uvc_\theta +wvs_\theta&
    m_\theta (w^2-u^2)+uw(c_\theta^2 -s_\theta^2) \\
    &v^2& -uvs_\theta + wvc_\theta\\
    && u^2s_\theta^2 + w^2 c_\theta^2 - 2uw m_\theta
    \end{array}\right) \\ 
  = \left(\begin{array}{ccc} 
    u^2c_\theta +w^2s_\theta
  \end{array}
\end{align} 
\end{document}

### 2020-12-08 plot/EkmanProfileClass.py

@@ -1,0 +1,380 @@
+#!/opt/local/bin/python 
+import my_pylib as mp 
+import numpy as np
+import scipy as sp 
+from scipy.special import lambertw as prodlog
+
+def build_grid(trgt,thick=1) :
+    vert_stretch=1.05   #grid stretching
+    dyloc=1             #grid resolution at wall
+    ny=1
+    yloc=dyloc
+    while yloc < thick*trgt:
+        dyloc*=vert_stretch 
+        yloc+=dyloc
+        ny+=1
+        
+    yp=np.zeros(ny)
+
+    dyp=1
+    yp[0]=0
+    for i in range(ny-1):
+        yp[i+1]=yp[i]+dyp
+        dyp*=vert_stretch 
+    ym=yp/trgt
+
+    return ny,yp,ym
+
+
+class Complex:
+    def __init__(self, real, imag=0.0):
+        self.r = real
+        self.i = imag
+
+    def __add__(self, other):
+        return Complex(self.r + other.r,
+                       self.i + other.i)
+
+    def __sub__(self, other):
+        return Complex(self.r - other.r,
+                       self.i - other.i)
+
+    def __mul__(self, other):
+        return Complex(self.r*other.r - self.i*other.i,
+                       self.i*other.r + self.r*other.i)
+
+    def __div__(self, other):
+        sr, si, ot, oi = self.r, self.i,other.r, other.i # short forms
+        r = float(ot**2 + oi**2)
+        return Complex((sr*ot+si*oi)/r, (si*ot-sr*oi)/r)
+
+    def __abs__(self):
+        return sqrt(self.r**2 + self.i**2)
+
+    def __neg__(self):   # defines -c (c is Complex)
+        return Complex(-self.r, -self.i)
+
+    def __eq__(self, other):
+        return self.r == other.r and self.i == other.i
+
+    def __ne__(self, other):
+        return not self.__eq__(other)
+        
+    def __str__(self):
+        return '(%g + %g i)' % (self.r, self.i)
+
+    def __repr__(self):
+        return 'Complex' + str(self)
+    
+class EkmanUniversalClass:
+    
+    def __init__(self):
[…]

### 2020-12-08 plot/my_pylib.py

@@ -1,0 +1,1 @@
+/Users/zrick/CODE/PYTHON/my_pylib.py

### 2020-12-08 plot/plot.py

@@ -1,0 +1,665 @@
+#!/opt/local/bin/python
+import my_pylib as mp 
+import numpy as np
+import scipy.interpolate as spip
+import argparse 
+import matplotlib.pyplot as plt
+import sys
+import warnings
+from scipy.special import lambertw as prodlog
+import scipy as sp
+from netCDF4 import Dataset
+from math import erf
+
+from EkmanProfileClass import Complex, EkmanUniversalClass, build_grid
+
+D2R = np.pi / 180. 
+
+use_data=True   # if true -- the netcdf files from DNS are needed 
+plot_ustar_alpha=False
+plot_summary=False
+plot_visc=True
+print_table=False
+plot_outer_log=False
+plot_total_rot=False
+plot_convergence_1000=True
+
+colors = {400 : 'gray',
+          500 : 'pink',
+          750 : 'cyan',
+          1000: 'blue',
+          1300: 'red',
+          1600: 'black',
+          5000: 'black',
+          10000:'gray',
+          30000:'black',
+          150000:'black',
+          300000:'black'} 
+
+base='/Users/zrick/WORK/research_projects/SIM_NEUTRAL/netcdf/' 
+
+files = { 500: base+'avg_flw_ri0.0_re500.0_co0.0_512x192x1024.nc',
+          750: base+'avg_flw_ri0.0_re750.0_3072x384x3072.nc',
+          1000:base+'avg_flw_ri0.0_re1000.0_co0.0_1536x512x3072.nc',
+          1300:base+'avg_flw_ri0.0_re1300.0_co0_2560x640x5120.nc',
+          1600:base+'avg_flw_ri0.0_re1600.0_co0_3840x960x7680.nc' } 
+
+sc=EkmanUniversalClass() # yp_ref,up_ref,wp_ref,deltap_ref,us_ref,al_ref*D2R-geo_rotate,plot=False)
+
+if use_data: 
+    re_ref=1600
+    nu_ref=2./(re_ref**2)
+
[…]

### 2020-12-08 plot/prepare.py

@@ -1,0 +1,180 @@
+#!/opt/local/bin/python
+
+from netCDF4 import Dataset
+import numpy as np 
+import matplotlib.pyplot as plt 
+import my_pylib as mp
+
+f_coriolis=1.0    # Coriolis parameter = 1
+U_G=1.0           # Geostrophic wind = 1 
+                  # => nu=1/re
+
+path='/Users/zrick/WORK/research_projects/SIM_NEUTRAL/netcdf/' 
+cases=[400,500,750,1000,1300,1600]
+
+files={ 400: 'avg_flw_ri0.0_re400.0_1024x192x1024.nc' , 
+        500: 'avg_flw_ri0.0_re500.0_2048x192x2048.nc' , 
+        750: 'avg_flw_ri0.0_re750.0_3072x384x3072.nc' ,
+        1000:'avg_flw_ri0.0_re1000.0_co0_3072x512x6144.nc' ,
+        1300:'avg_flw_ri0.0_re1300.0_co0_2560x640x5120.nc' ,
+        1600:'avg_flw_ri0.0_re1600.0_co0_3840x960x7680.nc'} 
+
+it_start= { 
+    400:100,
+    500:200,
+    750:0,
+    1000:100,
+    1300:120,
+    1600:90       
+    }
+
+angle={400:0.26177, 500:0., 750:0.35779, 1000:0.35779, 1300:0.35779, 1600:0.35779}
+color={400:'grey', 500:'blue', 750: 'purple',  1000:'red', 1300:'orange', 1600:'green'}
+ustar={} 
+alpha={} 
+retau={}
+nu={} 
+reg={}
+
+# check tensor rotation
+check_tensor=0
+if check_tensor != 0 :
+    a1=np.zeros([3,3])
+    print(a1) 
+    b1=mp.rotate_tensor(a1,1.0) 
+    print(b1)
+    print('=====')
+
+
+    a2=np.zeros([3,3]) 
+    a2[1,1]=1.0
+    print(a2)
+    b2=mp.rotate_tensor(a2,1.0) 
+    print(b2)
+    print('======')
+    
+    a3=np.zeros([3,3]) 
+    a3[0,0]=1.0
+    a3[1,1]=5.0 
+    a3[2,2]=1.0
+    a3[0,1]=0.1
+    a3[0,2]=0.5
+    a3[1,2]=0.0
+    a3[2,0]=0.0
+    print(a3)
+    b3=mp.rotate_tensor(a3,np.pi/4)
+    print(b3)
+    print('======')
+
[…]

### 2020-12-08 plot/replace_mean.py

@@ -1,0 +1,66 @@
+zaza#!/opt/local/bin/python 
+
+import numpy as np
+import my_pylib as mp
+import matplotlib.pyplot as plt
+from netCDF4 import Dataset 
+
+D2R = np.pi/180. 
+
+file='/Users/zrick/WORK/research_projects/SIM_NEUTRAL/netcdf/avg_flw_ri0.0_re1000.0_co0.0_1536x512x3072.nc'
+f=Dataset(file,'r')
+
+t=f['t'][:]
+i0=mp.geti(t,t[-1]-2*np.pi) -1
+
+u=f['rU'][i0:,:]
+w=f['rW'][i0:,:]
+al=f['FrictionAngle'][i0:]
+us=f['FrictionVelocity'][i0:] 
+t=f['t'][i0:]
+it=f['it'][i0:] 
+tke=f['Tke'][i0:,:]
+y=f['y'][:] 
+
+t_delta = t[-1]-t[0]- 2*np.pi
+
+al_mean=np.mean(al) 
+print('DIFFERENCE TO EXACT PREIODICITY:',t_delta/np.pi/2)
+print('MEAN VEERING:', al_mean,'deg / ', al_mean*D2R, 'rad' ) 
+
+tke_t=tke[10:,:]-tke[:-10,:]
+u_t = u[8:,:] - u[:-8,:]
+u_avg=np.mean(u,0)
+w_avg=np.mean(w,0)
+
+[u_rot,w_rot]= mp.rotate(u_avg,w_avg,0) #al_mean*D2R) 
+
+print('w1-w0 ROTATED:', w_rot[1]-w_rot[0])
+print('w1-w0 ORIGINAL:',w_avg[1]-w_avg[0])
+print('===')
+print('GEO rotation orig:',np.arctan(w_avg[-1]/u_avg[-1]))
+print('GEO rotation rotd:',np.arctan(w_rot[-1]/u_rot[-1])) 
+
+u_delta=u_rot-u[-1,:]
+w_delta=w_rot-w[-1,:]
+
+u_rotn=u_rot-u_avg
+w_rotn=w_rot-w_avg 
+
+#plt.plot(y,u_delta,label=r'$\Delta u$',ls='-',lw=1,c='red')
+#plt.plot(y,u_rotn,label=r'$\Delta u (rotate)$', ls=':',lw=1,c='red')
[…]

### 2020-12-08 plot/write_delta.py

@@ -1,0 +1,38 @@
+#!/opt/local/bin/python
+
+import numpy as np 
+
+
+def write_header(h,nx,ny,nz,it,t,re,fr,sc):
+    t_i4=np.dtype('<i4')
+    t_f8=np.dtype('<f8')
+    ipar=[52,nx,ny,nz,it]
+    fpar=[t,re,fr,sc] 
+    np.tofile(h,ipar,t_i4,count=5)
+    np.tofile(h,fpar,t_f8,count=4)
+    np.tofile(h,[52],t_i4,count=1)
+
+nx=10
+ny=512
+nz=10
+
+
+ih=open('delta_velo.dat','r')
+ou=open('delta_u.bin','ab')
+ov=open('delta_v.bin','ab') 
+ow=open('delta_w.bin','ab') 
+
+arr=np.zeros([3,ny,nx]) 
+
+i=0;
+for line in ih:
+    dat=line.split()[3:5]
+    arr[0,i,:] = dat[0]
+    arr[1,i,:] = 0.
+    arr[2,i,:] = dat[1] 
+    i=i+1 
+
+
+
+
+write_header(ou,0 )

### 2020-12-21 plot/EkmanProfileClass.py
updated outer-layer profile for stream-wise shear-aligned component

@@ -179,7 +179,7 @@ class EkmanUniversalClass:
                 break;
             i=i+1
 
-        p=p-2*C5/(re**2) * (z**2)
+        p=p-1.20*C5/(re**2) * (z**2)
             
         return(1./z,p) 
 
@@ -227,19 +227,25 @@ class EkmanUniversalClass:
 
 
         # outer-layer deficit [in outer units] 
-        scale_trans=1.20*1 #0.0495/us_loc #(0.0497)**(0.0)
-        ctr=0.321* 1#$ 0.0495/us_loc #(0.0495)**(0.0)
+        scale_trans=2#1.20*1 #0.0495/us_loc #(0.0497)**(0.0)
+        ctr= 0.24 - (70./re)
         arg=np.zeros(len(ym))
         arg[1:]=np.log(ym[1:]/ctr)
-        delta_log=z * np.cos(al_loc) 
-        us_ref,al_ref=self.ustar_alpha(1600) 
+        delta_log=z * np.cos(al_loc)
 
+        argz=1.4*np.pi*(ym+0.13)
+        damp=0.53
+        outer_u= (1-damp*np.cos(argz)*np.exp(-argz))/us_loc
+        outer_w=  damp*np.sin(argz)*np.exp(-argz)/us_loc
+        [ous,ows]=mp.rotate(outer_u,outer_w,al_loc)
+        
         wgt=(sp.special.erf(scale_trans*arg)+1)/2
 
-        fit2= ( wgt*(log_law-delta_log) ) # + 0*fit1 * (0.0497/us_loc)**2 ) 
+        fit2= ( wgt*(log_law-delta_log) ) # + 0*fit1 * (0.0497/us_loc)**2 )
+        fit2= ( wgt*(log_law-ous) ) 
         u[1:]-= fit2[1:] #* diff_u/dev_u
-        u-= 0.25*np.exp( -5.* ( (yp/dp_loc-0.85)**2) )
-
+        #u-= 0.25*np.exp( -5.* ( (yp/dp_loc-0.85)**2) )
+        #u[1:]=ous[1:]
         ################################################################################
         # OLD part for shear-spanwise velocity
         ################################################################################

### 2020-12-21 plot/plot.py

@@ -15,14 +15,14 @@ from EkmanProfileClass import Complex, EkmanUniversalClass, build_grid
 
 D2R = np.pi / 180. 
 
-use_data=True   # if true -- the netcdf files from DNS are needed 
-plot_ustar_alpha=False
-plot_summary=False
+use_data=False   # if true -- the netcdf files from DNS are needed 
+plot_ustar_alpha=True
+plot_summary=True
 plot_visc=True
-print_table=False
-plot_outer_log=False
-plot_total_rot=False
-plot_convergence_1000=True
+print_table=True
+plot_outer_log=True
+plot_total_rot=True
+plot_convergence_1000=False
 
 colors = {400 : 'gray',
           500 : 'pink',
@@ -38,7 +38,8 @@ colors = {400 : 'gray',
 
 base='/Users/zrick/WORK/research_projects/SIM_NEUTRAL/netcdf/' 
 
-files = { 500: base+'avg_flw_ri0.0_re500.0_co0.0_512x192x1024.nc',
+files = { 400: base+'avg_flw_ri0.0_re400.0_1024x384x1024.nc',
+          500: base+'avg_flw_ri0.0_re500.0_2048x192x2048.nc',
           750: base+'avg_flw_ri0.0_re750.0_3072x384x3072.nc',
           1000:base+'avg_flw_ri0.0_re1000.0_co0.0_1536x512x3072.nc',
           1300:base+'avg_flw_ri0.0_re1300.0_co0_2560x640x5120.nc',
@@ -87,7 +88,7 @@ if plot_visc:
     ax=fig.add_axes([0.1,0.1,0.87,0.8])
 
     fig2=plt.figure(figsize=(5,4))
-    ax2=fig2.add_axes([0.11,0.1,0.87,0.8]) 
+    ax2=fig2.add_axes([0.13,0.1,0.85,0.87]) 
     
     for re in [1600,1300,1000,750,500]:
         c=colors[re]
@@ -141,22 +142,21 @@ if plot_visc:
     fig.savefig('visc_layer.pdf',format='pdf' )
     plt.close('all')
 
-    # add high-re curve to outer_layer.pdf
-    re=3e5
-    nu=2./(re*re)
-    us,al=sc.ustar_alpha(re)
-    deltap=us*us/nu 
[…]

### 2020-12-31 write/manuscript.tex
updated outer-layer profile for stream-wise shear-aligned component
ekman-profile for u,w in outer layer; modifications in inner w profile

@@
 \usepackage{natbib}
+
+\newcommand{\p}{\partial}
+\newcommand{\gzm}{{\xi}}%{\gamma z_\text{match}}}
 
@@
 \begin{figure}
-  \centerline{\includegraphics[width=0.7\textwidth]{ustar_alpha.pdf}}
+  \centerline{\includegraphics[width=0.7\textwidth]{../plot/ustar_alpha.pdf}}
   \caption{Variation of $Z_\star$ and surface veering with Reynolds number according to
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{u_profile.pdf}
+  \includegraphics[width=\textwidth]{../plot/u_profile.pdf}
   \caption{Shear-aligned profiles of velocity components $U^{\alpha+}$ in inner (left) and outer (right) units.} 
@@
 \begin{figure}
-  \includegraphics[width=1.0\textwidth]{w_profile.pdf}
+  \includegraphics[width=1.0\textwidth]{../plot/w_profile.pdf}
   \caption{Profiles of shear-aligned span-wise velocity $(W^\alpha)^+$ versus inner and outer height.
@@
   \phantom{AAA}\textbf{(a)\hspace{0.3\textwidth}(b)\hspace{0.3\textwidth}(c)}\\
-  \includegraphics[width=\textwidth]{ekman_profiles_data.pdf}
+  \includegraphics[width=\textwidth]{../plot/ekman_profiles_data.pdf}
   \caption{ \textbf{(a)} Velocity deficity, \textbf{(b)} velocity profile in shear-aligned hodographs and \textbf{(c)} hodograph in geostrophy-aligned coordinates.
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{mag_dir.pdf} 
+  \includegraphics[width=\textwidth]{../plot/mag_dir.pdf} 
   \caption{Total velocity and veering (in degrees) vs inner and outer height.
@@
 \item
+\end{itemize}
+
+\appendix
+
+\section{Laminar Ekman solution with consideration of inner layer}
+
+\begin{subequations}
+\begin{eqnarray}
+  \left(\begin{matrix}
+    \p_t U\\
+    \p_t W  
+  \end{matrix}\right)&=&\left(\begin{matrix}
+     fW &+& \nu \p_z^2 U\\ 
+    -f(U-G) &+& \nu \p_z^2 W
+  \end{matrix}\right)\\ 
+  \Rightarrow \partial_t (U+iW) &=& f(W-i(U-G))) + \nu \p_z^2(U+iW)
+\end{eqnarray}
+In stationary conditions, this system is solved by
+\begin{eqnarray}
+  \hat{u}(z) &=& U_{\inf} + e^{-\gamma z} \left[A \cos\gamma z + B \sin\gamma z\right]   \\
+  \hat{w}(z) &=& W_{\inf} + e^{-\gamma z} \left[-A \sin\gamma z + B \cos\gamma z\right]
+\end{eqnarray} 
+\end{subequations}
+where the constants $U_\inf$, $W_\inf$ set the top boundary condition and $A$ and $B$ set the bottom boundary condition. 
+%
+The most common boundary condition for a surface Ekman layer is $A=U_\inf=G$, $B=0$, and $W_\inf=0$.
+%
+The lower boundary condition, however, neglects the existence of the surface layer, and it appears reasonable to define
+$A=c G$ where $c<1$ is a constant that icorporates the increased shear in the surface layer.
+%
+Given a 'matching height' $z_{match}$ and normalized matching height $\xi=\gamma z_\text{match}$ in the upper part of the inner layer, we can match the Ekman profile
+to the inner layer by letting
+%
+\begin{eqnarray}
+  \begin{matrix} 
+  u(z_\text{match}) \equiv u_\text{match} &=& U_{\inf} + e^{-\gzm}} \left [ A \cos\gzm + B\sin\gzm \right] \\
+    w(z_\text{match}) \equiv w_\text{match} &=& W_{\inf} + e^{-\gzm}} \left [-A \sin\gzm + B\cos\gzm \right]
+  \end{matrix} 
+\end{eqnarray}
+\begin{eqnarray}
+  e^\gzm  \left(\begin{matrix}
+     u_\text{match}-U_{\inf} \\ 
+     w_\text{match}-W_{\inf}
+  \end{matrix}\right) = \left(\begin{matrix}
+  \cos\gzm A &+ \sin\gzm B \\ -\sin\gzm A &+ \cos\gzm B
+  \end{matrix} \right) \\
+\end{eqnarray}
+Matching the profile at $\xi=0$, one obtains
+$A = \Delta u_\text{match} $ and $B=-\Delta w_\text{match} $. Otherwise, it is
+\begin{subequations} 
+\begin{eqnarray}
+  e^\gzm \Delta u_\text{match} = A \cos\gzm + B\sin\gzm \\
+  e^\gzm (\cos\gzm \Delta u_\text{match} - \sin\gzm \Delta w_\text{match}) = A (\cos^2\gzm + \sin^2\gzm)
+\end{eqnarray}
+\begin{eqnarray}
+  \Rightarrow A &=&e^\gzm (\cos\gzm\Delta u_\text{match} + \sin\gzm \Delta w_\text{match})\\
+  B&=&  \frac{e^\gzm \Delta u_\text{match} - A \cos \gzm}{\sin\gzm} \\
+  &=& e^\gzm \frac{ \sin^2\gzm \Delta u_\text{match}+\sin\xi\cos\xi \Delta w_\text{match} }{\sin\xi }  \\
+  &=& e^\gzm (\sin\xi\Delta u_\text{match} + \cos\xi\Delta w_\text{match}) 
+\end{eqnarray}
+\end{subequations} 
 \bibliographystyle{plainnat}
 \bibliography{ansorge.bib}
+
 \end{document}

### 2020-12-31 plot/EkmanProfileClass.py

@@ -133,7 +133,7 @@ class EkmanUniversalClass:
         A_i=np.average(f1.i[i1-2:i1+2])
         self.A=Complex(A_r,A_i) 
         self.yp=yp
-        print(self.A) 
+        #print(self.A) 
         # integrate constants C1 and C5 for higher-order theory
         # C1 = int_0^{\infty} f_1(y^- )dy^-
         C1_r=mp.integrate(yp/deltap,f1.r,len(yp))
@@ -143,7 +143,7 @@ class EkmanUniversalClass:
         # C5 = int_0^{50} f_5(y+)dy^+
         imax=mp.geti(yp,50)
         self.C5=Complex(mp.integrate(yp,f5.r,imax),mp.integrate(yp,f5.i,imax))
-        print(self.C5) 
+        #print(self.C5) 
 
         self.initialized=True
 
@@ -204,8 +204,22 @@ class EkmanUniversalClass:
 
         ym=np.array(yp/dp_loc)
 
-        u=np.zeros(len(yp))
+        # OUTER REFERENCE PROFILES
+        #
+        argz=0.66*(2.*np.pi*(ym+0.12)) 
+        dampA =  0.42*(us_loc/0.05- 0.07/re) #np.exp(zdum) * ( dU_mtc*np.cos(zdum) + dW_mtc*np.sin(zdum) )
+        dampB =  0.0#np.exp(zdum) * ( dU_mtc*np.sin(zdum) + dW_mtc*np.cos(zdum) )
+
+        outer_u= (1-(dampA*np.cos(argz)+dampB*np.sin(argz))*np.exp(-argz))/us_loc
+        outer_w=   ( dampA*np.sin(argz)-dampB*np.cos(argz))*np.exp(-argz)/us_loc
+        [ous,ows]=mp.rotate(outer_u,outer_w,al_loc)
         
+
+        ################################################################################
+        # STREAMWISE COMPONENT (SHEAR-ALGINED)
+        ################################################################################
+
+        u=np.zeros(len(yp))
         i_x=mp.geti(yp,self.LIMIT_INNER)
 
         # start with logarithmic law (everywhere)
@@ -219,117 +233,68 @@ class EkmanUniversalClass:
         scale=2.03
         u_ref = 1/0.07825
         ymatch = 19 
-        w=np.zeros(len(yp))
[…]

### 2020-12-31 plot/plot.py

@@ -117,12 +117,16 @@ if plot_visc:
         ax.plot(yp_dat,u_sfc/us_dat, c=c,lw=3,ls='-',label=r'$Re={}$'.format(re),alpha=0.5)
         ax.plot(yp,up,               c=c,lw=3,ls=':')
 
-        if  re ==1600: 
+        if  re ==1600:
             ax2.plot(ym_dat,(u_sfc-u_sfc[-1])/us_dat,c=c,lw=3,ls='-',label=r'$Re={}$ (data)'.format(re),alpha=0.5)
+            ax2.plot(ym_dat,(w_sfc-w_sfc[-1])/us_dat,c=c,lw=3,ls=':',alpha=0.5)
             ax2.plot(ym,(up-up[-1]),c=c,lw=1,ls='--',label=r'$Re={}$ (model)'.format(re))
+            ax2.plot(ym,(wp-wp[-1]),c=c,lw=1,ls=':',label=r'$Re={}$ (model)'.format(re))
         else:
             ax2.plot(ym_dat,(u_sfc-u_sfc[-1])/us_dat,c=c,lw=3,ls='-',label=r'$Re={}$'.format(re),alpha=0.5)
+            ax2.plot(ym_dat,(w_sfc-w_sfc[-1])/us_dat,c=c,lw=3,ls=':',alpha=0.5)
             ax2.plot(ym,(up-up[-1]),c=c,lw=1,ls='--')
+            ax2.plot(ym,(wp-wp[-1]),c=c,lw=1,ls=':')
 
 
             
@@ -142,19 +146,9 @@ if plot_visc:
     fig.savefig('visc_layer.pdf',format='pdf' )
     plt.close('all')
 
-    us,al=sc.ustar_alpha(1000)
-    argz=1.4*np.pi*(ym+0.14)
-    damp=np.sin(al)*1.7
-    outer_u = (1-damp*np.cos(argz)*np.exp(-argz))/us
-    outer_w =  damp*np.sin(argz)*np.exp(-argz)/us
-    [ous,ows]=mp.rotate(outer_u,outer_w,al) 
-
-    ax2.plot(ym,ous-ous[-1],c='black',ls=':',lw=0.5)
-    ax2.plot(ym,ows-ows[-1],c='black',ls='-',lw=0.5)
-    ax2.plot(ym,0*ym,ls='-',lw=0.5,c='black')
     
     ax2.set_xlim(0,1.5)
[…]

### 2020-12-31 plot/EkmanProfileClass.py
updated plots and changed pre-factor of C5 in ustar_alpha of EkmanProfileClass.py to 1.00

@@ -179,7 +179,7 @@ class EkmanUniversalClass:
                 break;
             i=i+1
 
-        p=p-1.20*C5/(re**2) * (z**2)
+        p=p-1.00*C5/(re**2) * (z**2)
             
         return(1./z,p) 
 

### 2021-03-03 write/manuscript.tex
reworked w-profile to eliminate sharp transition in surface vicinity for high Reynolds number

@@
 \documentclass[a4paper,11pt]{article}
+
 \usepackage[margin=2.5cm]{geometry}
@@
 \usepackage{natbib}
+\usepackage{xcolor}
 
@@
 \newcommand{\gzm}{{\xi}}%{\gamma z_\text{match}}}
+\newcommand{\todo}[1]{\textcolor{red}{$[$#1$]$}}
 
@@
   %
-  In both profiles, the wake region can be described by smooth error-function transition from a logarithmic layer
-  to the constant outer velocity.
-
+  In the outer layer, the velocity component profiles can be described by an Ekman-spiral with adapted
+  boundary conditions that result in a reduction of the spiral-like rotation. 
+  %
 \end{abstract}
@@
 And locally, directional sheer of the wind in the upper part of the surface layer may cause a yaw for large wind power generation devices (REFERENCE?).
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\section{Formulation}
-%
-\paragraph{Physical Problem.} We consider incompressible, turbulent Ekman flow, i.e. the turbulent flow over a flat rotating plate, as
-a physical model for the truly neutral atmospheric boundary layer (ABL).
-%
-That is, the f-plane approximation is used confining effects of rotation to the horizontal components of
-velocity and neglecting dynamical effects due to latitudinal changes of the rate of rotation. 
-%
-%%PAR
-%
-\paragraph{Notation.} The dimensional velocity vector of the numerical simulations is $\underline{U} = (U_1,U_2,U_3) = (U,V,W)$ over the coordinate system
-$Oxyz$, where an approximate alignment with the surface shear stress is achieved.
-%
-For analysis of the results, we use two coordinate systems that are
-(i)  exactly aligned with the surface shear stress and labelled by an upper index $\alpha$ as in $\underline{U}^\alpha$ for the velocity vector, and
-(ii) the coordinate system aligned with the free-atmosphere geostrophic wind labelled by an upper index $G$ as in $\underline{U}^G$.
-%
-We denote by $u_\star$ the modulus surface friction velocity and let $Z_\star=1/u_\star$ for brevity.
 
-\section{Towards a Universal Velocity Profile for the Turbulent Ekman Layer}
-\subsection{Drag-Law} 
 The drag law for Ekman flow, including prediction of surface veering, has received considerable attention in the past.
@@
   %
-
 \end{itemize} 
-\begin{figure}
-  \centerline{\includegraphics[width=0.7\textwidth]{../plot/ustar_alpha.pdf}}
-  \caption{Variation of $Z_\star$ and surface veering with Reynolds number according to
-    the theory by Spalart et al. (1989) and as estimated from DNS data}
-  \label{fig:drag_law}
-\end{figure} 
 %
@@
 %
-\subsection{Stream-wise Velocity} 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+\section{Formulation}
+%
+We consider here incompressible, turbulent Ekman flow, that is, the turbulent flow over a flat rotating plate, as
+a physical model for the truly neutral atmospheric boundary layer (ABL).
+%
+The f-plane approximation is applied such that rotation only acts on horizontal velocity components;
+we thus neglect rotational effects on the horizontal components of velocity and dynamical effects
+due to latitudinal variation of the rate of rotation.
+%
+\par
+%
+The dimensional velocity vector of the numerical simulations is $\underline{U} = (U_1,U_2,U_3) = (U,V,W)$ over the coordinate system
+$Oxyz$, where an approximate alignment (plus/minus few degrees centigrade) of the direction $Ox$ with the surface shear stress is achieved.
+The coordinate $O_y$ points away from the wall,  and $Oz$ points in the span-wise direction normal to $Oxy$. 
+%
+For analysis of the results, we use two coordinate systems that are
+(i)  exactly aligned with the surface shear stress and labelled by an upper index $\alpha$ as in $\underline{U}^\alpha$ for the velocity vector, and
+(ii) the coordinate system aligned with the free-atmosphere geostrophic wind labelled by an upper index $G$ as in $\underline{U}^G$.
+%
+We denote by $u_\star$ the modulus surface friction velocity and let $Z_\star=1/u_\star$ for brevity.
+
+\section{Towards a Universal Velocity Profile for the Turbulent Ekman Layer}
+
+\end{itemize}
+\begin{figure}
+  \centerline{\includegraphics[width=0.7\textwidth]{../plot/ustar_alpha.pdf}}
+  \caption{Variation of $Z_\star$ and surface veering with Reynolds number according to
+    the theory by Spalart et al. (1989) and as estimated from DNS data}
+  \label{fig:drag_law}
+\end{figure}
+% 
+\subsection{Drag-Law}
+Formulation of a drag-law for Ekman flow consists estimation of two key parameters in the flow as a function of Reynolds number.
+%
+First, the modulus of the surface friction normalized by the geostrophic wind forcing, $u_\star$ and,
+second, the direction of surface shear stress, $\alpha$, also termed wind veering. 
+%
+A non-zero veering of the wind is a rather special case in comparison with most turbulent flows conisdered in an
+engineering context, and it confronts us with a situation where the most appropriate coordinate system for analysis
+(namely that aligned with the surface shear stress) is a priori unknown. 
+%
+We compare our DNS data against a semi-empirical based on integral consideration \cite{Spalart:1989p2436}
+and find, as demonstrated in previous work \cite{Ansorge:2014hf} excellent agreement in the range
+ $400<\mathbf{Re}<1600$, representing a factor of 16 in variation of viscosity. 
+%
+\todo{Add LES points}
+% 
+\par
+%
+We also find that the solution of the transient equation involved in estimation of $u_\star$ for a given Reynolds number
+$Re_D$ is approximated reasonably by the formulation
+\begin{equation}
+  u_\star= 4\log(\mathbf{Re}_D))-8
+\end{equation}
+which quantifies the 'weak' dependence of $u_\star$ on the Reynolds number as an approximately logarithmic one.
+%
+
+\subsection{Stream-wise Velocity Component} 
 For the stream-wise velocity profile (that in non-rotating flows due to the geometry is always aligned with the surface shear stress),
@@
     where the direction of 
-    in absence of roughness elements are marginal linear regime / viscous sub-layer \cite{Foken:2002gv,Foken:1978bc} 
-  \item buffer layer transition from linear regime to log-law  
-  \item log-layer (K\'arm\'an, Prandtl, Jimenez, Marusic)
-  \item outer log-layer () 
-  \item wake region
-  \item outer region 
+    in absence of roughness elements are marginal linear regime / viscous sub-layer \cite{Foken:2002gv,Foken:1978bc}.
+    This viscous law has no degree of freedom given the drag-law defining surface friction and veering.
+    %
+    Theoretical foundation is lacking for the exact shape of the velocity profile in the buffer layer; though crucial for turbulence
+    production, it is commonly understood as a transition region between the linear profile at the surface and the logarithmic profile
+    aloft.
+    A pure blending from the linear velocity profile into the logarithmic one is, however, not reasonable as both the linear and logarithmic
+    profile overestimate the velocity in the buffer layer. We therefore introduce a higher-order compensating term into
+    the linear profile before blending it into the logarithmic law:
+    \begin{equation}
+      U_\text{inner}^{\alpha}^+ = \frac{ y^+ + \gamma_4 (y^+)^{4} + \gamma_6 (y^+)^{6} } { 1 + \gamma_6 /u_{ref} (y^+ )^6}, 
+    \end{equation} 
+    where the fourth- and sixth-order terms in the nominator are designed to fit the profile below $y^+\approx 10$ and
+    the sixth-order correction in the denominator limits the growth of $u_\text{inner}$ for large $y^+$. 
+  \item In the logarithmic region, we use the profile
+    \begin{equation}
+      U_\text{log}^{\alpha}^+ = \frac{1}{\kappa} \log y^+ + C 
+    \end{equation}
+    with $\kappa=0.416$ and $C=5.4605$. %  log-layer (K\'arm\'an, Prandtl, Jimenez, Marusic)
+  \item The super-geostrophic maximum of streamwise velocity and a secondary minimum aloft the bulk-turbulent part of
+    the boundary layer are well-described by a classic Ekman spiral with adapted boundary conditions and a shift in reference height. 
   \end{itemize}

### 2021-03-03 plot/EkmanProfileClass.py

@@ -96,8 +96,8 @@ class EkmanUniversalClass:
         self.KAPPA = k
         self.C = C
 
-        log_law=np.zeros(len(yp))
-        log_law[1:] = np.log(yp[1:])/k + C
+        #log_law=np.zeros(len(yp))
+        log_law = self.profile_log(yp,re) # np.log(yp[1:])/k + C
         log_dev_outer_u=up-log_law
         log_dev_outer_u[:i2]=0.
         log_dev_outer_w=wm
@@ -183,6 +183,16 @@ class EkmanUniversalClass:
             
         return(1./z,p) 
 
+    def profile_log(self,yp,re):
+        self.qinit()
+
+        x=np.zeros(len(yp))
+        x[1:] = np.log(yp[1:])/self.KAPPA  + self.C
+        return x 
+
+    def erf_transition(self,w,l,x):
+        return 0.5*(sp.special.erf(w*np.log(x/l))+1)
+    
     def profile_plus(self,yp,re,us=-1,al=-1):
         self.qinit()
 
@@ -197,111 +207,63 @@ class EkmanUniversalClass:
         else:
             us_loc,al_loc=us,al
 
-        #print('PPLUS: us_loc, al_loc:',us_loc,al_loc, us,al)
+        u=np.zeros(len(yp))
+        w=np.zeros(len(yp)) 
         
         z=1/us_loc 
         dp_loc=re**2/(2*z**2)
-
         ym=np.array(yp/dp_loc)
+        trans_scale=2.0
 
-        # OUTER REFERENCE PROFILES
-        #
-        argz=0.66*(2.*np.pi*(ym+0.12)) 
-        dampA =  0.42*(us_loc/0.05- 0.07/re) #np.exp(zdum) * ( dU_mtc*np.cos(zdum) + dW_mtc*np.sin(zdum) )
-        dampB =  0.0#np.exp(zdum) * ( dU_mtc*np.sin(zdum) + dW_mtc*np.cos(zdum) )
-
-        outer_u= (1-(dampA*np.cos(argz)+dampB*np.sin(argz))*np.exp(-argz))/us_loc
-        outer_w=   ( dampA*np.sin(argz)-dampB*np.cos(argz))*np.exp(-argz)/us_loc
+        ################################################################################
+        # 2-COMPONENT OUTER REFERENCE PROFILES
+        argz=    0.66*(2.*np.pi*(ym+0.12)) 
[…]

### 2021-03-03 plot/plot.py

@@ -16,13 +16,14 @@ from EkmanProfileClass import Complex, EkmanUniversalClass, build_grid
 D2R = np.pi / 180. 
 
 use_data=False   # if true -- the netcdf files from DNS are needed 
-plot_ustar_alpha=True
-plot_summary=True
-plot_visc=True
-print_table=True
-plot_outer_log=True
-plot_total_rot=True
+plot_ustar_alpha=False
+plot_summary=False
+plot_visc=False
+print_table=False
+plot_outer_log=False
+plot_total_rot=False
 plot_convergence_1000=False
+plot_les_comp=True
 
 colors = {400 : 'gray',
           500 : 'pink',
@@ -34,7 +35,8 @@ colors = {400 : 'gray',
           10000:'gray',
           30000:'black',
           150000:'black',
-          300000:'black'} 
+          300000:'black',
+          1000000:'black'} 
 
 base='/Users/zrick/WORK/research_projects/SIM_NEUTRAL/netcdf/' 
 
@@ -177,17 +179,38 @@ if plot_outer_log :
     fig1=plt.figure(figsize=(12,5))
     ax1=fig1.add_axes([0.05,0.1,0.4,0.8])
     ax2=fig1.add_axes([0.55,0.1,0.4,0.8])
-    ax5=fig1.add_axes([0.09,0.455,0.18,0.43]) 
+    ax5=fig1.add_axes([0.11,0.455,0.18,0.43]) 
 
     fig2=plt.figure(figsize=(12,5))
     ax3=fig2.add_axes([0.05,0.1,0.4,0.8])
     ax4=fig2.add_axes([0.55,0.1,0.4,0.8])
     ax6=fig2.add_axes([0.26,0.185,0.18,0.37]) 
 
-    for re in [500,750,1000,1300,1600,10000,150000]:
+    for re in [500,750,1000,1300,1600,5000,10000,150000,1000000]:
         c=colors[re]
         nu=2./re**2
-        us,al=sc.ustar_alpha(re)
+        if  re <= 1600 :
+            f=Dataset(files[re],'r')
+            t_use = -1
+            
+            u_dat=f['rU'][t_use,:]
[…]

### 2021-03-03 plot/plot.py
added additional info to les-comparison output in plot.py

@@ -734,7 +734,6 @@ if plot_convergence_1000:
 if plot_les_comp == True:
     les_path='../data/'
     f_cor=1.03e-04
-    nu_air=1.456e-05 
     les_ReD1000_10cm={'f':'N_ReD1000_10cm_big_pr.001.nc',
                       'z0m':0.05,
                       'z0':0.0021,
@@ -768,7 +767,7 @@ if plot_les_comp == True:
         t=t[i_srt:]
         print('')
         print('Plotting LES Data for Re=',c['re']) 
-        print('  - STARTING FROM INDEX:', i_srt, 'LEN:', len(t), 'RANGE: ', (t[-1]-t[0])/1.e4/np.pi/2 )
+        print('  STARTING FROM INDEX', i_srt, '(nt=', len(t), ') time-RANGE: ', (t[-1]-t[0])/1.e4/np.pi/2,'inertial periods' )
         u=f_h['u'][i_srt:,:]
         v=-f_h['v'][i_srt:,:]
         w=f_h['w'][i_srt:,:]
@@ -798,14 +797,18 @@ if plot_les_comp == True:
         z0 = c['z0']
         KAPPA=0.40
         log_level=1
-        us = avg[0][log_level] * KAPPA / np.log(z[log_level]/z0)
-        zp=z*us/nu_air
-        zm=z/us*f_cor
         nu_est=2*(t_geo)**2 / (c['re']**2 * f_cor) 
+        us = avg[0][log_level] * KAPPA / np.log(z[log_level]/z0)
+        zp=z*us/nu_est
+        zm=z/us*f_cor
         
-        print('  - roughness length:',c['z0'],'m') 
-        print('  - RE (CALCULATED):', t_geo*np.sqrt(2./(f_cor*nu_air)))
-        print('  - nu (estimated from exact Re):', nu_est )
+        print('  - roughness length                   [m]:',c['z0']) 
+        print('  - RE (CALCULATED)                    [1]:',t_geo*np.sqrt(2./(f_cor*nu_est)))
+        print('  - nu (estimated from exact Re)   [m^2/s]:',nu_est )
+        print('  - outer velocity (u,v,magnitude)   [m/s]:',u_geo,v_geo,t_geo)
[…]

### 2021-05-28 write/manuscript.tex
some edits on manuscript; bulk statistics for theory (mas trans[prt and turning); slight modification of theory to better match the super-geostrophic maximum in u

--- 2020-12-31 write/manuscript.tex
@@
-\documentclass[a4paper,11pt]{article}
-\usepackage[margin=2.5cm]{geometry}
+\documentclass[a4paper,11pt]{amsart}
+
+\usepackage[margin=3.5cm]{geometry}
+\usepackage{lineno}
 \usepackage{amsmath}
@@
 \usepackage{caption}
-\usepackage{natbib}
+\usepackage{xcolor}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% BIBLIOGRAPHY DETAILS
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\usepackage[style=authoryear-comp,firstinits=true,citestyle=authoryear,natbib=true,backend=bibtex]{biblatex}
+\renewcommand\nameyeardelim{, }
+\DeclareNameAlias{sortname}{family-given}
+\renewcommand*{\multinamedelim}{\addcomma\space}
+\renewcommand*{\finalnamedelim}{\addcomma\space}
+\setlength{\bibitemsep}{\baselineskip}
+\renewbibmacro{in:}{}
+
+\AtEveryBibitem{%
+  \clearfield{url}
+  \clearfield{urlday}
+  \clearfield{urlmonth}
+  \clearfield{urlyear}
+  \clearfield{issn}
+  \clearfield{month}
+%%%  \clearfield{doi}%
+  \clearfield{keywords}
+  \clearfield{note}
+}
+
+\bibliography{ansorge}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% CUSTOM DEFINITIONS
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \newcommand{\gzm}{{\xi}}%{\gamma z_\text{match}}}
+\newcommand{\todo}[1]{\textcolor{red}{$[$#1$]$}}
+\newcommand{\SFC}{\mathrm{sfc}}
 
-\date{Draft \today}
-\title{Towards a Universal Veering Profile for Turbulent Ekman Flow at arbitrary Reynolds number \\
-  \normalsize LES and DNS of Turbulent Ekman Flow \\}
+\date{\scriptsize Draft - Version \today}
+\title{Profiles of Wind Direction and Speed in Turbulent Ekman Flow}
+\author{Cedrick Ansorge, Hauke Wurps}
 
@@
 \begin{abstract}
-  An analytical formulation for the profiles of stream- and span-wise velocity in turbulent Ekman flow
-  is developed based on asymptotic theory as well as data from direct numerical simulation and large-eddy simulation.
+  The profiles of wind speed and direction in turbulent Ekman flow
+  are formulated based on asymptotic theory and data from direct numerical simulation. 
   %
@@
   %
+  In the outer layer, the velocity component profiles can be described by an Ekman-spiral with adapted
+  boundary conditions that result in a reduction of the spiral-like rotation. 
+  %
   The span-wise component poses a conceptual challenge to the channel-flow analogy
-  in the context of asymptotic mixing, and it exhibits a mixed scaling in the surface layer, but follows
+  in the context of asymptotic matching; it exhibits a mixed scaling in the surface layer, but follows
   outer scaling for most of the outer layer.
   %
-  In both profiles, the wake region can be described by smooth error-function transition from a logarithmic layer
-  to the constant outer velocity.
-
 \end{abstract}
@@
 
-The exact veering of the wind away from its geostrophic direction is due to direct and indirect (due to turbulence) frictional effects,
-and it plays a role for various practical problems: 
+Wind veer away from its geostrophic direction is
+(i) due to direct frictional effects in the very vicinity of the surface and
+(ii) due to turbulence which exerts indirect frictional effects; these effects
+cause a slow-down of the mean wind reducing the Coriolis force thus turning the wind
+in favor of the pressure gradient force. 
+%%
+Not only does the veering set the frame of reference for surface layer theory, it also has
+physical effects on large and small scales.
 %
-It sets the amount of cross-isobaric volume flow and is thus a key factor in the life-cycle of large-scale synoptic systems. This mechanism is commonly
-referred to as 'Ekman pumping' referring to the perturbation of the geostrophic equilibrium that causes the veering as it was first described
-and solved for analytically by \cite{Ekman:1905tl}.
+From a large-scale perspective, the veering of wind across the planetary boundary layer determines
+the amount of cross-isobaric mass-flux, commonly referred to as 'Ekman pumping' \citep{ekman:AMA1905},
+and it is thus a key factor in the life-cycle of large-scale synoptic systems.
 %
-On the mesoscale, shear in the surface layer is known to be crucial for the initiation and sustainment of strong convective events (REFERENCE?).
+% %% This shear is probably not related to Ekman-turning 
+% On the mesoscale, shear in the surface layer a crucial factor for the initiation and sustainment of strong convective events (REFERENCE?).
 % 
-And locally, directional sheer of the wind in the upper part of the surface layer may cause a yaw for large wind power generation devices (REFERENCE?).
+Locally, the directional shear of the wind in the upper part of the surface layer may cause a systmatic yaw
+for tall wind power generation devices where blades reach into the Ekman layer, i.e. that part of the boundary
+layer where the wind starts to turn. 
 %
+\par
+%
+Wind veer in the planetary boundary layer is commonly characterized by the surface veering angle $\alpha$
+defined as the angle between the negative surface shear stress $\tau_\SFC$ and the geostrophic wind.
+%
+Surface veering $\alpha$ and geostrophic drag $Z\equiv u_\star/G$, where $u_\star\equiv\sqrt{\nu|\tau_\SFC|}$,
+uniquely determine the surface drag $\tau_\SFC$ in a turbulent Ekman flow. 
+%
+In any quantitative description of the surface layer, the friction velocity $u_\star$ is the dynamic scale
+and $\alpha$ defines the alignment of the frame of reference.
+%
+Therefore, knowing these parameters is a pre-requisite for any quantitative theory of the surface layer and a first
+attempt in theoretical quantification was undertaken by \cite{rossby:PPO1935} using integral relations in the boundary layer.
+%
+Asymptotic similarity theory was later used by \cite{tennekes:JAS1973, blackadar:JAS1968},
+and--based on his seminal direct numerical simulations (DNS) of Ekman flow--,~\cite{spalart:JFM1989} suggested
+a modification to take into account effects of low to intermediate Reynolds numbers.
+Later on, constants were re-evaluated with a focus on the atmospheric boundary layer based on
+observations \cite{hogstrom:BM1988,hogstrom:BM1996}
+and numerical modelling \citep{spalart:PF2008, spalart:PF2009, ansorge:BM2014, ansorge:BM2019}. 
+%
+\par
+%
+Attempts were also undertaken to obtain profiles of the wind speed: 
+One approach is to match the inner and outer layer at a reference height;
+\cite{etling:2002, emeis:2018}  (Sec. 21.10; Eq. 21.48) choose the Prandtl-layer height to match
+the wind speed profiles, which, however, requires to externally prescribe the veering $\alpha$ at that height. 
+A one-dimensional profile with constant veering is given by \cite[Sec. 3; Eq. 3.1-3.19]{emeis:m2007}.
+
+\cite{gryning:BM2007} present an extension of the wind-speed profile beyond the surface layer
+using a neutral referece profile and a stability correction;
+\cite{kelly:BM2010}, based on a probabilistic representation of stratification,
+develop a model for the long-term mean windspeed
+in the atmospheric boundary layer and compare this with observation at different sites; 
+\cite{kelly:WE2016} demonstrate the effect of such improved model for wind-energy applications.
+%
+In consideration of the large scale separation in geophysical flow, the rotation of the wind in the
+surface layer is often assumed negligible, and above investigations merely focus on the wind speed; 
+that means, the veering of the wind with height is not described and there is little knowledge on the
+profile of the span-wise velocity component and the precise shape of the hodograph in the limit of a
+truly neutral Ekman boundary layer.
+%
+%%% \par  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% % No discussion of higher-order moments at this stage
+%%%
+%%% For non-rotating, non-external flows theoretical considerations by Townsend (cite to attached-eddy hypothesis)
+%%% have recently been confirmed from both wind-tunnel  experiments and numerical simulation
+%%% \cite{ng_comparison_2011,marusic_logarithmic_2013};
+%%% this implies that physical models for the second moments (Reynolds stresses) of the flow \ldots.
+%
+\par %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+Ekman-layer models and two-layer models of the ABL take into account rotational effects at higher altitudes,
+for instance when the wind speed needs to be evaluated at heights on the order of 100-200\,\mathrm{m},
+a particular concern when it comes to wind-power forecasting \citep{optis:BM2014}.
+%
+Despite rotational effects being considered, the formulation of these models for the outer layer
+and analysis of their performance primarily focuses on wind speed.
+% 
+\cite{jiang:JAS2018} recognized that the outer part of the Ekman boundary layer receives less attention in comparison
+with the surface layer and commence in studying the neutral problem by Large-Eddy simulation (LES).
+%
+They focus on the wind speed and find an extended logarithmic layer when considering the wind speed instead of
+the shear-aligned component, and they eventually demonstrate by means of an analytical model that this vertical
+extension of the logarithmic layer may be explained by a transfer of stress to the spanwise velocity component where
+it is assumed that the shear vector $\tau(z)$ and stress vectors $(\p_zU,\p_zV)$ are aligned. 
+%
+\par
+%
+Turbulent Ekman flow is considered here as a conceptual model of the
+homogeneous, stationary atmospheric boundary layer over a flat surface under neutral stratification. 
+%
+Universal profiles of the wind vector for turbulent Ekman flow can serve as well-described limit and reference
+for higher-order approaches taking into account possible effects of stratification, roughness or other physical
+complications encountered in the realgeophysical system.
+%
+While, on first sight, the study of such a strongly idealized case appears as an academic problem, it contains the essence
+of surface similarity as it is used in most atmospheric models, be it conceptual or numeric ones.
+%
+For instance, effects of roughness are commonly incorporated by a linear transformation of vertical scale involving the
+roughness parameter $z_0$ and for larger roughness also a displacement height~%
+\citep{monin:1975,jacobs:AFM1988,hogstrom:BM1988}; 
+effects of stability are commonly accounted for by a linearization around the neutrally stratified profile~%
+\citep{monin:ARF1970, monin:1975, hogstrom:BM1988, hogstrom:BM1996,sakagami:BM2020}.
+%
+Such generalized profiles may also serve as better initial condition for numerical simulation of the flow,
+to minimize the length of initial transient periods, or as benchmark for turbulence
+closures that can be tuned to reproduce the neutral limit case.
+%
+\par
+%
+Despite the strong simplifications implied by our choice of set-up, there is no straightforward approach to solving
+this conceptually well-defined problem.
+%
+Large-Eddy simulation not only needs to be tuned for the surface shear stress and veering angle, but it also
+relies on sub-grid closures that often assume alignment of the stress with gradients -- a questionable pre-requisite when the wind rotates with height.
+%
+While recent advances have been made in analysis of this simplified set-up \cite{jiang:JAS2018} (MORE?),
+there is not enough understanding for a quantitative generalization of the results to arbitrary external forcing.
+%
+At the same time, an increasing amount of high-quality and high-resolution data from turbulence-resolving approaches
+is emerging due to recent advances in high-performance computing and its application to geophysical problem sets;
+the geophysical range of scale separation, however, is--and it will remain so for the foreseeable future--out of reach
+for such simulation.
+%
+Here, the routinely employed concept of Reynolds-number similarity can help: it states that certain statistics (normally those of low order) become independent of the Reynolds number when scaled appropriately \todo{Dimotakis / da Silva / Pope referece?} .
+%
+It appears that for certain statistics in Ekman flow, this regime is reached with the Reynolds numbers that became
+possible with the increasing capabilities of supercomputering over the past years.
+%
+The robust features of mean velocity profiles across a range of Reynolds numbers
+are exploited in this work to develop a formulation for
+both the stream-wise and span-wise components of the mean velocity vector as a function of the
+Reynolds number.
+
+\todo{Numerical approach, DNS, Scale-separation, Re-similarity
+  \citep{dimotakis:JFM2000, dasilva:ARF2014, pope:2000, moin:ARF1998, mellado:JAM2018}}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
-\paragraph{Physical Problem.} We consider incompressible, turbulent Ekman flow, i.e. the turbulent flow over a flat rotating plate, as
+We consider here incompressible, turbulent Ekman flow, that is, the turbulent flow over a flat rotating plate, as
 a physical model for the truly neutral atmospheric boundary layer (ABL).
 %
-That is, the f-plane approximation is used confining effects of rotation to the horizontal components of
-velocity and neglecting dynamical effects due to latitudinal changes of the rate of rotation. 
+The f-plane approximation is applied such that rotation only acts on horizontal velocity components;
+we thus neglect rotational effects on the horizontal components of velocity and dynamical effects
+due to latitudinal variation of the rate of rotation.
 %
-%%PAR
+\par
 %
-\paragraph{Notation.} The dimensional velocity vector of the numerical simulations is $\underline{U} = (U_1,U_2,U_3) = (U,V,W)$ over the coordinate system
-$Oxyz$, where an approximate alignment with the surface shear stress is achieved.
+The dimensional velocity vector of the numerical simulations is $\underline{U} = (U_1,U_2,U_3) = (U,V,W)$ over the coordinate system
+$Oxyz$, where an approximate alignment (plus/ minus few degrees centigrade) of the direction $Ox$ with the surface shear stress is achieved.
+The coordinate $Oy$ points away from the wall,  and $Oz$ points in the span-wise direction normal to $Oxy$. 
 %
@@
 \section{Towards a Universal Velocity Profile for the Turbulent Ekman Layer}
-\subsection{Drag-Law} 
-The drag law for Ekman flow, including prediction of surface veering, has received considerable attention in the past.
-\begin{itemize}
-\item \cite{Rossby:1935wa} first approach;
-\item \cite{Tennekes:1973jh, Blackadar:1968ew} asymptotic similarity;
-\item \cite{Spalart:1989p2436, Spalart:2008p2432, Spalart:2009p2466} seminal LES studies in search for a universal log-law;
-\item \cite{Ansorge:2014hf, Ansorge:2019di} re-evaluation of the constants with geophysical focus;
-\item \cite{Hogstrom:1988wg, Hgstrm:1996kt} (re-)consideration of observational data.
+
+Formulations for the outer layer that take into account the rotation, and thus deviation from the channel-flow analogy
+need to be matched to the framework of surface similarity, and it
+occurs that a smooth transition from the inner to the outer layer is not easily achieved.
 %
+\cite{optis:BM2014}, for instance, define an ``effective geostrophic wind vector that has the same magnitude of the
+observed surface geostrophic wind and is rotated by the angle $\alpha$ [their nomenclature]'' to overcome the
+unsteady transition when coming approaching the Ekman layer from below.
+%
+While such rotation of the wind vector is \emph{a posteriori} justified by the observational data that the model outcomes are compared to,
+we believe that it is the manifestation of a mismatch between the theoretical treatment of the inner and outer layer of
+turbulent Ekman flow.
 
@@
 % 
-Attempts were also undertaken to obtain profiles of the wind speed: 
-\cite{Gryning:2007dy} present an extension of the wind-speed profile beyond the surface layer using a neutral referece profile
-and a stability correction;
-\cite{Kelly:2010ev}, based on a probabilistic representation of stratification, develop a model for the long-term mean windspeed
-in the atmospheric boundary layer and compare this with observation at different sites; 
-\cite{Kelly:2016bs} demonstrate the effect of such improved model for wind-energy applications.
+\subsection{Drag-Law}
+A drag-law for Ekman flow determines two key parameters of the flow as a function of Reynolds number alone.
 %
-These investigations merely focus on the wind speed and thus do not regard the actual vector, i.e. the veering of
-the wind with height is not described and there is little knowledge on the profile of the span-wise velocity component
-and the precise shape of the hodograph in the limit of a truly neutral Ekman boundary layer.
+These are, first, the modulus of the surface friction normalized by the geostrophic wind forcing, $u_\star$ and,
+second, the direction of surface shear stress, $\alpha$, also termed wind veering. 
 %
-While this is indeed an academic problem, the Ekman boundary layer often serves as limit for stratified boundary layers
-and thus not only needs to be understood but also well-described as reference for higher-order approaches that take into
-account possible effects of stratification, roughness or other physical complications of the geophysical system.
+A non-zero veering of the wind is a rather special case in comparison with most turbulent flows conisdered in an
+engineering context, and it confronts us with a situation where the most appropriate coordinate system for analysis
+(namely that aligned with the surface shear stress) is a priori unknown. 
 %
-\subsection{Stream-wise Velocity} 
+We compare our DNS data against a semi-empirical drag-law based on integral consideration \cite{spalart:JFM1989}
+and find, as demonstrated in previous work \cite{ansorge:BM2014}, excellent agreement in the range
+ $400<\mathbf{Re}<1600$, representing a factor of 16 in variation of viscosity. 
+%
+\par
+%
+We also find that the solution of the transient equation involved in estimation of $u_\star$ for a given Reynolds number
+$Re_D$ is approximated reasonably by the formulation
+\begin{equation}
+  u_\star= 4\log(\mathbf{Re}_D))-8
+\end{equation}
+which quantifies the 'weak' dependence of $u_\star$ on the Reynolds number as an approximately logarithmic one, at least
+for problems with a scale separation on the order that is relevant to geophysical problem ($Re_D<10^{8}$).  
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+\subsection{Profile in the outer layer}
+%
+The outer layer of Ekman flow is characterized by a turning of the wind velocity and the super-geostrophic maximum
+that is sustained by momentum convergence at the inflection point of the velocity profile. 
+The super-geostrophic maximum of streamwise velocity and a secondary minimum aloft the bulk-turbulent part of
+the boundary layer are well-described by a classic Ekman spiral with adapted boundary conditions and a shift
+in reference height.
+%
+The textbook solution for the Ekman layer reads as 
+\todo{equations / figures}
+\begin{figure}
+  \includegraphics[width=\textwidth]{../plot/outer_layer.pdf}
+  \caption{} 
+\end{figure}
+
+\subsection{Stream-wise Velocity Component} 
 For the stream-wise velocity profile (that in non-rotating flows due to the geometry is always aligned with the surface shear stress),
 well-established theories exist for various regimes according to their distance from the wall and the relative role of
-viscosity, turbulence and interaction with the outer region of the flow with the logarithmic law for the mean velocity as a central anchor point:
+viscosity, turbulence and interaction with the outer region of the flow with the logarithmic law for the mean velocity as a central anchor point.
+%
+\par
+%
 \begin{subequations}
-  \begin{itemize}
-  \item In immediate vicinity to the surface, local turbulent mixing cannot occur for the no-slip/no-penetration boundary condition,
+
+In immediate vicinity to the surface, local turbulent mixing cannot occur for the no-slip/no-penetration boundary condition,
 and the mean velocity is described by a viscous profile of the form
@@
 \end{equation}
-    where the direction of 
-    in absence of roughness elements are marginal linear regime / viscous sub-layer \cite{Foken:2002gv,Foken:1978bc} 
-  \item buffer layer transition from linear regime to log-law  
-  \item log-layer (K\'arm\'an, Prandtl, Jimenez, Marusic)
-  \item outer log-layer () 
-  \item wake region
-  \item outer region 
-  \end{itemize}
+where the direction of the velocity points into the exact opposite direction of the wall shear stress $\tau$. 
+In absence of roughness elements and for small roughness ($z_0^+<5$), this linear regime is known as
+viscous sub-layer \cite{foken:m2002,foken:BM1978}.
+In fact, this law of the wall has no degree of freedom given the drag, i.e. once $u_\star$ and $\alpha$ are defined. 
+%
+However, theoretical foundation is lacking for the exact shape of the velocity profile in the buffer layer; though crucial for turbulence
+production, it is commonly understood as a transition region between the linear profile at the surface and the logarithmic profile
+aloft.
+A pure blending from the linear velocity profile into the logarithmic one is, however, not reasonable as both the linear and logarithmic
+profile overestimate the velocity in the buffer layer. We therefore introduce a higher-order compensating term into
+the linear profile before blending it into the logarithmic law:
+\begin{equation}
+  U_\text{inner}^{\alpha}^+ = \frac{ y^+ + \gamma_4 (y^+)^{4} + \gamma_6 (y^+)^{6} } { 1 + \gamma_6 /u_{ref} (y^+ )^6}, 
+\end{equation} 
+where the fourth- and sixth-order terms in the nominator are designed to fit the profile below $y^+\approx 10$ and
+the sixth-order correction in the denominator limits the growth of $u_\text{inner}$ for large $y^+$.
+%
+\par
+%
+In the logarithmic region, we use the profile
+\begin{equation}
+  U_\text{log}^{\alpha}^+ = \frac{1}{\kappa} \log y^+ + C 
+\end{equation}
+with the von-K\'arm\'an constant $\kappa=0.416$ and the boundary condition $C=5.4605$. %  log-layer (K\'arm\'an, Prandtl, Jimenez, Marusic)
+%
+\par %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 \end{subequations} 
@@
 \end{figure} 
-
-Existing theories applied to the boundary layer:
-\begin{itemize}
-\item \cite{Etling:2002kx, Emeis:2018vq}  (Sec. 21.10; Eq. 21.48) -- ad-hoc matching at Prandtl-layer height; requires $\alpha$ 
-\item \cite{Emeis:2007cy} (Sec. 3; Eq. 3.1-3.19) -- 1D profile; considering constant (?) veering 
-\item \cite{Gryning:2007dy} (Mean wind profiles)
-\end{itemize}
-
-For non-rotating, non-external flows theoretical considerations by Townsend (cite to attached-eddy hypothesis)
-have recently been confirmed from both wind-tunnel  experiments and numerical simulation \cite{Ng:2011bk,Marusic:2013hf}
-implying physical models for the second moments (Reynolds stresses) of the flow.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 %
-The background rotation and associated veering of the surface wind implies a non-zero profile for the span-wise velocity that
-challenges the conventional assumptions related to the so-called channel-flow analogy:
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+The background rotation and associated veering of the surface wind implies a non-zero profile for the span-wise
+velocity which challenges the conventional assumptions related to the so-called channel-flow analogy:
 %
@@
 %
-This illustrates that the higher-order (in terms of the channel-flow analogy where $V=0$) correction follows a mixed scaling when the surface is approached which has been demonstrated for higher-order terms in other types of flow \citep{Mellado:2016hv}.
+This illustrates that the higher-order (in terms of the channel-flow analogy where $V=0$) correction follows a mixed scaling when the surface is
+approached which has been demonstrated for higher-order terms in other types of flow \citep{mellado:BM2016}. 
 %
@@
 by $Re_\tau^{-1/2}$ which is indeed known from the generalization of higher-order statistics, such as turbulent fluxes
-in the inner layer \citep{Marusic:2013hf} that also follow a mixed scaling in the inner layer.
+in the inner layer \citep{marusic:JFM2013} that also follow a mixed scaling in the inner layer.
 %
@@
 %
-Based on the physical extent of the viscous sub-layer in Ekman flow around five wall units \citep{Foken:2002gv, Ansorge:2019di},
+Based on the physical extent of the viscous sub-layer in Ekman flow around five wall units \citep{foken:m2002, ansorge:BM2019},
 we choose $z^+=5$ to transition from one to the other and note that $V$ is already very small at this height.
@@
   \phantom{AAA}\textbf{(a)\hspace{0.3\textwidth}(b)\hspace{0.3\textwidth}(c)}\\
-  \includegraphics[width=\textwidth]{../plot/ekman_profiles_data.pdf}
+  \includegraphics[width=\textwidth]{../plot/ekman_profiles.pdf}
   \caption{ \textbf{(a)} Velocity deficity, \textbf{(b)} velocity profile in shear-aligned hodographs and \textbf{(c)} hodograph in geostrophy-aligned coordinates.
@@
 \begin{eqnarray}
-  \hat{u}(z) &=& U_{\inf} + e^{-\gamma z} \left[A \cos\gamma z + B \sin\gamma z\right]   \\
-  \hat{w}(z) &=& W_{\inf} + e^{-\gamma z} \left[-A \sin\gamma z + B \cos\gamma z\right]
+  \hat{u}(z) &=& U_{\infty} + e^{-\gamma z} \left[A \cos\gamma z + B \sin\gamma z\right]   \\
+  \hat{w}(z) &=& W_{\infty} + e^{-\gamma z} \left[-A \sin\gamma z + B \cos\gamma z\right]
 \end{eqnarray} 
 \end{subequations}
-where the constants $U_\inf$, $W_\inf$ set the top boundary condition and $A$ and $B$ set the bottom boundary condition. 
+where the constants $U_\infty$, $W_\infty$ set the top boundary condition and $A$ and $B$ set the bottom boundary condition. 
 %
-The most common boundary condition for a surface Ekman layer is $A=U_\inf=G$, $B=0$, and $W_\inf=0$.
+The most common boundary condition for a surface Ekman layer is $A=U_{\infty}=G$, $B=0$, and $W_{\infty}=0$.
 %
@@
   \begin{matrix} 
-  u(z_\text{match}) \equiv u_\text{match} &=& U_{\inf} + e^{-\gzm}} \left [ A \cos\gzm + B\sin\gzm \right] \\
-    w(z_\text{match}) \equiv w_\text{match} &=& W_{\inf} + e^{-\gzm}} \left [-A \sin\gzm + B\cos\gzm \right]
+    u(z_\text{match}) \equiv u_\text{match} &=& U_{\infty} + e^{-\gzm}} \left [ A \cos\gzm + B\sin\gzm \right] \\
+    w(z_\text{match}) \equiv w_\text{match} &=& W_{\infty} + e^{-\gzm}} \left [-A \sin\gzm + B\cos\gzm \right]
   \end{matrix} 
@@
 \begin{eqnarray}
-  e^\gzm  \left(\begin{matrix}
-     u_\text{match}-U_{\inf} \\ 
-     w_\text{match}-W_{\inf}
-  \end{matrix}\right) = \left(\begin{matrix}
-  \cos\gzm A &+ \sin\gzm B \\ -\sin\gzm A &+ \cos\gzm B
+  \Rightarrow \left(\begin{matrix}
+    u_\text{match}-U_{\infty} \\ 
+    w_\text{match}-W_{\infty}
+  \end{matrix}\right) = e^{-\gzm}\left(\begin{matrix}A \\ B \end{matrix}\right)\left(\begin{matrix}
+    \cos\gzm &+ \sin\gzm \\ -\sin\gzm &+ \cos\gzm
   \end{matrix} \right) \\
@@
 \end{subequations}
-\bibliographystyle{plainnat}
-\bibliography{ansorge.bib}
+
+\printbibliography
 

### 2021-05-28 plot/EkmanProfileClass.py

@@ -5,7 +5,7 @@ import scipy as sp
 from scipy.special import lambertw as prodlog
 
 def build_grid(trgt,thick=1) :
-    vert_stretch=1.05   #grid stretching
+    vert_stretch=1.01   #grid stretching
     dyloc=1             #grid resolution at wall
     ny=1
     yloc=dyloc
@@ -96,7 +96,8 @@ class EkmanUniversalClass:
         self.KAPPA = k
         self.C = C
 
-        #log_law=np.zeros(len(yp))
+        re = np.sqrt(2*deltap/(us*us)) # nu = us*us/deltap
+
         log_law = self.profile_log(yp,re) # np.log(yp[1:])/k + C
         log_dev_outer_u=up-log_law
         log_dev_outer_u[:i2]=0.
@@ -107,8 +108,6 @@ class EkmanUniversalClass:
         
         f1=Complex(ud_align,wd_align) 
 
-        # self.outer_wp_itp=spip.interp1d(ym,wp*us)
-        
 
         imax=mp.geti(yp,50)
         f5_x=np.zeros(len(up))
@@ -119,10 +118,6 @@ class EkmanUniversalClass:
         f5_z[imax:]=0. 
         f5=Complex(f5_x,f5_z) 
         f1_high=Complex(f1.r-f5.r,f1.i) 
-
-        #self.outer_ud_itp=spip.interp1d(ym,log_dev_outer_u,'cubic')
-        #self.outer_wd_itp=spip.interp1d(ym,log_dev_outer_w,'cubic')
-        # self.inner_u_itp=spip.interp1d(yp,up,'cubic')
 
         #estimate intercept of f1 - log-law
         # Real part from f1_high u component at first grid point
@@ -157,7 +152,7 @@ class EkmanUniversalClass:
         Ar=self.A.r 
         k=self.KAPPA 
         C5=self.C5.r
-        k=0.415
+        k=0.416
         z=10
 
         #Ai=Ai+0.285
@@ -179,9 +174,9 @@ class EkmanUniversalClass:
                 break;
             i=i+1
 
-        p=p-1.00*C5/(re**2) * (z**2)
+        aleph=p-1.00*C5/(re**2) * (z**2)
             
-        return(1./z,p) 
+        return(1./z,aleph) 
 
     def profile_log(self,yp,re):
         self.qinit()
[…]

### 2021-05-28 plot/plot.py

@@ -18,12 +18,14 @@ D2R = np.pi / 180.
 use_data=False   # if true -- the netcdf files from DNS are needed 
 plot_ustar_alpha=False
 plot_summary=False
-plot_visc=False
+plot_visc_outer=False
 print_table=False
 plot_outer_log=False
 plot_total_rot=False
 plot_convergence_1000=False
-plot_les_comp=True
+plot_re1600=True
+plot_les_comp=False
+plot_applications=False
 
 colors = {400 : 'gray',
           500 : 'pink',
@@ -85,14 +87,193 @@ if use_data:
 
     sc.estimate(yp_ref,up_ref,wp_ref,deltap_ref,us_ref,al_ref*D2R-geo_rotate,plot=False)
 
-if plot_visc:
+if plot_re1600:
+    re=1600;
+    nu=2./(re*re) 
+
+
+    f=Dataset(files[re],'r')
+    u_datC=f['rU'][:,:]
+    w_datC=f['rW'][:,:]
+    y_dat=f['y'][:]
+    t_dat=f['t'][:] 
+    us=np.average(f['FrictionVelocity'][:])
+    al=np.average(f['FrictionAngle'][:])
+    
+    print(t_dat[-1]-t_dat[0]) 
+
+    file_name = '/Users/zrick/WORK/research_projects/SIM_NEUTRAL/netcdf/avg_flw_ri0.0_re1600.0_co0_3840x960x7680.nc'
+    f=Dataset(file_name,'r')
+    y=f['y'][:]
+    ny=len(y)
+    fig=plt.figure(figsize=(10,5))
+    ax1=fig.add_axes([0.06,0.1,0.4,0.85])
+    ax2=fig.add_axes([0.50,0.1,0.47,0.85])
+    NT=10
+    U=np.zeros([NT,ny])
+    W=np.zeros([NT,ny])
+    us=np.zeros([NT])
+    al=np.zeros([NT]) 
+
+
+    U = u_datC[28:,:]
+    W = w_datC[28:,:]
+
+    al = f['FrictionAngle'][:]
+    us = f['FrictionVelocity'][:]
+
+    Uavg=np.average(U,0)
+    Wavg=np.average(W,0)
+    Udsh=(U-Uavg)/np.average(us) 
+    Wdsh=(W-Wavg)/np.average(us)
+
+    
[…]

### 2021-06-02 write/manuscript.tex
added governing equations; definition of Reynolds number

@@
 
-\usepackage[style=authoryear-comp,firstinits=true,citestyle=authoryear,natbib=true,backend=bibtex]{biblatex}
-\renewcommand\nameyeardelim{, }
+\usepackage[maxcitenames=2,style=authoryear-comp,firstinits=true,citestyle=authoryear,natbib=true,backend=bibtex]{biblatex}
 \DeclareNameAlias{sortname}{family-given}
 \renewcommand*{\multinamedelim}{\addcomma\space}
-\renewcommand*{\finalnamedelim}{\addcomma\space}
+\renewcommand*{\finalnamedelim}{~and\space}
 \setlength{\bibitemsep}{\baselineskip}
@@
 
+\newcommand{\todo}[1]{\textcolor{red}{$[$#1$]$}}
+
+% TO BE USED IN MATH-ENVIRONMENT 
+\newcommand{\SFC}{\mathrm{sfc}}
 \newcommand{\p}{\partial}
 \newcommand{\gzm}{{\xi}}%{\gamma z_\text{match}}}
-\newcommand{\todo}[1]{\textcolor{red}{$[$#1$]$}}
-\newcommand{\SFC}{\mathrm{sfc}}
+\newcommand{\RO}{\mathrm{Ro}}  
+\newcommand{\RE}{\mathrm{Re}}
+\newcommand{\LR}{\Lambda_\RO}
 
@@
 %
-Therefore, knowing these parameters is a pre-requisite for any quantitative theory of the surface layer and a first
-attempt in theoretical quantification was undertaken by \cite{rossby:PPO1935} using integral relations in the boundary layer.
+Knowledge about $u_\star$ and $\alpha$ is thus a pre-requisite for any quantitative theory of the surface layer and a first
+attempt in theoretical quantification was undertaken by \citet{rossby:PPO1935} using integral relations in the boundary layer.
 %
-Asymptotic similarity theory was later used by \cite{tennekes:JAS1973, blackadar:JAS1968},
+Asymptotic similarity theory was later used by \citet{tennekes:JAS1973, blackadar:JAS1968},
 and--based on his seminal direct numerical simulations (DNS) of Ekman flow--,~\cite{spalart:JFM1989} suggested
@@
 One approach is to match the inner and outer layer at a reference height;
-\cite{etling:2002, emeis:2018}  (Sec. 21.10; Eq. 21.48) choose the Prandtl-layer height to match
+\citet{etling:2002, emeis:2018}  (Sec. 21.10; Eq. 21.48) choose the Prandtl-layer height to match
 the wind speed profiles, which, however, requires to externally prescribe the veering $\alpha$ at that height. 
-A one-dimensional profile with constant veering is given by \cite[Sec. 3; Eq. 3.1-3.19]{emeis:m2007}.
+A one-dimensional profile with constant veering is given by \citet[Sec. 3; Eq. 3.1-3.19]{emeis:m2007}.
 
-\cite{gryning:BM2007} present an extension of the wind-speed profile beyond the surface layer
+\citet{gryning:BM2007} present an extension of the wind-speed profile beyond the surface layer
 using a neutral referece profile and a stability correction;
@@
 % 
-\cite{jiang:JAS2018} recognized that the outer part of the Ekman boundary layer receives less attention in comparison
+\citet{jiang:JAS2018} recognized that the outer part of the Ekman boundary layer receives less attention in comparison
 with the surface layer and commence in studying the neutral problem by Large-Eddy simulation (LES).
@@
 %
-While recent advances have been made in analysis of this simplified set-up \cite{jiang:JAS2018} (MORE?),
-there is not enough understanding for a quantitative generalization of the results to arbitrary external forcing.
+Despite advances in analysis of this simplified set-up \citep{jiang:JAS2018} \todo{(MORE?)},
+there is yet insufficient understanding for a quantitative generalization of the results to
+arbitrary external forcing (manifest in variation of the Reynolds number).
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\section{Formulation}
+\section{Problem Formulation and Numerical Approach}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 %
-\par
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Notation and Governing Equations}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 The coordinate $Oy$ points away from the wall,  and $Oz$ points in the span-wise direction normal to $Oxy$.
-%
 For analysis of the results, we use two coordinate systems that are
@@
 We denote by $u_\star$ the modulus surface friction velocity and let $Z_\star=1/u_\star$ for brevity. 
+%
+\par
+%
+We consider the incompressible Navier--Stokes equations for the three velocity components on the f-plane in a framework
+that is governed by the physical parameters
+(i)   geostrophic wind magnitude $G=\sqrt{G_1^2+G_3^2}$ (where $\mathbf{G}=(G_1,0,G_3)$ is the geostrophic wind vector),
+(ii)  Coriolis parameter $f$ (representing the angular rotation), and
+(iii) kinematic viscosity $\nu$.
+   %
+These yield a Reynolds number $\RE$ as non-dimensional parameter governing the system.
+%
+We use the geostrophic wind as velocity and the Coriolis parameter $f$ as time scale for the non-dimensional framework which implies the Rossby radius $\LR=G/f$ as length scale, such that
+\begin{equation}
+  \RE=\frac{G \LR }{\nu}. 
+\end{equation}
+%
+\par
+%
+The non-dimensionalized set of governing equation reads as 
+\begin{subequations} 
+\begin{eqnarray}
+  \frac{\partial u_i}{\partial t} &=& \frac{\p \pi}{\p x_i} - u_j \frac{\p u_i}{\p x_j} + \epsilon_{i2j} (u_j -G_j)   + \frac{1}{\RE} \frac{\p^2 u_i}{\p x_{j}^2} \\ 
+  \frac{\partial u_j}{\partial x_j} &=& 0,   
+\end{eqnarray} 
+\label{eqn:governing} 
+\end{subequations}
+where $u_i=U_i/G$ are the non-dimensional components velocity, $\pi$ is non-dimensional pressure, and $g_j=G_j/G$ are non-dimensionalized components geostrophic wind (with $g_{1}^2+g_{3}^2=1$ by construction).
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Numerical Simulations}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\todo{DNS Code}
+\par
+\todo{(new and old) Simulations}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Towards a Universal Velocity Profile for the Turbulent Ekman Layer}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 Formulations for the outer layer that take into account the rotation, and thus deviation from the channel-flow analogy

### 2021-07-29 write/manuscript.tex
proceeded on manuscript

@@
 \usepackage{xcolor}
+\usepackage{dirtytalk}
+\usepackage{colortbl}
+\usepackage{booktabs}
 
@@
 \setlength{\bibitemsep}{\baselineskip}
+\setlength\parindent{3em}
+\setlength\parskip{0.5em}
 \renewbibmacro{in:}{}
@@
 \newcommand{\todo}[1]{\textcolor{red}{$[$#1$]$}}
+\newcommand{\quotation}[1]{"#1"}
 
@@
 Not only does the veering set the frame of reference for surface layer theory, it also has
-physical effects on large and small scales.
+effects at small and large scales.
 %
@@
 % 
-Locally, the directional shear of the wind in the upper part of the surface layer may cause a systmatic yaw
-for tall wind power generation devices where blades reach into the Ekman layer, i.e. that part of the boundary
-layer where the wind starts to turn. 
+At the small-scale end, the directional shear of the wind in the upper part of the surface layer
+may cause a systmatic yaw for tall wind power generation devices where blades reach into the Ekman layer,
+i.e. that part of the boundary layer where the wind starts to turn. 
 %
@@
 Later on, constants were re-evaluated with a focus on the atmospheric boundary layer based on
-observations \cite{hogstrom:BM1988,hogstrom:BM1996}
+observations \citep{hogstrom:BM1988,hogstrom:BM1996}
 and numerical modelling \citep{spalart:PF2008, spalart:PF2009, ansorge:BM2014, ansorge:BM2019}. 
@@
 % 
-\citet{jiang:JAS2018} recognized that the outer part of the Ekman boundary layer receives less attention in comparison
+Still, in \citeyear{jiang:JAS2018}, \citeauthor{jiang:JAS2018}
+recognized that the outer part of the Ekman boundary layer receives less attention in comparison
 with the surface layer and commence in studying the neutral problem by Large-Eddy simulation (LES).
@@
 %
-Universal profiles of the wind vector for turbulent Ekman flow can serve as well-described limit and reference
-for higher-order approaches taking into account possible effects of stratification, roughness or other physical
-complications encountered in the realgeophysical system.
+Universal profiles of the wind vector for turbulent Ekman flow not only are a well-described limit
+for theoretical exploration or higher-order approaches taking into account possible effects of stratification,
+roughness or other physical complications encountered in the realgeophysical system.
 %
@@
 %
-Such generalized profiles may also serve as better initial condition for numerical simulation of the flow,
+Furthermure, profiles can serve as better initial condition for numerical simulation of the flow,
 to minimize the length of initial transient periods, or as benchmark for turbulence
@@
 Large-Eddy simulation not only needs to be tuned for the surface shear stress and veering angle, but it also
-relies on sub-grid closures that often assume alignment of the stress with gradients -- a questionable pre-requisite when the wind rotates with height.
+relies on sub-grid closures that often assume alignment of the stress with gradients -- a questionable
+pre-requisite when the wind rotates with height.
 %
-Despite advances in analysis of this simplified set-up \citep{jiang:JAS2018} \todo{(MORE?)},
+\citet{esau:EFM2004} investigated the representation of the Ekman boundary layer by dynamical subgrid closures
+and \cite{zikanov:JFM2003} proposed a closure for the wind profile using a linearized representation of the eddy viscosity.
+%
+Despite advances in analysis of this simplified set-up \citep{jiang:JAS2018}, 
 there is yet insufficient understanding for a quantitative generalization of the results to
-arbitrary external forcing (manifest in variation of the Reynolds number).
+arbitrary external forcing (manifest in variation of the Reynolds number) -- and indeed the fundamental
+questions pertaining to such relatively simple dynamics of turbulence are not reflected in the
+research foci of LES for the ABL over the past 50 years \citep{stoll:BM2020}. 
+%
+\par
 % 
@@
 the geophysical range of scale separation, however, is--and it will remain so for the foreseeable future--out of reach
-for such simulation.
+for such simulation \citep{dimotakis:ARF2005}.
 %
-Here, the routinely employed concept of Reynolds-number similarity can help: it states that certain statistics (normally those of low order) become independent of the Reynolds number when scaled appropriately \todo{Dimotakis / da Silva / Pope referece?} .
+Here, the routinely employed concept of Reynolds-number similarity can help. It postulates the existence
+of \say{\emph{fully developed turbulence}  believed to occur for a sufficiently large but finite Reynolds number}
+\citep{barenblatt:PF1995}.
 %
-It appears that for certain statistics in Ekman flow, this regime is reached with the Reynolds numbers that became
-possible with the increasing capabilities of supercomputering over the past years.
+(Already in \citeyear{moin:ARF1998}, this in fact lead \citeauthor{moin:ARF1998} to the question \emph{how high a Re is high enough}?) 
 %
-The robust features of mean velocity profiles across a range of Reynolds numbers
-are exploited in this work to develop a formulation for
-both the stream-wise and span-wise components of the mean velocity vector as a function of the
-Reynolds number.
-
-\todo{Numerical approach, DNS, Scale-separation, Re-similarity
-  \citep{dimotakis:JFM2000, dasilva:ARF2014, pope:2000, moin:ARF1998, mellado:JAM2018}}
+In fully developed turbulence certain statistics such as dissipation \citep{dimotakis:ARF2005} or profiles of
+mean velocity \citep{barenblatt:JFM1993} become independent of the Reynolds number when scaled appropriately while
+other statistics, such as the near-wall maximum in velocity fluctuation depend on $\RE$ \citep{baars:JFM2020a} and external
+interfaces may exert an impact on near-wall scaling \citep{dasilva:ARF2014}. 
+%
+It appears that for certain statistics in Ekman flow, fully-developed turbulence is reached with the
+Reynolds numbers that became possible with the increasing capabilities of supercomputering over
+the past decades.
+%
+\par
+% 
+This paper exploits the robust features of mean velocity profiles from direct numerical
+simulation across a range of Reynolds numbers to formulate both the stream-wise and span-wise components
+of the mean velocity vector as a function of the Reynolds number.
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
-The non-dimensionalized set of governing equation reads as 
+The non-dimensionalized governing equations read as 
 \begin{subequations} 
@@
 \end{subequations}
-where $u_i=U_i/G$ are the non-dimensional components velocity, $\pi$ is non-dimensional pressure, and $g_j=G_j/G$ are non-dimensionalized components geostrophic wind (with $g_{1}^2+g_{3}^2=1$ by construction).
+where $u_i=U_i/G$ are the non-dimensional components velocity, $\pi$ is non-dimensional pressure, and $g_j=G_j/G$ are non-dimensionalized components
+geostrophic wind (with $g_{1}^2+g_{3}^2=1$ by construction).
 %
+These equations are solved inside a bounded cube of size $L_x \times L_y \times L_z$ with periodic boundary conditions in the lateral directions $Oxz$, a no-slip--no-penetration boundary at $y=0$ and
+a no-penetration, free-slip boundary at $y=L_y$.
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\todo{DNS Code}
+The problem is solved numerically by tLab\footnote{https://github.com/turbulencia/tlab},
+an open-source tool-suite to simulate and analyze turbulent flows.
+% 
+We use here a fourth-order--five-step Runge--Kutta integration and sixth-order compact schemes
+for spatial derivatives in all directions.
+%
+The incompressibility constraint is enforced by a fractional step approach where the Poisson equation
+for the pressure field is soled to machine accuracy using a combined spectral/compact approach as
+described in \citet{mellado:ZaM2012}.
+%
 \par
+%
+Simulations used here are shown in Tab.~\ref{tab:sim-setup}.
+%
+We extend an existing set of simulations for $\RE\in\{125\,000, 281\,250,\ 500\,000\}$ \citep[gray shading; cf.][]{ansorge:BM2014,ansorge:JFM2016} 
+by new simulations at higher Reynolds numbers up to $1.28\times10^6$.
+%
+In total, this yields one order of magnitude variation in terms of the scale separation in the boundary layer.
+%
 \todo{(new and old) Simulations}
+\begin{table}
+  \caption{\label{tab:sim-setup} Direct numerical simulation datasets used in this work}
+  \begin{tabular}{c r r c c c c c}
+    \toprule 
+    $\RE_\Lambda$ & $\RE_D$ & $L_{xy}/\Lambda$ & $N_x\times N_y\times N_z$ & $\Delta x^+$ & $\Delta y^+$ & $\left.\Delta z^+\right|_{z=0}$ \\ 
+    \midrule
+    \rowcolor{gray!30}
+       $125\,000$ &    $500$& 1.08 & $2048\times2048\times192$&4.1& 4.1& 1.05\\
+    \rowcolor{gray!30}
+       $281\,250$ & $   750$& 1.08 & $3072\times3072\times384$&5.6&11.2& 1.60\\
+    \rowcolor{gray!30}
+       $500\,000$ & $1\,000$& 1.08 & $3072\times6144\times512$&9.3& 4.7& 1.14\\
+    \midrule
+       $845\,000$ & $1\,300$& 0.54 & $2560\times5120\times640$&8.9& 4.5& 0.99\\
+
+    $1\,280\,000$ & $1\,600$& 0.54 & $3860\times7680\times960$&8.6& 4.3& 1.00\\ 
+    \bottomrule 
+  \end{tabular} 
+\end{table}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Overview of the new simulations at $\RE_\tau$ up to 3000}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+Uniqueness and necessity of new simulations
+%
+\par
+%
+PARAGRAPH ON BULK PARAMETERS AND SCALING (cf. \ref{tab:bulk_results})
+\begin{table}
+  \caption{Bulk characterization of the simulations \label{tab:bulk_results}}
+  \begin{tabular}{r r r r r r r }
+    \toprule
+    Re & $\delta^+=\RE_\tau$& $u_\star$ & $\alpha$ & $\delta_{95}/\delta$ & $f u_\star^{-3}\int_{0}^{\delta} e \mathrm{d} z $ & $ G^{-3}\int_{0}^{\delta} \epsilon \mathrm{d}z$\\
+      \midrule 
+      \rowcolor{gray!30} 
+      500 & 479 & 0.0619& 25.5 & 0.66& 0.88& 1.31 \\
+      \rowcolor{gray!30} 
+      750 & 886 & 0.0561& 21.0 & 0.65& 0.90& 1.34 \\
+      \rowcolor{gray!30}
+      1000& 1403& 0.0530& 18.8 & 0.62& 0.92& 1.30 \\
+      \midrule
+      1300& 2122& 0.0501& 17.9 & 0.59& 0.85& 1.24 \\
+      1600& 2978& 0.0482& 17.2 & 0.61& 0.91& 1.21 \\ 
+      \bottomrule
+    \end{tabular}
+\end{table} 
+%
+\par
+%
+PARAGRAPH ON VELOCITY AND STRESS PROFILES cf. \ref{fig:stresses}
+\begin{figure}
+  \includegraphics[width=0.66\textwidth]{../plot/eddy_viscosity.pdf}
+  \caption{Reynolds stresses and eddy viscosity across the Ekman layer plotted in ineer scaling for
+    various Reynolds number
+    \label{fig:stresses}} 
+\end{figure} 
+%
+\par
+%
+PARAGRAPH ON TURBULENCE STRUCTURE (2D SLICES, SPECTRA?) cf. Fig.~\ref{fig:slices}
+
+\begin{figure}
+  \centerline{\includegraphics[width=0.33\textwidth]{../plot/PLANES/tke_xz_10.pdf}
+  \includegraphics[width=0.33\textwidth]{../plot/PLANES/tke_xz_50.pdf}
+  \includegraphics[width=0.33\textwidth]{../plot/PLANES/tke_xz_400.pdf}}
+  \centerline{\includegraphics[width=\textwidth]{../plot/PLANES/tke_xy_1.pdf}}
+  \caption{horizontal slices of turbulence kinetic energy in the Buffer layer (i=10), Logarithmich Layer (i=100) and outer lzayer (i=400); coloring between percentiles 4 and 96 of the respective image.
+Lower panel:streamwise--vertical intersect through the domain}
+  \label{fig:slices} }
+
+\end{figure}
 
@@
   \centerline{\includegraphics[width=0.7\textwidth]{../plot/ustar_alpha.pdf}}
-  \caption{Variation of $Z_\star$ and surface veering with Reynolds number according to
+  \caption{Variation of geostrophic drag, $Z_\star$, and surface veering, $\alpha$, with Reynolds number according to
     the theory by Spalart et al. (1989) and as estimated from DNS data}
@@
 which quantifies the 'weak' dependence of $u_\star$ on the Reynolds number as an approximately logarithmic one, at least
-for problems with a scale separation on the order that is relevant to geophysical problem ($Re_D<10^{8}$).  
+for problems with a scale separation on the order that is relevant to geophysical problems ($Re_D<10^{8}$).  
 %
@@
 %
-The textbook solution for the Ekman layer reads as 
-\todo{equations / figures}
 \begin{figure}
-  \includegraphics[width=\textwidth]{../plot/outer_layer.pdf}
+  \includegraphics[width=0.5\textwidth]{../plot/outer_layer_u.pdf}\includegraphics[width=0.5\textwidth]{../plot/outer_layer_w.pdf}
   \caption{} 
@@
 %
-\paragraph{Inner Layer.} It is important to note that the logarithmic layer observed here for the spanwise velocity does not
+\subsub{Inner Layer.} It is important to note that the logarithmic layer observed here for the spanwise velocity does not
 conincide with the log-region of the streamwise velocity but is rather a 'continuation' through the outer layer.
@@
 Matching the profile at $\xi=0$, one obtains
-$A = \Delta u_\text{match} $ and $B=-\Delta w_\text{match} $. Otherwise, it is
-\begin{subequations} 
-\begin{eqnarray}
-  e^\gzm \Delta u_\text{match} = A \cos\gzm + B\sin\gzm \\
-  e^\gzm (\cos\gzm \Delta u_\text{match} - \sin\gzm \Delta w_\text{match}) = A (\cos^2\gzm + \sin^2\gzm)
-\end{eqnarray}
-\begin{eqnarray}
-  \Rightarrow A &=&e^\gzm (\cos\gzm\Delta u_\text{match} + \sin\gzm \Delta w_\text{match})\\
-  B&=&  \frac{e^\gzm \Delta u_\text{match} - A \cos \gzm}{\sin\gzm} \\
-  &=& e^\gzm \frac{ \sin^2\gzm \Delta u_\text{match}+\sin\xi\cos\xi \Delta w_\text{match} }{\sin\xi }  \\
-  &=& e^\gzm (\sin\xi\Delta u_\text{match} + \cos\xi\Delta w_\text{match}) 
-\end{eqnarray}
-\end{subequations}
+$A = \Delta u_\text{match} $ and $B=-\Delta w_\text{match} $; and when the direction $Ox$ is aligned
+with the geostrophic wind, we obtain the textbook-case $A=|\mathbf{G}|$ and $B=0$.
+
+Otherwise, choosing $B \ne 0$ allows to introduce a phase shift of the Ekman rotation with respect to
+the decay of the wind spiral.
+%
+As, however, in our context, the thickness and position of the spiral can already be controlled by
+the eddy viscosity and an offset in $\zeta$, here we let $B=0$.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Determining parameters A and B from profile values at a given height
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% 
+%% \begin{subequations} 
+%% \begin{eqnarray}
+%%   e^\gzm \Delta u_\text{match} = A \cos\gzm + B\sin\gzm \\
+%%   e^\gzm (\cos\gzm \Delta u_\text{match} - \sin\gzm \Delta w_\text{match}) = A (\cos^2\gzm + \sin^2\gzm)
+%% \end{eqnarray}
+%% \begin{eqnarray}
+%%   \Rightarrow A &=&e^\gzm (\cos\gzm\Delta u_\text{match} + \sin\gzm \Delta w_\text{match})\\
+%%   B&=&  \frac{e^\gzm \Delta u_\text{match} - A \cos \gzm}{\sin\gzm} \\
+%%   &=& e^\gzm \frac{ \sin^2\gzm \Delta u_\text{match}+\sin\xi\cos\xi \Delta w_\text{match} }{\sin\xi }  \\
+%%   &=& e^\gzm (\sin\xi\Delta u_\text{match} + \cos\xi\Delta w_\text{match}) 
+%% \end{eqnarray}
+%% \end{subequations}
 

### 2021-07-29 plot/plot.py

@@ -18,12 +18,13 @@ D2R = np.pi / 180.
 use_data=False   # if true -- the netcdf files from DNS are needed 
 plot_ustar_alpha=False
 plot_summary=False
+plot_evisc=True
 plot_visc_outer=False
-print_table=False
+print_table=True
 plot_outer_log=False
 plot_total_rot=False
-plot_convergence_1000=False
-plot_re1600=True
+plot_convergence=False
+plot_re1600=False
 plot_les_comp=False
 plot_applications=False
 
@@ -32,6 +33,7 @@ colors = {400 : 'gray',
           750 : 'cyan',
           1000: 'blue',
           1300: 'red',
+          1301: 'red',
           1600: 'black',
           5000: 'black',
           10000:'gray',
@@ -46,7 +48,8 @@ files = { 400: base+'avg_flw_ri0.0_re400.0_1024x384x1024.nc',
           500: base+'avg_flw_ri0.0_re500.0_2048x192x2048.nc',
           750: base+'avg_flw_ri0.0_re750.0_3072x384x3072.nc',
           1000:base+'avg_flw_ri0.0_re1000.0_co0.0_1536x512x3072.nc',
-          1300:base+'avg_flw_ri0.0_re1300.0_co0_2560x640x5120.nc',
+          1300:base+'avg_flw_ri0.0_re1300.0_co0_2560x640x5120aligned.nc',
+          1301:base+'avg_flw_ri0.0_re1300.0_co0_2560x640x5120aligned.nc',
           1600:base+'avg_flw_ri0.0_re1600.0_co0_3840x960x7680.nc' } 
 
 sc=EkmanUniversalClass() # yp_ref,up_ref,wp_ref,deltap_ref,us_ref,al_ref*D2R-geo_rotate,plot=False)
@@ -261,8 +264,89 @@ if plot_re1600:
     ax1.set_xlim(1,1e4)
     ax2.set_xlim(1,1e4)
     
+if plot_evisc:
+    fig=plt.figure(figsize=(7,4))
+    ax=fig.add_axes([0.1,0.11,0.8,0.87])
     
+    for re in [500,750,1000,1300,1600]:
+        nu=2./(re*re)
[…]

### 2021-09-09 write/manuscript.tex
fixed problems in velocity profile (divergence at z=0); added section 3 to manuscript

@@
   %
+  Viscous stress scales universally across the boundary layer in inner units while the total stress
+  becomes universal as a function of outer height.
+  %
+  This implies a mixed scaling for the turbulent stress and eddy viscosity across the inner layer
+  and convergence to a universal scaling as function of the outer height across the outer layer
+  for increasing scale separation vide Reynolds numbers. 
 \end{abstract}
@@
 The dimensional velocity vector of the numerical simulations is $\underline{U} = (U_1,U_2,U_3) = (U,V,W)$ over the coordinate system
-$Oxyz$, where an approximate alignment (plus/ minus few degrees centigrade) of the direction $Ox$ with the surface shear stress is achieved.
+$Oxyz$, where an approximate alignment (plus/ minus few degrees) of the direction $Ox$ with the surface shear stress is achieved.
 The coordinate $Oy$ points away from the wall,  and $Oz$ points in the span-wise direction normal to $Oxy$.
 For analysis of the results, we use two coordinate systems that are
-(i)  exactly aligned with the surface shear stress and labelled by an upper index $\alpha$ as in $\underline{U}^\alpha$ for the velocity vector, and
+(i)  exactly aligned with the surface shear stress
+\begin{subequations}
+\begin{equation}
+  \underline{\tau}_\mathrm{sfc} = \left(\begin{array}{c} \tau_x\\ \tau_y\\\tau_z\end{array}\right) = - \nu \left( \frac{\partial U}{\partial z} \hat{e}_x + \frac{\partial V}{\partial z} \hat{e}_y\right)
+\end{equation}
+and labelled by an upper index $\alpha$ as in $\underline{U}^\alpha$ for the velocity vector, and
 (ii) the coordinate system aligned with the free-atmosphere geostrophic wind labelled by an upper index $G$ as in $\underline{U}^G$.
 %
-We denote by $u_\star$ the modulus surface friction velocity and let $Z_\star=1/u_\star$ for brevity. 
+We denote the normalized modulus of the surface friction by
+\begin{equation}
+  u_\star = \sqrt{ || - \underline{\tau}_\mathrm{sfc} || } 
+  \label{eqn:ustar_def}
+\end{equation} and let $Z_\star=G/u_\star$;
+the surface veering angle $\alpha_\star$ is the angle between $\underline{\tau}$ and the geostrophic wind 
+\begin{equation}
+  \alpha_\star = \sphericalangle \left(\underline{G}, \underline{\tau}_\mathrm{sfc} \right). 
+\end{equation}
+\end{subequations}
+Analogously, we denote the height-local veering of the wind
+$\alpha(z)= \sphericalangle \left( \underline{G},\underline{U}(z)\right)$. 
 %
@@
 that is governed by the physical parameters
-(i)   geostrophic wind magnitude $G=\sqrt{G_1^2+G_3^2}$ (where $\mathbf{G}=(G_1,0,G_3)$ is the geostrophic wind vector),
+(i)   geostrophic wind magnitude $G=\sqrt{G_1^2+G_3^2}$ (where $\mathbf{G}=(G_1,G_2,0)$ is the geostrophic wind vector),
 (ii)  Coriolis parameter $f$ (representing the angular rotation), and
@@
 %
-We use the geostrophic wind as velocity and the Coriolis parameter $f$ as time scale for the non-dimensional framework which implies the Rossby radius $\LR=G/f$ as length scale, such that
+We use the geostrophic wind as velocity and the Coriolis parameter $f$ as time scale for the non-dimensional framework which implies the Rossby radius $\LR=G/f$ as length scale, such that the Reynolds number governing the problem reads as
 \begin{equation}
-  \RE=\frac{G \LR }{\nu}. 
+  \RE_\Lambda=\frac{G \LR }{\nu}. 
 \end{equation}
+%
+The scales used in defining $\RE_\Lambda$ loose there relevance in describing the turbulent state of the flow. The friction Reynolds number
+\begin{equation}
+  \RE_\tau = \frac{u_\star\delta}{\nu} = \frac{\RE_\Lambda}{Z_\star^2}, 
+\end{equation}
+where $\delta=u_\star/f$, gives a measure for the turbulence scale separation that is comparable to other
+wall-bounded flow geometries such as rotating or non-rotating channel, pipe, or Couette flow.
 %
@@
 where $u_i=U_i/G$ are the non-dimensional components velocity, $\pi$ is non-dimensional pressure, and $g_j=G_j/G$ are non-dimensionalized components
-geostrophic wind (with $g_{1}^2+g_{3}^2=1$ by construction).
+geostrophic wind (with $g_{1}^2+g_{2}^2=1$ by construction).
 %
-These equations are solved inside a bounded cube of size $L_x \times L_y \times L_z$ with periodic boundary conditions in the lateral directions $Oxz$, a no-slip--no-penetration boundary at $y=0$ and
-a no-penetration, free-slip boundary at $y=L_y$.
+These equations are solved inside a bounded cube of size $L_x \times L_y \times L_z$ with periodic boundary conditions in the lateral directions $Oxy$, a no-slip--no-penetration boundary at $z=0$, and
+a no-penetration, free-slip boundary at $z=L_z$.
 %
@@
 The incompressibility constraint is enforced by a fractional step approach where the Poisson equation
-for the pressure field is soled to machine accuracy using a combined spectral/compact approach as
+for the pressure field is solved to machine accuracy using a combined spectral/compact approach as
 described in \citet{mellado:ZaM2012}.
@@
 %
-We extend an existing set of simulations for $\RE\in\{125\,000, 281\,250,\ 500\,000\}$ \citep[gray shading; cf.][]{ansorge:BM2014,ansorge:JFM2016} 
-by new simulations at higher Reynolds numbers up to $1.28\times10^6$.
+We extend an existing set of simulations for $\RE_\Lambda\in\{125\,000; 281\,250;\ 500\,000\}$ \citep[gray shading; cf.][]{ansorge:BM2014,ansorge:JFM2016} 
+by new simulations at higher Reynolds numbers up to $1.28\times10^6$ with a horizontal domain extent up to $3.3\times 10^4$ viscous units. 
 %
@@
 %
-\todo{(new and old) Simulations}
 \begin{table}
-  \caption{\label{tab:sim-setup} Direct numerical simulation datasets used in this work}
+  \caption{Direct numerical simulation datasets used in this
+    work \label{tab:sim-setup}}
   \begin{tabular}{c r r c c c c c}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Overview of the new simulations at $\RE_\tau$ up to 3000}
+\section{Scaling behavior of the flow for $\RE_\tau$ up to 3000}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-Uniqueness and necessity of new simulations
+The generalization of profiles to arbitrary Reynolds numbers 
+requires sufficient scale separation in the simulations to not only quantify the effect of
+the Reynolds number on low-order flow statistics, but also assess the corresponding rate-of-change
+to eventually allow for an extrapolation of the findings.
+%
+While the simulations available give confidence in a first-order representation of the turbulent
+problem, the estimation of higher-order effects such as the dependency of the Reynolds number requires
+a broader scale separation that is made available by the two new simulations for 
+(cf. Tab.~\ref{tab:sim-setup}). 
+%
+%%% We thus require sufficient scale separation; while we are certain that even the lower Reynolds numbers
+%%% are well within the turbulent regime, the range of separation required to estimate $\RE$-effects should
+%%% be about one magnitude at the least, i.e. we anticipate that data for $\mathcal{O}(\RE)=10^6$ would 
+%%% allow us to extrapolate findings. 
+%
+Data at such scale separation has been obtained previously \citep[cf.][]{spalart:PF2008,spalart:PF2009}, 
+but we also need high confidence with respect to the convergence of simulation data for slow oscillations 
+and with respect to sampling convergence, which translates to two further requirements on the data: 
+%
+First, data should be free of artifacts from long-term oscillations across the vertical
+extent of the domain--primarily, simulations should be free of effects originating from the
+inertial oscillation; this is achieved here by replacing the mean value of the three-dimensional
+velocity fields by the time mean over a whole inertial oscillation. 
+%
+Second, high accuracy is also needed in terms of the statistical convergence of averages,
+bulk measures and large-scale structures; this requires a domain size $L_{x} > \mathcal{O}(\delta_{95})$. 
+%
+We use here $L_{x}=L_{y}=1.08\Lambda$ for cases with $\RE \le 1000$ and
+$L_{x}=L_{y}=0.54\Lambda$ for $\RE \ge 1300$ which corresponds to $L_{x}/\delta_{95}\approx 25$ for $Re_D=500$
+and $L_{x}/\delta_{95} \approx 18$ for $Re_D=1300$.
 %
@@
 %
-PARAGRAPH ON BULK PARAMETERS AND SCALING (cf. \ref{tab:bulk_results})
+Bulk parameters of the simulations are given in Tab.~\ref{tab:bulk_results}.
+%
+The surface stress is characterized by $u_\star$ and $\alpha_\star$ in relation to the geostrophic wind vector
+and discussed in more detail as the drag law below in Sec.~\ref{sec:drag-law}
+(we find the expected slight decrease of $u_\star$ and $\alpha_\star$ with increasing $\RE$).
+%
+The boundary-layer height estimated from the 95\% stress reduction, $\delta_{95}$, is around $0.6\delta$ to $0.66\delta$. 
+%
+Interestingly, the integrated TKE $\int_{0}^{\delta} e\mathrm{d}z$ stays constant when normalized by the friction velocity $u_\star$
+while the integrated dissipation $\int_{0}^{\delta} \epsilon\mathrm{d}z$ exhibits inviscid scaling when normalized
+by the magnitude $G$ of the geostrophic wind. 
+%
+(Alternatively normalized TKE and dissipation,
+$f G^{-3} \int e \mathrm{d} z$ and $u_\star^{-3} \int \epsilon\mathrm{d}z$, exhibit substantial
+dependence on $\RE$ for the variation of $u_s\tar$.)
+%
+This indicates that the bulk dissipation is governed by the forcing $G$--irrespective of $\RE$. 
+%
+Changes in $\RE$, however, affect the level and organization of turbulence, and the parameter
+representing this dependency is the friction velocity $u_\star$ which describes the turbulence
+production processes in the surface layer, in particular in the buffer layer. 
+% 
 \begin{table}
-  \caption{Bulk characterization of the simulations \label{tab:bulk_results}}
+  \caption{Bulk characterization of the simulations for different Reynolds numbers 
+    Viscous Reynolds number $\RE$,
+    friction Reynolds number  $\RE_\tau$,
+    friction velocity $u_\star$,
+    surface veering angle $\alpha_\star$,
+    normalized boundary layer depth $\delta_{95}/\delta$
+    inner normalization of vertically integrated TKE,
+    outer normalization of vertically integrated dissipation.
+    \label{tab:bulk_results}}
   \begin{tabular}{r r r r r r r }
     \toprule
-    Re & $\delta^+=\RE_\tau$& $u_\star$ & $\alpha$ & $\delta_{95}/\delta$ & $f u_\star^{-3}\int_{0}^{\delta} e \mathrm{d} z $ & $ G^{-3}\int_{0}^{\delta} \epsilon \mathrm{d}z$\\
+    Re & $\delta^+=\RE_\tau$& $u_\star$ & $\alpha_\star$ & $\delta_{95}/\delta$ & $f u_\star^{-3}\int_{0}^{\delta} e \mathrm{d} z $ & $ G^{-3}\int_{0}^{\delta} \epsilon \mathrm{d}z$\\
       \midrule 
@@
 %
-PARAGRAPH ON VELOCITY AND STRESS PROFILES cf. \ref{fig:stresses}
-\begin{figure}
-  \includegraphics[width=0.66\textwidth]{../plot/eddy_viscosity.pdf}
-  \caption{Reynolds stresses and eddy viscosity across the Ekman layer plotted in ineer scaling for
-    various Reynolds number
-    \label{fig:stresses}} 
-\end{figure} 
+Velocity profiles in inner and units (Fig.~\ref{fig:profiles}a,b) are in accordance
+with previous work \citep{coleman:JFM1992,spalart:PF2008,spalart:PF2009,ansorge:BM2014,ansorge:BM2019}:
+%
+The profile of the shear-aligned streamwise velocity component are well-collapsed for the Reynolds
+numbers investigated below $z^-\approx 0.15$ (circles in Fig.~\ref{fig:profiles}a); the logarithmic
+law is appropriate for $50 < z^+ < 0.15 Re_\tau$, where $z^+ = Re_\tau z^-$.
+%
+While the profiles $U^{\alpha_\star+}(z^+)$ diverge between different $\Re$ beyond $z^-=0.15$,
+the corresponding profiles $U^{\alpha_\star}(z^-)/u_\star$ agree closely,
+irrespective of $\RE$ which illustrates the
+inner--outer scale-duality in this external flow. 
+%
+It is noteworthy that also in the outer layer of the flow, $u_\star$--and not the magnitude the
+geostrophic wind speed $G$--is the governing velocity scale for inviscid, i.e. $\RE$-independent,
+normalization. \todo{WHY? cf. TKE/Dissipation normalization above}
+%
+No collapse in inner units is found for the profiles of spanwise velocity $V^{\alpha_\star+}(z^+)$;
+%
+When normalized in outer units, the profiles of spanwise velocity $V^{\alpha_\star}(z^-)/u_\star$
+agree well beyond $z^-\approx 0.3$ whereas the profiles of the streamwise component collapse also
+within the overlap layer that extends much closer down to the surface
+(circles in Fig.~\ref{fig:profiles}b).
+%
+While low-$\RE$ effects appear to be present for $\mathcal{O}(\RE)<10^3$, the spanwise component
+appears to converge to an $\RE$-independent limit within the range of scale separation considered
+here.
 %
@@
 %
-PARAGRAPH ON TURBULENCE STRUCTURE (2D SLICES, SPECTRA?) cf. Fig.~\ref{fig:slices}
+\begin{figure}
+  \begin{flushleft}
+    \textbf{a)}\hspace{0.47\textwidth} \textbf{b)}\\
+    \includegraphics[width=\textwidth]{../plot/uv_innerouter.pdf}\\
+    \caption{
+      \textbf{a)} Shear-aligned velocity profiles in inner units;
+      \textbf{b)} Shear-aligned velocity deficit in outer units; circles mark the
+      height $z^-=0.15$; dashed line gives the best logarithmic fit to the data. 
+      \label{fig:profiles}
+    }
+  \end{flushleft}
+\end{figure}
 
 \begin{figure}
-  \centerline{\includegraphics[width=0.33\textwidth]{../plot/PLANES/tke_xz_10.pdf}
-  \includegraphics[width=0.33\textwidth]{../plot/PLANES/tke_xz_50.pdf}
-  \includegraphics[width=0.33\textwidth]{../plot/PLANES/tke_xz_400.pdf}}
-  \centerline{\includegraphics[width=\textwidth]{../plot/PLANES/tke_xy_1.pdf}}
-  \caption{horizontal slices of turbulence kinetic energy in the Buffer layer (i=10), Logarithmich Layer (i=100) and outer lzayer (i=400); coloring between percentiles 4 and 96 of the respective image.
-Lower panel:streamwise--vertical intersect through the domain}
-  \label{fig:slices} }
+  \begin{flushleft} 
+    \textbf{a)}\hspace{0.47\textwidth} \textbf{b)}\\
+    \includegraphics[width=\textwidth]{../plot/stresses.pdf}
+    \textbf{c)}\hspace{0.47\textwidth} \textbf{d)}\\    
+    \includegraphics[width=\textwidth]{../plot/eddy_viscosity.pdf}\\
+  \end{flushleft} 
+  \caption{
+    \textbf{a-b)} Profiles of
+    the turbulent stress $S_\mathrm{turb}^+$ (solid), 
+    the viscous stress  $S_\mathrm{visc}^+$ (dotted), 
+    and the total stress, $S^+=S_\mathrm{visc}^++S_\mathrm{turb}^+$ (dashed)
+    as a function of inner height \textbf{(a)} and outer height \textbf{(b)}. 
+    \textbf{c-d)} Normalized eddy viscosity $\nu_E$ (solid) plotted versus
+    inner \textbf{(c)} and outer height \textbf{(d)}.
+    Different colors are for different Reynolds numbers (cf. Tab.~\ref{tab:sim-setup}).
+    Circles in \textbf{(a)} and \textbf{(b)} denote the height
+    $z^-=0.15$, diamonds in \textbf{(c)} and \textbf{(d)} are for $z^+=50$
+    as in Fig.~\ref{fig:profiles}.
+    \label{fig:stresses}}  
+\end{figure} 
+The normalized viscous stress
+\begin{subequations}
+\begin{equation}
+  S_\mathrm{visc}= \nu \sqrt\left(\frac{\p U}{\p z}\right)^2  + \left(\frac{\p V}{\p z}\right)^2}
+\end{equation}
+exhibits universal scaling when considered as $S_\mathrm{visc}^+(z^+)$ (Fig.~\ref{fig:stresses}c); this normalization is also appropriate in the outer layer where the viscous stress
+is, however, small. 
+%
+Small deviations from the universal profile are observed for the smallest Reynolds number $\RE=500$; 
+we attribute these to low-$\RE$ effects in the only transitionally turbulent flow ($\RE_\tau=479$). 
+%
+In contrast to the viscous stress, the total stress follows outer normalization, i.e.
+$S^+(z^-)$ is universal.
+%
+This implies a mixed scaling for the turbulent stress, 
+\begin{equation}
+  S_\mathrm{turb} = \sqrt{\overline{uw}^2 + \overline{vw}^2}. 
+\end{equation} 
+Indeed, $S_\mathrm{turb}$ only follows inner normalization below $z^+\lesssim 20$
+(where the turbulent contribution is negligible).
+%
+In the outer layer, where $S_\mathrm{visc}\rightarrow 0$, $S_\mathrm{turb}^+$
+follows outer normalization for $z^-\gtrsim0.15$---with increasing accuracy for larger $\RE$ and larger distance from
+the surface. 
+%
+In the overlap region, i.e. for $z^+>20$ and $z^-<0.15$, the mixed scaling for the
+turbulent stress can be expressed as
+\begin{equation}
+  S_\mathrm{turb}^+(z^+,\RE_\tau) = S^+(z^-) - S_\mathrm{visc}^+(z^+), 
+\end{equation}
+where $z^-=z^+/\RE_\tau$. 
+%
+\par
+%
+The Eddy viscosity plays a crucial part when modelling profiles and the vertical
+transport in turbulent flow. 
+%
+In analogy to the Fick-law for molecular diffusion, the eddy diffusivity is
+the effective diffusivity that yields the turbulent transport $S_\mathrm{turb}$ based
+on the strain rate.
+%
+For the symmetries in the flow (horizontal homoegeneity, and $W=0$), it is
+%
+\begin{equation}
+  \nu_{E}= \frac{S_\mathrm{turb}}{\sqrt{\left(\frac{\p U}{\p z} \right)^2 + \left(\frac{\p V}{\p z}\right)^2}} = \nu \frac{S_\mathrm{turb}}{S_\mathrm{visc}}.  
+\end{equation}
+%
+The inner normalization of $\nu_E$ s
+
+\begin{eqnarray}
+ \nu_E^+  \frac{Z_{\star}^{2}}{\delta+} &=& \frac{\nu_E^+}{\nu}
+\end{eqnarray} 
+\par
+%
+\todo{PARAGRAPH ON TURBULENCE STRUCTURE (2D SLICES, SPECTRA?) cf. Fig.~\ref{fig:slices}}
+
+\begin{figure}
+  \centerline{
+    \includegraphics[trim=0 0 1152 0, clip, width=0.33\textwidth]{../plot/PLANES/tke_xz_10.pdf}
+    \includegraphics[trim=1152 0 0 0, clip, width=0.33\textwidth]{../plot/PLANES/tke_xz_50.pdf}
+    \includegraphics[trim=0 0 1152 0, clip, width=0.33\textwidth]{../plot/PLANES/tke_xz_400.pdf}}
+  \centerline{
+    \includegraphics[trim=1152 0 0 0, clip, width=\textwidth]{../plot/PLANES/tke_xy_1.pdf}}
+  \caption{horizontal slices of turbulence kinetic energy in the
+    Buffer layer (i=10),
+    logarithmic Layer (i=100), and
+    outer layer (i=400);
+    coloring between percentiles 4 and 96 of the respective image. Lower panel:streamwise--vertical intersect through the domain 
+    \label{fig:slices}}
 
@@
   \centerline{\includegraphics[width=0.7\textwidth]{../plot/ustar_alpha.pdf}}
-  \caption{Variation of geostrophic drag, $Z_\star$, and surface veering, $\alpha$, with Reynolds number according to
-    the theory by Spalart et al. (1989) and as estimated from DNS data}
-  \label{fig:drag_law}
+  \caption{Variation of geostrophic drag, $Z_\star$, and surface veering, $\alpha_\star$, with Reynolds number according to
+    the theory by Spalart et al. (1989) and as estimated from DNS data
+    \label{fig:drag_law}}
 \end{figure}
@@
 \subsection{Drag-Law}
+\label{sec:drag-law}
 A drag-law for Ekman flow determines two key parameters of the flow as a function of Reynolds number alone.
@@
 These are, first, the modulus of the surface friction normalized by the geostrophic wind forcing, $u_\star$ and,
-second, the direction of surface shear stress, $\alpha$, also termed wind veering. 
+second, the direction of surface shear stress, $\alpha_\star$, also termed wind veering. 
 %
@@
 \begin{equation}
-  (U,W)^{\alpha+}  = (z^+,0)  
+  (U,W)^{\alpha_\star+}  = (z^+,0)  
 \end{equation}
@@
 viscous sub-layer \cite{foken:m2002,foken:BM1978}.
-In fact, this law of the wall has no degree of freedom given the drag, i.e. once $u_\star$ and $\alpha$ are defined. 
+In fact, this law of the wall has no degree of freedom given the drag, i.e. once $u_\star$ and $\alpha_\star$ are defined. 
 %
@@
 \begin{equation}
-  U_\text{inner}^{\alpha}^+ = \frac{ y^+ + \gamma_4 (y^+)^{4} + \gamma_6 (y^+)^{6} } { 1 + \gamma_6 /u_{ref} (y^+ )^6}, 
+  U_\text{inner}^{\alpha_\star}^+ = \frac{ y^+ + \gamma_4 (y^+)^{4} + \gamma_6 (y^+)^{6} } { 1 + \gamma_6 /u_{ref} (y^+ )^6}, 
 \end{equation} 
@@
 \begin{equation}
-  U_\text{log}^{\alpha}^+ = \frac{1}{\kappa} \log y^+ + C 
+  U_\text{log}^{\alpha_\star}^+ = \frac{1}{\kappa} \log y^+ + C 
 \end{equation}
@@
   \includegraphics[width=\textwidth]{../plot/u_profile.pdf}
-  \caption{Shear-aligned profiles of velocity components $U^{\alpha+}$ in inner (left) and outer (right) units.} 
+  \caption{Shear-aligned profiles of velocity components $U^{\alpha_\star+}$ in inner (left) and outer (right) units.} 
 \end{figure} 
@@
 While the universal profiles in vicinity of the wall implies that the profile be zero or at least small with respect to the stream-wise
-component, the veering requires a value of $V_{top}=U_G\sin\alpha$ in the free stream (and thus also at the top of the boundary layer if we assume
+component, the veering requires a value of $V_{top}=U_G\sin\alpha_\star$ in the free stream (and thus also at the top of the boundary layer if we assume
 that substantial velocity gradients are confined to the turbulent part of the flow). 
@@
  \underline{U}^{G}(z\mapsto\infty) &=& \underline{G}^G = \left(\begin{array}{c}1\\0\end{array}\right) \\
-   \left.\frac{\partial V^G}{\partial U^G}\right|_{z=0} &=& \sin\alpha
+   \left.\frac{\partial V^G}{\partial U^G}\right|_{z=0} &=& \sin\alpha_\star
 \end{eqnarray} 
@@
 %
-Outer scaling of the velocity profile further implies that the velocity deficity of the spanwise component $V^\alpha-G^\alpha$
+Outer scaling of the velocity profile further implies that the velocity deficity of the spanwise component $V^\alpha_star-G^\alpha_\star$
 be a universal function of the outer height $z^-$, i.e.  
 \begin{eqnarray}
-  V^\alpha-G^\alpha = f^\alpha_V(z^-).  
+  V^\alpha_\star-G^\alpha_\star = f^\alpha_\star_V(z^-).  
 \end{eqnarray}
 %
-Interpreting the shear-aligned spanwise velocity deficit $f_{V}^{\alpha}(z^-)$ as 
+Interpreting the shear-aligned spanwise velocity deficit $f_{V}^{\alpha_\star}(z^-)$ as 
 a signature of outer rotation may lead one to conclude that this universal function
@@
 %
-This cannot be the case for the variation of $f_{V}^{\alpha}(z^-)$ across the boundary layer (i.e. between $0<z^-<1$)
+This cannot be the case for the variation of $f_{V}^{\alpha_\star}(z^-)$ across the boundary layer (i.e. between $0<z^-<1$)
 must fulfill the drag law and boundary conditions;
-at the same time, the weak dependence of $\alpha$ and $u_\star$ on $Re$ implies this range depends on $Re$.
+at the same time, the weak dependence of $\alpha_\star$ and $u_\star$ on $Re$ implies this range depends on $Re$.
 %

### 2021-09-09 plot/EkmanProfileClass.py

@@ -98,7 +98,7 @@ class EkmanUniversalClass:
 
         re = np.sqrt(2*deltap/(us*us)) # nu = us*us/deltap
 
-        log_law = self.profile_log(yp,re) # np.log(yp[1:])/k + C
+        log_law = self.profile_log(yp) # np.log(yp[1:])/k + C
         log_dev_outer_u=up-log_law
         log_dev_outer_u[:i2]=0.
         log_dev_outer_w=wm
@@ -178,11 +178,13 @@ class EkmanUniversalClass:
             
         return(1./z,aleph) 
 
-    def profile_log(self,yp,re):
+    def profile_log(self,yp):
         self.qinit()
 
         x=np.zeros(len(yp))
         x[1:] = np.log(yp[1:])/self.KAPPA  + self.C
+        if yp[0] > 0:
+            x[0] =np.log(yp[0])/self.KAPPA  + self.C 
         return x 
 
     def erf_transition(self,w,l,x):
@@ -207,7 +209,7 @@ class EkmanUniversalClass:
         
         z=1/us_loc 
         dp_loc=re**2/(2*z**2)
-        ym=np.array(yp/dp_loc)
+        ym=np.array(yp)/dp_loc
         trans_scale=2.0
 
         ################################################################################
@@ -222,7 +224,7 @@ class EkmanUniversalClass:
         
         ################################################################################
         # STREAMWISE COMPONENT (shear-aligned) 
-        log_law=self.profile_log(yp,re) # np.log(yp)/self.KAPPA + self.C   
+        log_law=self.profile_log(yp) # np.log(yp)/self.KAPPA + self.C   
 
         # inner-layer (empirical profile for viscous and buffer layer) 
         c4=-0.0003825; c6=6.32e-6; u_ref = 0.07825
@@ -232,7 +234,7 @@ class EkmanUniversalClass:
         u[1:]=(1-wgt_i)*u_visc[1:] + wgt_i*log_law[1:]
 
         # outer-layer deficit
-        ctr= 0.33 - 120/re 
+        ctr=0.28- 2.25*np.sqrt(1./re)
[…]

### 2021-09-09 plot/check_ek.py

@@ -1,0 +1,23 @@
+#!/opt/local/bin/python
+import matplotlib.pyplot as plt 
+import numpy as np
+from EkmanProfileClass import Complex, EkmanUniversalClass, build_grid
+
+
+re=500
+ny,yp,ym=build_grid(1500,2)
+
+sc=EkmanUniversalClass()
+
+up,wp=sc.profile_plus(yp,re)
+
+fig=plt.figure()
+ax=fig.add_axes([0.1,0.1,0.8,0.8]) 
+ax.plot(yp,wp)
+ax.plot(yp,up)
+ax.set_xscale('log')
+plt.show()
+plt.close('all')
+
+quit() 
+

### 2021-09-09 plot/plot.py

@@ -19,8 +19,9 @@ use_data=False   # if true -- the netcdf files from DNS are needed
 plot_ustar_alpha=False
 plot_summary=False
 plot_evisc=True
+plot_profiles=True
 plot_visc_outer=False
-print_table=True
+print_table=False
 plot_outer_log=False
 plot_total_rot=False
 plot_convergence=False
@@ -49,7 +50,7 @@ files = { 400: base+'avg_flw_ri0.0_re400.0_1024x384x1024.nc',
           750: base+'avg_flw_ri0.0_re750.0_3072x384x3072.nc',
           1000:base+'avg_flw_ri0.0_re1000.0_co0.0_1536x512x3072.nc',
           1300:base+'avg_flw_ri0.0_re1300.0_co0_2560x640x5120aligned.nc',
-          1301:base+'avg_flw_ri0.0_re1300.0_co0_2560x640x5120aligned.nc',
+          1301:base+'avg_flw_ri0.0_re1300.0_co0_2560x640x5120.nc',
           1600:base+'avg_flw_ri0.0_re1600.0_co0_3840x960x7680.nc' } 
 
 sc=EkmanUniversalClass() # yp_ref,up_ref,wp_ref,deltap_ref,us_ref,al_ref*D2R-geo_rotate,plot=False)
@@ -265,9 +266,13 @@ if plot_re1600:
     ax2.set_xlim(1,1e4)
     
 if plot_evisc:
-    fig=plt.figure(figsize=(7,4))
-    ax=fig.add_axes([0.1,0.11,0.8,0.87])
-    
+    fig=plt.figure(figsize=(12,4.5))
+    ax2=fig.add_axes([0.06,0.11,0.43,0.87])
+    ax1=fig.add_axes([0.55,0.11,0.43,0.87])
+    fig2=plt.figure(figsize=(12,4.5)) 
+    ax3=fig2.add_axes([0.55,0.11,0.43,0.87])
+    ax4=fig2.add_axes([0.06,0.11,0.43,0.87])
+
     for re in [500,750,1000,1300,1600]:
         nu=2./(re*re)
         f=Dataset(files[re],'r',format='NETCDF4')
@@ -279,24 +284,24 @@ if plot_evisc:
         deltap=us*us/nu  # = delta * (us/nu) =  (us/f) * us /  nu [with f=1]
         ny=len(y)
         nt=len(us_arr)
-        u=f['rU'][:,:]/us
[…]

### 2021-09-09 plot/test_df.py

@@ -1,0 +1,21 @@
+#!/opt/local/bin/python
+import numpy as np
+import my_pylib as mp 
+us_loc=0.0619
+dp_loc=479
+
+yp=np.arange(0,10.5,0.25) 
+
+c_off=40/(dp_loc*us_loc)
+c1=26/(dp_loc*us_loc)
+dir_fit = c1*np.log(yp)**2+c_off
+
+a_off=0 #2.84/(dp_loc*us_loc)
+a_sqr=24.66/(dp_loc*us_loc)
+a_lin=10.24/(dp_loc*us_loc) 
+dir_fit2= a_off + a_sqr*np.sqrt(yp) + a_lin*yp
+
+
+for i in range(11):
+    ix=mp.geti(yp,i)
+    print(yp[ix],dir_fit[ix],dir_fit2[ix]) 

### 2021-09-09 write/manuscript.tex
finished section 3

@@
 %
-For the symmetries in the flow (horizontal homoegeneity, and $W=0$), it is
+For the symmetries in the flow (horizontal homogeneity, and $W=0$), it is
 %
@@
 %
-The inner normalization of $\nu_E$ s
-
+The inner normalization of $\nu_E$ is obtained by division through the molecular viscosity, and it is $\nu_E^+=\nu_E/\nu = S_\mathrm{turb}/S_\mathrm{visc}$.
+Under this normalization, the profiles of eddy viscosity collapse below $z^+\approx 20$ with a tendency towards better collapse
+at higher $z^+$ (up to $z^+\approx 50$ for $\RE=1600$; Fig.~\ref{fig:stresses}b).
+%
+In the outer layer, the eddy viscosity is characterized by a distinct minimum at $z^-\approx 0.6-0.8$,
+and we find that the following mixed normalization of $\nu_E$ by the geostrophic wind and friction velocity
+collapses the value of $\nu_E$ at this minimum (cf. Fig.~\ref{fig:stresses}d): 
+%
 \begin{eqnarray}
- \nu_E^+  \frac{Z_{\star}^{2}}{\delta+} &=& \frac{\nu_E^+}{\nu}
+  \nu_E^-=\nu_E^+  \frac{Z_{\star}}{\delta^+} = \nu_E \frac{1}{\nu} \frac{G}{u_\star}\frac{\nu}{u_\star \delta}  = \nu _E \frac{Gf}{u_\star^3}. 
 \end{eqnarray}
-\par
-%
-\todo{PARAGRAPH ON TURBULENCE STRUCTURE (2D SLICES, SPECTRA?) cf. Fig.~\ref{fig:slices}}
-
+%
+Substantial variation of the profiles is, however observed below and above this minimum for different $\RE$ which illustrates that
+this normalization is probably not generally appropriate across the outer layer.
+% 
+\par
+%
+The organization of the flow is depicted in terms of the turbulence kinetic energy in Fig.~\ref{fig:slices}.
+%
+In vicinity of the wall, at $y^+\approx 10$, (Fig.~\ref{fig:slices}a), elongated streaks aligned with the
+surface shear stress dominate.
+%
+Moving away from the wall, to $y^+\approx 150$ (well within the logarithmic region), the structures
+are larger and more isotropic, but they are still largely aligned with the surface shear stress.
+%
+Finally in the upper part of the outer layer, around $y^+\approx 1000$, no clear visual signature of the
+surface veering direction is found, and intense TKE structures (bright yellow)
+are organized on a large spatial scale with with weaker eddies (greenish structures) and quiescent regions
+in between.  
 \begin{figure}

### 2021-09-17 write/manuscript.tex
minor changes in Sections 3 and 4

@@
   %
-  The profile of the stream-wise component follows the classical viscous, logarithmic and wake scaling.
+  The profile of the streamwise component follows the classical viscous, logarithmic and wake scaling.
   %
@@
 This paper exploits the robust features of mean velocity profiles from direct numerical
-simulation across a range of Reynolds numbers to formulate both the stream-wise and span-wise components
+simulation across a range of Reynolds numbers to formulate both the streamwise and span-wise components
 of the mean velocity vector as a function of the Reynolds number.
@@
 $Oxyz$, where an approximate alignment (plus/ minus few degrees) of the direction $Ox$ with the surface shear stress is achieved.
-The coordinate $Oy$ points away from the wall,  and $Oz$ points in the span-wise direction normal to $Oxy$.
+The coordinate $Oz$ points away from the wall, and $Oy$ points in the span-wise direction normal to $Oxz$.
 For analysis of the results, we use two coordinate systems that are
@@
 that is governed by the physical parameters
-(i)   geostrophic wind magnitude $G=\sqrt{G_1^2+G_3^2}$ (where $\mathbf{G}=(G_1,G_2,0)$ is the geostrophic wind vector),
+(i)   geostrophic wind magnitude $G=\sqrt{G_1^2+G_2^2}$ (where $\mathbf{G}=(G_1,G_2,0)$ is the geostrophic wind vector),
 (ii)  Coriolis parameter $f$ (representing the angular rotation), and
@@
    %
-These yield a Reynolds number $\RE$ as non-dimensional parameter governing the system.
-%
-We use the geostrophic wind as velocity and the Coriolis parameter $f$ as time scale for the non-dimensional framework which implies the Rossby radius $\LR=G/f$ as length scale, such that the Reynolds number governing the problem reads as
+These yield a Reynolds number as non-dimensional parameter governing the system.
+%
+We use the geostrophic wind as velocity and the Coriolis parameter $f$ as time scale for the non-dimensional
+framework.
+%
+This implies the Rossby radius $\LR=G/f$ as length scale, such that one Reynolds number
+governing the problem reads as
+%
 \begin{equation}
@@
 %
-The scales used in defining $\RE_\Lambda$ loose there relevance in describing the turbulent state of the flow. The friction Reynolds number
-\begin{equation}
-  \RE_\tau = \frac{u_\star\delta}{\nu} = \frac{\RE_\Lambda}{Z_\star^2}, 
-\end{equation}
-where $\delta=u_\star/f$, gives a measure for the turbulence scale separation that is comparable to other
-wall-bounded flow geometries such as rotating or non-rotating channel, pipe, or Couette flow.
-%
-\par
-%
-The non-dimensionalized governing equations read as 
+The scales used in defining $\RE_\Lambda$ loose their relevance in describing the turbulent state of the flow.
+%
+The friction Reynolds number
+\begin{equation}
+  \RE_\tau = \frac{u_\star\delta}{\nu} = \delta^+ = \frac{\RE_\Lambda}{Z_\star^2}, 
+\end{equation}
+where $\delta=u_\star/f$ and we use a superscript $'+'$ to denote normalization by inner turbulence
+scales ($u_\star, \nu$).
+%
+$\RE_\tau$ is suitable to compare the turbulence scale separation also to other wall-bounded flow geometries
+such as rotating or non-rotating channel, pipe, or Couette flow.
+%
+\par
+%
+The governing equations  non-dimensionalized by $G$, $f$, and $\Lambda_\mathrm{\RO}$ read as 
 \begin{subequations} 
@@
 \section{Scaling behavior of the flow for $\RE_\tau$ up to 3000}
+\label{sec:scaling}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
-(Alternatively normalized TKE and dissipation,
+(TKE and dissipation normalized as
 $f G^{-3} \int e \mathrm{d} z$ and $u_\star^{-3} \int \epsilon\mathrm{d}z$, exhibit substantial
-dependence on $\RE$ for the variation of $u_s\tar$.)
+dependence on $\RE$ for the variation of $u_\star$.)
 %
@@
 %
-The profile of the shear-aligned streamwise velocity component are well-collapsed for the Reynolds
+The profiles of the shear-aligned streamwise velocity component are well-collapsed for the Reynolds
 numbers investigated below $z^-\approx 0.15$ (circles in Fig.~\ref{fig:profiles}a); the logarithmic
@@
 %
-While the profiles $U^{\alpha_\star+}(z^+)$ diverge between different $\Re$ beyond $z^-=0.15$,
+While the profiles $U^{\alpha_\star+}(z^+)$ diverge between different $\RE$ beyond $z^-=0.15$,
 the corresponding profiles $U^{\alpha_\star}(z^-)/u_\star$ agree closely,
@@
 %
-It is noteworthy that also in the outer layer of the flow, $u_\star$--and not the magnitude the
-geostrophic wind speed $G$--is the governing velocity scale for inviscid, i.e. $\RE$-independent,
+It is noteworthy that also in the outer layer of the flow, $u_\star$ (not the magnitude the
+geostrophic wind speed $G$) is the governing velocity scale for inviscid, i.e. $\RE$-independent,
 normalization. \todo{WHY? cf. TKE/Dissipation normalization above}
@@
 In contrast to the viscous stress, the total stress follows outer normalization, i.e.
-$S^+(z^-)$ is universal.
-%
-This implies a mixed scaling for the turbulent stress, 
+$S^+(z^-)$ is universal; a discrepancy in the inner layer does not occur as the total stress
+is approximately constant in the viscous and buffer layer, and a rescaling of the height would have
+no effect there; above, outer scaling is appropriate for the well-established dynamics 
+in the overlap region of inner and outer layer.
+%
+This, however, implies a mixed scaling for the turbulent stress, 
 \begin{equation}
@@
 %
-Finally in the upper part of the outer layer, around $y^+\approx 1000$, no clear visual signature of the
+In the upper part of the outer layer, around $y^+\approx 1000$, no clearly visible signature of the
 surface veering direction is found, and intense TKE structures (bright yellow)
@@
 \section{Towards a Universal Velocity Profile for the Turbulent Ekman Layer}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Formulations for the outer layer that take into account the rotation, and thus deviation from the channel-flow analogy
-need to be matched to the framework of surface similarity, and it
-occurs that a smooth transition from the inner to the outer layer is not easily achieved.
-%
-\cite{optis:BM2014}, for instance, define an ``effective geostrophic wind vector that has the same magnitude of the
-observed surface geostrophic wind and is rotated by the angle $\alpha$ [their nomenclature]'' to overcome the
-unsteady transition when coming approaching the Ekman layer from below.
-%
-While such rotation of the wind vector is \emph{a posteriori} justified by the observational data that the model outcomes are compared to,
-we believe that it is the manifestation of a mismatch between the theoretical treatment of the inner and outer layer of
-turbulent Ekman flow.
+\label{sec:profiles} 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+We now turn to the formulation of a general velocity profile that is fully determined by the only
+parameter of the idealized problem, namely a Reynolds number representing the scale separation or
+geometric size of the problem.
+%
+This precludes a drag law with wherewith we begin this section (\ref{sec:drag-law}).
+Based on the scaling arguments put forward in Sec.~\ref{sec:scaling}, we then develop
+a formulation of the wind vector in the Ekman layer (Sec~\ref{sec:ekman}) and separately
+for the streamwise (Sec.~\ref{sec:streamwise}) and spanwise components (Sec.~\ref{sec:spanwise})
+in the overlap and inner regions. 
+%
 
@@
 %
-\subsection{Profile in the outer layer}
+\subsection{Profile in the Ekman layer}
+\label{sec:ekman}
+%
+Formulations for the outer layer that take into account the rotation
+(and thus deviation from the channel-flow analogy)
+need to be matched to the framework of surface similarity.
+%
+A smooth transition from the inner layer to the Ekman layer, where the wind is characterized by a turning
+of its mean direction, is not easily achieved. 
+%
+\cite{optis:BM2014}, for instance, define an ``effective geostrophic wind vector that has the same magnitude of the
+observed surface geostrophic wind and is rotated by the angle $\alpha$ [their nomenclature]'' to overcome the
+unsteady transition when coming approaching the Ekman layer from below.
+%
+While such rotation of the wind vector is \emph{a posteriori} justified by the observational data that the model outcomes are compared to,
+we understand this as the manifestation of a mismatch between the theoretical treatment of the inner and outer layer of
+turbulent Ekman flow.
 %
@@
 that is sustained by momentum convergence at the inflection point of the velocity profile.
+%
 The super-geostrophic maximum of streamwise velocity and a secondary minimum aloft the bulk-turbulent part of
@@
 
-\subsection{Stream-wise Velocity Component} 
-For the stream-wise velocity profile (that in non-rotating flows due to the geometry is always aligned with the surface shear stress),
+\subsection{Streamwise Velocity Component} 
+For the streamwise velocity profile (that in non-rotating flows due to the geometry is always aligned with the surface shear stress),
 well-established theories exist for various regimes according to their distance from the wall and the relative role of
@@
 \begin{equation}
-  (U,W)^{\alpha_\star+}  = (z^+,0)  
+  (U,V)^{\alpha_\star+}  = (z^+,0)  
 \end{equation}
@@
 %
-\subsection{Span-wise velocity}
+\subsection{Spanwise velocity component}
 %
@@
 %
-While the universal profiles in vicinity of the wall implies that the profile be zero or at least small with respect to the stream-wise
+While the universal profiles in vicinity of the wall implies that the profile be zero or at least small with respect to the streamwise
 component, the veering requires a value of $V_{top}=U_G\sin\alpha_\star$ in the free stream (and thus also at the top of the boundary layer if we assume
@@
 %
-Outer scaling of the velocity profile further implies that the velocity deficity of the spanwise component $V^\alpha_star-G^\alpha_\star$
+Outer scaling of the velocity profile further implies that the velocity deficity of the spanwise component $V^{\alpha_\star}-G^{\alpha_\star}$
 be a universal function of the outer height $z^-$, i.e.  
 \begin{eqnarray}
-  V^\alpha_\star-G^\alpha_\star = f^\alpha_\star_V(z^-).  
+  V^{\alpha_\star}-G^{\alpha_\star} = f^{\alpha_\star}_V(z^-).  
 \end{eqnarray}
@@
 %
-While the stream-wise velocity follows a universal inner scaling that has acquired its universal, $Re$-independent shape for $Re_D> \mathcal{O}\left(10^3\right)$, the span-wise component that defines how the velocity vector veers when the surface is approached,
+While the streamwise velocity follows a universal inner scaling that has acquired its universal, $Re$-independent shape for $Re_D> \mathcal{O}\left(10^3\right)$, the span-wise component that defines how the velocity vector veers when the surface is approached,
 does not collapse in inner units, and there is, most importantly no sign of convergence even at the highest Reynolds numbers for

### 2021-09-17 plot/EkmanProfileClass.py

@@ -242,7 +242,6 @@ class EkmanUniversalClass:
         # SPANWISE COMPONENT (in-plane orthogonal to shear)
         #
         # inner region - empirical profile
-        #   - based on direction below y+~10
         w_norm = us_loc*dp_loc
         i1=mp.geti(yp,9)
         avisc=-18.852472992784048/w_norm

### 2021-09-29 write1/manuscript.tex
comments on current draft of part 2; worked on profile description in Ekman layer

--- 2021-09-17 write/manuscript.tex
@@
   \alpha_\star = \sphericalangle \left(\underline{G}, \underline{\tau}_\mathrm{sfc} \right).
+  \label{eqn:alpha_def} 
 \end{equation}
@@
 The generalization of profiles to arbitrary Reynolds numbers 
-requires sufficient scale separation in the simulations to not only quantify the effect of
-the Reynolds number on low-order flow statistics, but also assess the corresponding rate-of-change
+requires sufficient scale separation in the simulations, not only to quantify the effect of
+the Reynolds number on low-order flow statistics, but also to assess the corresponding rate-of-change
 to eventually allow for an extrapolation of the findings.
 %
-While the simulations available give confidence in a first-order representation of the turbulent
+While the simulations previously available (gray shading in Tab.~\ref{tab:sim-setup})
+give confidence in a first-order representation of the turbulent
 problem, the estimation of higher-order effects such as the dependency of the Reynolds number requires
@@
 %
-Velocity profiles in inner and units (Fig.~\ref{fig:profiles}a,b) are in accordance
+Velocity profiles in inner and outer units (Fig.~\ref{fig:profiles}a,b) are in accordance
 with previous work \citep{coleman:JFM1992,spalart:PF2008,spalart:PF2009,ansorge:BM2014,ansorge:BM2019}:
@@
 While the profiles $U^{\alpha_\star+}(z^+)$ diverge between different $\RE$ beyond $z^-=0.15$,
-the corresponding profiles $U^{\alpha_\star}(z^-)/u_\star$ agree closely,
-irrespective of $\RE$ which illustrates the
-inner--outer scale-duality in this external flow. 
-%
-It is noteworthy that also in the outer layer of the flow, $u_\star$ (not the magnitude the
+the corresponding profiles of the velocity deficit  $(U^{\alpha_\star}(z^-) - G_{1}^{\alpha} ) /u_\star$
+agree closely, irrespective of $\RE$.
+%
+This illustrates the inner--outer scale-duality in this external flow with inner scaling being
+appropriate in the inner layer and outer scaling in the outer layer. 
+%
+Also in the outer layer of the flow, $u_\star$ (not the magnitude the
 geostrophic wind speed $G$) is the governing velocity scale for inviscid, i.e. $\RE$-independent,
@@
 %
+\par
+%
 No collapse in inner units is found for the profiles of spanwise velocity $V^{\alpha_\star+}(z^+)$;
 %
-When normalized in outer units, the profiles of spanwise velocity $V^{\alpha_\star}(z^-)/u_\star$
+When normalized in outer units, the deficit profiles of spanwise velocity
+$(V^{\alpha_\star}(z^-)-G_{2}^{\alpha})/u_\star$
 agree well beyond $z^-\approx 0.3$ whereas the profiles of the streamwise component collapse also
@@
 %
+The value of $V^{\alpha_\star+}(z^-)-G_{2}^{\alpha}$ is sensitive to the wind veering for
+$z\rightarrow 0$ as---for use of the shear-aligned component---it has to approach the value
+$-G_2^\alpha=|G|\sin\alpha\ne0$ in view of the no-slip boundary condition.  
+%
 While low-$\RE$ effects appear to be present for $\mathcal{O}(\RE)<10^3$, the spanwise component
-appears to converge to an $\RE$-independent limit within the range of scale separation considered
-here.
+converges to an $\RE$-independent limit within the range of scale separation considered
+here, i.e.
+\begin{equation}
+  G_{2}^{\alpha}/u_\star = Z_\star \sin\alpha \rightarrow \text{const. for\ }\RE\rightarrow \infty, 
+\end{equation}
+which has indeed already been used by \cite{spalart:JFM1989}. 
 %
@@
 The normalized viscous stress
+%
 \begin{subequations}
@@
 \end{equation}
+\end{subequations} 
 where $z^-=z^+/\RE_\tau$. 
@@
 %
+\begin{subequations} 
 \begin{equation}
@@
 \begin{eqnarray}
-  \nu_E^-=\nu_E^+  \frac{Z_{\star}}{\delta^+} = \nu_E \frac{1}{\nu} \frac{G}{u_\star}\frac{\nu}{u_\star \delta}  = \nu _E \frac{Gf}{u_\star^3}. 
+  \nu_E^-=\nu_E^+  \frac{1}{\delta^+} = \nu_E \frac{1}{\nu} \frac{\nu}{u_\star \delta}  = \nu _E \frac{f}{u_\star^2}.
+  \label{eqn:evisc_outer_scaling}
 \end{eqnarray}
-%
-Substantial variation of the profiles is, however observed below and above this minimum for different $\RE$ which illustrates that
-this normalization is probably not generally appropriate across the outer layer.
+\end{subequations} 
+%
+Substantial variation of the profiles is, however observed below and above this minimum for
+different $\RE$ which illustrates that this normalization is probably not generally appropriate
+across the outer layer.
 % 
@@
 %
-In the upper part of the outer layer, around $y^+\approx 1000$, no clearly visible signature of the
+In the upper part of the outer layer, around $y^+\approx 1000$, no clear signature of the
 surface veering direction is found, and intense TKE structures (bright yellow)
@@
 %
-This precludes a drag law with wherewith we begin this section (\ref{sec:drag-law}).
+This precludes a drag law wherewith we begin this section (\ref{sec:drag-law}).
 Based on the scaling arguments put forward in Sec.~\ref{sec:scaling}, we then develop
 a formulation of the wind vector in the Ekman layer (Sec~\ref{sec:ekman}) and separately
-for the streamwise (Sec.~\ref{sec:streamwise}) and spanwise components (Sec.~\ref{sec:spanwise})
+for the streamwise (Sec.~\ref{sec:streamwise}) and spanwise (Sec.~\ref{sec:spanwise}) components
 in the overlap and inner regions. 
 %
-
-\end{itemize}
 \begin{figure}
@@
 \label{sec:drag-law}
-A drag-law for Ekman flow determines two key parameters of the flow as a function of Reynolds number alone.
-%
-These are, first, the modulus of the surface friction normalized by the geostrophic wind forcing, $u_\star$ and,
-second, the direction of surface shear stress, $\alpha_\star$, also termed wind veering. 
+A drag-law for Ekman flow determines---as a function of Reynolds number alone---the surface drag.
+%
+This can be formulated in terms of the modulus
+of the surface friction normalized by the geostrophic wind forcing, $u_\star$ (Eq.~(\ref{eqn:ustar_def}))
+and, the direction of surface shear stress, $\alpha_\star$ (Eq.~(\ref{eqn:alpha_def})),
+also termed wind veering. 
 %
 A non-zero veering of the wind is a rather special case in comparison with most turbulent flows conisdered in an
-engineering context, and it confronts us with a situation where the most appropriate coordinate system for analysis
-(namely that aligned with the surface shear stress) is a priori unknown. 
+engineering context, and it confronts us with a situation where the most appropriate coordinate system
+for analysis (namely that aligned with the surface shear stress) is a priori unknown. 
 %
@@
 observed surface geostrophic wind and is rotated by the angle $\alpha$ [their nomenclature]'' to overcome the
-unsteady transition when coming approaching the Ekman layer from below.
-%
-While such rotation of the wind vector is \emph{a posteriori} justified by the observational data that the model outcomes are compared to,
-we understand this as the manifestation of a mismatch between the theoretical treatment of the inner and outer layer of
+unsteady transition when approaching the Ekman layer from below.
+%
+While such rotation of the wind vector is \emph{a posteriori} justified by the observational data that the model outcomes are compared to, understand this as the manifestation of a mismatch between the theoretical treatment of the inner and outer layer of
 turbulent Ekman flow.
+%
+\par
+%
+The classic text-book solution of Ekman flow makes use of the physical boundary conditions (BCs) for the atmospheric
+boundary layer (no-slip at the bottom and geostrophic wind in the free atmosphere).
+%
+These two boundary conditions eliminate one mode of the analytical solution and determine the magnitude of
+the spiral.
+%
+In doing so, one has to assume that the solution is appropriate across the entire ABL, which is not the case:
+the dynamics put forth by \cite{ekman:AMA1905} are not appropriate in the surface layer of the ABL, where
+the logarithmic, buffer, and viscous sublayers are found. 
+%
+We to use a virtual boundary condition instead that takes into account the increased gradients in vicinity of
+the surface and matches the magnitude of the Ekman spiral to the values observed in our DNS:
+\begin{subequations} 
+\begin{eqnarray}
+  G^{-1}\left(\begin{array}{c} 
+    U_\mathrm{ek}\\ 
+    V_\mathrm{ek}
+  \end{array}\right)  &=& \left(\begin{array}{c} 1 \\ 0 \end{array}\right) 
+  + a_\mathrm{ek} e^{-z_\mathrm{ek}} \left(\begin{array}{r}
+    -\cos z_\mathrm{ek} \\ 
+    \sin z_\mathrm{ek}
+  \end{array}\right)  \\
+  z_\mathrm{ek} &=& \delta_{ek} (y^- -s_{ek}). 
+\end{eqnarray}
+\end{subequations}
+%
+This solution is characterized by three parameters,
+(i) an Ekman-layer depth scale $\delta_\mathrm{ek}$,
+(ii) the magnitude parameter of the spiral $a_{ek}$, and
+(iii) a zero-crossing point for the velocity $s_{ek}$.
+\begin{enumerate} 
+%
+\item[(i)] The Ekman-layer depth scale $\delta_\mathrm{ek}$ is fundamentally defined by the eddy diffusivity.
+  %
+  However, we have seen in Section \ref{sec:scaling} that a characteristic value for the eddy diffusivity
+  is not easily obtained for its strong dependence on the Reynolds number and distance from the surface.
+  %
+  We therefore resort to the physical manifestation of the eddy diffuysivity in an Ekman layer, and use the
+  boundary layer depth $\delta_\mathrm{ek} = 0.66 \delta \times 2\pi$.
+  %
+  For the relation $\delta_\mathrm{ek}=\sqrt{2\nu_\mathrm{ek}/f}$, this yields
+  $\nu_\mathrm{ek} \propto u_\star^2/f$ in accordance with the observations in
+  Sec.~\ref{sec:scaling} (Eq.~\ref{eqn:evisc_outer_scaling}). 
+  %
+\item[(ii)] The magnitude parameter of the spiral primarily defines the super-geostrophic maximum of the
+  wind profile aloft the logarithmic layer. Our simulations suggest this maximum of the velocity deficit
+  remains constant when normalized by $u_\star$ as shown in Fig.~\ref{fig:profiles_outer}.
+  %
+  The numerical value of $a_\mathrm{ek}$ is estimated from visual comparison, and we find $a_{ek}=8.4u_\star$;
+  while this appears rather large, it is pre-multiplied by $e^{-z_\mathrm{ek}}$ which has already decreased
+  to $\mathcal{O}(0.1)$ at the height of this maximum.
+  %
+  This choice ascertains that the velocity deficits $U/u_\star-Z_\star$ and $V/u_\star-Z_\star$ do not
+  depend on the velocity scale $u_\star$, but only on $G$ as
+  \begin{equation}
+    U_\mathrm{ek}/u_\star-Z_\star \propto a_\mathrm{ek} Z_\star = 8.4G.  \notag
+  \end{equation} 
+\item[(iii)] While $\delta_\mathrm{ek}$ sets the thickness across which the wind veering takes place,
+  the zero-crossing height $s_\mathrm{ek}$ defines an offset.
+  %
+  Physically, this offset can be understood as the height at which the surface was located assuming
+  a perfect Ekman flow down to the surface.
+  %
+  As this is not the case, and gradients are steeper in the highly turbulent boundary layer flow encountered
+  when approaching the surface, the offset is smaller than zero (the fully turbulent boundary layer is
+  actually thinner than an Ekman layer would be).
+  %
+  From our DNS data, we estimate $s_\mathrm{ek}=-0.12\delta$. 
+\end{enumerate} 
+%
+\todo{Paragraph and figure on idealized Ekman profile}
+%
+\par 
 % 
@@
   \includegraphics[width=0.5\textwidth]{../plot/outer_layer_u.pdf}\includegraphics[width=0.5\textwidth]{../plot/outer_layer_w.pdf}
-  \caption{} 
+  \caption{
+    \label{fig:profiles_outer} 
+  }
 \end{figure}
@@
 \subsection{Streamwise Velocity Component}
+\label{sec:streamwise} 
 For the streamwise velocity profile (that in non-rotating flows due to the geometry is always aligned with the surface shear stress),
@@
 \subsection{Spanwise velocity component}
+\label{sec:spanwise} 
 %

### 2021-09-29 write2/manuscript_part2_20210922_CA.tex
comments on current draft of part 2; worked on profile description in Ekman layer

\documentclass[a4paper,11pt]{article}

\usepackage[margin=2.5cm]{geometry}
\usepackage{booktabs} 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{natbib}
\usepackage{xcolor}

\newcommand{\p}{\partial}
\newcommand{\gzm}{{\xi}}%{\gamma z_\text{match}}}
\newcommand{\todo}[1]{\textcolor{red}{$[$#1$]$}}

\pdfsuppresswarningpagegroup=1

\author{hauke-zrick}

\date{Draft \today}
\title{Towards a Universal Veering Profile for Turbulent Ekman Flow at arbitrary Reynolds number - Part 2\\
  \normalsize LES and DNS of Turbulent Ekman Flow \\}

\begin{document}

\maketitle

\begin{abstract}
  An analytical formulation for the profiles of stream- and span-wise velocity in turbulent Ekman flow and large-eddy simulations are used to analyze the properties of the turbulent Ekman flow. 
%
	We show a comparison of a turbulent Ekman flow of $Re_D=1\,600$ simulated by DNS and LES.
\end{abstract}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\section{Introduction}

We here consider the turbulent Ekman flow as a homogeneous, stationary boundary layer flow over a rotating flat surface under neutral stratification. This flow is defined by only very few parameters. Nevertheless, it can serve as a simple model of the atmospheric boundary layer since it exhibits many characteristics that also shape the real atmosphere, namely the logarithmic layer and the Ekman spiral. Hence, the investigation of this case is suited to learn about the fundamental properties of the atmospheric boundary layer.

We investigate three different Reynolds numbers $Re_D$: $1\,600$, 150\,000, and 1\,000\,000, where $Re_D = GD/\nu$ and $D=\sqrt{2\nu/f}$, $Re_D = G/\sqrt{\frac{1}{2}\nu f}$. All cases are simulated by LES and the low Reynolds number case we also compare to DNS for comparison. The too high Reynolds number cases cannot be simulated by DNS due to the limitations of computational resources.

Short summary of part I? In part I of this publication, we presented a generic profile of wind direction and speed in turbulent Ekman flow.

The information how the wind direction changes with height is of great importance for wind-power forecasting \cite{optis2014moving}.

\todo{Wie ist die Grenzschicht? Was sagen die Profile und das LES darüber?}

Foundations: \cite{csanady1967resistance}, \cite{tennekes1973logarithmic}, \cite{spalart1989theoretical}

Neutral Ekman layer with LES: \cite{esau2004simulation}\cite{jiang2018large}

The same turbulent Ekman flow is simulated by DNS and LES and the results are compared.

Profile of wind direction from DNS - what can be learned from it for LES?

Limits of coarse resolution

The velocity profiles of both horizontal components of the turbulent Ekman flow provide a benchmark test for LES model. $\rightarrow$ First step towards a generic wind profile including stratification (and inversion?)?

This paper investigates the mean velocity profiles of the turbulent Ekman flow from moderate to high Reynolds numbers using the formulation from part I and LES.

\section{A Universal Velocity Profile for the Turbulent Ekman Layer}

The universal velocity profile for the turbulent Ekman layer consists of several parts for different layers of the boundary layer. Near the surface, viscous forces dominate the flow. In the middle part, the velocity profile has a logarithmic shape which passes over into an Ekman spiral. 
In the following, the distinct layers for both horizontal components of the velocity are described.

\subsection{Total profile}

\todo{In part II a summary of the theor. profile in minimalistic form? Without any deductions, just a description.}

From the layers of the velocity profile, a total profile for the whole boundary layer is formed. The profile of the stream-wise component of the velocity is seperated into three layers, which are the viscous layer $U_{visc}$, the logarithmic layer $U_{log}$, and the Ekman layer $U_{EK}$. The span-wise component of the velocity is seperated into two layers, namely the inner layer $V_{inner}$, and the Ekman layer $V_{EK}$.
\begin{align}
  U =& (1-w_{visc})U_{visc} + (w_{visc}-w_{outer})U_{log} + w_{outer}U_{EK}, \\
	V =& (1-w_{outer})V_{inner} + w_{outer}V_{EK}.
\end{align}
The transition between consecutive layers is formed by a transfer function:
\begin{align}\label{error}
  w_{*} = \frac{1}{2}\left(\textrm{erf}\left[\sigma_T\log\left(\frac{z}{z_{T}}\right)\right]+1\right),
\end{align}
where $\sigma_T$ is a transition scale that defines the width of the transition and $z_{T}$ is the height of the transition, where the upper and the lower layer equally contribute to the velocity ($w_{*}(z_{T})=0.5$).

\subsection{Drag-Law}

The geostrophic drag $Z \equiv u_\star/G$ and the direction between the shear stress and the geostrophic wind $\alpha$ are two key parameters of the Ekman flow. They can be estimated using a semi-empirical drag-law based on \cite{spalart1989theoretical}, which describes them as functions of only the Reynolds number:
\begin{subequations}\label{drag}
	\begin{align}
		\frac{G}{u_\star}\cos\phi^\star &= \frac{1}{\kappa}\log Re_\tau + C - A_r, \\
		\sin\phi^\star &= A_i\frac{u_\star}{G},\\
		\alpha &= \phi^\star - \frac{C_5}{Re_\tau},
	\end{align}
\end{subequations}
with $Re_\tau = \frac{Re_D^2}{2}\frac{u_\star^2}{G^2}$, $\kappa = 0.415$, $A_r = 4.80$, $A_i = -5.57$, $C = 5.4605$, $C_5 = -57.8$. This law is in excellent agreement with DNS in the range $400\leq Re_D\leq 1\,600$ as demonstrated by \cite{ansorge2014global}.

\subsection{Stream-wise Velocity Component}

In the viscous sublayer, the span-wise velocity is close to zero and the stream-wise velocity is described by the law of the wall:

\begin{align}
  U^{\alpha+} = z^+,
\end{align}
(the index $\alpha$ indicates the alignment with the direction of the shear stress). Around $z^+=5$, the velocity is beginning to deviate from its linear profile and the buffer layer forms the transition between viscous layer and logarithmic layer. From the surface up to the buffer layer, the stream-wise velocity is described by
\begin{align}
  U_{inner}^{\alpha+} = \frac{z^+ + \gamma_4(z^+)^4 + \gamma_6(z^+)^6}{1+\gamma_6/u_{ref}(z^+)^6},
\end{align}
where $\gamma_4=-3.825\cdot10^{-4}$, $\gamma_6=6.32\cdot10^{-6}$, and $u_{ref}=0.07825$.

The logarithmic region of the stream-wise velocity is 
\begin{align}
  U_{\log}^{\alpha+} = \frac{1}{\kappa}\log z^+ + C,
\end{align}
where $\kappa=0.416$, and $C=5.4605$. The transition between buffer layer and logarithmic layer is located at $z_T^+=19$ with a scale $\sigma_T=2$.

\subsection{Profile in Outer Layer}

Above the boundary layer, the horizontal pressure gradient is balanced by the coriolis force and the wind speed equals the geostrophic wind $G$. For the stationary case, the horizontal equations of motion can be written
\begin{subequations}
  \begin{align}
    0&=fV+\nu_e\partial_z^2U,\\
		0&=-f(U-G)+\nu_e\partial_z^2V,
  \end{align}
\end{subequations}
where $\nu_e$ is an eddy viscosity and the x-axis is aligned with the geostrophic wind. This is solved by
\begin{subequations}
  \begin{align}
    U_{EK} &= G + Ae^{-\tilde{z}}\cos\tilde{z},\\
		V_{EK} &= - Ae^{-\tilde{z}}\sin\tilde{z},
  \end{align}
\end{subequations}
where $A = 8.4u_\star - 150/Re_\tau$, $\tilde{z} = (z-z_r)/D_E$, $z_r = 0.12\delta$, and $D_E = 3\delta/4\pi\approx 0.24\delta$. The parameters are deduced from DNS.

The transition from the logarithmic layer to the Ekman layer is located at $z^-=0.3$ (talk: $z^-_T=0.3-120/Re_D$) with a transition scale of $\sigma_T=2$ for the stream-wise velocity.

\subsection{Span-wise Velocity}

For $z^+\leq10$, the knowledge on $U^\alpha$ is exploited to parametrize $V^\alpha$ in terms of direction:
\begin{align}
  \alpha_{visc} &= \frac{C_1+C_2(\log  z^+)^2}{Re_\tau Z},\\
	\Rightarrow V_{visc}^\alpha &= U^\alpha\tan(\alpha_{visc}),
\end{align}
where $C_1=40$, $C_2=26$, and $Z=u_\star/G$ \todo{(not valid for $z^+<1$!?)}. Above $z^+=10$, $V^\alpha$ is continued with
\begin{align}
  V_{inner}^\alpha = V_{10}^\alpha + A\log\left(\frac{z^+}{10}\right) + B(z^+-10),
\end{align}
which matches the value of $V_{10}^\alpha = V^\alpha_{visc}(z^+=10)$. To also match the gradient $m_{10}=\partial V_{visc}^\alpha/\partial z^+$ at $z^+=10$ and the value of $V_{EK}^\alpha$ at $z^-=0.27$, the values of $A$ and $B$ need to be
\begin{align}
  A &= \frac{V_{EK}^\alpha(z^-=0.27)-V_{10}^\alpha-m_{10}\left(0.27z^+/Re_{\tau}-10\right)}{\log(0.27z^+/(10Re_{\tau}))-(0.27z^+/Re_{\tau}-10)/10},\\
	B &= m_{10}-A/10.
\end{align}
At $z_T^-=0.27$, $V_{inner}^\alpha$ is blended into $V_{EK}$.

\section{Setup}

\subsection{Settings}

An incompressible, turbulent Ekman flow over a flat rotating plate is simulated for three different Reynolds numbers $Re_D = 10^3;1.5\cdot 10^5;10^6$, hereafter $Re1$, $Re2$, and $Re3$, respectively. The key input parameters of the three cases are presented in table \ref{simulation_parameters}.

\begin{table*}
	\centering
	\caption{Key parameters of the simulated cases. $f$, $\nu$, and $G$ are input parameters, while $u_\star$ and $\alpha$ are the expected results according to \cite{spalart1989theoretical}}
	\begin{tabular}{lccccccc}
          \toprule  
	  Name & $Re_D$ & $Re_\tau$ & $f$ [s$^-1$] & $\nu$ [m$^2$s$^-1$] & $G$ [ms$^-1$] & $u_\star$ [ms$^-1$] & $\alpha$ [$^\circ$] \\
          \midrule
          \midrule
	  Re1 & $10^3$ & $3\cdot10^3$ & $10^{-4}$ & $1.5\cdot10^{-5}$ & 0.0274 & 0.00144 & 18.96 \\
	  Re2 & $1.5\cdot10^5$ & $7.3\cdot10^6$ & $10^{-4}$ & $1.5\cdot10^{-5}$ & 4.108 & 0.1048 & 8.51 \\
	  Re3 & $10^6$ & $2.2\cdot10^8$ & $10^{-4}$ & $1.5\cdot10^{-5}$ & 27.39 & 0.5785 & 7.03 \\
          \bottomrule
	\end{tabular}
	\label{simulation_parameters}
\end{table*}

The domain is rotating around the z-axis with an angular velocity such that the Coriolis parameter is $f=10^{-4}$. The stratification of the flow is entirely neutral, i.e., the potential temperature $\Theta=const.$ for the whole domain. At the upper boundary, a no-penetration boundary condition is used and the horizontal components of the wind are forced to be equal to the geostrophic wind. At the bottom, a constant flux layer is assumed and the Monin-Obukhov Theory (MOST) is used to calculate the surface momentum fluxes $\overline{u''w''}_0$ and $\overline{v''w''}_0$.

To study the effect of resolution on the simulations, three different grid resolutions are chosen for each Reynolds number case. The grid cell sizes are around $\delta/50$, $\delta/100$, and $\delta/150$ for a coarse, a medium, and a fine resolution, respectively (see table \ref{simulation_parameters2}).

\begin{table*}
	\centering
	\caption{Simulations and grid parameters}
	\begin{tabular}{lccccccc}
          \toprule 
	  Name & $\Delta$\,[m] & $n_x$ & $n_y$ & $n_z$ & $L_x$\,[m] & $L_y$\,[m] & $L_z$\,[m]  \\ 
          \midrule
          \midrule 
	  Re1\_150 & 0.14 & 1536 & 1536 & 288 & 215.04 & 215.04 & 123 \\
		  Re1\_150\_dyn & 0.14 & 1536 & 1536 & 288 & 215.04 & 215.04 & 123 \\
		  Re1\_100 & 0.2 & 1080 & 1080 & 192 & 216 & 216 & 67.1 \\
		  Re1\_50 & 0.4 & 576 & 576 & 120 & 230.4 & 230.4 & 77.8 \\
                  \midrule 
		  Re2\_150 & 7 & 1536 & 1536 & 288 & 10752 & 10752 & 4245 \\
		  Re2\_150\_dyn & 7 & 1536 & 768 & 288 & 10752 & 5376 & 4245 \\
		  Re2\_100 & 10 & 1080 & 1080 & 192 & 10800 & 10800 & 3104 \\
		  Re2\_50 & 20 & 576 & 576 & 120 & 11520 & 11520 & 4113 \\
		  Re2\_50\_dyn & 20 & 576 & 576 & 120 & 11520 & 11520 & 4113 \\
                  \midrule 
		  Re2\_l\_sq & 7 & 1536 & 1536 & 288 & 10752 & 10752 & 4245 \\
		  Re2\_l & 7 & 1536 & 768 & 288 & 10752 & 5376 & 4245 \\
		  Re2\_m\_sq & 7 & 768 & 768 & 288 & 5376 & 5376 & 4245 \\
		  Re2\_m & 7 & 768 & 384 & 288 & 5376 & 2688 & 4245 \\
		  Re2\_s & 7 & 384 & 192 & 288 & 2688 & 1344 & 4245 \\
                  \midrule 
		  Re3\_150 & 40 & 1536 & 1536 & 240 & 61440 & 61440 & 2$1\,600$ \\
		  Re3\_150 & 40 & 1536 & 1536 & 240 & 61440 & 61440 & 17393 \\
		  Re3\_100 & 55 & 1080 & 1080 & 192 & 59400 & 59400 & 18687 \\
		  Re3\_50 & 110 & 576 & 576 & 120 & 63360 & 63360 & 6212 \\
		\end{tabular}
	\label{simulation_parameters2}
\end{table*}

The grid spacing inside of the boundary layer is isotropic up to the height $\delta = u_\star/f$, from where on the grid spacings in z-direction are stretched with the factor 1.02 each level, up to a maximum grid spacing of $6\Delta x$. The number of vertical grid points is chosen such that the domain height is at least three times $\delta$. Above two thirds of the total domain, Rayleigh damping is active to avoid reflections from the upper boundary.

\todo{The domain-size part rather in the results? Again a mix of setup and presentation of simulation results...} In part I of this publication, the DNS of low Reynolds number simulations ($Re_D\leq1\,600$) have a horizontal domain length of $L_x = 0.54\Lambda_{Ro}$ or $L_x = 1.08\Lambda_{Ro}$, where $\Lambda_{Ro}=G/f$ is the Rossby radius. (\todo{reason? include wavelengths that influence the profiles? citations?}). For these Reynolds numbers, $u_\star/G$ is around 0.05 so $L_x = 0.54G/f \approx 10 u_\star/f = 10 \delta$ (\todo{introduce $\delta_{95}$ as boundary layer height?}). However, $u_\star/G$ decreases with increasing Reynolds number and is around 0.02 for $Re_D=10^6$. A domain size of half the Rossby radius would then extend to $L_x\approx 25\delta$. Such a large domain would imply immense computational costs. To avoid this, the horizontal domain length was chosen to be $L_x\approx10\delta$ for all Reynolds numbers, which is $L_x/(G/f)\approx 0.48, 0.25,0.22$ for $Re1$, $Re2$, and $Re3$, respectively. An overview over the parameters of the settings is presented in table \ref{simulation_parameters2}.
\todo{Cite other domain sizes: \cite{jiang2018large} test domain size of 512x512 grid points or 2048x2048m, which means $L_x \approx 0.04\Lambda_{Ro}$, normal domain only half of this size:  $L_x \approx 0.02\Lambda_{Ro}$. \cite{spalart2008direct} $\Lambda_x = 2u_\star/f$, justified by \cite{csanady1967resistance} (``... it determines the size of the largest outer-layer
eddies.''). \cite{esau2004simulation} $L_x = 0.08\Lambda_{Ro}$.}

\begin{figure}
  \centerline{
	\includegraphics[width=0.4\textwidth]{figures/kappa_Re2_domain}
	\includegraphics[width=0.4\textwidth]{figures/hodograph_Re2_domain}
}
  \caption{Left: $\kappa=d\log(z^+)/dU^+$. Right: hodograph.}
  \label{domain}
\end{figure}

As LES uses a much coarser resolution than DNS, unresolved processes like the turbulent transport on the subgrid scale need to be modeled by a subgrid-scale model (SGS model). The model code of PALM offers two different SGS models: a 1.5-order closure after \cite{deardorff1980stratocumulus} and a dynamic closure after \cite{heinz2008realizability}. For most of the simulations, the 1.5-order closure is used, but several simulations are repeated using the dynamic closure for comparison.

The problem is solved using a three-step Runge-Kutta method. For scalar advection a 5th order Wicker-Skamarock scheme is employed. A comprehensive description of the LES model is given by \cite{maronga2020overview}.

\subsection{Correction of Mean Velocity}

The initial profile of the flow is calculated by a 1d-model with a Reynolds-average based turbulence parametrization. At the beginning of the 3d run, random perturbations are imposed to the horizontal velocity field to trigger the evolution of turbulent eddies. The following imbalance between pressure force and coriolis force results in an inertial oscillation of the period $T_{io}=2\pi/f$. The oscillation slowly decays over time and eventually vanishes. To avoid an overly long spin-up time, an ensemble average of the horizontal wind profiles $\langle U\rangle(z)$ and $\langle V\rangle(z)$ is taken over a period of $T_{avg} = T_{io}$ after an evolution of turbulence for $T_{spin-up} = T_{io}/2$. At the point in time $t = \frac{3}{2}T_{io}$, the momentary wind field is adjusted such that the instantaneous domain-averaged wind profiles $\overline{U}(z)$ and $\overline{V}(z)$ are shifted to the values of the ensemble average over a whole inertial period:
\begin{align}
  u^{new}_{ijk} = u_{ijk} - \overline{U}_k + \langle U\rangle_k,
\end{align}
where $i$ and $j$ indicate the horizontal grid points and $k$ indicates the vertical grid point. After this transformation of the flow field, the inertial oscillation is nearly gone and the simulation is continued for another period $T_{io}$ to collect data for the evaluation. The effect of this correction of the mean velocity on the inertial oscillation is shown in fig. \ref{E_ts}. This procedure seems to work better for low Reynolds numbers and for high Reynolds numbers simulated with coarse resolutions. This can be explained by the fact that the we only correct the mean velocity of the flow and while the form and orientation of the turbulent structures remains unchanged. Hence, after the correction, the mean velocity of both components corresponds the the mean of the the inertial period, whereas the current eddies correspond to a still swinging flow before the correction. So the structure and orientation of the turbulence is still off balance and might induce another inertial oscillation which is stronger for well resolved flows, since more turbulence is resolved. Nevertheless, through the procedure of correcting the mean profile after the recording over an inertial period we could remove the greater part of the inertial oscillation.

\begin{figure}
  \centerline{\includegraphics[width=0.9\textwidth]{figures/ts_ReD15e4_10m_0.177wu.000}}
  \caption{Left: domain-averaged kinetic energy of the simulation Re2\_100. Right: horizontal average of the friction velocity}
  \label{E_ts}
\end{figure}

\subsection{Choice of $z_0$}

In contrast to DNS, the viscosity of the fluid hardly influences the flow on the grid scale of the LES, so the terms in the Navier-Stokes equation containing $\nu$ are neglected and only the Euler equations are solved. Instead of a no-slip condition at the bottom, a constant flux layer is assumed below the first grid point and the Monin--Obukhov theory is used to compute the friction velocity and the stresses at the first grid point:
\begin{align}\label{most}
  u_\star = \frac{\kappa(U^2+V^2)^{0.5}}{\ln\left(z/z_0\right)},\\
  -\overline{u''w''}_0 = \frac{\kappa Uu_\star}{\ln\left(z/z_0\right)}.
\end{align}
Nevertheless, the viscosity is taken into account indirectly by choosing the roughness length $z_0$. This becomes clear by considering the law of the wall for a smooth surface:
\begin{align}
  u^+ = \frac{1}{\kappa}\ln(z^+) + C^+ = \frac{1}{\kappa}\ln\left(\frac{z^+}{z_0^+}\right)
\end{align}
with $\kappa\approx0.416$ and $C^+\approx 5.4605$ for a smooth wall (reference?). This leads to
\begin{align}
	z_0^+ = \exp\left(-\kappa C^+\right) \approx 0.1,
\end{align}
which is the roughness length for a smooth wall in inner units. So the roughness length in SI-units $z_0 = z_0^+\nu/u_\star\approx 0.1\nu/u_\star$ depends on the viscosity of the fluid, which means that the choice of $z_0$ is equivalent (\todo{?}) to a choice of the viscosity $\nu$.

\begin{figure}
  \centerline{\includegraphics[width=0.45\textwidth]{figures/3Re_ust_z0p}
	\includegraphics[width=0.45\textwidth]{figures/3Re_ust_z0p_detail}}
  \caption{The friction velocity as function of the roughness length with fit}
  \label{ust_z0p}
\end{figure}

The choice of $z_0$ and the turbulent flow (driven by the geostrophic wind $G$) determine the magnitude of $u_\star$ in a non trivial way. As will be demonstrated below, the choice of $z_0=0.1$ does not lead to the expected value of $u_\star$ predicted by Spalart. To reproduce the predicted $u_{\star pred}$, a sensitivity study was performed whose results are shown in fig. \ref{ust_z0p}. A variation of $z_0^+$ showed that the resulting $u_\star$ follow a power law $u_\star/u_{\star pred} = a(z^+_0)^m.$ In order to get the expected friction velocity, $z_0^+$ had to be 0.149, 0.174, and 0.196 for the cases Re1, Re2, and Re3, respectively. As these are the values for Reynolds numbers that differ by three orders of magnitude, this is only a slight dependence on the Reynolds number

\section{Results}

The theoretical profiles of both horizontal velocity components as well as corresponding LES solutions for all three Reynolds numbers are shown in fig. \ref{profiles_allred}. At first sight, the similar form of the profiles for all Reynolds numbers is apparent. In the viscous sublayer, the shear-aligned component increases linearly up to the buffer layer around $z^+\approx19$, where it (übergehen) slowly into the logarithmic layer. There the U-component increases logarithmically up to the supergeostrophic maximum and above decreases to its geostrophic value. The V-component plays nearly no role up to the middle of the logarithmic layer, where the transition to the Ekman layer takes place ($z^-\approx0.3-120/Re_D$). In inner scaling, the V-component of all Reynolds numbers is very similar, while the growth of the U-component leads to a smaller angle $\alpha$ for higher Reynolds numbers. In outer scaling, the profiles  of the shear-aligned velocity deficit collapse in the upper part of the boundary layer and the profiles of the V-component's velocity deficit collapse for the whole boundary layer.

The length of the profiles from the LES is identical for all Reynolds numbers. This is due to the fact that for all simulations, the ratio between boundary layer height and grid size is similar, which leads to similar lengths from the first grid point to the boundary layer top in a logarithmic plot.

\begin{figure}
  \centerline{
	\includegraphics[width=0.385\textwidth]{figures/uv_p_AllReD}
	\includegraphics[width=0.4\textwidth]{figures/uv_def_AllReD}
}
  \caption{Left: shear-aligned profiles in outer scaling. Right: shear-aligned velocity deficit in inner scaling}
  \label{profiles_allred}
\end{figure}

\todo{Show dependence of the profiles on the domain size.}

The (expression?) $\kappa = d\log(z^+)/dU^+$ is quite sensitive to the resolution as well as to the domain size.

\todo{Rating of the resolution as Jiang and Sullivan: value of $\kappa$, $z_{0,pr}$ estimated from profile vs. $z_0$, $\Delta W_{10}$}

Estimation of $\kappa$: maximum height: where the value of $\kappa = d\log(z^+)/dU^+$ of the theoretical profile deviates by more than 2\% of the used value $\kappa = 0.416$. minimum height: seventh grid point

\begin{table*}
	\centering
	\caption{Simulation results}
		\begin{tabular}{lccccc}
		  Name & $\kappa_{LES}$ & $z_{0,LES}/z_0$ & $\delta_{95}/\delta$ & $u_\star/u_{\star,dl}$ & $\alpha/\alpha_{dl}$ [$^\circ$] \\ \hline
			Re1\_150 & 0.449 & 0.36 & 0.58 & 1.00 & 0.91  \\
			Re1\_150\_dyn & 0.417 & 0.88 & 0.58 & 1.03 & 0.89 \\
			Re1\_100 & 0.485 & 0.18 & 0.61 & 1.00 & 0.89 \\
			Re1\_50 & - & - & 0.70 & 1.00 & 0.86 \\ \hline
			Re2\_150 & 0.464 & 0.09 & 0.59 & 1.00 & 0.92 \\
			Re2\_100 & 0.507 & 0.18 & 0.62 & 1.00 & 0.86 \\
			Re2\_50 & - & - & 0.73 & 1.00 & 0.85 \\ \hline
			Re2\_150\_a & 0.458 & 0.12 & 0.60 & 1.00 & 0.88 \\
			Re2\_150\_b & 0.460 & 0.11 & 0.59 & 1.00 & 0.86 \\
			Re2\_150\_c & 0.467 & 0.08 & 0.56 & 1.00 & 0.96 \\
			Re2\_150\_d & 0.477 & 0.06 & 0.57 & 1.00 & 0.97 \\ \hline
			Re2\_150\_dyn & 0.443 & 0.28 & 0.58 & 1.01 & 0.91 \\
			Re2\_50\_dyn & - & - & 0.7 & 1.01 & 0.81 \\ \hline
			Re3\_150 & 0.477 & 0.03 & 0.58 & 1.00 & 0.94 \\
			Re3\_150\_dyn & 0.453 & 0.12 & 0.62 & 1.01 & 0.86 \\
			Re3\_100 & 0.514 & 0.01 & 0.63 & 1.00 & 0.88 \\
			Re3\_50 & - & - & 0.72 & 1.00 & 0.85 \\
		\end{tabular}
	\label{simulation_results}
\end{table*}

The turbulent structures in boundary layers of high Reynolds numbers extend over a large range of scales. For the description of their vertical mean profiles, inner units ($z^+ = zu_*/\nu$, $U^+=U/u_*$) and outer units ($z^-=z/\delta$, $U^-=U/G$) are used. At the lower boundary, the x-axis of the inner units is aligned with the shear stress, whereas the x-axis of the outer units is aligned with the geostrophic wind. The angle between both axes is called $\alpha$.

A principal idea behind LES is to neglect the small scales of the flow and resolve only the large eddies, which carry most of the turbulent kinetic energy. Hence, the viscous sublayer and the buffer layer are not resolved by LES and cannot be compared. The first grid point of an LES usually lies inside of the logarithmic region of the boundary layer. Furthermore, the flow is usually underresolved in the lower layers of an LES, since near the bottom, the vertical component is massively restricted by the non-permeability of the wall.

$\delta_{95}$: height where the turbulent stress decreased to 5\% of its surface value. \todo{put $\delta_{95}/\delta$ in table. Which of both do we call boundary layer height?}

\begin{figure}
  \centerline{
	\includegraphics[width=0.3\textwidth]{figures/alt_uv_p_Re1a}
	\includegraphics[width=0.3\textwidth]{figures/alt_uv_p_Re2}
	\includegraphics[width=0.3\textwidth]{figures/alt_uv_p_Re3}
}
  \caption{Shear-aligned profiles in inner scaling}
  \label{3Re_uv_p}
\end{figure}

\subsection{Logarithmic layer stream-wise velocity}

\todo{include $50<z^+<0.15Re_\tau$, while $z^-=z^+Re_\tau$}

Figure \ref{3Re_kappa} shows $\kappa = z^+\partial U^+/\partial z^+$, which is constant in the logarithmic layer. The theoretical profile shows a constant $\kappa$ up to $z^-\approx0.1$ for the case Re1 and up to $z^-\approx0.15$ for the cases Re2 and Re3 \todo{The logarithmic layer already begins to feel the outer layer}. All cases have in common that an increase in resolution leads to a profile much closer to the theoretical curve.

A typical feature of LES is the log-layer mismatch of the first grid points above the bottom, comprehensively discussed by \cite{brasseur2010designing}. According to \cite{maronga2020improved}, mean profiles follow MOST at height levels starting from the seventh grid above the surface. \todo{\citep{jiang2018large} identify an SGS-buffer layer which starts at ...} \todo{define maximum deviation (in percent) from $\kappa$ to define the height of the (nearly) purely logarithmic region. Between the seventh grid point and this height, $\kappa$ of the LES is determined.}

\begin{figure}
  \centerline{
	\includegraphics[width=0.3\textwidth]{figures/kappa_Re1a}
	\includegraphics[width=0.3\textwidth]{figures/kappa_Re2}
	\includegraphics[width=0.3\textwidth]{figures/kappa_Re3}
}
  \caption{$\kappa$ in the logarithmic region and above for different Reynolds numbers and resolutions}
  \label{3Re_kappa}
\end{figure}

An obvious feature of the low Reynolds number case is that the viscous sublayer represents a notable share of the boundary layer, while this layer is hardly visible for the high Reynolds number cases.

A Reynolds number of $Re_D = 1\,600$ is a very unusual Reynolds number for an LES. The first grid point of the LES lies at $\Delta/2$ and should lie in the logarithmic region of the boundary layer (citation?). For the finer resolved simulations of the low Reynolds number, the first grid point lies well in the buffer layer or even in the viscous sublayer. That means that the simulation results of this point are unlikely to produce reasonable results since an LES does not comprehend the physics needed to describe the behavior of the flow in this region.

The transition from buffer layer to logarithmic layer lies around $z^+=19$. The first grid point of the simulation \textit{Re1} lies at $z^+=14$ for the 20\,cm resolution and at $z^+=10$ for the 14\,cm resolution.

The finest resolution ReX\_150 resolves well the boundary layer in agreement with the findings of Wurps2019 where the neutral simulation was well resolved with more than 100 grid levels inside of the boundary layer $\delta_95$. The ratio $\delta_95/\delta$ is roughly $2/3$ (slightly decreasing with Reynolds number). Hence, 150 grid levels within $\delta$ mean around 100 grid levels within $\delta_{95}$.

\begin{figure}
  \centerline{
	\includegraphics[width=0.3\textwidth]{figures/alt_uv_def_Re1a}
	\includegraphics[width=0.3\textwidth]{figures/alt_uv_def_Re2}
	\includegraphics[width=0.3\textwidth]{figures/alt_uv_def_Re3}
}
  \caption{Shear-aligned velocity deficit}
  \label{3Re_uv_def}
\end{figure}

\subsection{Logarithmic layer span-wise velocity}

Figure \ref{v-comp_log}
\begin{figure}
  \centerline{
	\includegraphics[width=0.3\textwidth]{figures/v-comp_log_Re1a}
	\includegraphics[width=0.3\textwidth]{figures/v-comp_log_Re2}
	\includegraphics[width=0.3\textwidth]{figures/v-comp_log_Re3}
}
  \caption{Shear-aligned velocity deficit of the v-component in the logarithmic layer}
  \label{v-comp_log}
\end{figure}

\subsection{Ekman layer}

\begin{figure}
  \centerline{
	\includegraphics[width=0.32\textwidth]{figures/alt_alpha_zm_Re1a}
	\includegraphics[width=0.3\textwidth]{figures/alt_alpha_zm_Re2}
	\includegraphics[width=0.3\textwidth]{figures/alt_alpha_zm_Re3}
}
  \caption{Direction of the mean flow with respect to the geostrophic wind. }
  \label{3Re_alpha}
\end{figure}

Hodograph: in reality (or DNS), the wind profile shows only a very little change of wind direction near the bottom, whereas LES starts veering from the first grid point on.

For the case Re1, the curves for all resolutions fall quite well onto the theoretical hodograph. For the cases Re2 and Re3, the hodographs clearly show that the veering is underestimated by the LES.

There is a clear tendency that finer resolutions lead to a bigger $\alpha$ which is strongest for case Re3. Nevertheless, the predicted $\alpha$ is reached by non of the LES.

The adaption of the theoretical curve by using the $\alpha$ from the fines LES seems to be beneficial only for high Reynolds numbers. $\alpha$ is defined as the direction of the wind at the bottom while in LES we only have the direction of the wind at the first grid point: further veering below the first grid point is not taken into account. For high Reynolds numbers, the direction of the flow stays almost constant for a much wider part of the boundary layer (around 1\% for Re2 and Re3), while for lower Reynolds number, the direction changes significantly much earlier (around 0.1\% for Re1). The total $\alpha$ predicted by the theoretical profile (and the DNS) is $16.8^\circ$ and the $\alpha$ of the LES with finest resolution of case Re1 is $15.3^\circ$. Despite this considerable difference, the curves match quite well and LES reproduces the correct course of wind directions at all heights above the first grid point.

\begin{figure}
  \centerline{
	\includegraphics[width=0.3\textwidth]{figures/alt_hodograph_Re1a}
	\includegraphics[width=0.3\textwidth]{figures/alt_hodograph_Re2}
	\includegraphics[width=0.3\textwidth]{figures/alt_hodograph_Re3}
}
  \caption{Geostrophy aligned hodograph}
  \label{3Re_hodograph}
\end{figure}

Limits of coarse resolution - where is the MOST-Point in Hodograph?

Vergleich SGS-Modelle

As stated before, eq. \ref{drag} is very well validated for the range $400<Re_D<1\,600$ \citep{ansorge2014global}. For Reynolds numbers like $Re_D=1.5\cdot10^5$ or even $Re_D=10^6$, there exist neither DNS nor experimental data. Hence, the solutions of eq. \ref{drag} are not to be taken as certainly correct and one might assume that the LES solution is not definitively incorrect. This being said, the theoretical profiles are recalculated with the values of $u_\star$ and $\alpha$ from the LES solution. \todo{$u_\star$ genau messbar, G nicht. Trotzdem Spalart verlässlich? Integrale? Herleitung angucken. Vielleicht ist Re für die LES anders? $\delta/z_0$ Indeed, \cite{jiang2018large} use $G$, $f$ and $z_0$ as input parameters -- and these \textit{are} the input parameters for an LES. Why are we adjusting $z_0$ to get the $u_\star$ predicted by \cite{spalart1989theoretical}?} Simulations of \cite{jiang2018large} are not the same as ours, since their choice of $z_0$ is a real roughness length and not just caused by the viscosity and a flat plate. Their roughness lengths are so high that the resulting viscosity would have to be considered in the Navier-Stokes equations.

\section{Discussion and Conclusions}

Interesting comparison between DNS and LES

Overall: good fit between theoretical profile and LES

Resolution of LES matters a lot

Helpful for rating LES profile: height where profile should deviate from logarithmic shape.

Theoretical profiles provide a very detailed benchmark for all aspects of the flow. A very good hit of the hodograph might coincide with a poor curve for $\kappa$.

\bibliographystyle{plainnat}
\bibliography{papers.bib}

\end{document}

### 2021-09-29 plot/plot.py

@@ -270,7 +270,7 @@ if plot_evisc:
     ax2=fig.add_axes([0.06,0.11,0.43,0.87])
     ax1=fig.add_axes([0.55,0.11,0.43,0.87])
     fig2=plt.figure(figsize=(12,4.5)) 
-    ax3=fig2.add_axes([0.55,0.11,0.43,0.87])
+    ax3=fig2.add_axes([0.56,0.11,0.43,0.87])
     ax4=fig2.add_axes([0.06,0.11,0.43,0.87])
 
     for re in [500,750,1000,1300,1600]:
@@ -331,7 +331,7 @@ if plot_evisc:
         print(evisc[10],evisc2[10])
         
         norm = 1./(us*us)
-        norm_v_outer=nu/(nu*us*us*us)
+        norm_v_outer=nu/(nu*us*us)#*us)
         norm_v_inner=1./nu
 
         ax3.plot(ym, evisc2*norm_v_outer,lw=2,ls='-',label='Re={}'.format(re),c=colors[re])
@@ -401,8 +401,8 @@ if plot_evisc:
     ax3.set_xlabel(r'$z^-$')
     ax3.set_xscale('log')
     ax3.set_xlim(1e-3,2)
-    ax3.set_ylim(0,0.7)
-    ax3.set_ylabel(r'$\dfrac{\nu_E}{\nu} \dfrac{Z_\star}{\delta^+}$')
+    ax3.set_ylim(0,0.1)
+    ax3.set_ylabel(r'$\dfrac{\nu_E}{\nu} \dfrac{1}{\delta^+}$')
     
     fig.savefig('stresses.pdf',format='pdf')
     fig2.savefig('eddy_viscosity.pdf',format='pdf')

### 2021-10-26 write1/manuscript.tex
continued writing on Section 4 -- ekman profile; spanwise profile; started on streamwise profile

@@
 \usepackage{graphicx}
-\usepackage{caption}
+\usepackage[footnotesize]{caption}
 \usepackage{xcolor}
@@
 %
-Wind veer in the planetary boundary layer is commonly characterized by the surface veering angle $\alpha$
+In the planetary boundary layer, wind veer is characterized by the surface veering angle $\alpha$
 defined as the angle between the negative surface shear stress $\tau_\SFC$ and the geostrophic wind.
@@
 %
-We denote the normalized modulus of the surface friction by
+We denote the modulus of the surface shear, the surface friction, by
 \begin{equation}
@@
 %
-These yield a Reynolds number as non-dimensional parameter governing the system.
+In absence of external variability, this system converges to a statistically steady state in the sense that flow
+statistics do not depend on time; and this state is defined by a Reynolds number,
+the only non-dimensional parameter that governs the system.
 %
@@
 \begin{subequations} 
+\label{eqn:governing} 
 \begin{eqnarray}
@@
 \end{eqnarray} 
-\label{eqn:governing} 
 \end{subequations}
@@
 We extend an existing set of simulations for $\RE_\Lambda\in\{125\,000; 281\,250;\ 500\,000\}$ \citep[gray shading; cf.][]{ansorge:BM2014,ansorge:JFM2016} 
-by new simulations at higher Reynolds numbers up to $1.28\times10^6$ with a horizontal domain extent up to $3.3\times 10^4$ viscous units. 
+by new simulations at higher Reynolds numbers up to $\RE_\Lambda=1.28\times10^6$ with a horizontal domain extent up to $3.3\times 10^4$ viscous units. 
 %
@@
 %
-The profiles of the shear-aligned streamwise velocity component are well-collapsed for the Reynolds
-numbers investigated below $z^-\approx 0.15$ (circles in Fig.~\ref{fig:profiles}a); the logarithmic
-law is appropriate for $50 < z^+ < 0.15 Re_\tau$, where $z^+ = Re_\tau z^-$.
+The profiles of the shear-aligned streamwise velocity component are well-collapsed for $\RE_D>500$
+below $z^-\approx 0.15$ (circles in Fig.~\ref{fig:profiles}a); the case with $\RE_D=500$ is only transitionally
+turbulent and there is no distinct inner-oouter scale separation.
+The logarithmic law is appropriate for $50 < z^+ < 0.15 Re_\tau$, where $z^+ = Re_\tau z^-$.
 %
@@
 geostrophic wind speed $G$) is the governing velocity scale for inviscid, i.e. $\RE$-independent,
-normalization. \todo{WHY? cf. TKE/Dissipation normalization above}
+normalization.
+%\todo{WHY? cf. TKE/Dissipation normalization above}
 %
@@
 $(V^{\alpha_\star}(z^-)-G_{2}^{\alpha})/u_\star$
-agree well beyond $z^-\approx 0.3$ whereas the profiles of the streamwise component collapse also
-within the overlap layer that extends much closer down to the surface
+agree well beyond $z^-\approx 0.3$. %
+This is a much higher level in comparison with the streamwise component that collapses
+also within the overlap layer, i.e. much closer to the surface
 (circles in Fig.~\ref{fig:profiles}b).
@@
 \end{equation}
-which has indeed already been used by \cite{spalart:JFM1989}. 
+which has indeed already been found by \cite{spalart:JFM1989}, who estimate the constant from
+an integral relation. 
 %
@@
 \end{figure} 
-The normalized viscous stress
+The viscous stress
 %
@@
 %
-The inner normalization of $\nu_E$ is obtained by division through the molecular viscosity, and it is $\nu_E^+=\nu_E/\nu = S_\mathrm{turb}/S_\mathrm{visc}$.
+The inner normalization of $\nu_E$ is obtained when dividing by the molecular viscosity:
+\begin{equation}
+  \nu_E^+=\nu_E/\nu = S_\mathrm{turb}/S_\mathrm{visc}.
+\end{equation}
+%
 Under this normalization, the profiles of eddy viscosity collapse below $z^+\approx 20$ with a tendency towards better collapse
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Towards a Universal Velocity Profile for the Turbulent Ekman Layer}
+\section{A universal velocity profile for the turbulent Ekman layer}
 \label{sec:profiles} 
@@
 %
-While such rotation of the wind vector is \emph{a posteriori} justified by the observational data that the model outcomes are compared to, understand this as the manifestation of a mismatch between the theoretical treatment of the inner and outer layer of
-turbulent Ekman flow.
-%
-\par
-%
-The classic text-book solution of Ekman flow makes use of the physical boundary conditions (BCs) for the atmospheric
-boundary layer (no-slip at the bottom and geostrophic wind in the free atmosphere).
-%
-These two boundary conditions eliminate one mode of the analytical solution and determine the magnitude of
-the spiral.
+Such rotation of the wind vector is \emph{a posteriori} justified by the observational data that the model outcomes are compared to.
+%
+This need for a connection of the two reference frames is a manifestation of a
+mismatch in the theoretical treatment of the inner and outer layer in this rotating flow configuration. 
+%
+\par
+%
+The text-book solution for Ekman flow makes use of the physical boundary conditions (BCs) for the atmospheric
+boundary layer (no-slip at the bottom and geostrophic wind in the free atmosphere) and a constant eddy
+viscosity.
+%
+Specifying the boundary conditions at top and bottom eliminates one mode of the analytical solution, and
+it determines the magnitude of the spiral.
 %
 In doing so, one has to assume that the solution is appropriate across the entire ABL, which is not the case:
-the dynamics put forth by \cite{ekman:AMA1905} are not appropriate in the surface layer of the ABL, where
-the logarithmic, buffer, and viscous sublayers are found. 
-%
-We to use a virtual boundary condition instead that takes into account the increased gradients in vicinity of
-the surface and matches the magnitude of the Ekman spiral to the values observed in our DNS:
+The dynamics put forth by \citeauthor{ekman:AMA1905} in \citeyear{ekman:AMA1905}
+are not appropriate in the surface layer of the ABL; better approximations exist for the logarithmic,
+buffer, and viscous sublayers. 
+%
+In view of this situation, we use an adapted Ekman spiral that does not enforce the
+boundary conditions at the surface but at a different height. 
+%
+This is achieved by considering the Ekman spiral only in the Ekman layer,
+thus giving way for the well-established logarithmic and viscous-layer profiles in the
+lower surface layer. 
+%
+Based on the derivation in App.~\ref{app:ekman_solution}, this profile is given by 
 \begin{subequations}
+  \label{eqn:profile_ekman}
   \begin{eqnarray}
-  G^{-1}\left(\begin{array}{c} 
+    \frac{1}{G}\left(\begin{array}{c} 
       U_\mathrm{ek}\\ 
@@
     \end{array}\right)  &=& \left(\begin{array}{c} 1 \\ 0 \end{array}\right) 
-  + a_\mathrm{ek} e^{-z_\mathrm{ek}} \left(\begin{array}{r}
+    + e^{-z_\mathrm{ek}} \left[ a_\mathrm{ek}  \left(\begin{array}{r}
       -\cos z_\mathrm{ek} \\ 
       \sin z_\mathrm{ek}
-  \end{array}\right)  \\
-  z_\mathrm{ek} &=& \delta_{ek} (y^- -s_{ek}). 
+      \end{array}\right)  + b_\mathrm{ek}\left(\begin{array}{r}
+        \sin z_\mathrm{ek} \\ \cos z_\mathrm{ek}
+        \end{array}\right)\right]\\
+    z_\mathrm{ek} &=& \delta_\mathrm{ek} (y^- -s_\mathrm{ek}). 
   \end{eqnarray}
@@
 %
-This solution is characterized by three parameters,
+The right-hand-side consists of two modes with magnitude $a_\mathrm{ek}$ and $b_\mathrm{ek}$
+shifted by $\pi/2$ with respect to each other. 
+%
+In the classic case, the second mode governed by $b_\mathrm{ek}$ is incompatible with
+the surface boundary condition. 
+%
+While this is not the case here for the general form of the profile, we
+phase shifts can also be captured by the parameter $s_\mathrm{ek}$, 
+and we stick with to a single-modal approach, i.e., we let $b_\mathrm{ek}=0$. 
+%
+\par 
+%
+This single-modal solution is characterized by three parameters,
 (i) an Ekman-layer depth scale $\delta_\mathrm{ek}$,
-(ii) the magnitude parameter of the spiral $a_{ek}$, and
-(iii) a zero-crossing point for the velocity $s_{ek}$.
-\begin{enumerate} 
-%
-\item[(i)] The Ekman-layer depth scale $\delta_\mathrm{ek}$ is fundamentally defined by the eddy diffusivity.
+(ii) the magnitude parameter of the spiral $a_\mathrm{ek}$, and
+(iii) a zero-crossing point for the velocity $s_\mathrm{ek}$.
+%
+The effects of varying these parameters are illustrated in Fig.~\ref{fig:ekman_ideal} where
+the classic Ekman solution is recovered by setting $a_\mathrm{ek}=1$, $s_\mathrm{ek}=0$ and
+$\delta_\mathrm{ek}=\sqrt{2\nu/f}$.
+% 
+These parameters are \emph{a priori} unknown as they need to conform to the turbulent state of the
+boundary layer; we use our DNS data to arrive at best estimates for them.
+%
+\par
+%
+\begin{figure}
+  \begin{flushleft}
+    \textbf{\ \ a)}\hspace{0.45\textwidth}\textbf{\ \ b)}\\
+  \end{flushleft} 
+  \includegraphics[width=\textwidth]{../plot/ekman_ideal.pdf}
+  \caption{\label{fig:ekman_ideal}
+    \textbf{a)} Generalized Ekman-profile of the geostrophic-aligned component $U_\mathrm{ek}$.
+    \textbf{b)} Hodograph for the geostrophic-aligned and pressure-gradient aligned components $U_\mathrm{ek}$ and $V_\mathrm{ek}$.
+    Thick, black dashed line shows the classic solution. The height corresponding to $z^-=0.30$ is marked by the dashed line in panel \textbf{(a)}
+    and by filled circles in panel \textbf{(b)}. The hodograph and profiles above this reference height are
+    shown as solid lines, below as opaque line.
+  } 
+\end{figure} 
+%
+\textbf{The Ekman-layer depth scale} $\delta_\mathrm{ek}$ is fundamentally defined by the eddy viscosity.
 %
@@
 %
-\item[(ii)] The magnitude parameter of the spiral primarily defines the super-geostrophic maximum of the
+\par
+%
+\textbf{The magnitude parameter of the Ekman spiral}, $a_\mathrm{ek}$, defines the super-geostrophic maximum of the
 wind profile aloft the logarithmic layer. Our simulations suggest this maximum of the velocity deficit
@@
 \begin{equation}
-    U_\mathrm{ek}/u_\star-Z_\star \propto a_\mathrm{ek} Z_\star = 8.4G.  \notag
-  \end{equation} 
-\item[(iii)] While $\delta_\mathrm{ek}$ sets the thickness across which the wind veering takes place,
-  the zero-crossing height $s_\mathrm{ek}$ defines an offset.
+  U_\mathrm{ek}/u_\star-Z_\star \propto a_\mathrm{ek} Z_\star = 8.4G.
+  \label{eqn:hodograph_scaling}
+\end{equation}
+Thus, the reduction of the area under the Ekman hodograph is directly linked to the variation
+in the friction velocity, which quantifies the well-known qualitative observation that a more turbulent boundary
+layer leads to a flatter hodograph in Ekman flow.
+%
+\par
+%
+\textbf{The offset parameter} $s_\mathrm{ek}$ defines the zero-crossing height of the profile
+(in contrast to $\delta_\mathrm{ek}$, which determines the thickness across which the wind veering takes place). 
 %
@@
 %
-  From our DNS data, we estimate $s_\mathrm{ek}=-0.12\delta$. 
-\end{enumerate} 
-%
-\todo{Paragraph and figure on idealized Ekman profile}
-%
-\par 
-% 
-The outer layer of Ekman flow is characterized by a turning of the wind velocity and the super-geostrophic maximum
-that is sustained by momentum convergence at the inflection point of the velocity profile.
+From our DNS data, we estimate $s_\mathrm{ek}=-0.12$. 
+%
+\par 
+% 
+In summary, the outer layer of Ekman flow is characterized by a turning of the wind
+velocity and the super-geostrophic maximum that is sustained by momentum convergence
+at the inflection point of the velocity profile.
 %
@@
 %
+Corresponding profiles are shown in comparison with data from three DNS runs in Fig.~\ref{fig:profiles_outer}.
+The idealized profiles capture the secondary minimum and convergence to the geostrophic equlibrium in the
+non-turbulent flow very well.
+%
+
 \begin{figure}
+  \begin{flushleft}
+    \textbf{\ \ a)}\hspace{0.45\textwidth}\textbf{\ \ b)}\\
+  \end{flushleft} 
   \includegraphics[width=0.5\textwidth]{../plot/outer_layer_u.pdf}\includegraphics[width=0.5\textwidth]{../plot/outer_layer_w.pdf}
@@
     \label{fig:profiles_outer}
+    Shear-aligned velocity deficit for the wall-streamwise (panel \textbf{(a)}) and wall-spanwise (panel \textbf{(b)})
+    components of the mean velocity $U^\alpha$ and $V^\alpha$.
+    Solid lines show DNS data, dashed lines the Ekman profiles $U_\mathrm{ek}$ and $V_\mathrm{ek}$ as defined in
+    Eq.~\ref{eqn:profile_ekman}. Region below $z^-=0.3$ is shown opaque. Variations in $U_\mathrm{ek}$ and $V_\mathrm{ek}$ are
+    a consequence of the normalization and related to chagnes in $u_\star$ and $\alpha_\star$ among the different $\RE_D$. 
   }
@@
 \begin{equation}
-  (U,V)^{\alpha_\star+}  = (z^+,0)  
+  U^{\alpha_\star+}  = z^+
 \end{equation}
@@
 viscous sub-layer \cite{foken:m2002,foken:BM1978}.
+%
 In fact, this law of the wall has no degree of freedom given the drag, i.e. once $u_\star$ and $\alpha_\star$ are defined. 
@@
 \begin{equation}
-  U_\text{inner}^{\alpha_\star}^+ = \frac{ y^+ + \gamma_4 (y^+)^{4} + \gamma_6 (y^+)^{6} } { 1 + \gamma_6 /u_{ref} (y^+ )^6}, 
+  U_\text{inner}^{\alpha_\star}^+ = f_{U,\mathrm{inner}}(z^+) = \frac{ z^+ + \gamma_4 (z^+)^{4} + \gamma_6 (z^+)^{6} } { 1 + \gamma_6 /u_{ref} (z^+ )^6}, 
 \end{equation} 
@@
 %
+These correction terms are small for $z^+<5$ where the linear law of the wall
+(shown in the inset in Fig.~\ref{fig:u_profiles}) yields an excellent fit of the data; we use here
+\begin{equation*}
+    \gamma_4 = -3.825 \times 10^{-4};\quad
+    \gamma_6 &=&  6.320 \times 10^{-6};\quad 
+    u_\mathrm{ref} &=& 1.278 \times 10^{1}.
+\end{equation*}
+
 \par
@@
 \begin{figure}
+  \begin{flushleft}
+    \textbf{\ \ a)}\hspace{0.45\textwidth} \textbf{\ \ b)} 
+  \end{flushleft} 
   \includegraphics[width=\textwidth]{../plot/u_profile.pdf}
-  \caption{Shear-aligned profiles of velocity components $U^{\alpha_\star+}$ in inner (left) and outer (right) units.} 
+  \caption{Shear-aligned profiles of velocity components $U^{\alpha_\star+}$ in inner (left) and outer (right) units.
+    \label{fig:u_profiles}} 
 \end{figure} 
@@
 The background rotation and associated veering of the surface wind implies a non-zero profile for the span-wise
-velocity which challenges the conventional assumptions related to the so-called channel-flow analogy:
-%
-While the universal profiles in vicinity of the wall implies that the profile be zero or at least small with respect to the streamwise
-component, the veering requires a value of $V_{top}=U_G\sin\alpha_\star$ in the free stream (and thus also at the top of the boundary layer if we assume
-that substantial velocity gradients are confined to the turbulent part of the flow). 
-%
-This is commonly shown in terms of velocity hodographs aligned with the outer, geostrophic flow (cf. Fig.~\ref{fig:hodograph}) and normalized by
+velocity which challenges the conventional assumptions related to the channel-flow analogy:
+%
+While the analogy with channel flow in vicinity of the wall implies that the streamwise component
+be zero or at least small, the veering requires a value of $V_{top}=U_G\sin\alpha_\star$ in the free stream
+(and thus also at the top of the boundary layer if we assume that any substantial velocity gradient is
+confined to the turbulent part of the flow). 
+%
+This continuous rotation of the wind vector is conveniently visualized by velocity hodographs
+aligned with the outer, geostrophic flow (cf. Fig.~\ref{fig:ekman_ideal}b) and normalized by
 the geostrophic wind. 
 %
-Thus, the drag-law already implies three important properties of these hodographs,
+The geometry of the flow and its drag imply the following for any hodograph: 
 (i) the boundary conditions at the surface,
@@
 \begin{eqnarray} 
- \underline{U}^{G}(z=0) &=& 0,\\ 
- \underline{U}^{G}(z\mapsto\infty) &=& \underline{G}^G = \left(\begin{array}{c}1\\0\end{array}\right) \\
-   \left.\frac{\partial V^G}{\partial U^G}\right|_{z=0} &=& \sin\alpha_\star
+  V^{\alpha_\star}(z=0) &=& 0,\\ 
+  \lim_{z\rightarrow\infty} V^{\alpha_\star} &=& G \sin\alpha_\star \\ 
+  \left.\partial_{z^+} V^{\alpha_\star+}}\right|_{z=0} &=& 0.  
 \end{eqnarray} 
@@
 %
-Outer scaling of the velocity profile further implies that the velocity deficity of the spanwise component $V^{\alpha_\star}-G^{\alpha_\star}$
-be a universal function of the outer height $z^-$, i.e.  
+\par
+%
+Outer scaling of the velocity profile further implies that the velocity deficit of
+$(V^{\alpha_\star}-G^{\alpha_\star})/u_\star$ be a universal function of the outer height $z^-$. 
+%
+In the outer region of the flow (for $z^-\mapsto 1$), $f_V(z^-)$, should govern the
+spanwise velocity profile, as is supported by our DNS data (Fig.\ref{fig:profiles}b);
+above $z^-\approx 0.3$, this profile is very well approximated by the Ekman-turning derived
+above (Eq.~(\ref{eqn:profile_ekman}); Fig.~\ref{fig:profiles_outer}b). 
+%
+While this deficit is a signature of outer rotation,
+it is inappropriate to extend this general relation to the surface where inner
+scales matter:
+%
+On the one hand, the variation of the spanwise velocity deficit across
+the boundary layer (i.e. between $0<z^-<1$) must match the difference implied by the drag law ($u_\star, \alpha_\star$)
+and the constant value of $V^{\alpha_\star}$ around $z^-=0.3$.
+%
+On the other hand,--provided the outer velocity deficit is $\RE$ independent--the $\RE$-dependence
+of $\alpha_\star$ and $u_\star$ implies that this difference cannot be constant as a function of $\RE$
+%
+\par 
+%
+We hence ask, how does the span-wise component scale when the surface is approached?
+%
+Clearly, the spanwise contribution is small in comparison with the streamwise component
+throughout much of the layer below $z^-\approx 0.3$. 
+%
+However, we cannot assume $V=0$ if a smooth matching between the inner and outer layers shall be achieved. 
+%
+In this context, we first realize that the velocity deficit $(V^{\alpha_\star}-G^{\alpha_\star})/u_\star$
+approaches a $\RE$-independent constant around $C_{V0}=Z_\star \sin\alpha -6.1$ at the surface; deviations from
+this constant are only found for the lowest Reynolds numbers which is in accordance with
+the low-$\RE$ correection suggested by \cite{spalart:JFM1989}.
+%
+This constrains the wind veer, and it quantitatively shows that the decreasing wall friction manifest in an
+increase of $Z_\star$ exactly compensates the decrease of wind turning measured by $\sin\alpha_\star$.
+
+If the difference across the boundary layer is constant ($C_{V0}$) across variations of $\RE$,
+the averaged gradient $\partial_{z^+} V^{\alpha_\star+}$ of the spanwise velocity component
+must decrease as $1/\delta^+$ as $Re_\tau$ increases.
+%
+Hence, it should--at a fixed height--be $V^{\alpha_\star} \propto \left(\delta^{+}\right)^{-1}$.
+%
+A profile that agrees with the constraints of the profile at the surface
+and exploits the dependence of $V^{\alpha_\star}$ on $\delta^+$ is 
+\begin{equation}
+  V^{\alpha_\star}\frac{\delta^+}{G} = f_{V,\mathrm{inner}}(z^+) = v_\mathrm{ref} \left( \omega_v z^+ -1 + \exp[-\omega_v z^+]\right),
+  \label{eqn:scaling_spanwise} 
+\end{equation}
+where $v_\mathrm{ref}$ controls the magnitude of the profile and $\omega_v$ sets the height at which the profile
+transitions into an approximately linear one.
+%
+We find excellent agreement with the DNS data for $500\le\RE_D\le1600$ below $z^+\approx 15$ with
+\[v_\mathrm{ref} = 18.85;  \quad \omega_v=0.2353\]
+(cf. Fig.~\ref{fig:inner_w}b). 
+%
+\par
+%
+Eq.~(\ref{eqn:scaling_spanwise}) establishes a universal mixed scaling for the spanwise velocity:
+%
+While it requires the vertical coordinate to be expressed in inner units,
+the velocity itself is normalized by the geostrophic wind, and becomes
+inversely proportional to the friction Reynolds number $\RE_\tau=\delta^+$ when considered at a fixed height.
+%
+In vicinity of the surface, such mixed scaling has already been identified for higher-order statistics in convective 
+flows \citep{mellado:BM2016,li:JAS2018}, where large scales leave their signatures in vicinity of the surface.
+%
+It is important to note here that, while $V$ is a first-order statistic from a statistical perspective,
+the spanwise velocity is a higher-order correction term from the view point of similarity theory and
+the channel-flow analogy that is routinely employed in the surface layer.
+%
+Further, this is consistent with the scaling for the velocity hodograph found in Eq.~(\ref{eqn:hodograph_scaling}) where
+the friction velocity also drops out.
+%
+\par
+%
+The universal scaling of $V^{\alpha_\star}/G\delta^+$ continues through the surface layer,
+and we find a very similar picture to the spanwise component with an increasing
+depth of a quasi-logarithmic layer as shown in Fig.~\ref{fig:inner_w}b.
+%
+We model this layer by a profile of the form
+%
+\begin{equation}
+  f_{V,\mathrm{log}}(z^+) = \frac{V_\mathrm{log}(z^+)}{G} \delta^+ = a_\mathrm{log} + b_\mathrm{log} \log z^+ + c_\mathrm{log} z^+
+\end{equation} 
+%
+This profile is constrained by
+\begin{align}
+  f_{V,\mathrm{log}}(z^+=10) &=& f_{V,\mathrm{inner}}(z^+=10) &=:& v_{10} \simeq 27.3\\
+  \frac{\p}{\p z^+} \left[ f_{V,\mathrm{log}}\right]_{z^+=10} &=& \frac{\p}{\p z^+}\left[ F_{V,\mathrm{inner}}\right]_{z^+=10} &=:&d_{10}\simeq 4.01 \\
+  f_{V,\mathrm{log}}(z^+=0.3\delta^+) &=& V^{\alpha_\star}_\mathrm{ek}(z^-=0.3)\delta^+&=:& v_{03} 
+\end{align}
+where $v_{03}$ is determined by $V_\mathrm{ek}(0.3)$ and $U_\mathrm{ek}(0.3)$ and depends on the Reynolds number.
+
+%
+\begin{figure}
+  \begin{flushleft}
+    \textbf{\ \ a}\hspace{0.45\textwidth}\textbf{\ \ b}\\
+  \end{flushleft} 
+  \includegraphics[width=1.0\textwidth]{../plot/w_profile.pdf}
+  \caption{Profiles of shear-aligned span-wise velocity $(W^\alpha)^+$ versus inner height. 
+    Dashed lines show DNS data, thick, opaque lines are from the semi-emprical theory developed above.
+    Left panel show standard inner normalization. Right panel shows the inviscid normalization yielding
+    a universal profile for the spanwise component of velocity in the inner layer. }
+  \label{fig:inner_w}
+\end{figure}
+\end{subequations}
+
+\begin{figure}
+  \phantom{AAA}\textbf{(a)\hspace{0.3\textwidth}(b)\hspace{0.3\textwidth}(c)}\\
+  \includegraphics[width=\textwidth]{../plot/ekman_profiles.pdf}
+  \caption{ \textbf{(a)} Velocity deficity, \textbf{(b)} velocity profile in shear-aligned hodographs and \textbf{(c)} hodograph in geostrophy-aligned coordinates.
+    Thick, solid lines show theory, dashed lines data from DNS. }
+  \label{fig:hodograph} 
+\end{figure}
+
+\section{Comparison with other theories}
+
+\textbf{Compare} to Emeis (2002) and Gryning (2007); highlight explicit knowledge on veering-profile $\rightarrow$ directional sheer;
+%
+\par%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+Implications for \textbf{K-theory} (we now can consider that shear and stress are not necessarily perfectly aligned).
+$\rightarrow$ can we do something to infer a K-profile from these theoretical considerations? 
+%
+\begin{figure}
+  \includegraphics[width=\textwidth]{../plot/mag_dir.pdf} 
+  \caption{Total velocity and veering (in degrees) vs inner and outer height.
+    Dashed lines show DNS data, thick lines are from semi-empirical theory.} 
+\end{figure} 
+%
+\section{Conclusions}
+%
+Applications:
+\begin{itemize}
+\item reference-shear for neutral profile approaches(systematic!) $\rightarrow$ wind engineering! 
+\item initial condition for LES/DNS to eliminate/minimize inertial oscillation in Benchmark simulations
+\item
+\end{itemize}
+
+\appendix
+
+\section{Laminar Ekman solution with consideration of inner layer}
+\label{app:ekman_solution}
+
+\begin{subequations}
 \begin{eqnarray}
-  V^{\alpha_\star}-G^{\alpha_\star} = f^{\alpha_\star}_V(z^-).  
+  \left(\begin{matrix}
+    \p_t U\\
+    \p_t W  
+  \end{matrix}\right)&=&\left(\begin{matrix}
+     fW &+& \nu \p_z^2 U\\ 
+    -f(U-G) &+& \nu \p_z^2 W
+  \end{matrix}\right)\\ 
+  \Rightarrow \partial_t (U+iW) &=& f(W-i(U-G))) + \nu \p_z^2(U+iW)
 \end{eqnarray}
-%
-Interpreting the shear-aligned spanwise velocity deficit $f_{V}^{\alpha_\star}(z^-)$ as 
-a signature of outer rotation may lead one to conclude that this universal function
-is appropriate down to the surface--irrespective of the Reynolds number.
-%
-This cannot be the case for the variation of $f_{V}^{\alpha_\star}(z^-)$ across the boundary layer (i.e. between $0<z^-<1$)
-must fulfill the drag law and boundary conditions;
-at the same time, the weak dependence of $\alpha_\star$ and $u_\star$ on $Re$ implies this range depends on $Re$.
-%
-%%%PAR
-%
-\paragraph{Outer Layer.} In the outer region of the flow (for $z^-\mapsto 1$), $f_V(z^-)$, should govern the
-spanwise velocity profile which indeed is supported by our DNS data (Fig.\ref{fig:outer_w}):
-%
-The deficit is close to zero for $z^-\gtrsim 1$, and can be approximated by a logarithmic profile in the
-range $0.15\lesssim z^- \lesssim 0.4$.
-%
-The slope of this profile is best fit by $\kappa_V=0.836 (1 + 150 / Re_D) \kappa$, where the correction term for low Reynolds number
-may be omitted for $Re_D \gg 1000$.
-%
-Based on Fig.~\ref{fig:outer_w}, we pick the offset as
-a constant fraction (0.906) of $V^\alpha_G=Z_\star \sin\alpha$, i.e
-\begin{subequations} 
-\begin{align}
-  V_\text{log}(z^-)  = \frac{1}{\kappa_V} \log\left(\frac{z^-}{0.4}\right) + 0.906V_G^{\alpha}. 
-\end{align} 
-%
-%%%PAR
-% 
-The eventually converged simulations at intermediate $Re_D$ (where the DNS could be run over several inertial periods)
-exhibit a super-geostrophic maximum (positiv velocity deficit) in the span-wise component.
-%
-This maximum is not present or less pronounced for the simulations at higher $Re$ where closer analysis reveals that the
-velocity profiles in the outer layer are still subject to a small though finite adaptation process on the inertial time
-scales $2\pi/f$ due to the relatively short time span for which these cases could be simulated.
-%
-While there is no physical model for this wake region that leads to an analytical velocity profile formulation in the span-wise direction,
-an error-function transition from the outer logarithmic law to the boundary condition provides a reasonable fit:
-%
-\begin{align}
-  V_\text{outer}(z^-) = (1-w_V)V_\text{log}(z^-) + w_V V_G^\alpha
-\end{align} 
-with
-\begin{align} 
-w_V = \frac{\mathrm{erf}\left[\xi_\text{outer} \log(z/z_\text{ref})\right]  + 1}{2}
-\end{align}
-where $\xi_\text{outer}$ is a transition scale that defines the width of the wake region and $z_\text{ref}$ defines the height.
-Based on the DNS data, we choose $\xi_\text{outer}=2$ and let $z_\text{ref}$ such that $V_\text{log}(z_\text{ref}) = 0.98V_G^\alpha$.
-There is, however, a degree of ambiguity in the exact choice of these parameters due to the imperfect convergence of the
-DNS data to the statistical equilibrium.
-%
+In stationary conditions, this system is solved by
+\begin{eqnarray}
+  \hat{u}(z) &=& U_{\infty} + e^{-\gamma z} \left[A \cos\gamma z + B \sin\gamma z\right]   \\
+  \hat{w}(z) &=& W_{\infty} + e^{-\gamma z} \left[-A \sin\gamma z + B \cos\gamma z\right]
+\end{eqnarray} 
 \end{subequations}
-%
-%%% PAR
-%
-\subsub{Inner Layer.} It is important to note that the logarithmic layer observed here for the spanwise velocity does not
-conincide with the log-region of the streamwise velocity but is rather a 'continuation' through the outer layer.
-%
-This is consistent with the recent finding of a second log-region in the outer region of Ekman flow
-(... et al. (incl. Sullivan), JAS 2019).
-%
+where the constants $U_\infty$, $W_\infty$ set the top boundary condition and $A$ and $B$ set the bottom boundary condition. 
+%
+The most common boundary condition for a surface Ekman layer is $A=U_{\infty}=G$, $B=0$, and $W_{\infty}=0$.
+%
+The lower boundary condition, however, neglects the existence of the surface layer, and it appears reasonable to define
+$A=c G$ where $c<1$ is a constant that icorporates the increased shear in the surface layer.
+%
+Given a 'matching height' $z_{match}$ and normalized matching height $\xi=\gamma z_\text{match}$ in the upper part of the inner layer, we can match the Ekman profile
+to the inner layer by letting
+%
+\begin{eqnarray}
+  \begin{matrix} 
+    u(z_\text{match}) \equiv u_\text{match} &=& U_{\infty} + e^{-\gzm}} \left [ A \cos\gzm + B\sin\gzm \right] \\
+    w(z_\text{match}) \equiv w_\text{match} &=& W_{\infty} + e^{-\gzm}} \left [-A \sin\gzm + B\cos\gzm \right]
+  \end{matrix} 
+\end{eqnarray}
+\begin{eqnarray}
+  \Rightarrow \left(\begin{matrix}
+    u_\text{match}-U_{\infty} \\ 
+    w_\text{match}-W_{\infty}
+  \end{matrix}\right) = e^{-\gzm}\left(\begin{matrix}A \\ B \end{matrix}\right)\left(\begin{matrix}
+    \cos\gzm &+ \sin\gzm \\ -\sin\gzm &+ \cos\gzm
+  \end{matrix} \right) \\
+\end{eqnarray}
+Matching the profile at $\xi=0$, one obtains
+$A = \Delta u_\text{match} $ and $B=-\Delta w_\text{match} $; and when the direction $Ox$ is aligned
+with the geostrophic wind, we obtain the textbook-case $A=|\mathbf{G}|$ and $B=0$.
+
+Otherwise, choosing $B \ne 0$ allows to introduce a phase shift of the Ekman rotation with respect to
+the decay of the wind spiral.
+%
+As, however, in our context, the thickness and position of the spiral can already be controlled by
+the eddy viscosity and an offset in $\zeta$, here we let $B=0$.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Determining parameters A and B from profile values at a given height
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% 
+%% \begin{subequations} 
+%% \begin{eqnarray}
+%%   e^\gzm \Delta u_\text{match} = A \cos\gzm + B\sin\gzm \\
+%%   e^\gzm (\cos\gzm \Delta u_\text{match} - \sin\gzm \Delta w_\text{match}) = A (\cos^2\gzm + \sin^2\gzm)
+%% \end{eqnarray}
+%% \begin{eqnarray}
+%%   \Rightarrow A &=&e^\gzm (\cos\gzm\Delta u_\text{match} + \sin\gzm \Delta w_\text{match})\\
+%%   B&=&  \frac{e^\gzm \Delta u_\text{match} - A \cos \gzm}{\sin\gzm} \\
+%%   &=& e^\gzm \frac{ \sin^2\gzm \Delta u_\text{match}+\sin\xi\cos\xi \Delta w_\text{match} }{\sin\xi }  \\
+%%   &=& e^\gzm (\sin\xi\Delta u_\text{match} + \cos\xi\Delta w_\text{match}) 
+%% \end{eqnarray}
+%% \end{subequations}
+
+\printbibliography
+
+\appendix
+\section{Old Stuff}
+\todo{OLD FROM HERE} 
 Below this region, the gradients in span-wise velocity are rather small and the span-wise velocity
@@
 %
-This illustrates that the higher-order (in terms of the channel-flow analogy where $V=0$) correction follows a mixed scaling when the surface is
-approached which has been demonstrated for higher-order terms in other types of flow \citep{mellado:BM2016}. 
 %
@@
 
-%
-\begin{figure}
-  \includegraphics[width=1.0\textwidth]{../plot/w_profile.pdf}
-  \caption{Profiles of shear-aligned span-wise velocity $(W^\alpha)^+$ versus inner and outer height.
-  Dashed lines show DNS data, thick, opaque lines are from the semi-emprical theory developed above. }
-  \label{fig:outer_w} 
-\end{figure}
-\end{subequations}
-
-\begin{figure}
-  \phantom{AAA}\textbf{(a)\hspace{0.3\textwidth}(b)\hspace{0.3\textwidth}(c)}\\
-  \includegraphics[width=\textwidth]{../plot/ekman_profiles.pdf}
-  \caption{ \textbf{(a)} Velocity deficity, \textbf{(b)} velocity profile in shear-aligned hodographs and \textbf{(c)} hodograph in geostrophy-aligned coordinates.
-    Thick, solid lines show theory, dashed lines data from DNS. }
-  \label{fig:hodograph} 
-\end{figure}
-
-\section{Comparison with other theories}
-
-\textbf{Compare} to Emeis (2002) and Gryning (2007); highlight explicit knowledge on veering-profile $\rightarrow$ directional sheer;
-%
-\par%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Implications for \textbf{K-theory} (we now can consider that shear and stress are not necessarily perfectly aligned).
-$\rightarrow$ can we do something to infer a K-profile from these theoretical considerations? 
-%
-\begin{figure}
-  \includegraphics[width=\textwidth]{../plot/mag_dir.pdf} 
-  \caption{Total velocity and veering (in degrees) vs inner and outer height.
-    Dashed lines show DNS data, thick lines are from semi-empirical theory.} 
-\end{figure} 
-%
-\section{Conclusions}
-%
-Applications:
-\begin{itemize}
-\item reference-shear for neutral profile approaches(systematic!) $\rightarrow$ wind engineering! 
-\item initial condition for LES/DNS to eliminate/minimize inertial oscillation in Benchmark simulations
-\item
-\end{itemize}
-
-\appendix
-
-\section{Laminar Ekman solution with consideration of inner layer}
-
-\begin{subequations}
-\begin{eqnarray}
-  \left(\begin{matrix}
-    \p_t U\\
-    \p_t W  
-  \end{matrix}\right)&=&\left(\begin{matrix}
-     fW &+& \nu \p_z^2 U\\ 
-    -f(U-G) &+& \nu \p_z^2 W
-  \end{matrix}\right)\\ 
-  \Rightarrow \partial_t (U+iW) &=& f(W-i(U-G))) + \nu \p_z^2(U+iW)
-\end{eqnarray}
-In stationary conditions, this system is solved by
-\begin{eqnarray}
-  \hat{u}(z) &=& U_{\infty} + e^{-\gamma z} \left[A \cos\gamma z + B \sin\gamma z\right]   \\
-  \hat{w}(z) &=& W_{\infty} + e^{-\gamma z} \left[-A \sin\gamma z + B \cos\gamma z\right]
-\end{eqnarray} 
-\end{subequations}
-where the constants $U_\infty$, $W_\infty$ set the top boundary condition and $A$ and $B$ set the bottom boundary condition. 
-%
-The most common boundary condition for a surface Ekman layer is $A=U_{\infty}=G$, $B=0$, and $W_{\infty}=0$.
-%
-The lower boundary condition, however, neglects the existence of the surface layer, and it appears reasonable to define
-$A=c G$ where $c<1$ is a constant that icorporates the increased shear in the surface layer.
-%
-Given a 'matching height' $z_{match}$ and normalized matching height $\xi=\gamma z_\text{match}$ in the upper part of the inner layer, we can match the Ekman profile
-to the inner layer by letting
-%
-\begin{eqnarray}
-  \begin{matrix} 
-    u(z_\text{match}) \equiv u_\text{match} &=& U_{\infty} + e^{-\gzm}} \left [ A \cos\gzm + B\sin\gzm \right] \\
-    w(z_\text{match}) \equiv w_\text{match} &=& W_{\infty} + e^{-\gzm}} \left [-A \sin\gzm + B\cos\gzm \right]
-  \end{matrix} 
-\end{eqnarray}
-\begin{eqnarray}
-  \Rightarrow \left(\begin{matrix}
-    u_\text{match}-U_{\infty} \\ 
-    w_\text{match}-W_{\infty}
-  \end{matrix}\right) = e^{-\gzm}\left(\begin{matrix}A \\ B \end{matrix}\right)\left(\begin{matrix}
-    \cos\gzm &+ \sin\gzm \\ -\sin\gzm &+ \cos\gzm
-  \end{matrix} \right) \\
-\end{eqnarray}
-Matching the profile at $\xi=0$, one obtains
-$A = \Delta u_\text{match} $ and $B=-\Delta w_\text{match} $; and when the direction $Ox$ is aligned
-with the geostrophic wind, we obtain the textbook-case $A=|\mathbf{G}|$ and $B=0$.
-
-Otherwise, choosing $B \ne 0$ allows to introduce a phase shift of the Ekman rotation with respect to
-the decay of the wind spiral.
-%
-As, however, in our context, the thickness and position of the spiral can already be controlled by
-the eddy viscosity and an offset in $\zeta$, here we let $B=0$.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Determining parameters A and B from profile values at a given height
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% 
-%% \begin{subequations} 
-%% \begin{eqnarray}
-%%   e^\gzm \Delta u_\text{match} = A \cos\gzm + B\sin\gzm \\
-%%   e^\gzm (\cos\gzm \Delta u_\text{match} - \sin\gzm \Delta w_\text{match}) = A (\cos^2\gzm + \sin^2\gzm)
-%% \end{eqnarray}
-%% \begin{eqnarray}
-%%   \Rightarrow A &=&e^\gzm (\cos\gzm\Delta u_\text{match} + \sin\gzm \Delta w_\text{match})\\
-%%   B&=&  \frac{e^\gzm \Delta u_\text{match} - A \cos \gzm}{\sin\gzm} \\
-%%   &=& e^\gzm \frac{ \sin^2\gzm \Delta u_\text{match}+\sin\xi\cos\xi \Delta w_\text{match} }{\sin\xi }  \\
-%%   &=& e^\gzm (\sin\xi\Delta u_\text{match} + \cos\xi\Delta w_\text{match}) 
-%% \end{eqnarray}
-%% \end{subequations}
-
-\printbibliography
-
 \end{document}

### 2021-10-26 plot/EkmanProfileClass.py

@@ -189,6 +189,19 @@ class EkmanUniversalClass:
 
     def erf_transition(self,w,l,x):
         return 0.5*(sp.special.erf(w*np.log(x/l))+1)
+
+    def profile_ekman(self,ym,us_loc):
+        z=1/us_loc
+        
+        c_argz=0.66*2.*np.pi 
+        argz=    c_argz*(ym +0.12) #0.66*(2.*np.pi*(ym+0.12)) 
+        dampA =  8.4*us_loc        #- 0./re #np.exp(zdum) * ( dU_mtc*np.cos(zdum) + dW_mtc*np.sin(zdum) )
+        #dampB= 0. 
+        e_u= (1-(dampA*np.cos(argz))*np.exp(-argz))*z  #-dampB*np.cos(argz)*np.exp(-argz)*z
+        e_w=   ( dampA*np.sin(argz))*np.exp(-argz)*z   #-dampB*np.cos(argz)*np.exp(-argz)*z
+
+        return [e_u,e_w]
+
     
     def profile_plus(self,yp,re,us=-1,al=-1):
         self.qinit()
@@ -207,20 +220,16 @@ class EkmanUniversalClass:
         u=np.zeros(len(yp))
         w=np.zeros(len(yp)) 
         
-        z=1/us_loc 
+        z=1./us_loc 
         dp_loc=re**2/(2*z**2)
         ym=np.array(yp)/dp_loc
         trans_scale=2.0
 
-        ################################################################################
-        # 2-COMPONENT OUTER REFERENCE PROFILES
-        c_argz=0.66*2.*np.pi 
-        argz=    c_argz*ym + c_argz*0.12 #0.66*(2.*np.pi*(ym+0.12)) 
-        dampA =  8.4*us_loc        - 0./re #np.exp(zdum) * ( dU_mtc*np.cos(zdum) + dW_mtc*np.sin(zdum) )
-        #dampB= 0. 
-        outer_u= (1-(dampA*np.cos(argz))*np.exp(-argz))*z  #-dampB*np.cos(argz)*np.exp(-argz)*z
-        outer_w=   ( dampA*np.sin(argz))*np.exp(-argz)*z   #-dampB*np.cos(argz)*np.exp(-argz)*z
-        [ous,ows]=mp.rotate(outer_u,outer_w,al_loc)
+        # Ekman profile -- geostrophic alignment
+        [ous,ows]=self.profile_ekman(ym,us_loc) 
+
+        # Shear-aligned Ekman profile
[…]

### 2021-10-26 plot/plot.py

@@ -17,12 +17,13 @@ D2R = np.pi / 180.
 
 use_data=False   # if true -- the netcdf files from DNS are needed 
 plot_ustar_alpha=False
+plot_ekman_ideal=True 
 plot_summary=False
-plot_evisc=True
+plot_evisc=False
 plot_profiles=True
-plot_visc_outer=False
+plot_visc_outer=True
 print_table=False
-plot_outer_log=False
+plot_outer_log=True
 plot_total_rot=False
 plot_convergence=False
 plot_re1600=False
@@ -264,6 +265,58 @@ if plot_re1600:
 
     ax1.set_xlim(1,1e4)
     ax2.set_xlim(1,1e4)
+
+if plot_ekman_ideal:
+    fig=plt.figure(figsize=(10,3.75))
+    ax1=fig.add_axes([0.06,0.11,0.41,0.87])
+    ax2=fig.add_axes([0.55,0.11,0.44,0.87])
+
+    z=np.arange(0,2*np.pi,0.02)
+    
+    sa_arr=[[+0.00, 0.4,'black', '-', 1.0],
+            [+0.12, 0.4,'red',   '-', 4.0],
+            [-0.12, 0.4,'blue',  '-', 1.0],
+            [+0.00, 1.0,'black', ':', 4.0], 
+            [+0.12, 1.0,'red',  '--', 1.0],
+            [-0.12, 1.0,'blue', '--', 1.0]
+            ]
+
+    for sa in sa_arr:
+        sek=sa[0]
+        aek=sa[1]
+        zek = z-sek
+        Uek=1-aek*np.exp(-zek)*np.cos(zek)
+        Vek=aek*np.exp(-zek)*np.sin(zek)
+        c=sa[2]
+        s=sa[3] 
+        w=sa[4]
+
+        i03=mp.geti(z,0.3*np.pi*2)
+        
+        ax1.plot(z[i03:],Uek[i03:],  c=c,ls=s,lw=w, label='a={1:4.1f},s={0:5.2f}'.format(sek,aek))
+        ax1.plot(z[:i03],Uek[:i03],  c=c,ls=s,lw=w, alpha=0.5)
+        
+        ax2.plot(Uek[i03:],Vek[i03:], c=c,ls=s,lw=w,label='a={1:4.1f} s={0:5.2f}'.format(sek,aek))
[…]

### 2021-10-27 write1/manuscript.tex
completed section 4.4

@@
 \begin{equation}
-  U_\text{inner}^{\alpha_\star}^+ = f_{U,\mathrm{inner}}(z^+) = \frac{ z^+ + \gamma_4 (z^+)^{4} + \gamma_6 (z^+)^{6} } { 1 + \gamma_6 /u_{ref} (z^+ )^6}, 
+  U_\text{inner}^{\alpha_\star}^+ = f_{U,\mathrm{visc}}(z^+) = \frac{ z^+ + \gamma_4 (z^+)^{4} + \gamma_6 (z^+)^{6} } { 1 + \gamma_6 /u_{ref} (z^+ )^6}, 
 \end{equation} 
@@
 
-If the difference across the boundary layer is constant ($C_{V0}$) across variations of $\RE$,
+If the difference across the boundary layer is constant ($C_{V0}$) vs. $\RE$,
 the averaged gradient $\partial_{z^+} V^{\alpha_\star+}$ of the spanwise velocity component
-must decrease as $1/\delta^+$ as $Re_\tau$ increases.
+must decrease as $1/\delta^+$ with increasing $Re_\tau$. 
 %
@@
 \begin{equation}
-  V^{\alpha_\star}\frac{\delta^+}{G} = f_{V,\mathrm{inner}}(z^+) = v_\mathrm{ref} \left( \omega_v z^+ -1 + \exp[-\omega_v z^+]\right),
+  V^{\alpha_\star}\frac{\delta^+}{G} = f_{V,\mathrm{visc}}(z^+) = v_\mathrm{ref} \left( \omega_v z^+ -1 + \exp[-\omega_v z^+]\right),
   \label{eqn:scaling_spanwise} 
@@
 (cf. Fig.~\ref{fig:inner_w}b). 
-%
-\par
-%
-Eq.~(\ref{eqn:scaling_spanwise}) establishes a universal mixed scaling for the spanwise velocity:
-%
-While it requires the vertical coordinate to be expressed in inner units,
-the velocity itself is normalized by the geostrophic wind, and becomes
-inversely proportional to the friction Reynolds number $\RE_\tau=\delta^+$ when considered at a fixed height.
-%
-In vicinity of the surface, such mixed scaling has already been identified for higher-order statistics in convective 
-flows \citep{mellado:BM2016,li:JAS2018}, where large scales leave their signatures in vicinity of the surface.
-%
-It is important to note here that, while $V$ is a first-order statistic from a statistical perspective,
-the spanwise velocity is a higher-order correction term from the view point of similarity theory and
-the channel-flow analogy that is routinely employed in the surface layer.
-%
-Further, this is consistent with the scaling for the velocity hodograph found in Eq.~(\ref{eqn:hodograph_scaling}) where
-the friction velocity also drops out.
-%
-\par
-%
-The universal scaling of $V^{\alpha_\star}/G\delta^+$ continues through the surface layer,
-and we find a very similar picture to the spanwise component with an increasing
-depth of a quasi-logarithmic layer as shown in Fig.~\ref{fig:inner_w}b.
-%
-We model this layer by a profile of the form
-%
-\begin{equation}
-  f_{V,\mathrm{log}}(z^+) = \frac{V_\mathrm{log}(z^+)}{G} \delta^+ = a_\mathrm{log} + b_\mathrm{log} \log z^+ + c_\mathrm{log} z^+
-\end{equation} 
-%
-This profile is constrained by
-\begin{align}
-  f_{V,\mathrm{log}}(z^+=10) &=& f_{V,\mathrm{inner}}(z^+=10) &=:& v_{10} \simeq 27.3\\
-  \frac{\p}{\p z^+} \left[ f_{V,\mathrm{log}}\right]_{z^+=10} &=& \frac{\p}{\p z^+}\left[ F_{V,\mathrm{inner}}\right]_{z^+=10} &=:&d_{10}\simeq 4.01 \\
-  f_{V,\mathrm{log}}(z^+=0.3\delta^+) &=& V^{\alpha_\star}_\mathrm{ek}(z^-=0.3)\delta^+&=:& v_{03} 
-\end{align}
-where $v_{03}$ is determined by $V_\mathrm{ek}(0.3)$ and $U_\mathrm{ek}(0.3)$ and depends on the Reynolds number.
-
 %
@@
 \end{figure}
+ %
+\par
+%
+For the adjacent surface layer, we find a log-like transition from the quasi-linear profile
+inner profile around $z^+=10$ to a linear profile with increasing $\RE$ (Fig.~\ref{fig:inner_w}b).
+%
+We model this transition by 
+%
+\begin{equation}
+  f_{V,\mathrm{log}}(z^+) = \frac{V_\mathrm{log}(z^+)}{G} \delta^+ = a_\mathrm{log} + b_\mathrm{log} \log z^+ + c_\mathrm{log} z^+. 
+\end{equation} 
+%
+This surface-layer profile matches the inner (viscous) scaling in vicinity of the surface to the outer
+(Ekman) scaling above $z^-=0.3$ when constrained by the viscous profile at the bottom and the
+Ekman profile at the top:
+%
+\begin{align}
+  f_{V,\mathrm{log}}(z^+=10) &=& f_{V,\mathrm{visc}}(z^+=10) &=:& v_{10} \simeq 27.3\\
+  \frac{\p}{\p z^+} \left[ f_{V,\mathrm{log}}\right]_{z^+=10} &=& \frac{\p}{\p z^+}\left[ F_{V,\mathrm{visc}}\right]_{z^+=10} &=:&d_{10}\simeq 4.01 \\
+  f_{V,\mathrm{log}}(z^+=0.3\delta^+) &=& V^{\alpha_\star}_\mathrm{ek}(z^-=0.3)\delta^+&=:& v_{03} 
+\end{align}
+where $v_{03}$ is determined by $V_\mathrm{ek}(0.3)$ and $U_\mathrm{ek}(0.3)$ and depends on $\RE$.
+%
+While the $\RE$-dependency of $a_\mathrm{log}$, $b_\mathrm{log}$, $c_\mathrm{log}$
+is small, it shows up in Fig.~\ref{fig:profiles} where the normalized profiles
+of spanwise velocity become more convex with increasing $\RE$.
+%
+We can now quantify this effect by means of the change of $c_\mathrm{log}$ versus $\RE$ which is shown in
+Fig.~\ref{fig:clog} (cf. Appendix~\ref{app:log-coefficients};
+$a_\mathrm{log}$ and $b_\mathrm{log}$ are then determined by the universal
+values of $v_{10}$ and $d_{10}$).
+%
+%
+\begin{figure}
+  \includegraphics[width=0.5\textwidth]{../plot/c_log.pdf}
+  \caption{c-log.
+    \label{fig:clog}}
+\end{figure} 
+%
+\subsubsection{Discussion}
+Eq.~(\ref{eqn:scaling_spanwise}) establishes a universal mixed scaling for the spanwise velocity
+in the viscous layer:
+%
+While it requires the vertical coordinate to be expressed in inner units,
+the velocity itself is normalized by the geostrophic wind, and becomes
+inversely proportional to the friction Reynolds number $\RE_\tau=\delta^+$
+when considered at a fixed height.
+%
+In vicinity of the surface, such mixed scaling has already been identified for higher-order statistics in convective 
+flows \citep{mellado:BM2016,li:JAS2018}, where large scales leave their signatures in vicinity of the surface.
+%
+It is important to note here that, while $V$ is a first-order statistic from a statistical perspective,
+the spanwise velocity is a higher-order correction term from the perspective of similarity theory and
+from the viewpoint of the channel-flow analogy that is routinely employed in the surface layer.
+%
+Further, this is consistent with the scaling for the velocity hodograph found in
+Eq.~(\ref{eqn:hodograph_scaling}) where the friction velocity also drops out.
+%
+\par
+%
+In the surface layer, there is not only a mixed scaling--as we had already identified
+in the viscous layer--, but we cannot find a universal function onto which the profiles
+of spanwise velocity collapse. 
+%
+This additional degree of freedom reflects the inner--outer matching problem for the
+spanwise velocity, and rather than giving a universal profile for this region,
+as is usually done, we resort here to a parametric description of the problem,
+namely in terms of the function $f_{V,\mathrm{log}}$ determined by the
+parameters $a_\mathrm{log}$, $b_\mathrm{log}$, $c_\mathrm{log}$ which can be estimated
+based on the above scaling considerations for any Reynolds number. 
+%
+We note that, once the parameter $a_\mathrm{log}$ is known, the parameters $b_\mathrm{log}$
+and $c_\mathrm{log}$ can be estimated solely based on $f_{V,\mathrm{visc}}$, i.e. using the
+value $v_{10}$ and $d_{10}$ found for the viscous region of the flow.
+%
+For the range of Reynolds number relevant to geophysical problems ($10^4\lesssim \RE_D \lesssim 10^6$),
+the variation of $c_\mathrm{log}$ is, however, rather small.
+%
 \end{subequations}
@@
 %% \end{subequations}
+
+\section{Matching the spanwise velocity profiles in the inner layer}
+\label{app:log-coefficients}
+The spanwise profile in vicinity of the surface is given by
+$V/G=f_{V,\mathrm{visc}}\delta ^+$ with
+\begin{subequations}
+\begin{eqnarray}
+  f_{V,\mathrm{visc}} &=& v_\mathrm{ref}\left(\omega_v z^+ -1 + e^{-\omega_v z^+}\right)\\
+  f_{V,\mathrm{log}}   &=& a_\mathrm{log} + b_\mathrm{log} \log z^+ + c_\mathrm{log} z^+ 
+\end{eqnarray}
+\end{subequations}
+Matching the profiles and gradient $z_0=10^+$ and the value at $z_1=0.3\dlta^+$ yields
+\begin{subequations} 
+\begin{eqnarray}
+  v_\mathrm{ref} \left(\omega_v z_{0} + e^{-\omega_v z_{0}}\right)
+  & = v_{0}=& a_\mathrm{log} + b_\mathrm{log} \log z_{0} +c_\mathrm{log} z_{0}  \\
+  &   v_{1}=& a_\mathrm{log} + b_\mathrm{log} \log z_{1} +c_\mathrm{log} z_{1} \\
+  v_\mathrm{ref}\omega_z \left( 1 - e^{-\omega_z z_{0}}\right)
+  & = d_{0}=& \frac{b_\mathrm{log}}{z_{10}} + c_\mathrm{log} 
+\end{eqnarray}
+\end{subequations} 
+The gradient condition implies $b_\mathrm{log}=(d_{0} -c_\mathrm{log})z_{0}$, and yields
+\begin{subequations}
+\begin{eqnarray}
+  %v_{0}&=& a_\mathrm{log} + z_{0}\left[ (d_0-c_\mathrm{log}) \log z_{0} +c_\mathrm{log}\right]\\
+  v_0-z_0d_0\log z_0&=& a_\mathrm{log} + c_\mathrm{log} (z_0- z_0\log z_0) \\
+  %v_1 &=& a_\mathrm{log} + z_0(d_0-c_\mathrm{log})\log z_1 +  c_\mathrm{log}z_1\\ 
+  v_1-z_0d_0\log z_1 &=& a_\mathrm{log} +c_\mathrm{log} (z_1- z_0\log z_0)  \\
+  \Rightarrow c_\mathrm{log}&=& \frac{ \Delta v - z_0d_0 \log z_1/z_0}{\Delta z} 
+\end{eqnarray}
+with $\Delta z=z_1-z_0$ and $\Delta v=v_1-v_0$.
+%
+Then, the coefficient $a_\mathrm{log}$ is estimated as 
+\begin{eqnarray} 
+  a_\mathrm{log} &=& v_0-z_0d_0\log z_0 -  \frac{ \Delta v - z_0d_0 \log z_1/z_0}{\Delta z}\left[
+    z_0-z_0\log z_0 
+    \right]. 
+\end{eqnarray}
+We note that $ \log(z_1/z_0) / (z_1-z_0) \rightarrow 0$ for large $z_1$, and as $z_1=0.3\delta^+$,
+this implies that the second term in $c_\mathrm{log}$ only plays a role at low and intermediate $\RE$.
+Then, $a_\mathrm{log}$ can be estimated as 
+\begin{eqnarray}
+  a_\mathrm{log} \simeq v_0 -z_0 \left[ d_0 \log  z_0 - \frac{\Delta v}{\Delta z} ( 1- \log z_0) \right] 
+\end{eqnarray}
+for large $\RE$. 
+\end{subequations}
 

### 2021-10-27 plot/plot.py

@@ -316,7 +316,7 @@ if plot_ekman_ideal:
     ax2.set_xlabel(r'$U_\mathrm{ek}/G$')
     ax2.set_ylabel(r'$V_\mathrm{ek}/G$') 
     plt.savefig('ekman_ideal.pdf',format='pdf') 
-    
+    plt.close('all')
     
 if plot_evisc:
     fig=plt.figure(figsize=(12,4.5))
@@ -461,20 +461,109 @@ if plot_evisc:
     fig2.savefig('eddy_viscosity.pdf',format='pdf')
 
 if plot_profiles:
+    # calculate and plot a_log
 
+    c_log = []
+    c1=[]
+    c2=[]
+    re_arr=np.arange(2.5,10,0.05)
+    re_arr=np.power(10,re_arr) 
+    for re in re_arr:
+        us,al=sc.ustar_alpha(re)
+        dp=us*us*re*re/2 
+        ym=np.array([0.3])
+        [e_u,e_w]=sc.profile_ekman(ym,us)
+        [ous,ows]=mp.rotate(e_u,e_w,al)
+        w0=27
+        d0=4.01
+        w1=-ows[0]*us*dp
+        deltav=w1-w0
+        deltaz=0.3*dp-10
+        c_log.append((deltav - 10*d0*np.log(0.03*dp)) / deltaz)
+        c1.append(deltav/deltaz)
+        c2.append(10*d0*np.log(0.03*dp)/deltaz)
+        #blog=(d0-clog)*10
+        #a_log.append(w0-blog*np.log(10) - clog*10)
+        print(re,us,al,c_log[-1],c1[-1],c2[-1])
+        
+    c_log=np.array(c_log).flatten()
+    c1=np.array(c1).flatten()
+    c2=np.array(c2).flatten() 
+    fig=plt.figure(figsize=(5,4))
+
+    ax=fig.add_axes([0.13,0.13,0.85,0.85])
+    i0=mp.geti(re_arr,4e4)
+    i1=mp.geti(re_arr,4e6)
+    ax.plot(re_arr[i0:i1],c_log[i0:i1],ls='-',lw=4,c='black',alpha=0.5)
+    ax.plot(re_arr,c_log,ls='-',lw=2,label=r'$c_\mathrm{log}$',c='black')
+    ax.plot(re_arr,c1,ls='--',lw=1,label=r'$\Delta v/ \Delta $') 
[…]

### 2021-11-12 write1/manuscript.tex
minor changes on manuscript

@@
 %
-The organization of the flow is depicted in terms of the turbulence kinetic energy in Fig.~\ref{fig:slices}.
-%
-In vicinity of the wall, at $y^+\approx 10$, (Fig.~\ref{fig:slices}a), elongated streaks aligned with the
-surface shear stress dominate.
-%
-Moving away from the wall, to $y^+\approx 150$ (well within the logarithmic region), the structures
-are larger and more isotropic, but they are still largely aligned with the surface shear stress.
-%
-In the upper part of the outer layer, around $y^+\approx 1000$, no clear signature of the
-surface veering direction is found, and intense TKE structures (bright yellow)
-are organized on a large spatial scale with with weaker eddies (greenish structures) and quiescent regions
-in between.  
 \begin{figure}
@@
 \end{figure} 
+The organization of the flow is depicted in terms of the turbulence kinetic energy in Fig.~\ref{fig:slices}.
+%
+In vicinity of the wall, at $y^+\approx 10$, (Fig.~\ref{fig:slices}a), elongated streaks aligned with the
+surface shear stress dominate.
+%
+Moving away from the wall, to $y^+\approx 150$ (well within the logarithmic region), the structures
+are larger and more isotropic, but they are still largely aligned with the surface shear stress.
+%
+In the upper part of the outer layer, around $y^+\approx 1000$, no clear signature of the
+surface veering direction is found, and intense TKE structures (bright yellow)
+are organized on a large spatial scale with with weaker eddies (greenish structures) and quiescent regions
+in between.
 
@@
 %
+% 
+\subsection{Drag-Law}
+\label{sec:drag-law}
 \begin{figure}
@@
 \end{figure}
-% 
-\subsection{Drag-Law}
-\label{sec:drag-law}
 A drag-law for Ekman flow determines---as a function of Reynolds number alone---the surface drag.

### 2022-01-11 write1/manuscript.tex
further work on manuscript

@@
 \date{\scriptsize Draft - Version \today}
-\title{Profiles of Wind Direction and Speed in Turbulent Ekman Flow}
+\title{Wind veer and and Speed in Turbulent Ekman Flow}
 \author{Cedrick Ansorge, Hauke Wurps}
@@
 Not only does the veering set the frame of reference for surface layer theory, it also has
-effects at small and large scales.
+effects at small and large scales \todo{Insert: Brown, 2005; Svensson and Holtslag, 2009}.
 %
@@
 % 
-At the small-scale end, the directional shear of the wind in the upper part of the surface layer
-may cause a systmatic yaw for tall wind power generation devices where blades reach into the Ekman layer,
-i.e. that part of the boundary layer where the wind starts to turn. 
+Within the atmospheric boundary layer (ABL), directional shear of the wind in
+the upper part of the surface layer may cause a systmatic yaw for tall wind power generation devices
+where blades reach into the Ekman layer, i.e. that part of the boundary layer where the wind starts to turn;
+an exact estimate of such effects is critical in the site assessments for wind farms
+\todo{References: Calaf et al., 2010 Phys Fluids; Mirocha et al., 2018 Wind Energy Science} . 
 %
@@
 %
-Knowledge about $u_\star$ and $\alpha$ is thus a pre-requisite for any quantitative theory of the surface layer and a first
-attempt in theoretical quantification was undertaken by \citet{rossby:PPO1935} using integral relations in the boundary layer.
+Knowledge about $u_\star$ and $\alpha$ is thus a pre-requisite for any quantitative theory of the surface layer
+and a \citet{rossby:PPO1935} constrained the two parameters based on integral relations in the ABL.
 %
@@
 and numerical modelling \citep{spalart:PF2008, spalart:PF2009, ansorge:BM2014, ansorge:BM2019}.
+\todo{Refer to Pena et al, 2010} 
 %
@@
 % 
+More recently, \citet{ghannam:jas2021} treated the horizontally averaged momentum budget to show that departures from
+shear-alignment in the vicinity of the surface result in an integral of the wind veer ($\alpha_M$ in their notation)
+over the height to very high accuracy ($\int_{z_0}^{H} \sin\alpha_M$ in their notation; their Eq.~(16)).
+%
+Classic surface-layer similarity is recovered when the angle $\alpha_M$ does not depend on height, i.e., the
+wind veer is constant across the surface layer. 
+%
+If, however, the wind veer depends on height, the profiles of stress and mean velocities depart from the
+scalings implied by classic surface-layer similarity.
+%
+\par
+%
 Turbulent Ekman flow is considered here as a conceptual model of the
@@
 %
-For instance, effects of roughness are commonly incorporated by a linear transformation of vertical scale involving the
+More complex accounts generally refer to the homogeneous stationary problem as a base state: 
+%
+(i) Roughness is commonly incorporated by a linear transformation of vertical scale involving the
 roughness parameter $z_0$ and for larger roughness also a displacement height~%
 \citep{monin:1975,jacobs:AFM1988,hogstrom:BM1988};
-effects of stability are commonly accounted for by a linearization around the neutrally stratified profile~%
-\citep{monin:ARF1970, monin:1975, hogstrom:BM1988, hogstrom:BM1996,sakagami:BM2020}.
-%
-Furthermure, profiles can serve as better initial condition for numerical simulation of the flow,
+%
+(ii) Stability can be accounted for by a linearization around the neutrally stratified profile~%
+\citep{monin:ARF1970, monin:1975, hogstrom:BM1988, hogstrom:BM1996,sakagami:BM2020};
+%
+(iii) Non-stationarity in the pressure-gradient forcing can be accounted for by a linear damped-oscillator
+approach around the base state\citep{momen:jas2016}; 
+%
+(iv) Barotropic and baroclinic effects on the velocity profile require to consider 
+the height-dependence of the veer and stress misalignment \citep{momen:jas2018, ghannam:jas2021}.
+% 
+Furthermore, such a solution can serve as better initial condition for numerical simulation of the flow,
 to minimize the length of initial transient periods, or as benchmark for turbulence
@@
 Despite the strong simplifications implied by our choice of set-up, there is no straightforward approach to solving
-this conceptually well-defined problem.
+this well-defined problem.
 %
@@
 %
-In fully developed turbulence certain statistics such as dissipation \citep{dimotakis:ARF2005} or profiles of
-mean velocity \citep{barenblatt:JFM1993} become independent of the Reynolds number when scaled appropriately while
-other statistics, such as the near-wall maximum in velocity fluctuation depend on $\RE$ \citep{baars:JFM2020a} and external
-interfaces may exert an impact on near-wall scaling \citep{dasilva:ARF2014}. 
+Certain statistics of fully developed turbulence, such as dissipation \citep{dimotakis:ARF2005} or profiles of
+mean velocity \citep{barenblatt:JFM1993}, become independent of the Reynolds number when appropriately scaled; 
+other statistics, such as the near-wall maximum in velocity fluctuation depend on $\RE$ \citep{baars:JFM2020a} and externality
+of the flow may exert an impact on near-wall scaling \citep{dasilva:ARF2014}. 
 %
 It appears that for certain statistics in Ekman flow, fully-developed turbulence is reached with the
-Reynolds numbers that became possible with the increasing capabilities of supercomputering over
+Reynolds numbers that became possible due to an increase of computing capabilities over
 the past decades.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Problem Formulation and Numerical Approach}
+\section{Problem Form
+  ulation and Numerical Approach}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-01-11 plot/plot.py

@@ -17,13 +17,13 @@ D2R = np.pi / 180.
 
 use_data=False   # if true -- the netcdf files from DNS are needed 
 plot_ustar_alpha=False
-plot_ekman_ideal=True 
+plot_ekman_ideal=False 
 plot_summary=False
 plot_evisc=False
-plot_profiles=True
+plot_profiles=False
 plot_visc_outer=True
 print_table=False
-plot_outer_log=True
+plot_outer_log=False
 plot_total_rot=False
 plot_convergence=False
 plot_re1600=False
@@ -271,6 +271,9 @@ if plot_ekman_ideal:
     ax1=fig.add_axes([0.06,0.11,0.41,0.87])
     ax2=fig.add_axes([0.55,0.11,0.44,0.87])
 
+    fig2=plt.figure(figsize=(5,3.25))
+    ax3=fig2.add_axes([0.18,0.13,0.8,0.84]) 
+    
     z=np.arange(0,2*np.pi,0.02)
     
     sa_arr=[[+0.00, 0.4,'black', '-', 1.0],
@@ -295,27 +298,44 @@ if plot_ekman_ideal:
         
         ax1.plot(z[i03:],Uek[i03:],  c=c,ls=s,lw=w, label='a={1:4.1f},s={0:5.2f}'.format(sek,aek))
         ax1.plot(z[:i03],Uek[:i03],  c=c,ls=s,lw=w, alpha=0.5)
-        
+       
         ax2.plot(Uek[i03:],Vek[i03:], c=c,ls=s,lw=w,label='a={1:4.1f} s={0:5.2f}'.format(sek,aek))
         ax2.plot(Uek[:i03],Vek[:i03], c=c,ls=s,lw=w,alpha=0.5)
         ax2.scatter(Uek[i03],Vek[i03],c=c)
+
+        ax3.plot(Uek[i03:],Vek[i03:], c=c,ls=s,lw=w,label=r'A={1:4.1f} $z_r$={0:5.2f}'.format(sek,aek))
+        ax3.plot(Uek[:i03],Vek[:i03], c=c,ls=s,lw=w,alpha=0.5)
+        ax3.scatter(Uek[i03],Vek[i03],c=c)
+        
     lg=ax1.legend(loc='best')
+    lg2=ax3.legend(loc='best') 
 
[…]

### 2022-01-11 write1/manuscript.tex
inserted some references; more to come

@@
 % 
-More recently, \citet{ghannam:jas2021} treated the horizontally averaged momentum budget to show that departures from
+More recently, \citet{ghannam:QJR2021} treated the horizontally averaged momentum budget to show that departures from
 shear-alignment in the vicinity of the surface result in an integral of the wind veer ($\alpha_M$ in their notation)
@@
 (iii) Non-stationarity in the pressure-gradient forcing can be accounted for by a linear damped-oscillator
-approach around the base state\citep{momen:jas2016}; 
+approach around the base state\citep{momen:JAS2016}; 
 %
 (iv) Barotropic and baroclinic effects on the velocity profile require to consider 
-the height-dependence of the veer and stress misalignment \citep{momen:jas2018, ghannam:jas2021}.
+the height-dependence of the veer and stress misalignment \citep{momen:JAS2018, ghannam:QJR2021}.
 % 

### 2022-03-08 write1/manuscript.tex
some updates in the manuscript (Hauke's comments) and further polishing); first results for ReD1600_50 from hawk.hlrs

@@
 \date{\scriptsize Draft - Version \today}
-\title{Wind veer and and Speed in Turbulent Ekman Flow}
+\title{Profiles of Wind Veer and Wind Speed in Turbulent Ekman Flow}
 \author{Cedrick Ansorge, Hauke Wurps}
@@
 Not only does the veering set the frame of reference for surface layer theory, it also has
-effects at small and large scales \todo{Insert: Brown, 2005; Svensson and Holtslag, 2009}.
+effects at small and large scales from large-scale dispersion via plume spreading to
+cyclone spin-down \citep{svensson:BM2009}
+and on the capabilities of data assimilation and accuracy of surface flux estimates \citep{brown:QJR2005}. 
 %
@@
 an exact estimate of such effects is critical in the site assessments for wind farms
-\todo{References: Calaf et al., 2010 Phys Fluids; Mirocha et al., 2018 Wind Energy Science} . 
+\citep{calaf:PF2010, mirocha:WES2018}. 
 %
@@
 %
-Knowledge about $u_\star$ and $\alpha$ is thus a pre-requisite for any quantitative theory of the surface layer
-and a \citet{rossby:PPO1935} constrained the two parameters based on integral relations in the ABL.
+Knowledge about $u_\star$ and $\alpha$ is thus a pre-requisite for any quantitative theory of the surface layer, 
+and \citet{rossby:PPO1935} constrained the two parameters based on integral relations in the ABL.
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Problem Form
-  ulation and Numerical Approach}
+\section{Problem Formulation and Numerical Approach}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 such as rotating or non-rotating channel, pipe, or Couette flow.
+%
+Another common measure of scale separation is the Reynolds number defined by the laminar Ekman layer
+thickness $D=\sqrt{2\nu/f}$, 
+\begin{equation}
+  \RE_D = \frac{G D} {\nu } = \sqrt{2 \RE_\Lambda}. 
+\end{equation}
 %
@@
 problem, the estimation of higher-order effects such as the dependency of the Reynolds number requires
-a broader scale separation that is made available by the two new simulations for 
+a broader scale separation that is made available by the two new simulations at increased Reynolds number  
 (cf. Tab.~\ref{tab:sim-setup}). 
@@
 below $z^-\approx 0.15$ (circles in Fig.~\ref{fig:profiles}a); the case with $\RE_D=500$ is only transitionally
-turbulent and there is no distinct inner-oouter scale separation.
+turbulent and there is no distinct inner-outer scale separation.
 The logarithmic law is appropriate for $50 < z^+ < 0.15 Re_\tau$, where $z^+ = Re_\tau z^-$.
@@
 %
-Also in the outer layer of the flow, $u_\star$ (not the magnitude the
-geostrophic wind speed $G$) is the governing velocity scale for inviscid, i.e. $\RE$-independent,
+Also in the outer layer of the flow, $u_\star$ (and not the magnitude of the
+geostrophic wind $G$) governs the inviscid, i.e. $\RE$-independent,
 normalization.
@@
 \end{equation}
-which has indeed already been found by \cite{spalart:JFM1989}, who estimate the constant from
+which has indeed already been found by \cite{spalart:JFM1989}, who estimates the constant from
 an integral relation. 
@@
     \textbf{a-b)} Profiles of
-    the turbulent stress $S_\mathrm{turb}^+$ (solid), 
+    the turbulent stress $S_\mathrm{turb}^+$ (dashed), 
     the viscous stress  $S_\mathrm{visc}^+$ (dotted), 
-    and the total stress, $S^+=S_\mathrm{visc}^++S_\mathrm{turb}^+$ (dashed)
+    and the total stress, $S^+=S_\mathrm{visc}^++S_\mathrm{turb}^+$ (solid)
     as a function of inner height \textbf{(a)} and outer height \textbf{(b)}. 
@@
 \end{equation}
-exhibits universal scaling when considered as $S_\mathrm{visc}^+(z^+)$ (Fig.~\ref{fig:stresses}c); this normalization is also appropriate in the outer layer where the viscous stress
+exhibits universal scaling when considered as $S_\mathrm{visc}^+(z^+)$ (Fig.~\ref{fig:stresses}a); this normalization is also appropriate in the outer layer where the viscous stress
 is, however, small. 
@@
 Under this normalization, the profiles of eddy viscosity collapse below $z^+\approx 20$ with a tendency towards better collapse
-at higher $z^+$ (up to $z^+\approx 50$ for $\RE=1600$; Fig.~\ref{fig:stresses}b).
+at higher $z^+$ for higher Reynolds number (up to $z^+\approx 50$ for $\RE=1600$; Fig.~\ref{fig:stresses}c).
 %
@@
 %
-This can be formulated in terms of the modulus
-of the surface friction normalized by the geostrophic wind forcing, $u_\star$ (Eq.~(\ref{eqn:ustar_def}))
-and, the direction of surface shear stress, $\alpha_\star$ (Eq.~(\ref{eqn:alpha_def})),
-also termed wind veering. 
-%
-A non-zero veering of the wind is a rather special case in comparison with most turbulent flows conisdered in an
+This can be formulated by the normalized surface friction, $u_\star$ (Eq.~(\ref{eqn:ustar_def}), also termed geostrophic drag), 
+and the direction of surface shear stress, $\alpha_\star$ (Eq.~(\ref{eqn:alpha_def}), also termed wind veer). 
+%
+A non-zero veering of the wind is a rather special case in comparison with most turbulent flows considered in an
 engineering context, and it confronts us with a situation where the most appropriate coordinate system
@@
 %
-We compare our DNS data against a semi-empirical drag-law based on integral consideration \cite{spalart:JFM1989}
-and find, as demonstrated in previous work \cite{ansorge:BM2014}, excellent agreement in the range
+We compare our DNS data against a semi-empirical drag-law based on integral consideration \citep{spalart:JFM1989}
+and find, as demonstrated in previous work \citep{ansorge:BM2014}, excellent agreement in the range
  $400<\mathbf{Re}<1600$, representing a factor of 16 in variation of viscosity. 
@@
 \begin{equation}
-  u_\star= 4\log(\mathbf{Re}_D))-8
+  Z_\star= 4\log(\mathbf{Re}_D))-8
 \end{equation}
@@
 %
-While this is not the case here for the general form of the profile, we
+While this is not the case here for the general form of the profile, the
 phase shifts can also be captured by the parameter $s_\mathrm{ek}$, 
@@
 A pure blending from the linear velocity profile into the logarithmic one is, however, not reasonable as both the linear and logarithmic
-profile overestimate the velocity in the buffer layer. We therefore introduce a higher-order compensating term into
-the linear profile before blending it into the logarithmic law:
-\begin{equation}
-  U_\text{inner}^{\alpha_\star}^+ = f_{U,\mathrm{visc}}(z^+) = \frac{ z^+ + \gamma_4 (z^+)^{4} + \gamma_6 (z^+)^{6} } { 1 + \gamma_6 /u_{ref} (z^+ )^6}, 
-\end{equation} 
-where the fourth- and sixth-order terms in the nominator are designed to fit the profile below $y^+\approx 10$ and
-the sixth-order correction in the denominator limits the growth of $u_\text{inner}$ for large $y^+$.
-%
-These correction terms are small for $z^+<5$ where the linear law of the wall
-(shown in the inset in Fig.~\ref{fig:u_profiles}) yields an excellent fit of the data; we use here
+profile overestimate the velocity in the buffer layer.
+%
+We therefore introduce a two-step correction procedure, accounting for the smaller-than linear growth beyond $y^+\approx 5$, and assuring smooth matching with the logarithmic law at $y^+=40$:
+\begin{equation}
+  U_\text{inner}^{\alpha_\star}^+ = \frac{z^+}{1+c_1 (y^+)^2} +  (c_2 y^+ - a_\mathrm{match} )\frac{1+tanh\left[0.2(y^+-22)\right]}{2} + c_3 e^{-c_4(y^+-22)^2}. \label{eqn:u_inner} 
+\end{equation} 
+%
+We use here
 \begin{equation*}
-    \gamma_4 = -3.825 \times 10^{-4};\quad
-    \gamma_6 &=&  6.320 \times 10^{-6};\quad 
-    u_\mathrm{ref} &=& 1.278 \times 10^{1}.
+  c_1 &=& 0.00185;\quad
+  c_2 &=& 0.195;\quad 
+  c_3 &=& 0.4 ; \quad
+  c_4 &=& 0.35. 
 \end{equation*}
-
+%
+The second and third terms on the right hand side vanish for $y^+\ll 22$, and $c_1=0.00185$
+implies an approximately 5\% correction at $y^+=5$ and an 18.5\% correction at $y^+=10$.
+%
+The second and third term on the R.H.S. of eq.~(\ref{eqn:u_inner}) are an empirical fit to
+the velocity profiles observed in the buffer layer and appear independent of the Reynolds number
+for the range observed here.
+%
+The coefficient $a_\mathrm{match}$, which has no effect in the viscous sublayer, is then used to
+match this formulation to the logarithmic law employed above. 
+%
 \par
@@
 with the von-K\'arm\'an constant $\kappa=0.416$ and the boundary condition $C=5.4605$. %  log-layer (K\'arm\'an, Prandtl, Jimenez, Marusic)
+For this logarithmic law, $a_\mathrm{match}=3.569861$ for a matching at $y^+=40$. 
 %
@@
 In this context, we first realize that the velocity deficit $(V^{\alpha_\star}-G^{\alpha_\star})/u_\star$
-approaches a $\RE$-independent constant around $C_{V0}=Z_\star \sin\alpha -6.1$ at the surface; deviations from
+approaches a $\RE$-independent constant around $C_{V0}=Z_\star \sin\alpha=6.1$ at the surface; deviations from
 this constant are only found for the lowest Reynolds numbers which is in accordance with
@@
 %
-\subsubsection{Discussion}
+\section{Discussion}
+\subsection{} 
 Eq.~(\ref{eqn:scaling_spanwise}) establishes a universal mixed scaling for the spanwise velocity

### 2022-03-08 plot/EkmanProfileClass.py

@@ -235,13 +235,34 @@ class EkmanUniversalClass:
         # STREAMWISE COMPONENT (shear-aligned) 
         log_law=self.profile_log(yp) # np.log(yp)/self.KAPPA + self.C   
 
-        # inner-layer (empirical profile for viscous and buffer layer) 
-        c4=-0.0003825; c6=6.32e-6; u_ref = 0.07825
-        yp_mtc = 19 
-        wgt_i = self.erf_transition(trans_scale,yp_mtc,yp[1:]) 
-        u_visc=( yp + c4*yp**4+c6*yp**6) / (1 + u_ref*c6*yp**6)
-        u[1:]=(1-wgt_i)*u_visc[1:] + wgt_i*log_law[1:]
+        # inner-layer (empirical profile for viscous and buffer layer)
+        # OLD - sixth order viscous term + blending 
+        #c4=-0.0003825; c6=6.32e-6; u_ref = 0.07825
+        #yp_mtc = 19 
+        #wgt_i = self.erf_transition(trans_scale,yp_mtc,yp[1:]) 
+        #u_visc=( yp + c4*yp**4+c6*yp**6) / (1 + u_ref*c6*yp**6)
+        #u[1:]=(1-wgt_i)*u_visc[1:] + wgt_i*log_law[1:]
 
+        # NEW - explicit viscous layer that matches the log formulation
+        delta=1; a_match=3.569861; iter=1
+        i_mtc=mp.geti(yp,40)
+        while np.abs(delta) > 1e-12 and iter<10:
+            u_visc=yp/(1+0.00185*yp**2)
+            u_visc=u_visc + (0.195*yp-a_match)*(1.+np.tanh(0.2*(yp-22)))/2.
+            u_visc=u_visc +  0.40*np.exp(-0.035*(yp-22)**2)
+            
+            delta = u_visc[i_mtc] - log_law[i_mtc]
+            a_match=a_match+delta
+            iter=iter+1
+
+        if iter >= 9:
+            print("ERROR: CONVERGENCE BETWEEN LOG AND VISCOUS LAYER COULD NOT BE ACHIEVED")
+            print("ERROR: delta:", delta )
+            quit() 
+
+        u[1:i_mtc] = u_visc[1:i_mtc]
+        u[i_mtc:]  = log_law[i_mtc:] 
+            
         # outer-layer deficit
         ctr=0.28- 2.25*np.sqrt(1./re)
[…]

### 2022-03-08 write2/note.tex
some updates in the manuscript (Hauke's comments) and further polishing); first results for ReD1600_50 from hawk.hlrs

\documentclass[a4paper,1pt]{amsart}
\usepackage{amssymb}
\usepackage{amsmath}
\newcommand{\RE}{\mathbf{Re}}

\date{\today}
\title{Note on Boundary condition and choice of grid} 
\begin{document}
\noindent By choice of the grid, it is for the first grid point ($z_1$)
\[ z_1 = c_\delta \delta \Rightarrow z_1^{-}=c_\delta.\]
The height argument of the log-law is, however, an inner height, and we need to use $z^+$.
Using the relation $z^-=z^+/\RE_\tau$, we get 
\[ z_{1}^{+} = c_\delta \RE_\tau\] 
for the height of the first grid node in inner units.
%
This is a pure consequence of the construction of the grid under changing $\RE$ (namely using a constant number of grid points per $\delta$.
%
A similar relation holds for the rough equivalent:
\[ \frac{z_1}{z_0} = \frac{c_\delta \delta}{z_0} = c_\delta \frac{u_\star}{f} \frac{u_\star}{c_\nu \nu} = \frac{c_\delta}{c_\nu} \RE_\tau.\]
This underlines the role of $c_\nu$ in linking the classic inner ('+') scale and the rough inner unit ($z_0$), it is namely
\[ \frac{z^+}{z/z_0} = c_\nu.\]
First, this shows that--apart from iny variation in $c_\nu$--the scaling with $z_0$ is the classic inner ('+') scaling.
%
Second, the small variation of $c_\nu$ (estimated/tuned against the DNS results[!!!])
across the Reynolds number (less than the alread weak variation in $u_\star$
and orders of magnitudes less than the variation in $Re_\tau$) shows that the analogy between aerodynamically rough
and smooth flows holds quite nice, but not exact.
%
In that sense, the collapse of profile for the vertical scale $z/z_0$ is to be understood as an 'approximate' inner scaling. 
%
It remains to be seen whether the small variation of $c_\nu$ with $\RE$ is due to technical details of the LES formulation under changing
scale separation and different action of the SGS scheme, or whether these are an actual physical property of the flow.
%
It may also be that by changing $c_\nu$ we capture the inadverted effect of $z_1$ coming too close to the surface for
some configurations. 
%
\par
%
With regards to the velocity boundary condition, we now see that
\[U_1=U(z_1) = \frac{u_\star}{\kappa} \ln \frac{z_1}{z_0}  = \frac{u_\star}{\delta} \ln\RE_\tau - \ln\frac{c_\delta}{c_\nu}\]
where the first term ($\ln\RE_\tau$) captures the dependency of the grid on the Reynolds number and the second term reflects the tuning parameter
used to get a correct $u_\star$ in the LES.
%
\par
%

\end{document}

### 2022-03-08 plot/plot.py

@@ -18,10 +18,10 @@ D2R = np.pi / 180.
 use_data=False   # if true -- the netcdf files from DNS are needed 
 plot_ustar_alpha=False
 plot_ekman_ideal=False 
-plot_summary=False
+plot_summary=True
 plot_evisc=False
 plot_profiles=False
-plot_visc_outer=True
+plot_visc_outer=False
 print_table=False
 plot_outer_log=False
 plot_total_rot=False
@@ -29,6 +29,7 @@ plot_convergence=False
 plot_re1600=False
 plot_les_comp=False
 plot_applications=False
+test_inner_streamwise=False
 
 colors = {400 : 'gray',
           500 : 'pink',
@@ -1077,8 +1078,11 @@ if plot_summary :
         up_loc,wp_loc=sc.profile_plus(yp_loc,re_loc)
         
         sqdp = 1.#/(deltap_loc)
-        
-        alpha_loc= re_loc/1601 % 1 
+
+        if re_loc < 1601: 
+            alpha_loc= re_loc/1601 % 1
+        else :
+            alpha_loc = 1. 
         
         match_i=0.10
         match_o=0.40
@@ -1095,7 +1099,7 @@ if plot_summary :
             al_geo = np.arctan(w_dat[-1]/u_dat[-1])
             al_sfc1 = f['FrictionAngle'][t_use]*D2R
             al_sfc2 = np.arctan(w_dat[1]/u_dat[1]) 
-            
+
             [u_geo,w_geo] = mp.rotate(u_dat,w_dat,al_geo)
             [u_sfc,w_sfc] = mp.rotate(u_dat,w_dat,al_sfc1) 
             yp_dat = y_dat*us_dat/nu_loc
@@ -1107,11 +1111,15 @@ if plot_summary :
             i_i=mp.geti(yp_dat,match_i*deltap_loc)
             ax2.scatter(yp_dat[i_i]*sqdp,(u_sfc[i_i])/us_dat,c='black')
 
+        print('RE:',  re_loc, 'u*:', us_loc, 'transparency:',alpha_loc) 
[…]

### 2023-05-22 write1/manuscript.tex
Data and modifications from 2022

@@
 \usepackage{xcolor}
-\usepackage{dirtytalk}
 \usepackage{colortbl}
@@
 \newcommand{\todo}[1]{\textcolor{red}{$[$#1$]$}}
-\newcommand{\quotation}[1]{"#1"}
+\renewcommand{\quotation}[1]{"#1"}
 
@@
 A one-dimensional profile with constant veering is given by \citet[Sec. 3; Eq. 3.1-3.19]{emeis:m2007}. 
-
+%%%
+%%% ADD REFERENCE TO Klein and Maier (PAMM, 2021) for failure of matching of profile in 1D ODT-LES
+%%% ADD REFERENCE TO Lindevall and Svensson (QJRMS, 2019, vol 145)
+%%%
 \citet{gryning:BM2007} present an extension of the wind-speed profile beyond the surface layer
@@
 Ekman-layer models and two-layer models of the ABL take into account rotational effects at higher altitudes,
-for instance when the wind speed needs to be evaluated at heights on the order of 100-200\,\mathrm{m},
+for instance when the wind speed needs to be evaluated at heights on the order of $100-200\,\mathrm{m}$,
 a particular concern when it comes to wind-power forecasting \citep{optis:BM2014}.
@@
 Here, the routinely employed concept of Reynolds-number similarity can help. It postulates the existence
-of \say{\emph{fully developed turbulence}  believed to occur for a sufficiently large but finite Reynolds number}
+of \quotation{\emph{fully developed turbulence}  believed to occur for a sufficiently large but finite Reynolds number}
 \citep{barenblatt:PF1995}.
@@
 \begin{equation}
-  S_\mathrm{visc}= \nu \sqrt\left(\frac{\p U}{\p z}\right)^2  + \left(\frac{\p V}{\p z}\right)^2}
+  S_\mathrm{visc}= \nu \sqrt{\left(\frac{\p U}{\p z}\right)^2  + \left(\frac{\p V}{\p z}\right)^2}
 \end{equation}
@@
 \begin{equation}
-  U_\text{inner}^{\alpha_\star}^+ = \frac{z^+}{1+c_1 (y^+)^2} +  (c_2 y^+ - a_\mathrm{match} )\frac{1+tanh\left[0.2(y^+-22)\right]}{2} + c_3 e^{-c_4(y^+-22)^2}. \label{eqn:u_inner} 
+  U_\text{inner}^{{\alpha_\star}^+} = \frac{z^+}{1+c_1 (y^+)^2} +  (c_2 y^+ - a_\mathrm{match} )\frac{1+tanh\left[0.2(y^+-22)\right]}{2} + c_3 e^{-c_4(y^+-22)^2}. \label{eqn:u_inner} 
 \end{equation} 
@@
 \begin{equation*}
-  c_1 &=& 0.00185;\quad
-  c_2 &=& 0.195;\quad 
-  c_3 &=& 0.4 ; \quad
-  c_4 &=& 0.35. 
+  c_1 = 0.00185;\quad
+  c_2 = 0.195;\quad 
+  c_3 = 0.4 ; \quad
+  c_4 = 0.35. 
 \end{equation*}
@@
 \begin{equation}
-  U_\text{log}^{\alpha_\star}^+ = \frac{1}{\kappa} \log y^+ + C 
+  {U_\text{log}^{\alpha_\star}}^+ = \frac{1}{\kappa} \log y^+ + C 
 \end{equation}
@@
   \lim_{z\rightarrow\infty} V^{\alpha_\star} &=& G \sin\alpha_\star \\ 
-  \left.\partial_{z^+} V^{\alpha_\star+}}\right|_{z=0} &=& 0.  
+  \left.\partial_{z^+} V^{\alpha_\star+}\right|_{z=0} &=& 0.  
 \end{eqnarray} 
@@
 %
-\end{subequations}
-
 \begin{figure}
@@
 \textbf{Compare} to Emeis (2002) and Gryning (2007); highlight explicit knowledge on veering-profile $\rightarrow$ directional sheer;
+
+Consider Townsend ``Turbulent Shear Flow''  Chapter 7.18: p. 319 
 %
@@
   \begin{matrix} 
-    u(z_\text{match}) \equiv u_\text{match} &=& U_{\infty} + e^{-\gzm}} \left [ A \cos\gzm + B\sin\gzm \right] \\
-    w(z_\text{match}) \equiv w_\text{match} &=& W_{\infty} + e^{-\gzm}} \left [-A \sin\gzm + B\cos\gzm \right]
+    u(z_\text{match}) \equiv u_\text{match} &=& U_{\infty} + e^{-\gzm} \left [ A \cos\gzm + B\sin\gzm \right] \\
+    w(z_\text{match}) \equiv w_\text{match} &=& W_{\infty} + e^{-\gzm} \left [-A \sin\gzm + B\cos\gzm \right]
   \end{matrix} 
@@
 \end{subequations}
-Matching the profiles and gradient $z_0=10^+$ and the value at $z_1=0.3\dlta^+$ yields
+Matching the profiles and gradient $z_0=10^+$ and the value at $z_1=0.3\delta^+$ yields
 \begin{subequations} 
@@
   \label{eqn:w_profile_inner}
-  \end{array} \right, 
+  \end{array} \right. , 
 \end{align}
@@
 \end{align}
-with $\Delta V &=& V_\text{outer}(z^-_\text{up})-V_\text{inner}(z^+_\text{low})$
+with $\Delta V = V_\text{outer}(z^-_\text{up})-V_\text{inner}(z^+_\text{low})$
 and  $z_{arg} = (z - z_\text{low})/(z_\text{up}-z_\text{low})$.
@@
 
+\end{subequations}
+
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2023-05-22 palm_hawk/plot_profiles.py

@@ -5,22 +5,53 @@ from netCDF4 import Dataset
 import matplotlib.pyplot as plt
 import glob
 
-files = glob.glob('ReD1600_*_pr.*.nc')
-files = sorted(files, key=lambda x: x.split('.')[1])
-files = sorted(files, key=lambda x: x.split('_')[1])
-
+files_pr = glob.glob('ReD1600_*_pr.all.nc')
+files_pr = sorted(files_pr, key=lambda x: x.split('.')[1])
+files_pr = sorted(files_pr, key=lambda x: x.split('_')[1])
 
 fig=plt.figure(figsize=(6,5))
+fig2=plt.figure(figsize=(6,5))
 ax=fig.add_axes([0.1,0.1,0.85,0.85])
+ax2=fig2.add_axes([0.1,0.1,0.85,0.85]) 
 
-for f in files:
-    f_nc=Dataset(f,'r') 
-    u = f_nc['u'][-1,:]
-    z_u=f_nc['zu'][:]
-    ax.plot(u[1:],z_u[1:],'o-',label=f)
+for f in files_pr:
+    f_pr=Dataset(f,'r')
+    res=int(f.split('_')[1] )
+    red=int(f.split('D')[1].split('_')[0] )
+    f_ts_name=('ReD{0}_{1:03d}_ts.all.nc'.format(red,res))
+    print(f_ts_name)
+    f_ts=Dataset(f_ts_name,'r') 
 
-    lg=plt.legend(loc='best') 
-ax.set_yscale('log')
-ax.set_ylim(0.01,2e2)
-plt.show() 
+    nu = 2./(red*red)
+    u = f_pr['u'][-1,:]
+    z_u=f_pr['zu'][:]
+    v = f_pr['v'][-1,:]
+    z_v=f_pr['zv'][:]
+    ww=f_pr['w*2'][-1,:] 
+    us_palm=f_ts['us*'][-1] 
+    us =np.sqrt(nu*np.sqrt(u[1]**2+v[1]**2)/z_u[1])
+    zp = z_u*us/nu 
+    G = np.sqrt(u[-1]**2 + v[-1]**2)
+
+    ax.plot(zp[1:],u[1:]/us_palm,'-',label=f,c='black',alpha=np.sqrt(res/500))
+    ax.plot(zp[1:],-v[1:]/us_palm,'-',c='red',alpha=np.sqrt(res/500))
+
[…]

### 2023-05-22 plot/my_pylib.py

@@ -1,1 +1,144 @@
-/Users/zrick/CODE/PYTHON/my_pylib.py
+#!/opt/local/bin/python 
+
+from numpy import zeros
+
+def read_fortran_record(f_h,t): 
+   t_i4=np.dtype('<i4') 
+   dum1,=np.fromfile(f_h,t_i4,count=1)  
+   n=int(dum1/t.itemsize) 
+   rec=np.fromfile(f_h,t,count=n) 
+   dum2=np.fromfile(f_h,t_i4,count=1) 
+   
+   if dum1 == dum2 : 
+      return [n,rec]
+   else :  
+      print('ERROR READING RECORD',dum1,dum2) 
+      quit() 
+
+
+class DnsGrid : 
+   def __init__(self,fname): 
+      f_h=open(fname,'rb') 
+      t_i4=np.dtype('<i4') 
+      t_f8=np.dtype('<f8') 
+
+      [n_isize,rec_isize]=read_fortran_record(f_h,t_i4) 
+      [n_fsize,rec_fsize]=read_fortran_record(f_h,t_f8) 
+
+      self.nx=rec_isize[0];  self.lx=rec_fsize[0]
+      self.ny=rec_isize[1];  self.ly=rec_fsize[1]
+      self.nz=rec_isize[2];  self.lz=rec_fsize[2] 
+
+      [nx_test,self.x]=read_fortran_record(f_h,t_f8)  
+      [ny_test,self.y]=read_fortran_record(f_h,t_f8) 
+      [nz_test,self.z]=read_fortran_record(f_h,t_f8) 
+
+      if not ( nx_test == self.nx and ny_test == self.ny and nz_test == self.nz ): 
+         print('Grid Dimensions do not match') 
+         quit() 
+
+def cumsum(x,y,n) : 
+   res=np.zeros(n)
+   for i in range(n-1):   
+      res[i+1] = res[i] + y[i]*(x[i]-x[i-1]) 
+      
+   return res; 
+
+def rotate(a,b,alpha) : 
+   from math import sin,cos 
+   c= cos(alpha)*a + sin(alpha)*b 
+   d=-sin(alpha)*a + cos(alpha)*b
+
+   return [c,d] 
+
+def rotate_tensor(a,alpha) :
+   from math import sin,cos  
+   #check shape 
[…]

### 2024-06-27 etling_model.tex
update on part 2

\documentclass[a4paper,10pt]{amsart}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}

\author{Cedrick Ansorge}
\date{\today}
\title{Notes on Etling velocity profile for Ekman/Prandtl layer} 
\begin{document}
\maketitle

\newcommand{\eqnref}[1]{Eq.~(\ref{eqn:#1})}
\section{Basic velocity profiles} 
\subsection{Velocity profile in the Prandtl layer} 
\begin{align}
  u_p(z) = \frac{u_\star}{\kappa} log \frac{z^+}{z_{0}^{+}}
\end{align}
In coordinates of the geostrophic wind (for a veering angle of $\alpha_0=\alpha(z=0)$
\begin{align}
  u_p(z) =\frac{u_\star}{\kappa } \log\frac{z^+}{z_{0}^{+}}\begin{pmatrix}\cos\alpha_0\\\sin\alpha_0\end{pmatrix} 
\end{align}

\subsection{Velocity profile in the Ekman layer}
\begin{align}
  \begin{pmatrix} u_\text{ek} \\ v_\text{ek}\end{pmatrix}(z) = G\sqrt{2}\sin\alpha_0
  \begin{pmatrix}\frac{1}{\sqrt{2}\sin\alpha_0}- &e^{-\tilde{z}} \cos (\tilde{z} +\frac{\pi}{4} - \alpha_0) \\ 
                                                 &e^{-\tilde{z}} \sin (\tilde{z} +\frac{\pi}{4} - \alpha_0)
  \end{pmatrix} 
\end{align}
with the scaled Ekman height
\[\tilde{z} = \frac{z-z_P}{D}, \]
where $z_P$ is the Prandtl layer thickness (here assumed as 10\% of the BL thicknes)
and $D$ the Ekman layer thickness. For D, there exists a matching based on the constant diffusivity within the Prandtl layer 
\begin{align}
  D = \sqrt{2\frac{K_m}{f}}  = \sqrt{2\frac{\kappa u_\star z_p}{f}} = \sqrt{2\kappa\delta z_p}  
\end{align}
with $\delta=u_\star/f$. Using $z_p=0.1\delta$, we obtain
\[D=\sqrt{0.2\kappa}\delta\approx 0.3\delta\]

\section{Matching point}
The profiles shall be matched at the upper end of the Prandtl layer ($z=z_P$). Given the geostrophic
wind, Prandtl layer height $z_p$ and height offset (in the case of a rough surface, the roughness height; for aerodynamically smooth surfaces a scaled version of the profile offset parameter), the profiles can be matched to yields the remaining parameters $u_\star$ and $\alpha_0$.

\begin{subequations} 
\begin{align}
  \frac{u_\star}{\kappa} \log \frac{z_{P}}{z_{0}} \cos\alpha_0 &=& G\sqrt{2}\sin\alpha_0 \left[\frac{1}{\sqrt{2}\sin\alpha_0} - e^{-\tilde{z}} \cos\phi \right]  \\
  \frac{u_\star}{\kappa} \log \frac{z_{P}}{z_{0}} \sin\alpha_0 &=& G \sqrt{2} \sin\alpha_0 e^{-\tilde{z}}\sin\phi  
\end{align}
\end{subequations}
where -- utilizing the condition $z=z_P$, it is  $\phi=\pi/4-\alpha_0$. We can then eliminate $\sin\alpha_0$ from the equation for the span-wise component to obtain
\begin{subequations} 
  \begin{align}
    \cos\alpha_0 &=& \frac{G}{u_\star}\sqrt{2}\lambda_z&\sin\alpha_0&\left[\frac{1}{\sqrt{2}\sin\alpha_0} - \cos\phi \right]  \label{eqn:alpha_estimator}\\
    \frac{u_\star}{G}   &=& \sqrt{2}\lambda_z &&\sin\phi \label{eqn:ustar_estimator}
  \end{align}
\end{subequations}
where
\[\lambda_z = \frac{\kappa}{\log\frac{z_P}{z_0}}.\]
Next, we rewrite \eqnref{alpha_estimator} using \eqnref{ustar_estimator} as
\begin{subequations} 
\begin{align}
  \cos\alpha_0 = \sqrt{2}\lambda_z\frac{G}{u_\star} \left(\frac{1}{\sqrt{2}\sin\phi}- \frac{\sin\alpha_0\cos\phi}{\sin\phi}\right) \\
  \sin\phi\cos\alpha_0 + \cos\phi\sin\alpha_0=\frac{1}{\sqrt{2}}
\end{align}
The LHS of the latter equation can be rewritten as
\[\sin(\phi+\alpha_0) = \sin(\pi/4)\] 
such that the equation holds for any $\alpha_0$. Thus, $\alpha_0$ becomes a parameter of the solution, and we can simply write
\begin{align}
  \frac{u_\star}{G} = \frac{\sqrt{2}\kappa}{\log(z_P/z_0)} \sin\left(\frac{\pi}{4}-\alpha_0\right) 
\end{align} 
\end{subequations}

\subsection{Relation to the Reynolds number}
In a smooth flow, the scaling parameter of the logarithmic height is
\[\frac{1}{\kappa}\log z^+ + A =\frac{1}{\kappa}\left[ \log z^+ - ( -\kappa A)\right]  = \frac{1}{\kappa}\left[\log z^+ - \log(e^{-\kappa A})\right] = \log(z^+/z_{0}^+) \]
where $z_{0}^+ = e^{-\kappa A}\approx 1/8$.
This implies that $z_P/z_0=z_{P}^{+}/z_{0}^+ = e^{\kappa A} z_{P}^{+} = e^{\kappa A} \gamma_P Re_\tau$. Hence, the ratio $z_P/z_0$ implicitly defines the Reynolds number $Re_\tau$ as
\begin{align}
  Re_\tau = \frac{z_P}{z_0} \frac{1}{e^{\kappa A} \gamma_P} \sim Re_\tau.
\end{align}

\subsection{Linearization for $\alpha_0$ in the trigonometric terms} 
Based on a first guess for $u_\star$ and $\alpha_0$, we can then estimate $u_\star/G$  from \eqnref{ustar_estimator} and $\alpha_0$ from \eqnref{alpha_estimator} and iterate this procedure
to arrive at a solution for $\alpha_0$ and $u_\star$.

A fist guess may be obtained from a linearized version of the equations in the trigonometric terms.
It is 
\begin{subequations}
  \begin{align} 
    \sin\phi &=& \sin(\pi/4-\alpha_0) \approx \sin(\pi/4) - \cos(\pi/4) \alpha_0 &=& \frac{1}{\sqrt{2}} (1-\alpha_0) \\
    \cos\phi &=& \cos(\pi/4-\alpha_0) \approx \cos(\pi/4) + \sin(\pi/4) \alpha_0 &=& \frac{1}{\sqrt{2}} (1+\alpha_0) \\ 
    \sin\alpha_0 &\approx& \sin(0) + \alpha_0 &=& \alpha_0.  
  \end{align}
\end{subequations}
We thus obtain
\begin{align}
  -1 &=& \frac{G\lambda_z}{u_\star}
         \left[\frac{1}{\alpha_0} -(1-\alpha_0)  \right] \\ 
  \Rightarrow 0 &=& \alpha_0 + \frac{G\lambda_z}{u_\star} \left[ 1- (\alpha_0-\alpha_0^{2})\right] 
\end{align} 
\end{document}

### 2024-06-27 write1/manuscript.tex
update on part 2

--- 2022-03-08 write1/manuscript.tex
@@
-\documentclass[a4paper,11pt]{amsart}
-
-\usepackage[margin=3.5cm]{geometry}
+\RequirePackage{fix-cm} 
+\documentclass[smallcondensed,final]{svjour3}
+\smartqed
+\usepackage{appendix}
+\usepackage{amsmath}
+\usepackage{graphicx}
 \usepackage{lineno}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage[footnotesize]{caption}
+\usepackage{array}
+\usepackage{longtable}
+\usepackage{natbib}
+
+% additional packages
 \usepackage{xcolor}
-\usepackage{dirtytalk}
 \usepackage{colortbl}
 \usepackage{booktabs} 
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% BIBLIOGRAPHY DETAILS
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\usepackage[maxcitenames=2,style=authoryear-comp,firstinits=true,citestyle=authoryear,natbib=true,backend=bibtex]{biblatex}
-\DeclareNameAlias{sortname}{family-given}
-\renewcommand*{\multinamedelim}{\addcomma\space}
-\renewcommand*{\finalnamedelim}{~and\space}
-\setlength{\bibitemsep}{\baselineskip}
-\setlength\parindent{3em}
-\setlength\parskip{0.5em}
-\renewbibmacro{in:}{}
-
-\AtEveryBibitem{%
-  \clearfield{url}
-  \clearfield{urlday}
-  \clearfield{urlmonth}
-  \clearfield{urlyear}
-  \clearfield{issn}
-  \clearfield{month}
-%%%  \clearfield{doi}%
-  \clearfield{keywords}
-  \clearfield{note}
+\usepackage{amssymb}
+\usepackage{hyperref}
+\setcitestyle{aysep={}}
+\linenumbers
+
+\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
+\expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
+\expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
+\renewenvironment{#1}%
+{\linenomath\csname old#1\endcsname}%
+{\csname oldend#1\endcsname\endlinenomath}}% 
+\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
+\patchAmsMathEnvironmentForLineno{#1}%
+\patchAmsMathEnvironmentForLineno{#1*}}%
+\AtBeginDocument{%
+\patchBothAmsMathEnvironmentsForLineno{equation}%
+\patchBothAmsMathEnvironmentsForLineno{align}%
+\patchBothAmsMathEnvironmentsForLineno{flalign}%
+\patchBothAmsMathEnvironmentsForLineno{alignat}%
+\patchBothAmsMathEnvironmentsForLineno{gather}%
+\patchBothAmsMathEnvironmentsForLineno{multline}%
 }
 
-\bibliography{ansorge}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand{\LR}{\Lambda_\RO} 
+\newcommand{\DOI}[1]{\href{https://dx.doi.org/#1}{#1}}
+
+\hypersetup{
+  colorlinks=true,
+  urlcolor=blue,
+  citecolor=.,
+  linkcolor=.
+}
 
@@
 
-Wind veer away from its geostrophic direction is
+The Coriolis force bends the path of motion on a rotating sphere and establishes geostrophic equilibrium when in balance with a pressure gradient force.
+Wind veer away from the wind direction in geostrophic equilibrium is
 (i) due to direct frictional effects in the very vicinity of the surface and
@@
 % %% This shear is probably not related to Ekman-turning 
-% On the mesoscale, shear in the surface layer a crucial factor for the initiation and sustainment of strong convective events (REFERENCE?).
+% On the meso-scale, shear in the surface layer a crucial factor for the initiation and sustained strong convective events (REFERENCE?).
 % 
 Within the atmospheric boundary layer (ABL), directional shear of the wind in
-the upper part of the surface layer may cause a systmatic yaw for tall wind power generation devices
+the upper part of the surface layer may cause a systematic yaw for tall wind power generation devices
 where blades reach into the Ekman layer, i.e. that part of the boundary layer where the wind starts to turn;
@@
 %
-Surface veering $\alpha$ and geostrophic drag $Z\equiv u_\star/G$, where $u_\star\equiv\sqrt{\nu|\tau_\SFC|}$,
+Surface veering $\alpha$ and geostrophic drag $Z\equiv u_\star/G$, where the friction velocity $u_\star\equiv\sqrt{|\tau_\SFC|/\rho}$,
 uniquely determine the surface drag $\tau_\SFC$ in a turbulent Ekman flow. 
@@
 %
-Knowledge about $u_\star$ and $\alpha$ is thus a pre-requisite for any quantitative theory of the surface layer, 
+Knowledge about $u_\star$ and $\alpha$ is thus a prerequisite for any quantitative theory of the surface layer, 
 and \citet{rossby:PPO1935} constrained the two parameters based on integral relations in the ABL.
 %
-Asymptotic similarity theory was later used by \citet{tennekes:JAS1973, blackadar:JAS1968},
+Asymptotic similarity theory was later used by \cite{tennekes:JAS1973, blackadar:JAS1968},
 and--based on his seminal direct numerical simulations (DNS) of Ekman flow--,~\cite{spalart:JFM1989} suggested
 a modification to take into account effects of low to intermediate Reynolds numbers.
-Later on, constants were re-evaluated with a focus on the atmospheric boundary layer based on
+Later on, constants were re-evaluated with a focus on the ABL based on
 observations \citep{hogstrom:BM1988,hogstrom:BM1996}
 and numerical modelling \citep{spalart:PF2008, spalart:PF2009, ansorge:BM2014, ansorge:BM2019}.
-\todo{Refer to Pena et al, 2010} 
 %
@@
 One approach is to match the inner and outer layer at a reference height;
-\citet{etling:2002, emeis:2018}  (Sec. 21.10; Eq. 21.48) choose the Prandtl-layer height to match
-the wind speed profiles, which, however, requires to externally prescribe the veering $\alpha$ at that height. 
+\cite{etling:2002, emeis:2018}  (Sec. 21.10; Eq. 21.48) choose the Prandtl-layer height $z_\text{Prandtl}$ to match
+the wind speed profiles, which, however, requires external prescription of $\alpha(z_\text{prandtl})$, the veering at that height. 
 A one-dimensional profile with constant veering is given by \citet[Sec. 3; Eq. 3.1-3.19]{emeis:m2007}. 
+%%%
 
 \citet{gryning:BM2007} present an extension of the wind-speed profile beyond the surface layer
-using a neutral referece profile and a stability correction;
+using a neutral reference profile and a stability correction;
 \cite{kelly:BM2010}, based on a probabilistic representation of stratification,
-develop a model for the long-term mean windspeed
-in the atmospheric boundary layer and compare this with observation at different sites; 
+develop a model for the long-term mean wind speed
+in the ABL and compare this with observation at different sites; 
 \cite{kelly:WE2016} demonstrate the effect of such improved model for wind-energy applications.
@@
 truly neutral Ekman boundary layer.
+%
+A climatology of wind turning in the ABL is given by \cite{lindvall:QJR2019} 
+%
+\cite{klein:PAM2021a} use a statistical turbulence modelling approach that yields a two-component velocity profile,
+but they also find that the exact representation of turning is challenging.
+%%%
+
 %
@@
 %
-Ekman-layer models and two-layer models of the ABL take into account rotational effects at higher altitudes,
-for instance when the wind speed needs to be evaluated at heights on the order of 100-200\,\mathrm{m},
+Ekman-layer models are roughly based on Ekman's seminal \citeyear{ekman:AMA1905} paper in combinations with
+additional assumptions, such as a prescribed profile shape for eddy viscosity \cite{ellison:QJR1955} and two-layer models of the ABL take into account rotational effects at higher altitudes,
+for instance when the wind speed needs to be evaluated at heights on the order of $100-200\,\mathrm{m}$,
 a particular concern when it comes to wind-power forecasting \citep{optis:BM2014}.
@@
 % 
-Still, in \citeyear{jiang:JAS2018}, \citeauthor{jiang:JAS2018}
-recognized that the outer part of the Ekman boundary layer receives less attention in comparison
-with the surface layer and commence in studying the neutral problem by Large-Eddy simulation (LES).
-%
-They focus on the wind speed and find an extended logarithmic layer when considering the wind speed instead of
-the shear-aligned component, and they eventually demonstrate by means of an analytical model that this vertical
-extension of the logarithmic layer may be explained by a transfer of stress to the spanwise velocity component where
-it is assumed that the shear vector $\tau(z)$ and stress vectors $(\p_zU,\p_zV)$ are aligned. 
-%
-\par
-% 
-More recently, \citet{ghannam:QJR2021} treated the horizontally averaged momentum budget to show that departures from
-shear-alignment in the vicinity of the surface result in an integral of the wind veer ($\alpha_M$ in their notation)
-over the height to very high accuracy ($\int_{z_0}^{H} \sin\alpha_M$ in their notation; their Eq.~(16)).
+Still, in \citeyear{jiang:JAS2018}, \citeauthor{jiang:JAS2018} recognized that the outer
+part of the Ekman boundary layer receives less attention in comparison with the surface layer
+and study the neutral problem by Large-Eddy simulation (LES).
+%
+They focus on the wind speed and find an extended logarithmic layer when considering the wind
+speed instead of the shear-aligned component, and they eventually demonstrate by means of
+an analytical model that this vertical extension of the logarithmic layer may be explained
+by a transfer of stress to the span-wise velocity component where it is assumed that the shear
+vector $\tau(z)$ and stress vectors $(\p_zU,\p_zV)$ are aligned. 
+%
+\par
+% 
+More recently, \citet{ghannam:QJR2021} treated the horizontally averaged momentum budget to show
+that departures from shear-alignment in the vicinity of the surface result in an integral of
+the wind veer ($\alpha_M$ in their notation) over the height to very high accuracy ($\int_{z_0}^{H} \sin\alpha_M$ in their notation; their Eq.~(16)).
 %
@@
 Turbulent Ekman flow is considered here as a conceptual model of the
-homogeneous, stationary atmospheric boundary layer over a flat surface under neutral stratification. 
+homogeneous, stationary ABL over a flat surface under neutral stratification. 
 %
@@
 for theoretical exploration or higher-order approaches taking into account possible effects of stratification,
-roughness or other physical complications encountered in the realgeophysical system.
+roughness or other physical complications encountered in the real geophysical system.
 %
@@
 Large-Eddy simulation not only needs to be tuned for the surface shear stress and veering angle, but it also
-relies on sub-grid closures that often assume alignment of the stress with gradients -- a questionable
-pre-requisite when the wind rotates with height.
+relies on sub-grid closures that commonly assume alignment of the turbulent stress with gradients.
+%
+This pre-requisite is not fulfilled when the wind rotates with height.
 %
@@
 questions pertaining to such relatively simple dynamics of turbulence are not reflected in the
-research foci of LES for the ABL over the past 50 years \citep{stoll:BM2020}. 
+research on LES for the ABL over the past 50 years \citep{stoll:BM2020}. 
 %
@@
 Here, the routinely employed concept of Reynolds-number similarity can help. It postulates the existence
-of \say{\emph{fully developed turbulence}  believed to occur for a sufficiently large but finite Reynolds number}
+of \emph{fully developed turbulence}  believed to occur for a sufficiently large but finite Reynolds number
 \citep{barenblatt:PF1995}.
@@
 %
-We consider here incompressible, turbulent Ekman flow, that is, the turbulent flow over a flat rotating plate, as
-a physical model for the truly neutral atmospheric boundary layer (ABL).
+We consider here incompressible, turbulent Ekman flow, that is, the turbulent flow over a flat rotating plate, as a physical model for the truly neutral ABL.
 %
@@
 Analogously, we denote the height-local veering of the wind
-$\alpha(z)= \sphericalangle \left( \underline{G},\underline{U}(z)\right)$. 
+$\alpha(z)= \sphericalangle \left( \underline{G},\underline{U}(z)\right)$, where $\underline{G}=(G_1,G_2,0)$ is the geostrophic wind vector.
 %
@@
 We consider the incompressible Navier--Stokes equations for the three velocity components on the f-plane in a framework
-that is governed by the physical parameters
-(i)   geostrophic wind magnitude $G=\sqrt{G_1^2+G_2^2}$ (where $\mathbf{G}=(G_1,G_2,0)$ is the geostrophic wind vector),
+that is governed by % the physical parameters
+(i)   geostrophic wind magnitude $G=\sqrt{G_1^2+G_2^2}$,
 (ii)  Coriolis parameter $f$ (representing the angular rotation), and
@@
 %
-The scales used in defining $\RE_\Lambda$ loose their relevance in describing the turbulent state of the flow.
-%
-The friction Reynolds number
+The scales used in defining $\RE_\Lambda$ are of limited relevance for description of the turbulent flow state.
+%
+The turbulence scale separation in a wall-bounded flow is commonly characterized by the friction Reynolds number
+\citep{jimenez:ARF2014}:
 \begin{equation}
@@
 scales ($u_\star, \nu$).
-%
-$\RE_\tau$ is suitable to compare the turbulence scale separation also to other wall-bounded flow geometries
-such as rotating or non-rotating channel, pipe, or Couette flow.
 %
@@
 \end{subequations}
-where $u_i=U_i/G$ are the non-dimensional components velocity, $\pi$ is non-dimensional pressure, and $g_j=G_j/G$ are non-dimensionalized components
-geostrophic wind (with $g_{1}^2+g_{2}^2=1$ by construction).
-%
-These equations are solved inside a bounded cube of size $L_x \times L_y \times L_z$ with periodic boundary conditions in the lateral directions $Oxy$, a no-slip--no-penetration boundary at $z=0$, and
+where $u_i=U_i/G$ are the non-dimensional components velocity, $\pi$ is non-dimensional pressure, $g_j=G_j/G$ are non-dimensionalized components geostrophic wind (with $g_{1}^2+g_{2}^2=1$ by construction), and $\epsilon$ is the Levi--Civita tensor.
+%
+These equations are solved inside a bounded cube of size $L_x \times L_y \times L_z$ with periodic boundary conditions in the lateral (streamwise and spanwise) directions, a no-slip--no-penetration boundary at $z=0$, and
 a no-penetration, free-slip boundary at $z=L_z$.
@@
 \begin{table}
-  \caption{Direct numerical simulation datasets used in this
-    work \label{tab:sim-setup}}
-  \begin{tabular}{c r r c c c c c}
+  \caption{\label{tab:sim-setup} Direct numerical simulation data sets used in this work.
+$Re_\Lambda$ and $Re_D$ refer to the Reynolds number defined in terms of the Rossby radius $\Lambda$ and Ekman-layer thickness $D$ respectively.
+$L_{xy}$ is the domain size in the stream- and span-wise direction.
+The grid is given by the number of grid points in the stream-wise ($N_x$), span-wise ($N_y$) and vertical ($N_z$) directions respectively.
+The resolution in the span-wise and stream-wise directions are given as $\Delta x^+$ and $\Delta y^+$.
+The grid in the vertical is stretched, and  resolution at the wall is given by $\Delta z^+$.}
+  \centering\begin{tabular}{r r r c c c c c}
     \toprule 
@@
     \rowcolor{gray!30}
-       $125\,000$ &    $500$& 1.08 & $2048\times2048\times192$&4.1& 4.1& 1.05\\
+       $125\,000$ &    $500$& 1.08 & $2048\times2048\times192$&4.1& 4.1& 1.05 \\
     \rowcolor{gray!30}
-       $281\,250$ & $   750$& 1.08 & $3072\times3072\times384$&5.6&11.2& 1.60\\
+       $281\,250$ & $   750$& 1.08 & $3072\times3072\times384$&5.6& 5.6& 1.60 \\ 
     \rowcolor{gray!30}
        $500\,000$ & $1\,000$& 1.08 & $3072\times6144\times512$&9.3& 4.7& 1.14\\ 
+       $845\,000$ & $1\,300$& 0.54 & $2560\times5120\times640$&8.9& 4.5& 0.99\\
+
+    $1\,280\,000$ & $1\,600$& 0.54 & $3860\times7680\times960$&8.6& 4.3& 1.00\\
+    \bottomrule 
+  \end{tabular}
+
+ \caption{\label{tab:data} DOIs and reference to the openly accesible data set at refubium repository} 
+  \centering\begin{tabular}{r r r }
+    \toprule
+    $\RE_D$ & DOI & Reference \\
     \midrule
-       $845\,000$ & $1\,300$& 0.54 & $2560\times5120\times640$&8.9& 4.5& 0.99\\
-
-    $1\,280\,000$ & $1\,600$& 0.54 & $3860\times7680\times960$&8.6& 4.3& 1.00\\ 
+    \rowcolor{gray!30} 500& \DOI{10.17169/refubium-42505} & \cite{ansorge:2024c}\\
+    \rowcolor{gray!30} 1000& \DOI{10.17169/refubium-42507}& \cite{ansorge:2024a}\\
+
+    1300& \DOI{10.17169/refubium-42508} & \cite{ansorge:2024}  \\
+    1600& \DOI{10.17169/refubium-42509} & \cite{ansorge:2024b} \\
     \bottomrule 
@@
     surface veering angle $\alpha_\star$,
-    normalized boundary layer depth $\delta_{95}/\delta$
+    normalized boundary layer depth $\delta_{95}/\delta$, 
     inner normalization of vertically integrated TKE, 
@@
     \label{tab:bulk_results}}
-  \begin{tabular}{r r r r r r r }
+  \centering\begin{tabular}{r r r r r r r }
       \toprule
@@
 %
-Velocity profiles in inner and outer units (Fig.~\ref{fig:profiles}a,b) are in accordance
-with previous work \citep{coleman:JFM1992,spalart:PF2008,spalart:PF2009,ansorge:BM2014,ansorge:BM2019}:
+Velocity profiles in inner units ($U^+(z^+)$, Fig.~\ref{fig:profiles}a)and outer units ($U^-(z^-)$, Fig.~\ref{fig:profiles}b)
+are in accordance with previous work \citep{coleman:JFM1992,spalart:PF2008,spalart:PF2009,ansorge:BM2014,ansorge:BM2019}:
 %
@@
 Also in the outer layer of the flow, $u_\star$ (and not the magnitude of the
-geostrophic wind $G$) governs the inviscid, i.e. $\RE$-independent,
-normalization.
+geostrophic wind $G$) governs the inviscid normalization, i.e. a scaling independent of the Reynolds-number.
 %\todo{WHY? cf. TKE/Dissipation normalization above}
@@
 %
-No collapse in inner units is found for the profiles of spanwise velocity $V^{\alpha_\star+}(z^+)$;
+No collapse in inner units is found for the profiles of spanwise velocity $V^{\alpha_\star+}(z^+)$. 
 %
@@
 \begin{figure}
-  \begin{flushleft} 
   \textbf{a)}\hspace{0.47\textwidth} \textbf{b)}\\
@@
   \includegraphics[width=\textwidth]{../plot/eddy_viscosity.pdf}\\
-  \end{flushleft} 
   \caption{
@@
     Different colors are for different Reynolds numbers (cf. Tab.~\ref{tab:sim-setup}).
-    Circles in \textbf{(a)} and \textbf{(b)} denote the height
-    $z^-=0.15$, diamonds in \textbf{(c)} and \textbf{(d)} are for $z^+=50$
+    Circles in \textbf{(a)} and \textbf{(c)} denote the height
+    $z^-=0.15$, diamonds in \textbf{(b)} and \textbf{(d)} are for $z^+=50$
     as in Fig.~\ref{fig:profiles}.
@@
 surface veering direction is found, and intense TKE structures (bright yellow)
-are organized on a large spatial scale with with weaker eddies (greenish structures) and quiescent regions
+are organized on a large spatial scale with weaker eddies (greenish structures) and quiescent regions
 in between.
@@
 %
-This precludes a drag law wherewith we begin this section (\ref{sec:drag-law}).
-Based on the scaling arguments put forward in Sec.~\ref{sec:scaling}, we then develop
-a formulation of the wind vector in the Ekman layer (Sec~\ref{sec:ekman}) and separately
-for the streamwise (Sec.~\ref{sec:streamwise}) and spanwise (Sec.~\ref{sec:spanwise}) components
-in the overlap and inner regions. 
+This precludes, first, a drag law wherewith we begin this section (\ref{sec:drag-law}).
+Based on the scaling arguments put forward in Sec.~\ref{sec:scaling}, we then develop, second,
+a formulation of the wind vector in the Ekman layer (Sec~\ref{sec:ekman}). Finally, we come up
+with a separate formulation of the, third, stream-wise and, fourth, span-wise velocity components in the
+overlap and inner regions of the flow. 
 %
@@
   Z_\star= 4\log(\mathbf{Re}_D))-8
+  \label{eqn:zstar}
 \end{equation}
@@
 %
-\cite{optis:BM2014}, for instance, define an ``effective geostrophic wind vector that has the same magnitude of the
-observed surface geostrophic wind and is rotated by the angle $\alpha$ [their nomenclature]'' to overcome the
+\cite{optis:BM2014}, for instance, define an \emph{``effective geostrophic wind vector that has the same magnitude of the
+observed surface geostrophic wind and is rotated by the angle $\alpha$ [their nomenclature]''} to overcome the
 unsteady transition when approaching the Ekman layer from below.
@@
 %
-The text-book solution for Ekman flow makes use of the physical boundary conditions (BCs) for the atmospheric
-boundary layer (no-slip at the bottom and geostrophic wind in the free atmosphere) and a constant eddy
+The text-book solution for Ekman flow makes use of the physical boundary conditions (BCs) for
+the~ABL (no-slip at the bottom and geostrophic wind in the free atmosphere) and a constant eddy
 viscosity.
@@
         \sin z_\mathrm{ek} \\ \cos z_\mathrm{ek}
-        \end{array}\right)\right]\\
-    z_\mathrm{ek} &=& \delta_\mathrm{ek} (y^- -s_\mathrm{ek}). 
+        \end{array}\right)\right]. 
   \end{eqnarray}
 \end{subequations}
+with  $z_\mathrm{ek} = \delta_\mathrm{ek} (y^- -s_\mathrm{ek})$.
 % 
@@
 this constant are only found for the lowest Reynolds numbers which is in accordance with
-the low-$\RE$ correection suggested by \cite{spalart:JFM1989}.
+the low-$\RE$ correction suggested by \cite{spalart:JFM1989}.
 %
@@
 %
+Given the Ekman formulation of the velocity profile introduced in Sec.~\ref{sec:ekman},
+one may express $v_{03}$ using the Ekman profile introduced in Sec.~\ref{sec:ekman} together
+with the approximation for $u_\star(Re)$ found in Eq.~(\ref{eqn:zstar}). 
+% 
+%
 While the $\RE$-dependency of $a_\mathrm{log}$, $b_\mathrm{log}$, $c_\mathrm{log}$
@@
 \section{Discussion}
-\subsection{} 
+\subsection{Implications for surface-layer scaling} 
 Eq.~(\ref{eqn:scaling_spanwise}) establishes a universal mixed scaling for the spanwise velocity
@@
 
-\section{Comparison with other theories}
+\subsection{Comparison with other theories}
 
 \textbf{Compare} to Emeis (2002) and Gryning (2007); highlight explicit knowledge on veering-profile $\rightarrow$ directional sheer;
+
+\textbf{Compare} to van Driest scaling
+
+\textbf{Remember} of interpretation in the context of eddy viscosity (Fig.~\ref{fig:stresses}) 
+Consider Townsend ``Turbulent Shear Flow''  Chapter 7.18: p. 319 
 %
@@
 \end{itemize}
+
+\bibliographystyle{spbasic_updated} 
+\bibliography{ansorge}
 
@@
 
-\printbibliography
-
 \appendix

### 2024-06-27 write1/vandriest.tex
update on part 2

\documentclass[a4paper,12pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{geometry}

\newcommand{\p}{\partial}

\title{van Driest's model for wall scaling}

\begin{document}

We identify the total shear stress $\tau$ as
\begin{equation}
  \tau = \underbrace{\mu \frac{\p \overline{u}}{\p y}}_\text{viscous stress} - \underbrace{\rho \overline{u^\prime v^\prime}\vphantom{\frac{\p \overline{u}}{\p y}}}_\text{Reynolds' stress}, 
\end{equation}
where $y$ is distance from the wall, $\mu$ dynamic viscosity ($\nu=\mu/\rho$) and $\rho$ is density. 
Next, observe that the amplitude of motion due to oscillation of an infinite plate decays as $\exp\left[-y/A\right]$. Hence, we use for the damping of fluid oscillation due to a fixed wall, the model
\begin{equation}
  1- e^{-\tilde{y}},
\end{equation}
where $\tilde{y}=y/A$. Now, we express the stress, according to Prandtl
\begin{subequations} 
\begin{eqnarray}
  \frac{\tau}{\rho} &=& \nu \frac{\p \overline{u}}{\p y} +  r \sqrt{\overline{u^{\prime2}}} \sqrt{\overline{v^{\prime2}}}  \\
       &=& \nu \frac{\p \overline{u}}{\p y} + r l_1 l_2 \left(\frac{\p \overline{u}}{\p y} \right)^2 \\
       &=& \nu \frac{\p \overline{u}}{\p y} + \kappa^2 y^2 \left(\frac{\p \overline{u}}{\p y} \right)^2, 
\end{eqnarray}
\end{subequations}
where $\kappa$ is the von-K\'arm\'an constant and $l$ Prandtl's mixing length. This model is well-known to hold appropriately in fully developed turbulent flow.

\par

Near a wall, however, the turbulence is not fully developed, but damped by the presence of that very wall such that the prefactor $1-e^{-\tilde{y}}$ can be taken into account in Reynolds' stres term:
\begin{eqnarray}
  \label{eqn:stress_model_dimensional}
  \frac{\tau}{\rho} &=& \nu \frac{\p \overline{u}}{\p y} + \kappa^2 l^2 \left(1- e^{-\tilde{y}}\right)^2\left(\frac{\p \overline{u}}{\p y}\right)^2
\end{eqnarray} 
Non-dimensionalize Eq.~(\ref{eqn:stress_model_dimensional}) using $u_\star=\sqrt{\tau_\text{wall}/{\rho}}$ and the wall unit $y_+=\nu/\sqrt{\tau_\text{wall}/\rho}$:
\begin{eqnarray}
  \tau^+ \left( = \frac{\tau}{\tau_\text{wall}} \right) = \frac{\p u^+}{\p y^+} + \kappa^2 y^{+2} \left(1- e^{-\widetilde{y^+}}\right)^2 \left(\frac{\p u^+}{\p y^+} \right)^2
\end{eqnarray}

In the constant-flux layer, it is $\tau=\tau_\text{Wall}$, such that
\begin{eqnarray}
  0 &=&  \frac{\p u^+}{\p y^+} + \kappa^2 y^{+2} \left(1- e^{-\widetilde{y^+}}\right)^2 \left(\frac{\p u^+}{\p y^+} \right)^2 - 1 \\ 
  0 &=&  \left(\frac{\p u^+}{\p y} \right)^2 + \frac{1}{\kappa^2 y^{+2}\left(1-e^{-\widetilde{y^+}}\right)^2} \left(\frac{\p u^+}{\p y^+} -1\right)
\end{eqnarray}
and we solve for $\p_{y^+}u^+$ to obtain 
\begin{eqnarray} 
  \frac{\p u^+}{\p y^+} = \frac{2}{1+\sqrt{1+4 \kappa^2 y^{+2}\left(1-e^{-\widetilde{y^+}}\right)}} 
\end{eqnarray}

\end{document}

### 2024-06-27 write2/manuscript_part2_2024.tex
update on part 2

\documentclass[a4paper,11pt]{article}

\usepackage[margin=2.5cm]{geometry}
\usepackage{booktabs} 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{natbib}
\usepackage{xcolor}
\usepackage{tikz}

\newcommand{\p}{\partial}
\newcommand{\gzm}{{\xi}}%{\gamma z_\text{match}}}
\newcommand{\todo}[1]{\textcolor{red}{$[$#1$]$}}
\newcommand{\RE}{\mathbf{Re}}
\hyphenation{Ol-den-burg}

\DeclareRobustCommand\sampleline[1]{%
  \tikz\draw[#1] (0,0) (0,\the\dimexpr\fontdimen22\textfont2\relax)
  -- (2em,\the\dimexpr\fontdimen22\textfont2\relax);%
}

\providecommand{\keywords}[1]
{
  \small	
  \textbf{\textit{Keywords---}} #1
}

\pdfsuppresswarningpagegroup=1

\author{Hauke Wurps and Cedrick Ansorge}

\date{Draft \today}
\title{Towards a Universal Veering Profile for Turbulent Ekman Flow at arbitrary Reynolds number - Part 2\\
  \normalsize LES and DNS of Turbulent Ekman Flow \\}

\begin{document}

\maketitle

\begin{abstract}
  % Introduction % Theoretical profile
  Turbulent Ekman flow is a canonical flow representation of the atmospheric boundary layer, characterized by a logarithmic increase of the wind speed near the surface and a turning of the wind vector aloft. In part I of this work, we derived a formulation of the mean wind velocity profiles of the stationary turbulent Ekman flow based on direct numerical simulation (DNS). Here, we explore the extrapolation of these profiles to atmospheric Reynolds numbers using large-eddy simulation (LES).
  % Case desription and setup
  We compare theoretical profiles of the wind vector for low Reynolds-number turbulent Ekman flow to results from DNS and LES. This necessitates the inclusion of viscous effects, usually neglected in LES, along with modifications to the standard bottom boundary condition.
  % Results and Discussion
  Analysis of the grid-convergence for higher resolution unveils a convergence of LES data towards the theoretical profiles for both intermediate and high Reynolds numbers. Our LES thus confirm that--within the logarithmic layer (the lowest 10\,\% of the boundary layer)--- the turning of the wind vector rotates by approximately one-third of its total rotation across the entirety of the boundary layer.
  % Conclusions
  Such agreement of our theoretical formulation and LES data raises our confidence in the underlying assumptions of the theory and thus reinforces the utility of the theoretical profiles as a reference for intermediate and a quasi-reference for higher Reynolds number simulations.
%

\keywords{Large-Eddy Simulation $\cdot$ Scale separation $\cdot$ Ekman layer $\cdot$ Prandtl layer $\cdot$ Hodograph}
\end{abstract}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\section{Introduction}

% \todo{start introduction rather with TEF than LES (rather subject than method)}
The interaction of atmospheric boundary layer  (ABL) flow with its lower boundary condition, the surface, is a defining property of the ABL \citep{stull:1988}. Relevant aspects of the interaction with the boundary not only include its heterogeneity \citep{avissar:MWR1989,giorgi:RG1997,claussen:BM1991,garratt:BM1990} and roughness \citep{monin:ARF1970,brutsaert:WRR1975,raupach:AMR1991,kostelecky:JFM2024}, but also a wind veer in favor of the pressure gradient force when the surface is approaced \citep{ekman:AMA1905}. An important tool with capabilities to model the flow-surface interaction is Large-Eddy Simulation (LES). Indeed, LES is widely used to model turbulent flow in the atmospheric boundary layer (ABL) \citep{stoll2020large}. This includes various applications, such as simulating wind turbine and wind farm wakes, and interactions within wind farm clusters \citep{porte2011large,mehta2014large,breton2017survey}. Furthermore, LES is applied in complex environments like mountainous or urban areas \citep{stoll2020large,garcia2018predictive} and to investigate the dispersion of pollution \cite{han2018large}. These applications typically focus on the lower segment of the ABL on the rotating earth, known as the Prandtl or surface layer. In this layer, the vertical wind speed profile commonly exhibits a logarithmic increase, followed by a prominent change of wind direction within the Ekman layer above. Accurate characterization of how wind speed and direction vary with height is of great importance for wind-power forecasting and projection \citep{optis2014moving}.

The simplest representation of the ABL taking into account rotation \emph{and} the quasi non-turbulent, free atmosphere aloft is the turbulent Ekman flow. It is a horizontally homogeneous, statistically stationary boundary-layer flow over a rotating flat surface. Here we consider the problem under neutral stratification, where the potential temperature is constant across the whole domain. Turbulent Ekman flow shows the key characteristics shaping the real ABL: a logarithmic layer and the Ekman spiral. Near the surface, viscous forces dominate the flow. The stationary ABL-height is then defined by the interplay between turbulent growth due to the shear instability of the configuration \citep{lilly1966instability} and rotational suppression of turbulence due to the Coriolis effect. The system's statistical equilibrium is defined by one single parameter, the Reynolds number $\RE$, signifying the scale separation between the largest and smallest scale within the problem. The largest scales are given by shear-induced eddies constrained by the boundary layer's height, while the smallest scales reside within the dissipative range, where strong viscous forces suppress turbulence. The dimensional mean solution of the turbulent Ekman layer is influenced by three parameters: the geostrophic wind $G$, the Coriolis parameter $f$, and the kinematic viscosity $\nu$. These three parameters define the Reynolds number $Re_D = GD/\nu$. With $D=\sqrt{2\nu/f}$ the Ekman layer depth of a laminar flow we write

\begin{align}\label{red}
 Re_D = \sqrt{2}\frac{G}{\sqrt{\nu f}}
\end{align}

The mean solutions of the velocity components are functions of only the Reynolds number and the height above the ground $z$: $u(\RE_D,z)$, $v(\RE_D,z)$ (see e.g. \cite{csanady1967resistance}).

Part I of this publication, presents a framework to predict wind direction and speed across the turbulent Ekman flow, drawing from the works of \cite{csanady1967resistance}, \cite{tennekes1973logarithmic}, and \cite{spalart1989theoretical}. As \cite{spalart1989theoretical}, we use $G$, $f$, and $\nu$ to derive the friction velocity $u_\star$ and the angle $\alpha_\star$ between geostrophic wind and shear stress at the wall. While earlier descriptions of the turbulent Ekman layer were limited to specific parts of the boundary layer, we derive mean velocity profiles covering the entire ABL with due regard of changes in the Reynolds number. We have shown that the theoretical velocity profiles are in excellent agreement with DNS for intermediate Reynolds numbers.

However, it remains uncertain, if the extrapolation of the above framework towards atmospheric $\RE$ is valid. The substantial separation of scales in the ABL renders direct numerical simulation (DNS) of the entire turbulent flow at scale impossible. Within the framework of LES, only the larger eddies are resolved, that contain the main share of kinetic energy, while turbulent mixing below the subfilter-scale is parameterized. This technique allows for the simulation of the ABL at very high (atmospheric) $\RE$.

The accurate representation of mean profiles of velocity, i.e. of first-order statistics of the wind field, is commonly assumed when using LES in a predictive manner  \citep{fedorovich2004convective}. Here, we compare LES outcomes to theoretical solutions to determine the validity of this assumption under different conditions. As LES introduces the grid size as an additional parameter (often understood in relation to a filter scale, cf. \cite{pope2004ten}), we conduct a thorough analysis of its impact on simulation results. \cite{esau2004simulation} and \cite{jiang2018large} have executed similar LES analyses of neutral Ekman layers, however, they introduce an additional parameter by selecting an arbitrary roughness length $z_0$, whereas we derive the parameter $z_0$ corresponding to a fluid with a kinematic viscosity of $\nu = 1.5\cdot10^{-5}$ over flat surface.

Here, we examine three different Reynolds numbers $Re_D$: $1600$, $1.5\cdot10^5$, and $10^6$. While the case $Re_D=1600$ is also investigated by DNS (cf. part 1 of this study), $Re_D=1.5 \cdot 10^5$ and $10^6$ correspond to the scale separation found in typical atmospheric boundary layers. For $Re_D=1600$, we compare DNS and LES results using the exact same forcing and settings. This is the first comparison of this kind for the turbulent Ekman flow. The LES results converge to the DNS results with decreasing grid cell size. This approach raises our confidence in the LES results to a new level for the direct comparison with a resolving simulation. Furthermore, it allows us to extend the insight from DNS to atmospheric scale where an approach through DNS is clearly infeasible for computational constraints. We observe a clear $\RE$ dependency in the LES results, matching the predictions of the theoretical framework.

The content is structured as follows. Section \ref{universal_profile} provides a complete mathematical description of the turbulent Ekman layer's velocity profiles. A description of the simulated cases and the numerical set-up is given in section \ref{setup}, followed by the presentation of results and its comparison to the theoretical profiles in section \ref{results}. We conclude in section \ref{conclusion}.

\section{A Universal Velocity Profile for the Turbulent Ekman Layer}
\label{universal_profile}

In this chapter, we introduce the universal velocity profile for the turbulent Ekman layer developed in part I. 
%Definitions
We align the coordinate Ox with the surface stress, Oz points normal to the surface and normal to Ox, and Oy in the span-wise direction, normal to Oxz.
%The mean velocity profiles for all $\RE$ exhibit a dual scaling nature: in proximity to the lower boundary, the velocity profiles collapse when scaled using
In proximity to the lower boundary, the boundary layer scales in viscous (inner) units, denoted by the index ``$+$'' ($z^+ = zu_\star/\nu$, $U^+=U/u_\star$, where $U$ is the velocity component in x-direction). From the top perspective, the profiles converge when scaled using outer units, denoted by the index ``$-$'' ($z^-=z/\delta$, $U^-=U/G$, with $\delta=u_\star/f$ representing the boundary layer height). 
At the lower boundary, the x-axis of the inner units aligns with the shear stress, while the x-axis of the outer units aligns with the geostrophic wind. The angle between both axes is denoted as $\alpha_\star$, the surface veer of the wind across the boundary layer.

% \subsection{Drag-Law}

The surface stress is linked to the geostrophic wind via the geostrophic drag $Z\equiv u_\star/G$ and the angle between the shear stress and the geostrophic wind $\alpha_\star$. These two parameters form the basis of the boundary-layer scaling and are  estimated using the semi-empirical drag-law introduced by \cite{spalart1989theoretical}, which describes them as functions of only the Reynolds number:
\begin{subequations}\label{eqn:drag}
	\begin{align}
		\frac{G}{u_\star}\cos\phi^\star &= \frac{1}{\kappa}\log Re_\tau + C - A_r, \\
		\sin\phi^\star &= A_i\frac{u_\star}{G},\\
		\alpha_\star &= \phi^\star - \frac{C_5}{Re_\tau},\\
		Re_\tau &= \frac{u_\star^2}{\nu f} \left( = \frac{1}{2} Re_{D}^2 \frac{u_{\star}^2}{G^2}\right),
	\end{align}
\end{subequations}
where we use $\kappa = 0.416$, $A_r = 4.80$, $A_i = -5.57$, $C = 5.4605$, $C_5 = -57.8$ based on the DNS data available (cf. Part I, \cite{ansorge2014global}).

% \subsection{Total profile}

Profiles of the velocity components are studied in three different layers reflecting the change of dominant balance when moving away from the surface. These are (i) the viscous sub-layer, denoted by an index $_\text{visc}$ (ii) the logarithmic layer denoted by $_\text{log}$, and (iii) the Ekman layer denoted $_\text{Ek}$. We match $U_{visc}$ and $U_{log}$ according to their respective formulations and combine them to $U_{inner}$. The span-wise component of the velocity is separated into two layers, namely the inner layer $V_{inner}$, and the Ekman layer $V_{EK}$.

\subsection{Ekman layer}

The outer layer of the ABL is characterized by a triadic balance between turbulent flux, pressure gradient and the Coriolis; the vertical change of the Coriolis force causes a pronounced height dependence of the wind direction resulting in the Ekman spiral \citep{ekman:AMA1905}. The classic solution employs Ekman dynamics down to the surface---and thus also in the surface layer, where this is a rather poor representation of turbulent mixing. Here, we model the surface layer by a lower boundary condition for the Ekman spiral determined from DNS:
\begin{subequations}
  \begin{align}
    U_{EK} &= G + Ae^{-\tilde{z}}\cos\tilde{z},\\
		V_{EK} &= - Ae^{-\tilde{z}}\sin\tilde{z},
  \end{align}
\end{subequations}
where the x-axis is aligned with the geostrophic wind and $A = 8.4u_\star$, $\tilde{z} = (z-z_r)/D_E$, $z_r = 0.12\delta$, and $D_E = 3\delta/4\pi\approx 0.24\delta$. The transition from the logarithmic layer to the Ekman layer is located at $z^{-}=0.28- 2.25 Re_{D}^{-1/2}$ with a transition scale of $\sigma_T=2$ for the stream-wise velocity.

\subsection{Inner and viscous layers}

In the viscous sublayer, the shear-aligned velocity $U^{\alpha_\star}$ is described by the law of the wall $U^{\alpha_\star+} = z^+$ and the span-wise velocity $V^{\alpha_\star}$ is zero by choice of the reference frame 
(the index $\alpha_\star$ indicates the alignment with the direction of the shear stress). Around $z^+=5$, the velocity begins to deviate from linearity, and the buffer layer forms the transition from the viscous to the  logarithmic layer. From the surface up to the buffer layer, the stream-wise velocity is described by
\begin{subequations}
  \begin{align}
    U_\text{buffer}^{\alpha_\star+} = \frac{z^+}{1+c_1(z^+)^2} + (c_2z^+-a_\text{match})\frac{1+\tanh[0.2(z^+-22)]}{2}+c_3e^{-c_4(z^+-22)^2}.
    \label{eqn:ubuffer}
  \end{align}
With $c_1 = 0.00185$, $c_2 = 0.195$, $c_3 = 0.4$, $c_4 = 0.35$. The coefficient $a_{match}=3.5727$ is chosen to match the u-profile in the logarithmic layer above at $z^+=40$, such that 
\begin{align}
  U_\text{inner}^{\alpha_\star+} = \begin{cases}
	  U_\text{buffer}^{\alpha_\star+}, & z^+ \le 40.\\
		\kappa^{-1}\log z^+ + C &   z^+>40.
	\end{cases}
  \label{eqn:log-law}
\end{align}
\end{subequations}
with the von-Kármán constant $\kappa=0.416$, and $C=5.4605$ as in the estimation of the geostrophic drag in Eq.~(\ref{eqn:drag}).

For the span-wise velocity, the boundary conditions and Ekman dynamics imply that 
\begin{subequations}
\begin{align}
V^{\alpha_\star}_\text{visc} \delta^+\sim G f_V(z^+),
\end{align} where $f_V$ is a universal, non-dimensional function (cf. Part I). Above the viscous 
layer, scaling arguments are scarce, but there is evidence for a logarithmic scaling of also the 
span-wise velocity component. We  hence use  
\begin{align}
  V^{\alpha_\star}_\text{inner} = \left\{ \begin{matrix}
      V^{\alpha_\star}_\text{visc} &=& \dfrac{G}{\delta^+}v_{ref}(w_vz^+-1+exp[-w_vz^+])  & z^+ \le 10\\ 
      V_\text{log}^{\alpha_\star}        &=& \dfrac{G}{\delta^+}(a_{log} + b_{log}\log(z^+) + c_{log}z^+) & z^+>10. 
  \end{matrix}\right.
\end{align}
\end{subequations}
with $v_\text{ref} = 18.85$ and $w_v = 0.2353$ leads to excellent agreement with the DNS data. The coefficients $a_\text{log}, b_\text{log}$, and $c_\text{log}$ are determined by (i)-(ii) a smooth transition to $V_\text{visc}$ at $z^+=10$, and (iii) the condition $V_\text{log}(z^-=0.3) = V_\text{EK}(z^-=0.3) =: v_{03}$ (cf. Part I). For the Reynolds-number dependency of the Ekman-layer profiles, $V_\text{Ek}(z^-=0.3)$ depends on $Re$, such that the coefficients exhibit a weak dependence on $Re$.

\subsection{Blending of inner and outer profiles, synopsis}

A smooth transition between consecutive layers is achieved using a transfer function:
\begin{align}\label{eqn:error}
  w_{*} = \frac{1}{2}\left(\textrm{erf}\left[\sigma_T\log\left(\frac{z}{z_{T}}\right)\right]+1\right),
\end{align}
where $erf$ is the error function, $\sigma_T$ is a transition scale that defines the width of the transition and $z_{T}$ is the height of the transition, where the upper and the lower layer equally contribute to the velocity ($w_{*}(z_{T})=0.5$). The inner profiles $U_\text{inner}$ and $V_\text{inner}$ are blended to the Ekman profile using eq.~(\ref{eqn:blend}) with $\sigma_T = 2$ and $z_T^- = 0.28- 2.25\sqrt{1/Re_D}$.

\begin{subequations}\label{eqn:blend}
  \begin{align}
    U &= (1-w_{outer})U_{inner} + w_{outer}U_{EK},\\
	  V &= (1-w_{outer})V_{inner} + w_{outer}V_{EK}.
	\end{align}
\end{subequations}

\begin{figure}
  \centerline{
	\includegraphics[width=\textwidth]{figures_2024/alt_single_theor_Ekman_profiles_2022.pdf}}
  \caption{a) Theoretical shear-aligned velocity profile ($U^+$,$-V^+$) of the turbulent Ekman flow for $Re_D = 2000$ (\sampleline{}) and $Re_D = 4000$ (\sampleline{dashed}) and the different layers (colors). b) Hodograph of the geostrophy-aligned velocity components}
  \label{single_ekman}. 
\end{figure}

The velocity profiles are given for two Reynolds numbers in Fig. \ref{single_ekman}. 
In the viscous sublayer, the shear-aligned component $U^+$ increases linearly with height up 
$z^+\approx5$. In between $5 \lesssim z^+ \lesssim 30$, we find the buffer layer with a transition from the linear law-of-the-wall to the logarithmic law. Above $z^+\approx30$, the logarithmic layer begins. Then, $U^+$ increases logarithmically up to the Ekman layer and reaches its supergeostrophic maximum. Above, it decreases to its free-stream geostrophic value. The depth of the logarithmic layer increases with $\RE$. The spanwise component $V^+$ remains close to zero up to the middle of the logarithmic layer, where the transition to the Ekman layer takes place ($z^-\approx0.28-2.25\sqrt{1/Re_D}$). The profiles of $V^+$ of all Reynolds numbers have a similar shape but are shifted in $z^+$. The similar $V^+$ and the growth of $U^+$ leads to a smaller angle $\alpha_\star$ between surface shear stress and geostrophic wind for higher $\RE$, which is visible in the hodograph.

\section{Case description and numerical set-up}
\label{setup}
%
%\subsection{Notation}
%
%\begin{table*}
	%\centering
%%	\caption{}
	%\begin{tabular}{ll}
    %\toprule
		%$\Delta$ & grid cell size in LES \\
		%$\overline{u''w''}_0$ &  \\
		%$\overline{v''w''}_0$ &  \\
		%$\delta_{95}$ & height where the turbulent stress decreases to 5\% of its surface value\\
		%\bottomrule
	%\end{tabular}
%\end{table*}

\subsection{Settings}

An incompressible, turbulent Ekman flow over a hydrodynamically smooth surface is simulated for three different Reynolds numbers $Re_D = 10^3;1.5\times 10^5;10^6$, hereafter $Re1$, $Re2$, and $Re3$, respectively. The input parameters are given in table \ref{simulation_parameters}.

\begin{table*}
	\centering
	\caption{Parameters of the simulated cases. $f$, $\nu$, and $G$ are input parameters and define the Reynolds numbers $Re_D$ and $Re_\tau$, while $u_\star$, $\alpha_\star$, and $\delta = u_\star/f$ are resulting properties of the flow}
	\begin{tabular}{lcccccccc}
          \toprule  
	  Name & $Re_D$ & $Re_\tau$ & $f$ [s$^{-1}$] & $\nu$ [m$^2$s$^{-1}$] & $G$ [ms$^{-1}$] & $u_\star$ [ms$^{-1}$] & $\alpha_\star$ [$^\circ$]& $\delta$ [m] \\
          \midrule
	  Re1 & $1.6\times10^3$ & $3.0\times10^3$ &  &  & 0.0438 & 0.00211 & 16.8 & 21.1 \\
	  Re2 & $1.5\times10^5$ & $7.3\times10^6$ & $10^{-4}$ & $1.5\times10^{-5}$ & 4.108 & 0.1048 & 8.5 & 1048\\
	  Re3 & $1\times10^6$ & $2.2\times10^8$ &  &  & 27.39 & 0.5785 & 7.0 & 5785 \\
          \bottomrule
	\end{tabular}
	\label{simulation_parameters}
\end{table*}

The domain is rotating around the z-axis with an angular velocity corresponding to the Coriolis parameter $f=10^{-4}s^{-1}$. The stratification of the flow is truly neutral, i.e., the potential temperature is constant for the whole domain. Nevertheless, the boundary layer height is not growing infinitely but results from a balance between shear production and rotational suppression of turbulence. At the upper boundary, sufficiently far aloft of the boundary-layer height, a no-penetration boundary condition is used and the horizontal components of the wind are geostrophic, which is a Dirichlet-type boundary condition. At the bottom, a constant-flux layer is assumed and Monin--Obukhov similarity theory (MOST) is used to calculate the surface momentum fluxes. The Navier-Stokes equations are integrated using a 3rd-order low-storage Runge-Kutta method. For scalar advection a 5th-order Wicker--Skamarock scheme is employed. The Poisson equation is solved using a direct fast Fourier transform (FFT). In LES, the turbulent transport on the subgrid scale (SGS) needs to be modeled by a dedicated model, the SGS model. We use two different SGS models to check whether the subgrid closure significantly influences the LES solution: a 1.5-order closure after \cite{deardorff1980stratocumulus} and a dynamic closure after \cite{heinz2008realizability}. For most of the simulations, the 1.5-order closure is used, since the dynamic closure needs more computational resources, but the simulations with $\Delta^-=200^{-1}$ are repeated using the dynamic closure.  A description of the LES model is given by \cite{maronga2020overview}.

\begin{table*}
	\centering
	\caption{Simulations and grid parameters: ReX stands for one of the Reynolds numbers Re1, Re2, and Re3. $\Delta$ is the grid cell size, $\delta=u_\star/f$ is the boundary layer height, $n_i$ is the number of grid cells in the direction $O_i$, $L_x$ and $L_z$ are the domain sizes in the horizontal and vertical direction, respectively}
	  \begin{tabular}{lccccccc}
          \toprule 
	    Name & $\Delta^-$ & $n_x$ & $n_y$ & $n_z$ & $L_x/\delta$ & $L_z/\delta$  \\ 
          \midrule
  	  ReX\_50 & 1/50 & 144 & 144 & 128 & 2.88 & 5.0 \\
  	  ReX\_100 & 1/100 & 288 & 288 & 216 & 2.88 & 4.5 \\
	    ReX\_150 & 1/150 & 432 & 432 & 288 & 2.88 & 3.7 \\
  	  ReX\_200 & 1/200 & 576 & 576 & 384 & 2.88 & 4.1 \\
  	  ReX\_dyn & 1/200 & 576 & 576 & 384 & 2.88 & 4.1 \\
           \bottomrule
		\end{tabular}
	\label{simulation_parameters2}
\end{table*}

To study the effect of resolution on the simulations, four different grid resolutions are chosen for each Reynolds number case. The grid cell size $\Delta$ is around $\delta/50$, $\delta/100$, $\delta/150$, and $\delta/200$, one coarse, two medium, and one fine resolution, respectively (see table \ref{simulation_parameters2}). In total, 15 simulations are carried out. The grid spacing inside the boundary layer is isotropic up to $z = 1.3\delta$. Aloft, the grid spacing along Oz is stretched by 3\% per grid point until a maximum spacing of $(\Delta z )_\mathrm{max} = 6\Delta x$ is reached. The number of vertical grid points is chosen such that $L_z \ge 3 \delta$. Different domain heights are caused by numerical requirements of the FFT-solver. In the upper third of the domain, Rayleigh damping is active to avoid wave reflections from the top boundary.

The flow is initialized with wind speed profiles based on a one-dimensional model with a Reynolds-average based turbulence parametrization. At the beginning of the simulation, random perturbations are imposed on the velocity field to trigger turbulence. The resulting imbalance between pressure force and Coriolis force results in an inertial oscillation of the period $T_{io}=2\pi/f$, where a part of the flow's mean kinetic energy oscillates between U- and V-component. The oscillation decays over time and would eventually vanish for large time. In order to obtain profiles in statistical equilibrium of the flow, we use a spin-up time of 1.5 $T_{io}$ and perform a horizontal domain average over 2 $T_{io}$.

In part I of this publication, the DNS of $Re_D\leq1\,600$ is carried out for a horizontal domain size of $(0.54\Lambda_{Ro})^2$ - $(1.08\Lambda_{Ro})^2$, where $\Lambda_{Ro}=G/f$ is the Rossby radius. For these Reynolds numbers, $u_\star/G$ is around 0.05 so $L_x \approx 10 \delta$ However, $u_\star/G$ decreases with increasing Reynolds number and is around 0.02 for $Re_D=10^6$. A domain size of half the Rossby radius would then extend to $L_x\approx 25\delta$. Such a large domain would imply immense computational costs. \cite{spalart2008direct} used a horizontal domain of $L_x = 2\delta$, arguing that this length allows the resolution of the largest outer-layer eddies according to \cite{csanady1967resistance}. During a sensitivity test of the domain size we observed that simulations with domain sizes $L\geq 4\delta$ often tend to accumulate turbulence kinetic energy in the upper half of the the boundary layer. This TKE increases over several inertial oscillations with energy mostly on the scale of the domain size. Such a development was not observed in the DNS. We could successfully avoid such a behaviour by using a domain of size $L \approx 3\delta$ in combination with a shifted periodic boundary condition in y-direction, as described by \cite{munters2016shifted}. Although for the Ekman flow the direction of the mean flow is only aligned with the x-direction of the grid near the surface, a shift of the boundary condition by $L/3$ significantly helped to suppress the accumulation of TKE in the upper half of the boundary layer.

\subsection{Viscosity and roughness length}

In LES, one postulates that a sufficient part of the largest eddies is resolved so as to represent the dominant non-linear effects of turbulent mixing \citep{pope2004ten}. Below these resolved scales, turbulence is modeled as a more or less isotropically acting diffusive agent by a closure model (dynamic, Deardorff, see above). Thus, molecular friction is not considered directly, but only by virtue of a turbulence model linking the resolved and dissipative scales. In their seminal works on the spectral energy transfer in homogeneous isotropic turbulence, Kolmogorov and Obukhov (\cite{kolmogorov1941dissipation},\cite{obukhov1941distribution}) showed that the energy transfer rate across the spectrum is in fact a constant. This implies that the transfer rate across the cut-off scale in LES does not depend on the magnitude of the viscous scale, presupposed that (i) the cut-off scale of the LES is well within the inertial range and (ii) the LES turbulence is approximately isotropic and homogeneous at the smallest resolved scales. Consequently, SGS-models of LES do not necessarily require explicit information about the actual viscosity of the fluid or other viscous parameters.

In LES at low $\RE$ or very high resolution the subgrid eddy viscosities may fall far below the molecular viscosity of air. In the context of PALM's Deardorff closure, it is $K_m = c_0 \Delta \sqrt{e}$, where $c_0 = 0.1$ \citep{deardorff1980stratocumulus}, $\Delta$ is the grid size and $e$ is the SGS-TKE, calculated by a prognostic equation. Hence, very low $e$ as well as fine resolution can lead to $K_m<\nu$. When this is the case, $\nu$ cannot be ignored anymore, and we let $K_m = c_0 \Delta \sqrt{e} + \nu$. The governing equation of the velocity components in PALM reads
\begin{align}\label{gov_palm}
  \frac{\partial u_i}{\partial t} = - \frac{\partial u_i u_j}{\partial x_j} -\epsilon_{ijk}f_ju_k + \epsilon_{i3j} f_3 u_{g,j} - \frac{1}{\rho_0}\frac{\partial\pi^*}{\partial x_i} + \frac{\partial}{\partial x_j}\left( K_m\left(\frac{\partial u_i}{\partial x_j} + \frac{\partial u_j}{\partial x_i} \right)\right)
\end{align}
for a neutrally stratified flow. In the limit of either very well resolved simulations or very low $\RE$ (or both), $K_m \rightarrow \nu$, so the last term on the right hand side of eq. (\ref{gov_palm}) becomes $\nu\frac{\partial^2u_i}{\partial x_j^2}$, which complies with the Navier-Stokes equations for an incompressible fluid.

\label{low-bound}

In contrast to the interior closure of the LES where the direct effect of $\nu$ is of small relevance compared to the eddy viscosity---at least for high $\RE$---, this is not true for the viscid effects at the bottom boundary. Here, a constant flux layer is usually assumed below the first grid point and the Monin--Obukhov similarity theory is used to compute the friction velocity and the stresses at the first half grid point (where the horizontal velocities are located on the Arakawa-C staggered grid):
\begin{align}\label{most}
  u_\star = \frac{\kappa(U^2+V^2)^{0.5}}{\ln\left(z/z_0\right)},\\
  -\overline{u''w''}_0 = \frac{\kappa Uu_\star}{\ln\left(z/z_0\right)}.
\end{align}
In these expressions, viscosity enters indirectly by virtue of the roughness length $z_0$ when considering the law of the wall for a smooth surface:
\begin{align}
  u^+ = \frac{1}{\kappa}\ln(z^+) + C^+ = \frac{1}{\kappa}\ln\left(\frac{z^+}{z_0^+}\right).
\end{align}
This leads to the expression
\begin{align}
	z_0^+ = z_0\frac{u_\star}{\nu} = \exp\left(-\kappa C^+\right),
\end{align}
and for an aerodynamically smooth flow, $z_0^+ = 0.1031$ (using the same values as in eq. \ref{eqn:log-law}). This is known to be the minimal roughness length of a turbulent boundary layer (see e.g. \cite{kraus2008grundlagen}). The roughness length in SI-units 
\begin{align}
	z_0 = z_0^+\frac{\nu}{u_\star}\approx 0.1\frac{\nu}{u_\star}
\end{align}
hence depends on the viscosity of the fluid, which means that---given fixed surface properties---the choice of $z_0$ implies a particular value for the viscosity $\nu$.

When using MOST for the surface fluxes it is assumed that the height of the first grid point lies inside of the logarithmic layer. Again, the limit of low $\RE$ and high resolution requires adaptations to this boundary condition. In the case of very fine resolution, the first grid point might fall into the buffer layer or even the viscous layer of the flow, so the equations of MOST are no longer adequate to calculate the local stress. To avoid that, we follow the recommendation of \cite{kawai2012wall} to use the horizontal velocity from a higher layer $z_{sl}$ to compute the mean stress in the constant flux layer. Furthermore, we adopt the boundary condition suggested by \cite{maronga2020improved} and use the domain averaged velocities:
\begin{align}\label{us_mean}
  u_{*,mean} = \frac{\kappa<u_h(z_{sl})>}{\ln(z_{sl}/z_0)},
\end{align}
where $u_h= \sqrt{u^2+v^2}$ and angle brackets refer to the horizontal average over the entire domain. The mean stress is then used as a boundary condition at the first grid point ($z=z_1$). It is distributed locally to the stresses in x- and y-direction via
\begin{align}
  \overline{u'w'}_0(x,y,z_1) = -u_{*,mean}^2 \frac{u(x,y,z_1)}{\sqrt{<u>^2(z_1)+<v>^2(z_1)}},
\end{align}
and accordingly for $\overline{v'w'}_0$. This way, the domain average of the stress components yield the total stress of eq. \ref{us_mean} ($u_{*,mean}^2 = \sqrt{<\overline{u'w'}_0>^2+<\overline{v'w'}_0>^2}$). As reference height we use $z_{sl}^- \approx 0.1$ (depending on the height of the closest grid level). By using this higher reference height for the boundary condition we solve two problems: First, the reference height is inside of the logarithmic layer, second, we use a velocity from a region where the flow is much better resolved than close to the surface.

\section{Results and Discussion}
\label{results}

In this chapter, we compare the results from the LES to the theoretical bulk parameters (\ref{bulk}) and velocity profiles (\ref{vel_profiles}) and discuss the dynamics of the turbulent flow in LES.

\subsection{Bulk parameters}
\label{bulk}

\begin{table*}
	\centering
	\caption{LES results: $\Delta^+$ is the grid cell size in wall units, $u_\star$ the friction veloctiy, $\alpha_\star$ the angle between geostrophic wind and negative surface shear stress, $\kappa_{LES}$ the Kármán-measure of the logarithmic layer (estimated by linear regression), $C^+$ the intercept (see eq. \ref{eqn:log-law}), and $\delta_{95}$ the height estimated by a linear interpolation of the shear stress reduction to 95\% of its surface value}
		\begin{tabular}{lcccccc}
		  Name & $\Delta^+$ & $u_\star/u_{\star,th}$ & $\alpha_{\star,th}-\alpha_\star$ & $\kappa_{LES}$ & $C^+$ & $\delta_{95}/\delta$ \\
			\hline
Re1\_50	&	59	&	1.021	&	1.5°	&	-	&	-	&	0.71	\\
Re1\_100	&	30	&	1.008	&	1.1°	&	0.52	&	8.3	&	0.66	\\
Re1\_150	&	20	&	1.004	&	1.0°	&	0.48	&	7.0	&	0.64	\\
Re1\_200	&	15	&	1.003	&	0.7°	&	0.46	&	6.5	&	0.62	\\
Re1\_dyn	&	15	&	1.002	&	0.9°	&	0.43	&	5.7	&	0.58	\\
\hline													
Re2\_50	&	$1.5\times10^5$	&	1.008	&	1.1°	&	-	&	-	&	0.77	\\
Re2\_100	&	$7.3\times10^4$	&	1.001	&	0.5°	&	0.53	&	12.2	&	0.66	\\
Re2\_150	&	$4.9\times10^4$	&	1.000	&	0.2°	&	0.47	&	8.9	&	0.60	\\
Re2\_200	&	$3.7\times10^4$	&	1.000	&	0.1°	&	0.44	&	6.8	&	0.59	\\
Re2\_dyn	&	$3.7\times10^4$	&	1.000	&	0.0°	&	0.42	&	5.6	&	0.55	\\
\hline													
Re3\_50	&	$4.5\times10^6$	&	1.007	&	0.7°	&	-	&	-	&	0.76	\\
Re3\_100	&	$2.2\times10^6$	&	1.001	&	0.5°	&	0.53	&	14.3	&	0.65	\\
Re3\_150	&	$1.5\times10^6$	&	1.000	&	0.3°	&	0.47	&	10.3	&	0.60	\\
Re3\_200	&	$1.1\times10^6$	&	1.001	&	0.1°	&	0.44	&	7.7	&	0.59	\\
Re3\_dyn	&	$1.1\times10^6$	&	1.000	&	0.2°	&	0.43	&	6.7	&	0.56	\\
			\hline
		\end{tabular}
	\label{simulation_results}
\end{table*}

% z0
A key parameter governing the turbulent state of the flow is the geostrophic drag: it puts the surface friction in relation to the geostrophic wind and thus illustrates the conversion of mean-flow kinetic energy to turbulence. The ratio of $u_{\star}$ resulting from the simulations to the theoretical value $u_{\star,th}$ are shown in Tbl. \ref{simulation_results}. All values are close to one, while the strongest deviations are observed for the simulations with the coarsest resolutions. A slight dependence on the resolution can be observed for Re1, as $u_\star$ steadily approaches $u_{\star,th}$ with increasing resolution. For Re2 and Re3, all but the coarsest resolution nearly exactly match the theoretical value. The choice of $z_0$ and the geostrophic wind $G$ determine the magnitude of $u_\star$ in a non-trivial way. From a top perspective, the horizontal velocity increases from its geostrophic value to the supergeostrophic maximum and then decreases with decreasing height. The horizontal mean velocity at the grid point closest to $z^-=0.1$ is used to calculate $u_\star$ according to eq. \ref{us_mean}. It is remarkable that the choice of $z_0^+=0.1031$ leads to a value of $u_{\star}$ very close to the prediction by semi-empirical considerations (cf. \cite{spalart1989theoretical}).

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.7\textwidth]{figures_2024/d3y_3Re_hodograph}
  \caption{Geostrophy aligned hodographs of the LES in comparison to the theoretical and laminar hodographs. A $\times$ indicates the first grid point of the LES. At the points marked with $+$, the velocity is used to determine $u_\star$}
  \label{3Re_hodograph}
\end{figure}

%alpha
Another key parameter of the Ekman flow is the surface veering angle $\alpha_\star$ between the directions of the negative surface shear stress and the geostrophic wind. For the LES, the direction of the negative stress is represented by the first derivatives of the velocity components calculated by 3rd order forward finite differences: $\alpha_\star = \tan^{-1}(\langle \partial v/\partial z\rangle_0/\langle \partial u/\partial z\rangle_0)-\tan^{-1}(V_G/U_G)$. Table \ref{simulation_results} shows that nearly all LES yield an $\alpha_\star$ smaller than the theoretical value $\alpha_{\star,th}$. In general, we observe an increase of $\alpha_\star$ with resolution. The finest resolution of $Re2$ and $Re3$ generally align closely with the theoretical direction while for $Re1\_200$ still 0.9° is missing. Fig. \ref{3Re_hodograph} illustrates that this is not caused by poor quality of the simulation since all of the simulations of $Re1$ follow the theoretical hodograph quite closely and even overestimate the angle at the respective height by a little. In contrast to the higher $\RE$, the veering of the wind vector continues in the lower parts of the boundary layer. Hence, for the low-$\RE$ case, the general observation of a higher $\alpha_\star$ with finer resolution is also caused by the approach of the final $\alpha_\star$ with decreasing height of the first grid point. In general, the finer resolved simulations show veering angles that are very close to the theoretic predictions.

\subsection{Logarithmic layer stream-wise velocity}
\label{vel_profiles}

\begin{figure}[ht]
  \centerline{
	\includegraphics[width=\textwidth]{figures_2024/d3y_3Re_profiles_outer.png}
	}
  \caption{Shear-aligned velocity profiles in outer scaling. Solid lines: u-component, dashed lines: v-component. The lowest grid point is marked by $x$}
  \label{velocity_outer}
\end{figure}

% profiles
Figure \ref{velocity_outer} shows the velocity profiles from the LES and the theoretical profiles for each of the $\RE$. In general, the velocity profiles of the LES agree well with the theoretical profiles: The course of the theoretical Ekman layer is matched and the simulations exhibit a logarithmic layer for the U-component. The lowest points of the LES of the higher $\RE$ fall into the logarithmic layer. For Re1, the lowest grid point falls into the buffer layer, which is visible as the curved course of the U-component below the logarithmic layer in the theoretical profile. The best resolved simulation with Deardorff-closure even seems to follow the course of the upper part of the buffer layer, but with a resolution of $\Delta^+=15$ this is a coincidence caused by a well-known S-shape of velocity profiles close to rigid walls \cite{brasseur2010designing}. This log-layer mismatch arises from a competition between the scales $u_\star$ and $z$ and other velocity and length scales introduced by the discretization of the dynamical system \citep{mason1992stochastic,brasseur2010designing}. In other words, at the lower boundary, the relevant eddies are too small to be resolved by the grid and their contribution to the flow has to be modeled. Also the vertical component is massively restricted by the non-permeability of the wall, known as blocking effect. On the contrary, the SGS-closure assumes isotropic turbulence, which is not the case on the grid scale near the wall.

\begin{figure}[ht]
  \centerline{
	\includegraphics[width=\textwidth]{figures_2024/d3y_3Re_kappa}
}
  \caption{K\'arm\'an measure $\kappa$ in the logarithmic region and above for different Reynolds numbers and resolutions}
  \label{3Re_kappa}
\end{figure}

% kappa
In the logarithmic region, the profile of the u-component should follow the logarithmic law of eq. \ref{eqn:log-law} with the K\'arm\'an measure $\kappa = (z^+\partial U^+/\partial z^+)^{-1}=const.$. In the low-Reynolds-number case, the viscous sublayer represents about $0.5\%$ of the boundary layer, while this layer is not visible for the high-Reynolds-number cases, where the share drops to $~10^{-5}-10^{-7}$. Above the viscous sublayer, the theoretical profile shows a near-constant value for $\kappa$ up to $z^-\approx0.1$ for the case Re1 and up to $z^-\approx0.12$ for the cases Re2 and Re3. In order to estimate $\kappa_{LES}$ from the simulations, we perform a linear regression of the mean velocity in x-direction and the logarithm of height between the seventh grid point and the grid point corresponding to the height $z^-=0.1$ for $Re1\_X$ and $z^-=0.12$ for $Re2\_X$ and $Re3\_X$. We consider only points above the 7th grid point following arguments of \cite{maronga2020improved}, that in PALM, the mean velocity profiles follow MOST above the seventh grid point. The number of values for each regression is 6, 12, and 18 for $ReX\_100$, $ReX\_150$ and $ReX\_200$, respectively.  We do not compute $\kappa$ for $ReX\_50$ since there are only six grid points below $z^-\approx 0.12$. The resulting $\kappa_{LES}$ for the other simulations are shown in Tbl. \ref{simulation_results}. The coefficient of determination is above 0.99 for all fits. $\kappa_{LES}$ is decreasing with finer resolution for all $\RE$. While the values for $ReX\_100$ are rather high with a value around $0.53$, finer resolution leads to values closer to measurements around $0.46$ for $Re1\_200$ and around $0.44$ for $Re2\_200$ and $Re3\_200$. The dynamic subgrid closure yields lower values for $\kappa_{LES}$: in $Re1\_dyn$ we even see $\kappa_{LES}=0.39$, while $Re2\_dyn$ and $Re3\_dyn$ yield $0.42$ and $0.43$, respectively. Figure \ref{3Re_kappa} shows the K\'arm\'an measure in the logarithmic layer. For the simulations $ReX\_50$, no constant $\kappa$ can be observed. An increase in resolution leads to a profile of the K\'arm\'an measure closer to the theoretical curve for all cases. Close to the bottom, the K\'arm\'an measure is heavily influenced by the proximity of the wall and rather a function of the vertical index than of the physical distance from the wall. In accordance with the observations of \cite{maronga2014monin} and \cite{maronga2017formulation}, the kinks in the K\'arm\'an measure diminish around the seventh grid point for all simulations. Above, the curve smoothens and approaches the expected value of $\kappa$ (at least for the finer resolutions).

\begin{figure}[ht]
  \centerline{
	\includegraphics[width=\textwidth]{figures_2024/d3y_2pc_ww_3Re_0.079.png}
}
  \caption{Two-point correlation of w-component in y-direction at $z^-\approx 0.08$. The horizontal line is at $B_{ww}=0.3$}
  \label{2pc_008}
\end{figure}

As expected for an LES with isotropic grid spacing, the resolution near the surface is a critical part of the simulation. We can support our above observations by taking a look at the two-point correlation in the critical layers of the simulations. The two point correlation is defined as 
\begin{align}
  B_{xx}\left(x^*\right) = \langle u\left(x-x^*\right)u\left(x\right)\rangle/\sigma_u^2.
\end{align}
As suggested by \cite{wurps2020grid}, in an isotropic grid with periodic horizontal boundary conditions the turbulent structures of the velocity's w-component in y-direction can serve as an indicator how well the flow is resolved. Figure \ref{2pc_008} shows the two-point correlation $B_{ww}(y^*)$ at a height of $z^-\approx 0.08$, which is right inside of the logarithmic layer. We define the average size of a structure in the flow $\sigma$ as the distance where the two-point correlation drops to 0.3 (horizontal line in figure). The number of cells by which the average structure is resolved is equal to $\sigma/\Delta$. All $\RE$ show a similar behaviour: the average size of the structure decreases with finer resolution and seems to approach a limit. The structure size of ReX\_150 and ReX\_200 are very close together, which indicates beginning convergence. Furthermore, the number of cells by which the structures are resolved exceed 4 from the resolution of ReX\_150 on. This corresponds to the resolution from which on a logarithmic layer with a constant $\kappa$ can be seen. It also fits to the rule of thumb given in \cite{wurps2020grid}, that in a sufficiently resolved part of an LES the average structures of the w-component in y-direction should be resolved by at least 4 cells. The dynamic closure shows structures that are slightly smaller than the structures of $ReX\_200$, which is caused by a tendency to lower eddy-viscosities and, hence, a weaker coupling of neighboring grid cells.

\begin{figure}[ht]
  \centerline{
	\includegraphics[width=\textwidth]{figures_2024/d3y_2pc_ww_3Re_4.png}
}
  \caption{Two-point correlation of w-component in y-direction at the 8th grid point above the surface}
  \label{2pc_4}
\end{figure}

Figure \ref{2pc_4} shows $B_{ww}(y^*)$ at the 8th grid point above the bottom. Therefore, the corresponding heights differ between the simulations, that is $z^-=0.16, 0.08, 0.053, 0.04, 0.04$ for $ReX\_50$, $ReX\_100$, $ReX\_150$, $ReX\_200$, $ReX\_dyn$, respectively. For $Re1$, the curves almost collapse perfectly while for $Re2$ and $Re3$ we see slightly larger structures for $ReX\_50$ and slightly smaller structures for $ReX\_dyn$. This means that even 8 points above the lower boundary the smallest size of the turbulent structures rather depends on the grid cell size than on the actual height above the ground. According to the above findings and \cite{maronga2014monin}, the flow should start to be well resolved from here on (above the seventh grid point). And indeed, the number of resolving cells is very close to 4 for $Re1$ and between 3 and 4 for $Re2$ and $Re3$, which is close to the recommended 4 cells.

%
\subsection{Logarithmic layer span-wise velocity}

\begin{figure}[ht]
  \centerline{
	\includegraphics[width=\textwidth]{figures_2024/d3y_3Re_v_outer_lin.png}
}
  \caption{Span-wise velocity component in outer scaling}
  \label{3Re_v_lin}
\end{figure}

Figure \ref{3Re_v_lin} shows the span-wise velocity component in the lower part of the boundary layer. \todo{if we keep $V(z=\Delta)$ in the conclusion, add this here}. In the lowest part of $Re1$ ($z^-\lessapprox 0.05$), a part of the viscous layer is visible, where the velocity is slightly underestimated by the the LES. Above, the simulations $Re1\_200$ and $Re1\_dyn$ are very close to the theoretical curve. Lower resolution leads to an overestimation of the velocity in the logarithmic layer, which is consistent with a lack of large eddies that translates in a lack of mixing and thus a reduced turbulence level and increased velocity gradients. Around $z^-=0.25$, while blending into the Ekman layer, the curves cross each other and the coarse simulations begin to underestimate the velocity. For the higher $Re$ all simulations overestimate the velocity in the logarithmic layer, while coarser resolution leads to a higher velocity and finer resolution leads to a lower velocity and better agreement with the theoretical curve. Again, we see a crossing of the curves, between $z^-=0.2$ and $z^-=0.25$. A possible explanation of the fast increase and the steeper course of the coarser simulations is that the layers are coupled via less cells.

\subsection{Ekman layer}

Above the logarithmic layer, the Ekman layer follows, characterized by a change of wind direction. The course of the wind velocity vector is visualized by hodographs, as plotted in Fig. \ref{3Re_hodograph}. The hodograph of Re1\_X is followed quite closely by all resolutions. Hence, all simulations of the low-$\RE$ case---even Re1\_50---are resolved sufficiently to closely capture the course of the wind vector in the Ekman layer. The higher $\RE$ behave differently in the sense that the hodographs lie inside of the theoretical hodograph. An increased resolution ensures that at least the lowest grid point approaches the hodograph while the course of the hodograph's upper right part still does not reach the theoretical curve.

In Fig. \ref{3Re_hodograph}, the cross indicates the lowest grid point and the plus indicates the height where the velocity for the boundary condition is taken from. To avoid taking a velocity from the first layer of the simulation, where the turbulent flow is poorly resolved, we took the horizontal velocity near $z^-=0.1$, where the mean velocity already veered away from the direction of the surface stress by around one third of $\alpha$. However, the veering does not seem to influence the resulting bulk stress $u_\star$ at the bottom: all but the coarsest resolutions yield a $u_\star$ very close to the theoretic reference.

\begin{figure}[ht]
  \centerline{
	\includegraphics[width=\textwidth]{figures_2024/d3y_3Re_Ekman_lin.png}
}
  \caption{Shear-aligned velocity deficit in outer scaling. Solid lines: U-component, dashed lines: V-component}
  \label{3Re_Ekman}
\end{figure}

Figure \ref{3Re_Ekman} shows the shear-aligned velocity deficit $U^--U^-_G$ (v-component accordingly) of the whole boundary layer. The strongest deviations from the theoretical profiles can be seen for the coarsest LES near the super-geostrophic maximum of the U-component. Finer resolution leads to a very good agreement with the extent of the maximum, which also holds for the higher $\RE$. The super-geostrophic maximum is a consequence of vertical redistribution of momentum across the ABL; less redistribution leads to a less pronounced maximum. At the same time, of course, one would expect an increased level of turbulence to cause a flatter hodograph. BUT, the surface veering is set rather by the interplay of MOST and the prescribed roughness while the profile in the Ekman layer is a resolved quantity. However, for higher Reynolds numbers $Re2$ and $Re3$, the maximum velocity in x-direction is still not reached by the LES and the approach of the geostrophic wind in the upper part of the boundary layer takes place at lower heights than in theory. This is even more obvious for the V-component, where the maximum value is at a significantly lower height than in the theoretical curve. In general, a finer resolution leads to a lower location of the maximum. The maximum at lower heights might also explain the too high velocities in the logarithmic layer for the higher $Re$.

\section{Conclusions}
\label{conclusion}

We studied truly neutral Ekman boundary layers---absent of capping inversions, non-stationarity, and surface heterogeneity. For the first time, an LES of the Ekman layer is \emph{a posteriori} validated against a DNS for an identical configuration. Despite the common assumption that viscosity as a parameter drops out in LES as a consequence of the turbulence closure, we considered a range of scale separations (Reynolds numbers) and found that for an exact representation of the surface turning and wall friction as well as the wind-turning profile, viscous effects need consideration. We derive an analytical relation demonstrating that --- given fixed surface properties --- the specification of $z_0$ implicitly defines a viscosity. The LES suffer from the dilemma, that in Ekman flow, some aspects of crucial importance happen on relatively small scales, such as the rotation-surface interaction which is confined to the inner layer.

With the wind-profile formulation developed in Part 1 of this work, we have a reference based on first principles for intermediate-Re simulations, and a quasi-reference for the higher-Re simulations. While we acknowledge that both the LES and the wind-profile model suffer from assumptions for high Reynolds number, their quantitative agreement across a wide range of scale separations hints towards the consistency. In particular, we understand that the grid convergence of LES towards the theoretical profiles underpins the inviscid scaling hypotheses in the development of the theory underlying part 1 of this paper.

We simulated three different Reynolds numbers while having DNS results for the lowest $\RE$. For the low $\RE$, viscous forces have a significant contribution to the balance of forces on the grid scale of the LES, hence we adapted the LES code to consider the fluid's viscosity in addition to the modeled eddy-viscosity. We did not introduce a roughness length $z_0$ at the lower boundary as additional parameter but deduced $z_0$ according to the law of the wall and the shear velocity expected according to the semi-empirical law by \cite{spalart1989theoretical}. The interplay of geostrophic wind, shear velocity, and roughness length in the simulation showed remarkable consistency, which supports the value of the adapted boundary condition at the bottom \citep{maronga2017formulation}. The dependence of the LES solution on the grid cell size was investigated through a comparison of four different resolutions. The setups of all $\RE$ use similar grid sizes in terms of the outer scale ($\Delta^-=\Delta/\delta=const.$) but different grid sizes in terms of the inner scale ($\Delta^+=\delta^+\cdot const. = Re_\tau\cdot const.$). This means that from an inner scale perspective the high Reynolds numbers were much less well resolved.

The convergence towards the theoretical profile expressed itself in different major aspects of the flow. To reach the expected total rotation $\alpha_\star$, a sufficient resolution was necessary. Furthermore, a sufficient vertical resolution was needed to simulate a logarithmic layer with a constant K\'arm\'an measure in some vertical extent.

A resolution of 150 grid levels below $\delta$ resolves well the boundary layer across all $\RE$, in agreement with the findings of \cite{wurps2020grid}. Their study demonstrated successful resolution of the neutral simulation with more than 100 grid levels within the boundary layer $\delta_{95}$. The ratio $\delta_{95}/\delta$ is approximately $2/3$ (gradually decreasing with Reynolds number). Therefore, having 150 grid levels within $\delta$ roughly corresponds to around 100 grid levels within $\delta_{95}$.

In summary, we synthesize some technical recommendations for the correct simulation of Ekman layer dynamics on a process level:
\begin{itemize}
\item The roughness parameter ($z_0$) essentially defines a Reynolds number of the LES Problem
\item $V|_{z=\Delta} < u_\star$ for an accurate matching of the hodograph and proper quantification of $\alpha_\star$. \todo{where is this in the results?}
\item There should exist a logarithmic layer, or explicit consideration of viscous interaction with the surface; otherwise the assumptions of the surface closure (MOST / dynamic wall model) will fail.
\item The parameters $\alpha_\star$ and $u_\star$ characterizing the bulk Ekman dynamics (across the vertical extent of the boundary layer) are matched by the LES if---in a three-dimensionally isotropic grid---more than 150 grid points are used in the vertical direction.
\item For very high resolution or relatively small Reynolds number, the viscous friction needs to be taken into account as the modeled eddy viscosity may locally drop to zero.
\end{itemize}

When these considerations are taken into account, LES becomes possible at uncommonly low Reynolds number and resolution, which allows a quantitative comparison to state-of-the art DNS.

Some resolution and best-practice constraints developed in this work are strong, in some cases even prohibitive. They result from the externality of the flow, i.e. the presence of non-turbulent fluid aloft which leads to a duality of scales (cf. section \ref{universal_profile}): both the inner and outer dynamics need to be visible to the resolved LES scales, at least to some extent. Despite these relatively strong constraints, we appreciate the capability of LES which, for example in case Re3, can appropriately match both $u_\star$, $\alpha_\star$ and the hodograph with the first grid point located at about $10^6$ wall units. This means, there is a gap in resolution in comparison to a true DNS of five to six orders of magnitude per direction; illustrating that a DNS at this scale would be prohibitive and will remain so for a foreseeable time. From this inner, or small-scale perspective, a requirement of few hundred grid points across the boundary layers is not a lot.

We did not see a second logarithmic layer as \cite{jiang2018large} neither in the theoretical formulation nor in our simulations. A reason might be their introduction of the additional parameter $z_0$. \todo{keep this paragraph? After our discussion about the change of Re when introducing a roughness length I don't believe this explanation anymore...}

The theoretical formulation of mean velocity profiles within the turbulent Ekman layer can serve as benchmark for model frameworks such as LES. These reference profiles offer a deeper analysis beyond the conventional assessment of the logarithmic increase in wind speed within the Prandtl layer. Comparing simulation results to the expected bulk parameters $u_\star$ and $\alpha_\star$ and to the hodograph and conducting a detailed evaluation of the wind speed profiles can yield valuable insights into the correct interplay among the model's grid resolution, turbulence closure, and boundary conditions. In the future, expanding the theoretical profiles to include additional aspects, such as temperature stratification, could potentially provide an even more valuable reference.

\section{Acknowledgements}

The simulations were performed at the HPC Cluster EDDY, located at the University of Oldenburg (Germany) and funded by the Federal Ministry for Economic Affairs and Energy (Bundesministeriums für Wirtschaft und Energie) under grant number 0324005.

\bibliographystyle{plainnat}
\bibliography{papers.bib,ansorge.bib}

\end{document}

### 2024-06-27 plot/EkmanProfileClass.py

@@ -9,7 +9,7 @@ def build_grid(trgt,thick=1) :
     dyloc=1             #grid resolution at wall
     ny=1
     yloc=dyloc
-    while yloc < thick*trgt:
+    while yloc-dyloc < thick*trgt:
         dyloc*=vert_stretch 
         yloc+=dyloc
         ny+=1
@@ -73,9 +73,12 @@ class EkmanUniversalClass:
 
         self.LIMIT_INNER=50
         self.LIMIT_OUTER=0.10
-        
-        self.KAPPA=0.416 # k
-        self.C = 5.4605 # C
+        # Old Values from estimate routine (see below) 
+        # self.KAPPA= 0.446 #0.416 # k
+        # self.C =    5.98 #5.4605 # C
+        # Values from Jonathan (20230601; ReD=1000; Fit 38+ -> 0.12-)
+        self.KAPPA= 0.416 #0.416 # k
+        self.C =    5.4605 #5.98 # C
         self.A = Complex(4.79823,-5.56645)
         self.C5= Complex(-57.7728,19.5046)
         self.initialized=True
@@ -152,7 +155,7 @@ class EkmanUniversalClass:
         Ar=self.A.r 
         k=self.KAPPA 
         C5=self.C5.r
-        k=0.416
+        #k=0.416
         z=10
 
         #Ai=Ai+0.285
@@ -192,7 +195,7 @@ class EkmanUniversalClass:
 
     def profile_ekman(self,ym,us_loc):
         z=1/us_loc
-        
+        # z~ = (z-z_r) / D_E = (z  + 0.12 \delta ) / ( 0.66 * 2*pi *delta)
         c_argz=0.66*2.*np.pi 
         argz=    c_argz*(ym +0.12) #0.66*(2.*np.pi*(ym+0.12)) 
         dampA =  8.4*us_loc        #- 0./re #np.exp(zdum) * ( dU_mtc*np.cos(zdum) + dW_mtc*np.sin(zdum) )

### 2024-06-27 plot/plot.py

@@ -1,4 +1,4 @@
-#!/opt/local/bin/python
+#!/usr/bin/python3
 import my_pylib as mp 
 import numpy as np
 import scipy.interpolate as spip
@@ -23,14 +23,15 @@ plot_evisc=False
 plot_profiles=False
 plot_visc_outer=False
 print_table=False
-plot_outer_log=False
-plot_total_rot=False
+plot_outer_log=True
 plot_convergence=False
 plot_re1600=False
 plot_les_comp=False
 plot_applications=False
 test_inner_streamwise=False
 plot_profile_comparison=True
+plot_vandriest=False
+plot_total_rot=False
 
 colors = {400 : 'gray',
           500 : 'pink',
@@ -46,7 +47,7 @@ colors = {400 : 'gray',
           300000:'black',
           1000000:'black'} 
 
-base='/Users/zrick/WORK/research_projects/SIM_NEUTRAL/netcdf/' 
+base='/home/cedria87/WORK/RESEARCH/SIM_NEUTRAL/netcdf/' 
 
 files = { 400: base+'avg_flw_ri0.0_re400.0_1024x384x1024.nc',
           500: base+'avg_flw_ri0.0_re500.0_2048x192x2048.nc',
@@ -56,13 +57,56 @@ files = { 400: base+'avg_flw_ri0.0_re400.0_1024x384x1024.nc',
           1301:base+'avg_flw_ri0.0_re1300.0_co0_2560x640x5120.nc',
           1600:base+'avg_flw_ri0.0_re1600.0_co0_3840x960x7680.nc' } 
 
+
+def etling_us(z0,zp,al):
+    # universal parameter lz = lambda_z * 1/sqrt(2)
+    lz=np.sqrt(2)*sc.KAPPA / np.log(zp/z0)
+    us = lz*np.sin(np.pi/4.-al)
+    
+    # checking solution
+    #print('SOLUTION: {1:8.6f}, {2:8.6f} {3:8.6f}'.format(0,us,cos_al,al/np.pi*180) )
+    #print(us/kappa*np.log(zp/z0)*np.cos(al_0))
+    #print(G*(1-np.sqrt(2)*np.sin(al_0)*np.cos(pi4-al_0)))
[…]

### 2024-06-27 plot/test_df.py

@@ -1,21 +1,0 @@
-#!/opt/local/bin/python
-import numpy as np
-import my_pylib as mp 
-us_loc=0.0619
-dp_loc=479
-
-yp=np.arange(0,10.5,0.25) 
-
-c_off=40/(dp_loc*us_loc)
-c1=26/(dp_loc*us_loc)
-dir_fit = c1*np.log(yp)**2+c_off
-
-a_off=0 #2.84/(dp_loc*us_loc)
-a_sqr=24.66/(dp_loc*us_loc)
-a_lin=10.24/(dp_loc*us_loc) 
-dir_fit2= a_off + a_sqr*np.sqrt(yp) + a_lin*yp
-
-
-for i in range(11):
-    ix=mp.geti(yp,i)
-    print(yp[ix],dir_fit[ix],dir_fit2[ix]) 

### 2024-06-28 write2/manuscript_part2_2024.tex
worked on chapter 3

@@
-\documentclass[a4paper,11pt]{article}
+\documentclass[draft,a4paper,11pt]{article}
 
@@
 
-In this chapter, we introduce the universal velocity profile for the turbulent Ekman layer developed in part I. 
+In this chapter, we introduce the velocity profile for the turbulent Ekman layer developed in part I. 
 %Definitions
@@
 %The mean velocity profiles for all $\RE$ exhibit a dual scaling nature: in proximity to the lower boundary, the velocity profiles collapse when scaled using
-In proximity to the lower boundary, the boundary layer scales in viscous (inner) units, denoted by the index ``$+$'' ($z^+ = zu_\star/\nu$, $U^+=U/u_\star$, where $U$ is the velocity component in x-direction).
-From the top perspective, the profiles converge when scaled using outer units, denoted by the index ``$-$'' ($z^-=z/\delta$, $U^-=U/G$, with $\delta=u_\star/f$ representing the boundary layer height).
+In proximity to the lower boundary, the boundary layer scales in viscous (inner) units, denoted by the index ``$+$'' ($z^+ = zu_\star/\nu$, $U^+=U/u_\star$, where $U$ is the velocity component in x-direction and $u_\star = \sqrt{\nu \p_z U}$ the friction velocity).
+From the top perspective, the profiles converge when scaled using outer units, denoted by the index ``$-$'' ($z^-=z/\delta$, $U^-=U/G$, where $\delta=u_\star/f$ is the boundary-layer depth scale, $f$ the Coriolis parameter and $G$ the geostrophic wind speed).
 At the lower boundary, the x-axis of the inner units aligns with the shear stress, while the x-axis of the outer units aligns with the geostrophic wind.
 The angle between both axes is denoted as $\alpha_\star$, the surface veer of the wind across the boundary layer.
+%
+Profiles of the velocity components are studied in three different layers reflecting the change of dominant balance when moving away from the surface. These are
+(i) the Ekman layer denoted $(\cdot)_\text{Ek}$ (Sec.~\ref{sec:ekmanlayer}), 
+(ii) the viscous sub-layer, denoted by an index $(\cdot)_\text{visc}$, 
+and (iii) the logarithmic layer denoted by $(\cdot)_\text{log}$
+We match $U_\text{visc}$ and $U_\text{log}$ according to their respective formulations and combine them to $U_\text{inner}$ (Sec.~\ref{sec:innerlayer}).
+Inner and outer profiles are then matched by a weighting transfer function based on the error function. 
+%The span-wise component of the velocity is separated into two layers, namely the inner layer $V_\text{inner}$, and the Ekman layer $V_{EK}$.
 
@@
 
-The surface stress is linked to the geostrophic wind via the geostrophic drag $Z\equiv u_\star/G$ and the angle between the shear stress and the geostrophic wind $\alpha_\star$.
-These two parameters form the basis of the boundary-layer scaling and are  estimated using the semi-empirical drag-law introduced by \cite{spalart1989theoretical}, which describes them as functions of only the Reynolds number:
+The geostrophic drag, i.e.
+normalized friction $Z\equiv u_\star/G$, and $\alpha_\star$ form the basis of the boundary-layer scaling, and they are estimated using the semi-empirical drag-law introduced by \cite{spalart1989theoretical}, which describes them as functions of only the Reynolds number:
 \begin{subequations}\label{eqn:drag}
@@
 
-Profiles of the velocity components are studied in three different layers reflecting the change of dominant balance when moving away from the surface.
-These are (i) the viscous sub-layer, denoted by an index $_\text{visc}$ (ii) the logarithmic layer denoted by $_\text{log}$, and (iii) the Ekman layer denoted $_\text{Ek}$.
-We match $U_{visc}$ and $U_{log}$ according to their respective formulations and combine them to $U_{inner}$.
-The span-wise component of the velocity is separated into two layers, namely the inner layer $V_{inner}$, and the Ekman layer $V_{EK}$.
-
 \subsection{Ekman layer}
-
+\label{sec:ekmanlayer}
 The outer layer of the ABL is characterized by a triadic balance between turbulent flux, pressure gradient and the Coriolis; the vertical change of the Coriolis force causes a pronounced height dependence of the wind direction resulting in the Ekman spiral \citep{ekman:AMA1905}.
@@
 \subsection{Inner and viscous layers}
+\label{sec:innerlayer}
 
@@
 
-\subsection{Blending of inner and outer profiles, synopsis}
+\subsection{Transition from inner to outer profiles, synopsis}
 
@@
 
-An incompressible, turbulent Ekman flow over a hydrodynamically smooth surface is simulated for three different Reynolds numbers $Re_D = 10^3;1.5\times 10^5;10^6$, hereafter $Re1$, $Re2$, and $Re3$, respectively.
-The input parameters are given in table \ref{simulation_parameters}.
-
 \begin{table*}
@@
 \end{table*}
-
+An incompressible, turbulent Ekman flow over a hydrodynamically smooth surface is simulated using the Parallel Large Eddy Simulation model (PALM, \citealp{maronga2020overview}).
+We study three different Reynolds numbers Re$_D = 10^3;1.5\times 10^5;10^6$, hereafter Re1, Re2, and Re3, respectively (Tab.~\ref{simulation_parameters}).
+%
 The domain is rotating around the z-axis with an angular velocity corresponding to the Coriolis parameter $f=10^{-4}s^{-1}$.
-The stratification of the flow is truly neutral, i.e., the potential temperature is constant for the whole domain.
-Nevertheless, the boundary layer height is not growing infinitely but results from a balance between shear production and rotational suppression of turbulence.
-At the upper boundary, sufficiently far aloft of the boundary-layer height, a no-penetration boundary condition is used and the horizontal components of the wind are geostrophic, which is a Dirichlet-type boundary condition.
+The stratification of the flow is truly neutral, i.e., the potential temperature is constant across the whole domain.
+A constant boundary layer height on the order of $\delta=u_\star/f$ forms due to the balance between shear production and rotational suppression of turbulence.
+At the upper boundary, the flow is non-turbulent.
+Hence, a no-penetration boundary condition is used for the vertical velocity component while the horizontal components are prescribed according to geostrophic balance--a Dirichlet-type boundary condition.
 At the bottom, a constant-flux layer is assumed and Monin--Obukhov similarity theory (MOST) is used to calculate the surface momentum fluxes.
-The Navier-Stokes equations are integrated using a 3rd-order low-storage Runge-Kutta method.
-For scalar advection a 5th-order Wicker--Skamarock scheme is employed.
+The Navier-Stokes equations are integrated using a 3$^\text{rd}$-order low-storage Runge--Kutta method.
+For scalar advection a 5$^\text{th}$-order Wicker--Skamarock scheme is employed.
 The Poisson equation is solved using a direct fast Fourier transform (FFT).
-In LES, the turbulent transport on the subgrid scale (SGS) needs to be modeled by a dedicated model, the SGS model.
-We use two different SGS models to check whether the subgrid closure significantly influences the LES solution: a 1.5-order closure after \cite{deardorff1980stratocumulus} and a dynamic closure after \cite{heinz2008realizability}.
+In LES, the turbulent transport on the subgrid scale (SGS) needs to be modeled by an SGS closure model.
+We use two kinds of SGS closures to assess their impact on the LES solution: a 1.5-order closure after \cite{deardorff1980stratocumulus} and a dynamic closure after \cite{heinz2008realizability}.
 For most of the simulations, the 1.5-order closure is used, since the dynamic closure needs more computational resources, but the simulations with $\Delta^-=200^{-1}$ are repeated using the dynamic closure.
-A description of the LES model is given by \cite{maronga2020overview}.
 
@@
 In part I of this publication, the DNS of $Re_D\leq1\,600$ is carried out for a horizontal domain size of $(0.54\Lambda_{Ro})^2$ - $(1.08\Lambda_{Ro})^2$, where $\Lambda_{Ro}=G/f$ is the Rossby radius.
-For these Reynolds numbers, $u_\star/G$ is around 0.05 so $L_x \approx 10 \delta$ However, $u_\star/G$ decreases with increasing Reynolds number and is around 0.02 for $Re_D=10^6$.
+For these Reynolds numbers, $Z\approx 0.05$, such that $L_x \approx 10 \delta$.
+However, $Z$ decreases logarithmically with increasing Reynolds number and it is $Z\approx 0.02$ for $Re_D=10^6$.
 A domain size of half the Rossby radius would then extend to $L_x\approx 25\delta$.
-Such a large domain would imply immense computational costs.
+Such a large domain would imply immense computational cost.
 \cite{spalart2008direct} used a horizontal domain of $L_x = 2\delta$, arguing that this length allows the resolution of the largest outer-layer eddies according to \cite{csanady1967resistance}.
@@
 Thus, molecular friction is not considered directly, but only by virtue of a turbulence model linking the resolved and dissipative scales.
-In their seminal works on the spectral energy transfer in homogeneous isotropic turbulence, Kolmogorov and Obukhov (\cite{kolmogorov1941dissipation},\cite{obukhov1941distribution}) showed that the energy transfer rate across the spectrum is in fact a constant.
+In their seminal works on the spectral energy transfer in homogeneous isotropic turbulence, \cite{kolmogorov1941dissipation} and \cite{obukhov1941distribution} showed that the energy transfer rate across the spectrum is in fact constant provided that both dissipation and production terms can be neglected.
 This implies that the transfer rate across the cut-off scale in LES does not depend on the magnitude of the viscous scale, presupposed that (i) the cut-off scale of the LES is well within the inertial range and (ii) the LES turbulence is approximately isotropic and homogeneous at the smallest resolved scales.
@@
 When this is the case, $\nu$ cannot be ignored anymore, and we let $K_m = c_0 \Delta \sqrt{e} + \nu$.
-The governing equation of the velocity components in PALM reads
+Hence, (absent stratification) the governing equation of for momentum in our version of PALM is
 \begin{align}\label{gov_palm}
-  \frac{\partial u_i}{\partial t} = - \frac{\partial u_i u_j}{\partial x_j} -\epsilon_{ijk}f_ju_k + \epsilon_{i3j} f_3 u_{g,j} - \frac{1}{\rho_0}\frac{\partial\pi^*}{\partial x_i} + \frac{\partial}{\partial x_j}\left( K_m\left(\frac{\partial u_i}{\partial x_j} + \frac{\partial u_j}{\partial x_i} \right)\right)
+  \frac{\partial u_i}{\partial t} = - \frac{\partial u_i u_j}{\partial x_j} -\epsilon_{ijk}f_ju_k + \epsilon_{i3j} f_3 u_{g,j} - \frac{1}{\rho_0}\frac{\partial\pi^*}{\partial x_i} + \frac{\partial}{\partial x_j}\left( K_m\left(\frac{\partial u_i}{\partial x_j} + \frac{\partial u_j}{\partial x_i} \right)\right).
 \end{align}
-for a neutrally stratified flow.
-In the limit of either very well resolved simulations or very low $\RE$ (or both), $K_m \rightarrow \nu$, so the last term on the right hand side of eq.
-(\ref{gov_palm}) becomes $\nu\frac{\partial^2u_i}{\partial x_j^2}$, which complies with the Navier-Stokes equations for an incompressible fluid.
+In the limit of either well-resolved simulations or low $\RE$ (or both), $K_m \rightarrow \nu$, so the last term on the right hand side of eq.
+(\ref{gov_palm}) becomes $\nu\frac{\partial^2u_i}{\partial x_j^2}$, which complies with the Navier-Stokes equations for an incompressible fluid underlying our reference solution from DNS.
 
@@
 
-In contrast to the interior closure of the LES where the direct effect of $\nu$ is of small relevance compared to the eddy viscosity---at least for high $\RE$---, this is not true for the viscid effects at the bottom boundary.
-Here, a constant flux layer is usually assumed below the first grid point and the Monin--Obukhov similarity theory is used to compute the friction velocity and the stresses at the first half grid point (where the horizontal velocities are located on the Arakawa-C staggered grid):
+In contrast to the interior closure of the LES where the direct effect of $\nu$ is of small relevance compared to the eddy viscosity---at least for high $\RE$---, this is not true for the viscid effects at the bottom boundary:  A constant flux layer is assumed in between the first grid point and the lowermost half point (on the Arakawa-C staggered grid) to estimate friction velocity and surface stresses by MOST:
 \begin{align}\label{most}
   u_\star = \frac{\kappa(U^2+V^2)^{0.5}}{\ln\left(z/z_0\right)},\\
-  -\overline{u''w''}_0 = \frac{\kappa Uu_\star}{\ln\left(z/z_0\right)}.
+  -\overline{u''w''}_0 = \frac{\kappa Uu_\star}{\ln\left(z/z_0\right)}, 
 \end{align}
+where double dashes refer to the unresolved, i.e.
+SGS-parameterized, fluctuations.
 In these expressions, viscosity enters indirectly by virtue of the roughness length $z_0$ when considering the law of the wall for a smooth surface:
@@
 \end{align}
-This leads to the expression
-\begin{align}
-	z_0^+ = z_0\frac{u_\star}{\nu} = \exp\left(-\kappa C^+\right),
-\end{align}
-and for an aerodynamically smooth flow, $z_0^+ = 0.1031$ (using the same values as in eq.
-\ref{eqn:log-law}).
-This is known to be the minimal roughness length of a turbulent boundary layer (see e.g.
-\cite{kraus2008grundlagen}).
-The roughness length in SI-units
+%This leads to the expression
+%\begin{align}
+%	z_0^+ = z_0\frac{u_\star}{\nu} = \exp\left(-\kappa C^+\right),
+%\end{align}
+For aerodynamically smooth flow, it is $z_0^+ = 0.1031$ based on Eq.~\ref{eqn:log-law}.
+(This is known to be the minimal roughness length of a turbulent boundary layer; see e.g.
+\citealp{kraus2008grundlagen}.) Hence,
 \begin{align}
@@
 \end{align}
-hence depends on the viscosity of the fluid, which means that---given fixed surface properties---the choice of $z_0$ implies a particular value for the viscosity $\nu$.
+depends on the viscosity $\nu$ of the fluid. In other words, given fixed surface properties, a choice for $z_0$ implies a particular value for the viscosity $\nu$.
 
@@
 In the case of very fine resolution, the first grid point might fall into the buffer layer or even the viscous layer of the flow, so the equations of MOST are no longer adequate to calculate the local stress.
-To avoid that, we follow the recommendation of \cite{kawai2012wall} to use the horizontal velocity from a higher layer $z_{sl}$ to compute the mean stress in the constant flux layer.
-Furthermore, we adopt the boundary condition suggested by \cite{maronga2020improved} and use the domain averaged velocities:
+To avoid that, we follow the recommendation of \cite{kawai2012wall} and use the horizontal velocity from a higher layer $z_{sl}$ to compute the mean stress in the constant flux layer.
+Further, we adopt the boundary condition suggested by \cite{maronga2020improved} and use domain averaged velocities for the estimation $u_\star$:
 \begin{align}\label{us_mean}
-  u_{*,mean} = \frac{\kappa<u_h(z_{sl})>}{\ln(z_{sl}/z_0)},
+  u_\star \approx u_{*,mean} = \frac{\kappa<u_h(z_{sl})>}{\ln(z_{sl}/z_0)},
 \end{align}
@@
 The mean stress is then used as a boundary condition at the first grid point ($z=z_1$).
-It is distributed locally to the stresses in x- and y-direction via
+It is distributed locally to the unresolved stresses in x- and y-direction via
 \begin{align}
-  \overline{u'w'}_0(x,y,z_1) = -u_{*,mean}^2 \frac{u(x,y,z_1)}{\sqrt{<u>^2(z_1)+<v>^2(z_1)}},
+  \overline{u''w''}_0(x,y,z_1) = -u_{*,mean}^2 \frac{u(x,y,z_1)}{\sqrt{<u>^2(z_1)+<v>^2(z_1)}},
 \end{align}
-and accordingly for $\overline{v'w'}_0$.
+and accordingly for $\overline{v''w''}_0$.
 This way, the domain average of the stress components yield the total stress of eq.
-\ref{us_mean} ($u_{*,mean}^2 = \sqrt{<\overline{u'w'}_0>^2+<\overline{v'w'}_0>^2}$).
-As reference height we use $z_{sl}^- \approx 0.1$ (depending on the height of the closest grid level).
-By using this higher reference height for the boundary condition we solve two problems: First, the reference height is inside of the logarithmic layer, second, we use a velocity from a region where the flow is much better resolved than close to the surface.
+\ref{us_mean} ($u_{*,mean}^2 = \sqrt{<\overline{u''w''}_0>^2+<\overline{v''w''}_0>^2}$).
+For the reference height we use the grid level closest to $z_{sl}^- \approx 0.1$.
+This alleviates two problems: First, $z_{sl}$ is within log-layer.
+Second, the flow at $z=z_{sl}$ is better resolved than closer to the surface.
 

### 2024-06-28 write2/manuscript_part2_2024.tex
started work on Sec. 4

@@
 % z0
-A key parameter governing the turbulent state of the flow is the geostrophic drag: it puts the surface friction in relation to the geostrophic wind and thus illustrates the conversion of mean-flow kinetic energy to turbulence.
-The ratio of $u_{\star}$ resulting from the simulations to the theoretical value $u_{\star,th}$ are shown in Tbl.
+The geostrophic drag $Z=u_\star/G$ quantifies the conversion of mean-flow kinetic energy ($G^2$) to turbulence ($u_{\star}^2$).
+The ratio of $u_{\star}$ resulting from the simulations to the theoretical value $u_{\star,th}$ are shown in Tab.
 \ref{simulation_results}.
@@
 From a top perspective, the horizontal velocity increases from its geostrophic value to the supergeostrophic maximum and then decreases with decreasing height.
-The horizontal mean velocity at the grid point closest to $z^-=0.1$ is used to calculate $u_\star$ according to eq.
-\ref{us_mean}.
+The horizontal mean velocity at the grid point closest to $z^-=0.1$ is used to calculate $u_\star$ according to Eq.~(\ref{us_mean}).
 It is remarkable that the choice of $z_0^+=0.1031$ leads to a value of $u_{\star}$ very close to the prediction by semi-empirical considerations (cf.
@@
 %alpha
-Another key parameter of the Ekman flow is the surface veering angle $\alpha_\star$ between the directions of the negative surface shear stress and the geostrophic wind.
-For the LES, the direction of the negative stress is represented by the first derivatives of the velocity components calculated by 3rd order forward finite differences: $\alpha_\star = \tan^{-1}(\langle \partial v/\partial z\rangle_0/\langle \partial u/\partial z\rangle_0)-\tan^{-1}(V_G/U_G)$.
-Table \ref{simulation_results} shows that nearly all LES yield an $\alpha_\star$ smaller than the theoretical value $\alpha_{\star,th}$.
+The surface veering $\alpha_\star$ is key to quantify the Ekman pumping and of central relevance for larger-scale applications.
+For the LES, we use the first derivatives of the velocity components calculated by 3$^\text{rd}$-order forward finite differences to estimate $\alpha_\star$:
+\begin{align}
+  \alpha_\star = \arctan\left(\left.\left\langle \frac{\partial v}{\partial z}\right\rangle_0\right/\left\langle \frac{\partial u}{\partial z}\right\rangle_0\right)-\arctan\left(\frac{V_G}{U_G}\right).
+\end{align}
+ Table \ref{simulation_results} shows that nearly all LES yield an $\alpha_\star$ smaller than the theoretical value $\alpha_{\star,\text{th}}$.
 In general, we observe an increase of $\alpha_\star$ with resolution.
-The finest resolution of $Re2$ and $Re3$ generally align closely with the theoretical direction while for $Re1\_200$ still 0.9° is missing.
+The finest resolution of Re2 and Re3 generally align closely with the theoretical direction while for Re1\_200 still 0.9° is missing.
 Fig.
-\ref{3Re_hodograph} illustrates that this is not caused by poor quality of the simulation since all of the simulations of $Re1$ follow the theoretical hodograph quite closely and even overestimate the angle at the respective height by a little.
+\ref{3Re_hodograph} illustrates that this is not caused by poor quality of the simulation since all of the simulations of Re1 follow the theoretical hodograph quite closely and even overestimate the angle at the respective height by a little.
 In contrast to the higher $\RE$, the veering of the wind vector continues in the lower parts of the boundary layer.
 Hence, for the low-$\RE$ case, the general observation of a higher $\alpha_\star$ with finer resolution is also caused by the approach of the final $\alpha_\star$ with decreasing height of the first grid point.
-In general, the finer resolved simulations show veering angles that are very close to the theoretic predictions.
+In general, the finer resolved simulations show veering angles that are very close to the theoretical prediction.
 

### 2024-07-08 write2/manuscript_part2_2024.tex
worked on part 4

@@
 \end{align}
- Table \ref{simulation_results} shows that nearly all LES yield an $\alpha_\star$ smaller than the theoretical value $\alpha_{\star,\text{th}}$.
-In general, we observe an increase of $\alpha_\star$ with resolution.
-The finest resolution of Re2 and Re3 generally align closely with the theoretical direction while for Re1\_200 still 0.9° is missing.
-Fig.
-\ref{3Re_hodograph} illustrates that this is not caused by poor quality of the simulation since all of the simulations of Re1 follow the theoretical hodograph quite closely and even overestimate the angle at the respective height by a little.
-In contrast to the higher $\RE$, the veering of the wind vector continues in the lower parts of the boundary layer.
+Nearly all LES yield an $\alpha_\star$ smaller than the theoretical value $\alpha_{\star,\text{th}}$ (cf.
+Tab.~\ref{simulation_results}), where we observe a slight increase of $\alpha_\star$ with resolution; The finest resolution of Re2 and Re3 generally align closely with the theoretical direction while for Re1\_200 the offset is 0.9°.
+The hodograph (Fig.~\ref{3Re_hodograph}) illustrates that this is not caused by poor quality of the simulation since all of the simulations of Re1 closely follow the theoretical hodograph; in fact, the turning angle at the respective height is actually slightly overestimated.
+Rather, the offset is due to the veering of the wind vector continues in the lower part of the boundary layer that is not present in the LES.
 Hence, for the low-$\RE$ case, the general observation of a higher $\alpha_\star$ with finer resolution is also caused by the approach of the final $\alpha_\star$ with decreasing height of the first grid point.
@@
 	}
-  \caption{Shear-aligned velocity profiles in outer scaling. Solid lines: u-component, dashed lines: v-component. The lowest grid point is marked by $x$}
+  \caption{Shear-aligned velocity profiles in outer scaling.
+Solid lines: u-component, dashed lines: v-component.
+The lowest grid point is marked by a cross ($\times$) colored according to the corresponding resolution.}
   \label{velocity_outer}
@@
 % profiles
-Figure \ref{velocity_outer} shows the velocity profiles from the LES and the theoretical profiles for each of the $\RE$.
-In general, the velocity profiles of the LES agree well with the theoretical profiles: The course of the theoretical Ekman layer is matched and the simulations exhibit a logarithmic layer for the U-component.
+Velocity profiles from LES and our theoretical framework agree well for the different $\RE$ in Fig.\,\ref{velocity_outer}: Most prominently, the rotation across the Ekman layer agrees and a logarithmic layer for the U-component is present.
 The lowest points of the LES of the higher $\RE$ fall into the logarithmic layer.
 For Re1, the lowest grid point falls into the buffer layer, which is visible as the curved course of the U-component below the logarithmic layer in the theoretical profile.
-The best resolved simulation with Deardorff-closure even seems to follow the course of the upper part of the buffer layer, but with a resolution of $\Delta^+=15$ this is a coincidence caused by a well-known S-shape of velocity profiles close to rigid walls \cite{brasseur2010designing}.
+The best resolved simulation with Deardorff-closure even seems to follow the course of the upper part of the buffer layer, but with a resolution of $\Delta^+=15$ this is a coincidence caused by a well-known S-shape of velocity profiles close to rigid wall \citep{brasseur2010designing}.
 This log-layer mismatch arises from a competition between the scales $u_\star$ and $z$ and other velocity and length scales introduced by the discretization of the dynamical system \citep{mason1992stochastic,brasseur2010designing}.
 In other words, at the lower boundary, the relevant eddies are too small to be resolved by the grid and their contribution to the flow has to be modeled.
-Also the vertical component is massively restricted by the non-permeability of the wall, known as blocking effect.
-On the contrary, the SGS-closure assumes isotropic turbulence, which is not the case on the grid scale near the wall.
+Also the vertical component is restricted by the wall's impermeability, known as blocking effect-- both effects illustrate a violation of isotropy near the wall and thus counter the SGS-closure assumptions.
 
@@
   \caption{K\'arm\'an measure $\kappa$ in the logarithmic region and above for different Reynolds numbers and resolutions}
-  \label{3Re_kappa}
+  \label{fig:3Re_kappa}
 \end{figure}
@@
 % kappa
-In the logarithmic region, the profile of the u-component should follow the logarithmic law of eq.
-\ref{eqn:log-law} with the K\'arm\'an measure $\kappa = (z^+\partial U^+/\partial z^+)^{-1}=const.$.
-In the low-Reynolds-number case, the viscous sublayer represents about $0.5\%$ of the boundary layer, while this layer is not visible for the high-Reynolds-number cases, where the share drops to $~10^{-5}-10^{-7}$.
+The goodness-of-fit in the logarithmic layer (Eq.\,\ref{eqn:log-law}) is assessed via the K\'arm\'an measure~(Fig.\ref{fig:3Re_kappa})
+\begin{align}
+  \kappa(z)=(z^+\partial U^+/\partial z^+)^{-1}\simeq const.
+\end{align}
+which is estimated here by linear regression of the mean velocity in x-direction against the logarithm of height between the seventh grid point and the $z^-=0.1$ for Re1 and $z^-=0.12$ for Re2 and Re3.
+The seventh grid point is chosen as lower boundary for the logarithmic layer as, in PALM, the mean velocity profile follows MOST aloft \citep{maronga2020improved}.
+The number of values for each regression is 6, 12, and 18 for ReX\_100, ReX\_150 and ReX\_200, respectively.
+In case Re1, the viscous sublayer occupies about $0.5\%$ of the boundary layer, but this share drops to $~10^{-5}-10^{-7}$ for cases Re2 and Re3 respectively.
 Above the viscous sublayer, the theoretical profile shows a near-constant value for $\kappa$ up to $z^-\approx0.1$ for the case Re1 and up to $z^-\approx0.12$ for the cases Re2 and Re3.
-In order to estimate $\kappa_{LES}$ from the simulations, we perform a linear regression of the mean velocity in x-direction and the logarithm of height between the seventh grid point and the grid point corresponding to the height $z^-=0.1$ for $Re1\_X$ and $z^-=0.12$ for $Re2\_X$ and $Re3\_X$.
-We consider only points above the 7th grid point following arguments of \cite{maronga2020improved}, that in PALM, the mean velocity profiles follow MOST above the seventh grid point.
-The number of values for each regression is 6, 12, and 18 for $ReX\_100$, $ReX\_150$ and $ReX\_200$, respectively.
-We do not compute $\kappa$ for $ReX\_50$ since there are only six grid points below $z^-\approx 0.12$.
-The resulting $\kappa_{LES}$ for the other simulations are shown in Tbl.
-\ref{simulation_results}.
-The coefficient of determination is above 0.99 for all fits.
-$\kappa_{LES}$ is decreasing with finer resolution for all $\RE$.
-While the values for $ReX\_100$ are rather high with a value around $0.53$, finer resolution leads to values closer to measurements around $0.46$ for $Re1\_200$ and around $0.44$ for $Re2\_200$ and $Re3\_200$.
-The dynamic subgrid closure yields lower values for $\kappa_{LES}$: in $Re1\_dyn$ we even see $\kappa_{LES}=0.39$, while $Re2\_dyn$ and $Re3\_dyn$ yield $0.42$ and $0.43$, respectively.
-Figure \ref{3Re_kappa} shows the K\'arm\'an measure in the logarithmic layer.
-For the simulations $ReX\_50$, no constant $\kappa$ can be observed.
-An increase in resolution leads to a profile of the K\'arm\'an measure closer to the theoretical curve for all cases.
+We do not compute $\kappa$ for ReX\_50 since there are only six grid points below $z^-\approx 0.12$, and o constant K\'arm\'an measure $\kappa(z)$ can be observed.
+The resulting best estimates for $\kappa_{LES}$ are shown in Tab.\,\ref{simulation_results}, where the coefficient of determination ('r-squared') is above 0.99 for all cases.
+
+For finer resolution, $\kappa_{LES}$ decreases with increasing Re for the range of scale separation investigated here.
+ReX\_100 yields a value around $0.53$ (which is outside the range agreed on by atmospheric observation and laboratory measurements), the finer resolution yields values around $0.46$ for Re1\_200 and around $0.44$ for Re2\_200 and Re3\_200, which is in excellenet agreement with DNS data and laboratory measurements.
+The dynamic subgrid closure yields lower values for $\kappa_{LES}$: in Re1\_dyn we even see $\kappa_{LES}=0.39$, while Re2\_dyn and Re3\_dyn yield $0.42$ and $0.43$, respectively.
 Close to the bottom, the K\'arm\'an measure is heavily influenced by the proximity of the wall and rather a function of the vertical index than of the physical distance from the wall.
-In accordance with the observations of \cite{maronga2014monin} and \cite{maronga2017formulation}, the kinks in the K\'arm\'an measure diminish around the seventh grid point for all simulations.
-Above, the curve smoothens and approaches the expected value of $\kappa$ (at least for the finer resolutions).
+In accordance with \cite{maronga2014monin} and \cite{maronga2017formulation}, the kinks in the K\'arm\'an measure diminish around the seventh grid point for all simulations.
+Above, the curve smoothens and approaches the expected value of $\kappa$.
 
@@
 
-As expected for an LES with isotropic grid spacing, the resolution near the surface is a critical part of the simulation.
-We can support our above observations by taking a look at the two-point correlation in the critical layers of the simulations.
-The two point correlation is defined as
+The resolution near the surface is critical for an LES with isotropic grid spacing. \todo{Missing transition from resolution to two-point correlation!} The two-point correlation is defined as 
 \begin{align}
-  B_{xx}\left(x^*\right) = \langle u\left(x-x^*\right)u\left(x\right)\rangle/\sigma_u^2.
+  B_{xx}\left(x^*\right) = \langle u\left(x-x^*\right)u\left(x\right)\rangle/\sigma_u^2,
 \end{align}
-As suggested by \cite{wurps2020grid}, in an isotropic grid with periodic horizontal boundary conditions the turbulent structures of the velocity's w-component in y-direction can serve as an indicator how well the flow is resolved.
-Figure \ref{2pc_008} shows the two-point correlation $B_{ww}(y^*)$ at a height of $z^-\approx 0.08$, which is right inside of the logarithmic layer.
+and it supports the above findings \todo{which findings?}. 
+For an LES with isotropic grid and periodic horizontal boundaries, the turbulent structures of the velocity's w-component in the y-direction indicate how well the flow is resolved \citep{wurps2020grid}.
+Fig.\,\ref{2pc_008} shows the two-point correlation $B_{ww}(y^*)$ at a height of $z^-\approx 0.08$, which is right inside of the logarithmic layer.
 We define the average size of a structure in the flow $\sigma$ as the distance where the two-point correlation drops to 0.3 (horizontal line in figure).

### 2024-07-08 write2/manuscript_part2_2024.tex
finished comments on Section 4

@@
-\documentclass[draft,a4paper,11pt]{article}
+\documentclass[a4paper,11pt]{article}
 
@@
 
-\subsection{Bulk parameters}
+\subsection{Geostrophic drag and surface wind veer}
 \label{bulk}
@@
 \begin{figure}[ht]
-  \centerline{
-	\includegraphics[width=\textwidth]{figures_2024/d3y_2pc_ww_3Re_0.079.png}
-}
-  \caption{Two-point correlation of w-component in y-direction at $z^-\approx 0.08$. The horizontal line is at $B_{ww}=0.3$}
-  \label{2pc_008}
+  \centerline{\includegraphics[width=\textwidth]{figures_2024/d3y_2pc_ww_3Re_0.079.png}}
+  \caption{Two-point correlation of w-component in y-direction at $z^-\approx 0.08$. The horizontal line at $B_{ww}=0.3$ is indicative of the average size of turbulent structures in the flow.}
+  \label{fig:2pc_008}
 \end{figure}
@@
 and it supports the above findings \todo{which findings?}. 
-For an LES with isotropic grid and periodic horizontal boundaries, the turbulent structures of the velocity's w-component in the y-direction indicate how well the flow is resolved \citep{wurps2020grid}.
-Fig.\,\ref{2pc_008} shows the two-point correlation $B_{ww}(y^*)$ at a height of $z^-\approx 0.08$, which is right inside of the logarithmic layer.
-We define the average size of a structure in the flow $\sigma$ as the distance where the two-point correlation drops to 0.3 (horizontal line in figure).
+For an LES with isotropic grid and periodic horizontal boundaries, the turbulent structures of the vertical velocity component in the y-direction (spanwise near the surface) indicate how well the flow is resolved \citep{wurps2020grid}.
+Fig.\,\ref{fig:2pc_008} shows the two-point correlation $B_{ww}(y^*)$ at a height of $z^-\approx 0.08$, right inside the logarithmic layer.
+We define the average size of a structure in the flow $\sigma$ as the separation at which the two-point correlation drops to 0.3 (horizontal line in Fig.~\ref{fig:2pc_008}).
 The number of cells by which the average structure is resolved is equal to $\sigma/\Delta$.
-All $\RE$ show a similar behaviour: the average size of the structure decreases with finer resolution and seems to approach a limit.
-The structure size of ReX\_150 and ReX\_200 are very close together, which indicates beginning convergence.
+All cases show a similar behaviour, irrespective of the Reynolds number: the average size of the structure decreases with finer resolution and approaches a limit.
+The structure size of cases ReX\_150 and ReX\_200 agrees very well, which indicates beginning convergence.
 Furthermore, the number of cells by which the structures are resolved exceed 4 from the resolution of ReX\_150 on.
@@
 \begin{figure}[ht]
-  \centerline{
-	\includegraphics[width=\textwidth]{figures_2024/d3y_2pc_ww_3Re_4.png}
-}
+  \centerline{\includegraphics[width=\textwidth]{figures_2024/d3y_2pc_ww_3Re_4.png}}
   \caption{Two-point correlation of w-component in y-direction at the 8th grid point above the surface}
-  \label{2pc_4}
+  \label{fig:2pc_4}
 \end{figure}
 
-Figure \ref{2pc_4} shows $B_{ww}(y^*)$ at the 8th grid point above the bottom.
+Figure \ref{fig:2pc_4} shows $B_{ww}(y^*)$ at the 8th grid point above the bottom.
 Therefore, the corresponding heights differ between the simulations, that is $z^-=0.16, 0.08, 0.053, 0.04, 0.04$ for $ReX\_50$, $ReX\_100$, $ReX\_150$, $ReX\_200$, $ReX\_dyn$, respectively.
@@
   \caption{Span-wise velocity component in outer scaling}
-  \label{3Re_v_lin}
+  \label{fig:3Re_v_lin}
 \end{figure}
 
-Figure \ref{3Re_v_lin} shows the span-wise velocity component in the lower part of the boundary layer.
+Figure \ref{fig:3Re_v_lin} shows the span-wise velocity component in the lower part of the boundary layer.
 \todo{if we keep $V(z=\Delta)$ in the conclusion, add this here}.

### 2024-07-08 write2/manuscript_part2_2024.tex
comments and edits on manuscript from Jun 25th 2024 by Cedrick

@@
 % kappa
-The goodness-of-fit in the logarithmic layer (Eq.\,\ref{eqn:log-law}) is assessed via the K\'arm\'an measure~(Fig.\ref{fig:3Re_kappa})
+The goodness-of-fit in the loga\-rith\-mic layer (Eq.\,\ref{eqn:log-law}) is assessed through the von-K\'ar\-m\'an mea\-sure (Fig.\,\ref{fig:3Re_kappa})
 \begin{align}
-  \kappa(z)=(z^+\partial U^+/\partial z^+)^{-1}\simeq const.
+  \kappa(z)=(z^+\partial U^+/\partial z^+)^{-1}\simeq const.,
 \end{align}
@@
 and it supports the above findings \todo{which findings?}. 
-For an LES with isotropic grid and periodic horizontal boundaries, the turbulent structures of the vertical velocity component in the y-direction (spanwise near the surface) indicate how well the flow is resolved \citep{wurps2020grid}.
-Fig.\,\ref{fig:2pc_008} shows the two-point correlation $B_{ww}(y^*)$ at a height of $z^-\approx 0.08$, right inside the logarithmic layer.
-We define the average size of a structure in the flow $\sigma$ as the separation at which the two-point correlation drops to 0.3 (horizontal line in Fig.~\ref{fig:2pc_008}).
-The number of cells by which the average structure is resolved is equal to $\sigma/\Delta$.
-All cases show a similar behaviour, irrespective of the Reynolds number: the average size of the structure decreases with finer resolution and approaches a limit.
-The structure size of cases ReX\_150 and ReX\_200 agrees very well, which indicates beginning convergence.
-Furthermore, the number of cells by which the structures are resolved exceed 4 from the resolution of ReX\_150 on.
+For an LES with isotropic grid and periodic horizontal boundaries, the the span-wise two-point correlation of the vertical velocity component $B_{ww}(y^*)$ indicates how well the flow is resolved \citep{wurps2020grid}.
+We define $\sigma$, the average size of a structure in the flow, as the separation at which the two-point correlation at a height $z^-\approx 0.08$, right inside the logarithmic layer, drops to 0.3, i.e.
+$B_{ww}(\sigma^\star) = 0.3$ (Fig.\,\ref{fig:2pc_008}).
+The number of cells by which the corresponding structure is resolved is then $\sigma/\Delta$.
+All cases show a similar behaviour, irrespective of the Reynolds number: the average size of the structure decreases with finer resolution.
+and it approaches a limit.
+The structure size of cases ReX\_150 and ReX\_200 agrees well, indicating convergence.
+Furthermore, the number of cells by which the structures are resolved exceeds four from the resolution of ReX\_150 on.
 This corresponds to the resolution from which on a logarithmic layer with a constant $\kappa$ can be seen.
@@
 
-Figure \ref{fig:2pc_4} shows $B_{ww}(y^*)$ at the 8th grid point above the bottom.
-Therefore, the corresponding heights differ between the simulations, that is $z^-=0.16, 0.08, 0.053, 0.04, 0.04$ for $ReX\_50$, $ReX\_100$, $ReX\_150$, $ReX\_200$, $ReX\_dyn$, respectively.
+In vicinity of the surface, the two-point statistics crucially depend on how the surface-interaction is modelled \citep{maronga2014monin}.
+which may cause strong sensitivity on the resolution.
+Indeed, the two-point correlations $B_{ww}(y^*)$ evaluated at the 8th grid point above the bottom (Fig.~\ref{fig:2pc_4}) appear rather universal.
+This is in spite of different heights the simulations ($z^-=0.16, 0.08, 0.053, 0.04, 0.04$ for $ReX\_50$, $ReX\_100$, $ReX\_150$, $ReX\_200$, $ReX\_dyn$, respectively).
 For $Re1$, the curves almost collapse perfectly while for $Re2$ and $Re3$ we see slightly larger structures for $ReX\_50$ and slightly smaller structures for $ReX\_dyn$.
@@
 \begin{figure}[ht]
-  \centerline{
-	\includegraphics[width=\textwidth]{figures_2024/d3y_3Re_v_outer_lin.png}
-}
+  \centerline{\includegraphics[width=\textwidth]{figures_2024/d3y_3Re_v_outer_lin.png}}
   \caption{Span-wise velocity component in outer scaling}
@@
 
-Figure \ref{fig:3Re_v_lin} shows the span-wise velocity component in the lower part of the boundary layer.
+he span-wise velocity component in the lower part of the boundary layer is shown in Fog.~\ref{fig:3Re_v_lin}.
 \todo{if we keep $V(z=\Delta)$ in the conclusion, add this here}.
 In the lowest part of $Re1$ ($z^-\lessapprox 0.05$), a part of the viscous layer is visible, where the velocity is slightly underestimated by the the LES.
-Above, the simulations $Re1\_200$ and $Re1\_dyn$ are very close to the theoretical curve.
-Lower resolution leads to an overestimation of the velocity in the logarithmic layer, which is consistent with a lack of large eddies that translates in a lack of mixing and thus a reduced turbulence level and increased velocity gradients.
-Around $z^-=0.25$, while blending into the Ekman layer, the curves cross each other and the coarse simulations begin to underestimate the velocity.
-For the higher $Re$ all simulations overestimate the velocity in the logarithmic layer, while coarser resolution leads to a higher velocity and finer resolution leads to a lower velocity and better agreement with the theoretical curve.
+Above, the simulations Re1\_200 and Re1\_dyn are very close to the theoretical curve.
+Lower resolution leads to an overestimation of the velocity in the logarithmic layer, which is consistent with a lack of large eddies and thus too little mixing, reduced turbulence intensity and increased velocity gradients.
+Around $z^-=0.25$, while blending into the Ekman layer, the curves cross each other and aloft the coarse simulations underestimate the velocity.
+For the higher $Re$ all simulations overestimate the velocity in the logarithmic layer while finer resolution loeads to lower velocity and better agreement with the theoretical curve.
 Again, we see a crossing of the curves, between $z^-=0.2$ and $z^-=0.25$.
@@
 
-In Fig.
+\todo{translational statement to hodograph missing} In Fig.
 \ref{3Re_hodograph}, the cross indicates the lowest grid point and the plus indicates the height where the velocity for the boundary condition is taken from.
@@
   \caption{Shear-aligned velocity deficit in outer scaling. Solid lines: U-component, dashed lines: V-component}
-  \label{3Re_Ekman}
+  \label{fig:3Re_Ekman}
 \end{figure}
 
-Figure \ref{3Re_Ekman} shows the shear-aligned velocity deficit $U^--U^-_G$ (v-component accordingly) of the whole boundary layer.
-The strongest deviations from the theoretical profiles can be seen for the coarsest LES near the super-geostrophic maximum of the U-component.
-Finer resolution leads to a very good agreement with the extent of the maximum, which also holds for the higher $\RE$.
+The shear-aligned velocity deficit $U^--U^-_G$ (v-component accordingly) agrres well with the expected DNS-based profiles.
+The largest deviations occur for the coarsest LES near the super-geostrophic maximum of the U-component.
+Finer resolution leads to excellent agreement, also holds for the higher $\RE$.
 The super-geostrophic maximum is a consequence of vertical redistribution of momentum across the ABL; less redistribution leads to a less pronounced maximum.
 At the same time, of course, one would expect an increased level of turbulence to cause a flatter hodograph.
-BUT, the surface veering is set rather by the interplay of MOST and the prescribed roughness while the profile in the Ekman layer is a resolved quantity.
+But, the surface veering is set rather by the interplay of surface-layer turbulence and the prescribed roughness while the profile in the Ekman layer is resolved.
 However, for higher Reynolds numbers $Re2$ and $Re3$, the maximum velocity in x-direction is still not reached by the LES and the approach of the geostrophic wind in the upper part of the boundary layer takes place at lower heights than in theory.

### 2024-07-08 write2/manuscript_part2_2024_hw.tex
comments and edits on manuscript from Jun 25th 2024 by Cedrick

--- 2024-07-08 write2/manuscript_part2_2024.tex
@@
 
-\author{Hauke Wurps and Cedrick Ansorge}
+\author{hauke-zrick}
 
@@
   % Introduction % Theoretical profile
-  Turbulent Ekman flow is a canonical flow representation of the atmospheric boundary layer, characterized by a logarithmic increase of the wind speed near the surface and a turning of the wind vector aloft.
+	Turbulent Ekman flow is the simplest representation of the atmospheric boundary layer that exhibits two key properties: the logarithmic increase of the wind speed near the surface and a turning of the wind vector aloft.
 In part I of this work, we derived a formulation of the mean wind velocity profiles of the stationary turbulent Ekman flow based on direct numerical simulation (DNS).
-Here, we explore the extrapolation of these profiles to atmospheric Reynolds numbers using large-eddy simulation (LES).
+Here, we explore the feasibility of extrapolating these profiles to atmospheric Reynolds numbers using large-eddy simulation (LES).
 	% Case desription and setup
-  We compare theoretical profiles of the wind vector for low Reynolds-number turbulent Ekman flow to results from DNS and LES.
+  We compare theoretical wind speed profiles of a low Reynolds-number turbulent Ekman flow to results from DNS and LES.
 This necessitates the inclusion of viscous effects, usually neglected in LES, along with modifications to the standard bottom boundary condition.
 	% Results and Discussion
-  Analysis of the grid-convergence for higher resolution unveils a convergence of LES data towards the theoretical profiles for both intermediate and high Reynolds numbers.
-Our LES thus confirm that--within the logarithmic layer (the lowest 10\,\% of the boundary layer)--- the turning of the wind vector rotates by approximately one-third of its total rotation across the entirety of the boundary layer.
+	Analysis of the grid resolution dependency showed a convergence of LES results towards the theoretical profiles for low and high Reynolds numbers.
+LES results confirmed that within the logarithmic layer (the lowest 10\,\% of the boundary layer), the turning of the wind vector contributes approximately one-third to the total rotation.
 	% Conclusions
-  Such agreement of our theoretical formulation and LES data raises our confidence in the underlying assumptions of the theory and thus reinforces the utility of the theoretical profiles as a reference for intermediate and a quasi-reference for higher Reynolds number simulations.
+  The notable agreement of the theoretical formulation and LES  results raised our confidence in its reliability for high Reynolds numbers.
+This reinforces the utility of the theoretical profiles as a true reference for intermediate and a quasi-reference for higher Reynolds number simulations.
 %
@@
 
-% \todo{start introduction rather with TEF than LES (rather subject than method)}
-The interaction of atmospheric boundary layer  (ABL) flow with its lower boundary condition, the surface, is a defining property of the ABL \citep{stull:1988}.
-Relevant aspects of the interaction with the boundary not only include its heterogeneity \citep{avissar:MWR1989,giorgi:RG1997,claussen:BM1991,garratt:BM1990} and roughness \citep{monin:ARF1970,brutsaert:WRR1975,raupach:AMR1991,kostelecky:JFM2024}, but also a wind veer in favor of the pressure gradient force when the surface is approaced \citep{ekman:AMA1905}.
-An important tool with capabilities to model the flow-surface interaction is Large-Eddy Simulation (LES).
-Indeed, LES is widely used to model turbulent flow in the atmospheric boundary layer (ABL) \citep{stoll2020large}.
+\todo{start introduction rather with TEF than LES (rather subject than method)}
+Large Eddy simulation (LES) is widely used to model turbulent flow in the atmospheric boundary layer (ABL) \citep{stoll2020large}.
 This includes various applications, such as simulating wind turbine and wind farm wakes, and interactions within wind farm clusters \citep{porte2011large,mehta2014large,breton2017survey}.
@@
 \begin{align}\label{red}
- Re_D = \sqrt{2}\frac{G}{\sqrt{\nu f}}
+ Re_D = \frac{G}{\sqrt{\frac{1}{2}\nu f}}
 \end{align}
@@
 
-Part I of this publication, presents a framework to predict wind direction and speed across the turbulent Ekman flow, drawing from the works of \cite{csanady1967resistance}, \cite{tennekes1973logarithmic}, and \cite{spalart1989theoretical}.
+Part I of this publication, presents a theoretical framework to predict wind direction and speed across the turbulent Ekman flow, drawing from the works of \cite{csanady1967resistance}, \cite{tennekes1973logarithmic}, and \cite{spalart1989theoretical}.
 As \cite{spalart1989theoretical}, we use $G$, $f$, and $\nu$ to derive the friction velocity $u_\star$ and the angle $\alpha_\star$ between geostrophic wind and shear stress at the wall.
-While earlier descriptions of the turbulent Ekman layer were limited to specific parts of the boundary layer, we derive mean velocity profiles covering the entire ABL with due regard of changes in the Reynolds number.
-We have shown that the theoretical velocity profiles are in excellent agreement with DNS for intermediate Reynolds numbers.
+While earlier descriptions of the turbulent Ekman layer were limited to specific parts of the boundary layer, we give a comprehensive theoretical description of the mean velocity profiles that covers the complete ABL and quantifies the Reynolds number.
+We have shown that the theoretical velocity profiles are in excellent agreement with DNS for low and intermediate Reynolds number.
 
-However, it remains uncertain, if the extrapolation of the above framework towards atmospheric $\RE$ is valid.
+However, it remains uncertain, if the extrapolation of the theory towards atmospheric $\RE$ is valid.
 The substantial separation of scales in the ABL renders direct numerical simulation (DNS) of the entire turbulent flow at scale impossible.
@@
 
-The accurate representation of mean profiles of velocity, i.e.
-of first-order statistics of the wind field, is commonly assumed when using LES in a predictive manner  \citep{fedorovich2004convective}.
+When using LES in a predictive manner, a common assumption is the accurate reproduction of velocity profiles, i.e.
+first-order statistics of the wind field \citep{fedorovich2004convective}.
 Here, we compare LES outcomes to theoretical solutions to determine the validity of this assumption under different conditions.
-As LES introduces the grid size as an additional parameter (often understood in relation to a filter scale, cf.
-\cite{pope2004ten}), we conduct a thorough analysis of its impact on simulation results.
+As LES inherently introduces the grid size as an additional (artificial) parameter, we conduct a thorough analysis of its impact on simulation results.
 \cite{esau2004simulation} and \cite{jiang2018large} have executed similar LES analyses of neutral Ekman layers, however, they introduce an additional parameter by selecting an arbitrary roughness length $z_0$, whereas we derive the parameter $z_0$ corresponding to a fluid with a kinematic viscosity of $\nu = 1.5\cdot10^{-5}$ over flat surface.
@@
 Here, we examine three different Reynolds numbers $Re_D$: $1600$, $1.5\cdot10^5$, and $10^6$.
-While the case $Re_D=1600$ is also investigated by DNS (cf.
+While the low-Re case ($Re_D=1600$) is also investigated by DNS (cf.
 part 1 of this study), $Re_D=1.5 \cdot 10^5$ and $10^6$ correspond to the scale separation found in typical atmospheric boundary layers.
-For $Re_D=1600$, we compare DNS and LES results using the exact same forcing and settings.
+For the low-Re case, we compare DNS and LES results using the exact same forcing and settings.
 This is the first comparison of this kind for the turbulent Ekman flow.
@@
 
-In this chapter, we introduce the velocity profile for the turbulent Ekman layer developed in part I. 
+A universal velocity profile for the turbulent Ekman layer is developed in part I of this paper. In this chapter, we give a brief mathematical description.
+
 %Definitions
-We align the coordinate Ox with the surface stress, Oz points normal to the surface and normal to Ox, and Oy in the span-wise direction, normal to Oxz.
-%The mean velocity profiles for all $\RE$ exhibit a dual scaling nature: in proximity to the lower boundary, the velocity profiles collapse when scaled using
-In proximity to the lower boundary, the boundary layer scales in viscous (inner) units, denoted by the index ``$+$'' ($z^+ = zu_\star/\nu$, $U^+=U/u_\star$, where $U$ is the velocity component in x-direction and $u_\star = \sqrt{\nu \p_z U}$ the friction velocity).
-From the top perspective, the profiles converge when scaled using outer units, denoted by the index ``$-$'' ($z^-=z/\delta$, $U^-=U/G$, where $\delta=u_\star/f$ is the boundary-layer depth scale, $f$ the Coriolis parameter and $G$ the geostrophic wind speed).
+The mean velocity profiles for all $\RE$ exhibit a dual scaling nature: in proximity to the lower boundary, the velocity profiles collapse when scaled using viscous (inner) units, denoted by the index ``$+$'' ($z^+ = zu_\star/\nu$, $U^+=U/u_\star$, where $U$ is the velocity component in x-direction).
+From the top perspective, the profiles converge when scaled using outer units, denoted by the index ``$-$'' ($z^-=z/\delta$, $U^-=U/G$, with $\delta=u_\star/f$ representing the boundary layer height).
+The outer and inner velocities $G$ and $u_\star$, respectively, are linked by a semi-empirical law \cite{spalart1989theoretical}.
 At the lower boundary, the x-axis of the inner units aligns with the shear stress, while the x-axis of the outer units aligns with the geostrophic wind.
-The angle between both axes is denoted as $\alpha_\star$, the surface veer of the wind across the boundary layer.
-%
-Profiles of the velocity components are studied in three different layers reflecting the change of dominant balance when moving away from the surface. These are
-(i) the Ekman layer denoted $(\cdot)_\text{Ek}$ (Sec.~\ref{sec:ekmanlayer}), 
-(ii) the viscous sub-layer, denoted by an index $(\cdot)_\text{visc}$, 
-and (iii) the logarithmic layer denoted by $(\cdot)_\text{log}$
-We match $U_\text{visc}$ and $U_\text{log}$ according to their respective formulations and combine them to $U_\text{inner}$ (Sec.~\ref{sec:innerlayer}).
-Inner and outer profiles are then matched by a weighting transfer function based on the error function. 
-%The span-wise component of the velocity is separated into two layers, namely the inner layer $V_\text{inner}$, and the Ekman layer $V_{EK}$.
+The angle between both axes is denoted by $\alpha_\star$.
 
-% \subsection{Drag-Law}
+\subsection{Drag-Law}
 
-The geostrophic drag, i.e.
-normalized friction $Z\equiv u_\star/G$, and $\alpha_\star$ form the basis of the boundary-layer scaling, and they are estimated using the semi-empirical drag-law introduced by \cite{spalart1989theoretical}, which describes them as functions of only the Reynolds number:
-\begin{subequations}\label{eqn:drag}
+The geostrophic drag $Z \equiv u_\star/G$ and the angle between the shear stress and the geostrophic wind $\alpha_\star$ are two key parameters of the Ekman flow.
+They can be estimated using a semi-empirical drag-law based on \cite{spalart1989theoretical}, which describes them as functions of only the Reynolds number:
+\begin{subequations}\label{drag}
 	\begin{align}
@@
 		\alpha_\star &= \phi^\star - \frac{C_5}{Re_\tau},\\
-		Re_\tau &= \frac{u_\star^2}{\nu f} \left( = \frac{1}{2} Re_{D}^2 \frac{u_{\star}^2}{G^2}\right),
+		Re_\tau &= \frac{u_\star^2}{\nu f},
 	\end{align}
 \end{subequations}
-where we use $\kappa = 0.416$, $A_r = 4.80$, $A_i = -5.57$, $C = 5.4605$, $C_5 = -57.8$ based on the DNS data available (cf. Part I, \cite{ansorge2014global}).
+with $Re_\tau = \frac{Re_D^2}{2}\frac{u_\star^2}{G^2}$ the turbulent Reynolds number, $\kappa = 0.416$, $A_r = 4.80$, $A_i = -5.57$, $C = 5.4605$, $C_5 = -57.8$.
+This law is in excellent agreement with DNS in the range $400\leq Re_D\leq 1\,600$ as demonstrated by \cite{ansorge2014global}.
 
-% \subsection{Total profile}
+\subsection{Total profile}
 
-\subsection{Ekman layer}
-\label{sec:ekmanlayer}
-The outer layer of the ABL is characterized by a triadic balance between turbulent flux, pressure gradient and the Coriolis; the vertical change of the Coriolis force causes a pronounced height dependence of the wind direction resulting in the Ekman spiral \citep{ekman:AMA1905}.
-The classic solution employs Ekman dynamics down to the surface---and thus also in the surface layer, where this is a rather poor representation of turbulent mixing.
-Here, we model the surface layer by a lower boundary condition for the Ekman spiral determined from DNS:
+As frame of reference we use the coordinate system Oxyz with Ox in the direction of the surface stress, Oz normal to the surface, and Oy in the span-wise direction normal to Oxz.
+The profile of the stream-wise component of the velocity is seperated into three layers, which are the viscous layer $U_{visc}$, the logarithmic layer $U_{log}$, and the Ekman layer $U_{EK}$.
+$U_{visc}$ and $U_{log}$ are matched by their formulation and combined to $U_{inner}$.
+The span-wise component of the velocity is separated into two layers, namely the inner layer $V_{inner}$, and the Ekman layer $V_{EK}$.
+The smooth transition between consecutive layers is achieved using a transfer function:
+\begin{align}\label{error}
+  w_{*} = \frac{1}{2}\left(\textrm{erf}\left[\sigma_T\log\left(\frac{z}{z_{T}}\right)\right]+1\right),
+\end{align}
+where $erf$ is the error function, $\sigma_T$ is a transition scale that defines the width of the transition and $z_{T}$ is the height of the transition, where the upper and the lower layer equally contribute to the velocity ($w_{*}(z_{T})=0.5$).
+
+\begin{subequations}\label{blend}
+  \begin{align}
+    U &= (1-w_{outer})U_{inner} + w_{outer}U_{EK},\\
+	  V &= (1-w_{outer})V_{inner} + w_{outer}V_{EK}.
+	\end{align}
+\end{subequations}
+
+\subsection{Profiles in Outer Layer}
+
+In the outer layer, the vertical change of the Coriolis force causes a  pronounced height dependence of the wind direction resulting in the Ekman spiral \cite{ekman1905influence}.
+While the classic solution employs Ekman dynamics down to the surface --- and thus also in the surface layer, where the pure Ekman dynamics are a rather poor representation of turbulent mixing ---, we model the surface layer by a lower boundary condition for the Ekman spiral that we determine from DNS data:
 \begin{subequations}
@@
 where the x-axis is aligned with the geostrophic wind and $A = 8.4u_\star$, $\tilde{z} = (z-z_r)/D_E$, $z_r = 0.12\delta$, and $D_E = 3\delta/4\pi\approx 0.24\delta$.
-The transition from the logarithmic layer to the Ekman layer is located at $z^{-}=0.28- 2.25 Re_{D}^{-1/2}$ with a transition scale of $\sigma_T=2$ for the stream-wise velocity.
+The transition from the logarithmic layer to the Ekman layer is located at $z^-_T=0.3-120/Re_D$ \todo{in script at 0.28- 2.25*np.sqrt(1./re)} with a transition scale of $\sigma_T=2$ for the stream-wise velocity.
 
-\subsection{Inner and viscous layers}
-\label{sec:innerlayer}
+\subsection{Shear-Aligned Velocity}
 
-In the viscous sublayer, the shear-aligned velocity $U^{\alpha_\star}$ is described by the law of the wall $U^{\alpha_\star+} = z^+$ and the span-wise velocity $V^{\alpha_\star}$ is zero by choice of the reference frame
+In the viscous sublayer, the span-wise velocity is close to zero and the shear-aligned velocity is described by the law of the wall:
+
+\begin{align}
+  U^{\alpha_\star+} = z^+,
+\end{align}
 (the index $\alpha_\star$ indicates the alignment with the direction of the shear stress).
-Around $z^+=5$, the velocity begins to deviate from linearity, and the buffer layer forms the transition from the viscous to the  logarithmic layer.
+Around $z^+=5$, the velocity is beginning to deviate from its linear profile and the buffer layer forms the transition between viscous layer and logarithmic layer.
 From the surface up to the buffer layer, the stream-wise velocity is described by
-\begin{subequations}
 \begin{align}
-    U_\text{buffer}^{\alpha_\star+} = \frac{z^+}{1+c_1(z^+)^2} + (c_2z^+-a_\text{match})\frac{1+\tanh[0.2(z^+-22)]}{2}+c_3e^{-c_4(z^+-22)^2}.
-    \label{eqn:ubuffer}
+  U_{visc}^{\alpha_\star+} = \frac{z^+}{1+c_1(z^+)^2} + (c_2z^+-a_{match})\frac{1+\tanh[0.2(z^+-22)]}{2}+c_3e^{-c_4(z^+-22)^2}.
 \end{align}
-With $c_1 = 0.00185$, $c_2 = 0.195$, $c_3 = 0.4$, $c_4 = 0.35$. The coefficient $a_{match}=3.5727$ is chosen to match the u-profile in the logarithmic layer above at $z^+=40$, such that 
+With $c_1 = 0.00185$, $c_2 = 0.195$ \todo{adapt $c_2$?}, $c_3 = 0.4$, $c_4 = 0.35$. The coefficient $a_{match}=3.5727$ is chosen to match the u-profile in the logarithmic layer above at $z^+=40$.
+
+The logarithmic region of the stream-wise velocity is 
+\begin{align}\label{log-law}
+  U_{\log}^{\alpha_\star+} = \frac{1}{\kappa}\log z^+ + C,
+\end{align}
+with the von-Kármán constant $\kappa=0.416$, and $C=5.4605$. The lower part of $U^{\alpha_\star+}$ is described by
 \begin{align}
-  U_\text{inner}^{\alpha_\star+} = \begin{cases}
-	  U_\text{buffer}^{\alpha_\star+}, & z^+ \le 40.\\
-		\kappa^{-1}\log z^+ + C &   z^+>40.
+  U_{inner}^{\alpha_\star+} = \begin{cases}
+	  U_{visc}^{\alpha_\star+}, & z^+<40.\\
+		U_{log}^{\alpha_\star+}, & z^+>40.
 	\end{cases}
-  \label{eqn:log-law}
 \end{align}
-\end{subequations}
-with the von-Kármán constant $\kappa=0.416$, and $C=5.4605$ as in the estimation of the geostrophic drag in Eq.~(\ref{eqn:drag}).
 
-For the span-wise velocity, the boundary conditions and Ekman dynamics imply that 
-\begin{subequations}
+The inner profile is then blended to the Ekman profile using eq. \ref{blend} with $\sigma_T = 2$ and $z_T^- = 0.28- 2.25\sqrt{1/Re_D}$.
+
+\subsection{Span-wise Velocity}
+
+In the viscous layer, the shear-aligned span-wise velocity is described as
 \begin{align}
-V^{\alpha_\star}_\text{visc} \delta^+\sim G f_V(z^+),
-\end{align} where $f_V$ is a universal, non-dimensional function (cf. Part I). Above the viscous 
-layer, scaling arguments are scarce, but there is evidence for a logarithmic scaling of also the 
-span-wise velocity component. We  hence use  
+  V^{\alpha_\star}_{visc} = \frac{G}{\delta^+}v_{ref}(w_vz^+-1+exp[-w_vz^+]).
+\end{align}
+The choice of $v_{ref} = 18.85$ and $w_v = 0.2353$ leads to excellent agreement with the DNS data.
+
+Above the viscous layer a log-like transition to the Ekman layer is modeled by
 \begin{align}
-  V^{\alpha_\star}_\text{inner} = \left\{ \begin{matrix}
-      V^{\alpha_\star}_\text{visc} &=& \dfrac{G}{\delta^+}v_{ref}(w_vz^+-1+exp[-w_vz^+])  & z^+ \le 10\\ 
-      V_\text{log}^{\alpha_\star}        &=& \dfrac{G}{\delta^+}(a_{log} + b_{log}\log(z^+) + c_{log}z^+) & z^+>10. 
-  \end{matrix}\right.
+  V_{log}^{\alpha_\star} = \frac{G}{\delta^+}(a_{log} + b_{log}\log(z^+) + c_{log}z^+).
 \end{align}
-\end{subequations}
-with $v_\text{ref} = 18.85$ and $w_v = 0.2353$ leads to excellent agreement with the DNS data.
-The coefficients $a_\text{log}, b_\text{log}$, and $c_\text{log}$ are determined by (i)-(ii) a smooth transition to $V_\text{visc}$ at $z^+=10$, and (iii) the condition $V_\text{log}(z^-=0.3) = V_\text{EK}(z^-=0.3) =: v_{03}$ (cf.
-Part I).
-For the Reynolds-number dependency of the Ekman-layer profiles, $V_\text{Ek}(z^-=0.3)$ depends on $Re$, such that the coefficients exhibit a weak dependence on $Re$.
+A smooth transition to $V_{visc}$ at $z^+=10$ and the condition $V_{log}(z^-=0.3) = V_{EK}(z^-=0.3) =: v_{03}$ lead to
+\begin{align}
+  c_{log} &= \frac{v_{03}-v_{10} - d_{10}z_{10}\log(z_{03}/z_{10})}{z_{03}-z_{10}},\\
+	a_{log} &= v_{10}-d_{10}z_{10}\log z_{10}+c_{log}z_{10}(1-\log z_{10}),\\
+	b_{log} &= (d_{10}-c_{log})z_{10}.
+\end{align}
+Here, $d_{10}=4.01$, $v_{10}=27.3$.
+Because of the $\RE$ dependence of $V_{Ek}$, $v_{03}$ also depends on $\RE$.
+$V_{log}^{\alpha_\star}$ is blended into $V_{EK}$ using eq.
+\ref{blend} with $\sigma_T = 2$ and $z_T^- = 0.28- 2.25\sqrt{1/Re_D}$.
 
-\subsection{Transition from inner to outer profiles, synopsis}
-
-A smooth transition between consecutive layers is achieved using a transfer function:
-\begin{align}\label{eqn:error}
-  w_{*} = \frac{1}{2}\left(\textrm{erf}\left[\sigma_T\log\left(\frac{z}{z_{T}}\right)\right]+1\right),
-\end{align}
-where $erf$ is the error function, $\sigma_T$ is a transition scale that defines the width of the transition and $z_{T}$ is the height of the transition, where the upper and the lower layer equally contribute to the velocity ($w_{*}(z_{T})=0.5$).
-The inner profiles $U_\text{inner}$ and $V_\text{inner}$ are blended to the Ekman profile using eq.~(\ref{eqn:blend}) with $\sigma_T = 2$ and $z_T^- = 0.28- 2.25\sqrt{1/Re_D}$.
-
-\begin{subequations}\label{eqn:blend}
-  \begin{align}
-    U &= (1-w_{outer})U_{inner} + w_{outer}U_{EK},\\
-	  V &= (1-w_{outer})V_{inner} + w_{outer}V_{EK}.
-	\end{align}
-\end{subequations}
+\subsection{Profile}
 
@@
 
-The velocity profiles are given for two Reynolds numbers in Fig. \ref{single_ekman}. 
-In the viscous sublayer, the shear-aligned component $U^+$ increases linearly with height up 
-$z^+\approx5$.
-In between $5 \lesssim z^+ \lesssim 30$, we find the buffer layer with a transition from the linear law-of-the-wall to the logarithmic law.
-Above $z^+\approx30$, the logarithmic layer begins.
+The velocity profiles are given for two Reynolds numbers in Fig.
+\ref{single_ekman}.
+In the viscous sublayer, the shear-aligned component $U^+$ increases linearly with height up to the buffer layer between $z^+\approx5$ and $z^+\approx30$, where it transitions into the logarithmic layer.
 Then, $U^+$ increases logarithmically up to the Ekman layer and reaches its supergeostrophic maximum.
-Above, it decreases to its free-stream geostrophic value.
-The depth of the logarithmic layer increases with $\RE$.
+Above, it decreases to its geostrophic value.
+The length of the logarithmic layer increases with $\RE$.
 The spanwise component $V^+$ remains close to zero up to the middle of the logarithmic layer, where the transition to the Ekman layer takes place ($z^-\approx0.28-2.25\sqrt{1/Re_D}$).
@@
 
+An incompressible, turbulent Ekman flow over a hydrodynamically smooth surface is simulated for three different Reynolds numbers $Re_D = 10^3;1.5\times 10^5;10^6$, hereafter $Re1$, $Re2$, and $Re3$, respectively.
+The input parameters are given in table \ref{simulation_parameters}.
+
 \begin{table*}
@@
 \end{table*}
-An incompressible, turbulent Ekman flow over a hydrodynamically smooth surface is simulated using the Parallel Large Eddy Simulation model (PALM, \citealp{maronga2020overview}).
-We study three different Reynolds numbers Re$_D = 10^3;1.5\times 10^5;10^6$, hereafter Re1, Re2, and Re3, respectively (Tab.~\ref{simulation_parameters}).
-%
+
 The domain is rotating around the z-axis with an angular velocity corresponding to the Coriolis parameter $f=10^{-4}s^{-1}$.
-The stratification of the flow is truly neutral, i.e., the potential temperature is constant across the whole domain.
-A constant boundary layer height on the order of $\delta=u_\star/f$ forms due to the balance between shear production and rotational suppression of turbulence.
-At the upper boundary, the flow is non-turbulent.
-Hence, a no-penetration boundary condition is used for the vertical velocity component while the horizontal components are prescribed according to geostrophic balance--a Dirichlet-type boundary condition.
+The stratification of the flow is truly neutral, i.e., the potential temperature is constant for the whole domain.
+Nevertheless, the boundary layer height is not growing infinitely but results from a balance between shear production and rotational suppression of turbulence.
+At the upper boundary, sufficiently far aloft of the boundary-layer height, a no-penetration boundary condition is used and the horizontal components of the wind are geostrophic, which is a Dirichlet-type boundary condition.
 At the bottom, a constant-flux layer is assumed and Monin--Obukhov similarity theory (MOST) is used to calculate the surface momentum fluxes.
-The Navier-Stokes equations are integrated using a 3$^\text{rd}$-order low-storage Runge--Kutta method.
-For scalar advection a 5$^\text{th}$-order Wicker--Skamarock scheme is employed.
+The Navier-Stokes equations are integrated using a 3rd-order low-storage Runge-Kutta method.
+For scalar advection a 5th-order Wicker--Skamarock scheme is employed.
 The Poisson equation is solved using a direct fast Fourier transform (FFT).
-In LES, the turbulent transport on the subgrid scale (SGS) needs to be modeled by an SGS closure model.
-We use two kinds of SGS closures to assess their impact on the LES solution: a 1.5-order closure after \cite{deardorff1980stratocumulus} and a dynamic closure after \cite{heinz2008realizability}.
+In LES, the turbulent transport on the subgrid scale (SGS) needs to be modeled by a dedicated model, the SGS model.
+We use two different SGS models to check whether the subgrid closure significantly influences the LES solution: a 1.5-order closure after \cite{deardorff1980stratocumulus} and a dynamic closure after \cite{heinz2008realizability}.
 For most of the simulations, the 1.5-order closure is used, since the dynamic closure needs more computational resources, but the simulations with $\Delta^-=200^{-1}$ are repeated using the dynamic closure.
+A comprehensive description of the LES model is given by \cite{maronga2020overview}.
 
@@
 In part I of this publication, the DNS of $Re_D\leq1\,600$ is carried out for a horizontal domain size of $(0.54\Lambda_{Ro})^2$ - $(1.08\Lambda_{Ro})^2$, where $\Lambda_{Ro}=G/f$ is the Rossby radius.
-For these Reynolds numbers, $Z\approx 0.05$, such that $L_x \approx 10 \delta$.
-However, $Z$ decreases logarithmically with increasing Reynolds number and it is $Z\approx 0.02$ for $Re_D=10^6$.
+For these Reynolds numbers, $u_\star/G$ is around 0.05 so $L_x \approx 10 \delta$ However, $u_\star/G$ decreases with increasing Reynolds number and is around 0.02 for $Re_D=10^6$.
 A domain size of half the Rossby radius would then extend to $L_x\approx 25\delta$.
-Such a large domain would imply immense computational cost.
+Such a large domain would imply immense computational costs.
 \cite{spalart2008direct} used a horizontal domain of $L_x = 2\delta$, arguing that this length allows the resolution of the largest outer-layer eddies according to \cite{csanady1967resistance}.
@@
 Thus, molecular friction is not considered directly, but only by virtue of a turbulence model linking the resolved and dissipative scales.
-In their seminal works on the spectral energy transfer in homogeneous isotropic turbulence, \cite{kolmogorov1941dissipation} and \cite{obukhov1941distribution} showed that the energy transfer rate across the spectrum is in fact constant provided that both dissipation and production terms can be neglected.
+In their seminal works on the spectral energy transfer in homogeneous isotropic turbulence, Kolmogorov and Obukhov (\cite{kolmogorov1941dissipation},\cite{obukhov1941distribution}) showed that the energy transfer rate across the spectrum is in fact a constant.
 This implies that the transfer rate across the cut-off scale in LES does not depend on the magnitude of the viscous scale, presupposed that (i) the cut-off scale of the LES is well within the inertial range and (ii) the LES turbulence is approximately isotropic and homogeneous at the smallest resolved scales.
@@
 When this is the case, $\nu$ cannot be ignored anymore, and we let $K_m = c_0 \Delta \sqrt{e} + \nu$.
-Hence, (absent stratification) the governing equation of for momentum in our version of PALM is
+The governing equation of the velocity components in PALM reads
 \begin{align}\label{gov_palm}
-  \frac{\partial u_i}{\partial t} = - \frac{\partial u_i u_j}{\partial x_j} -\epsilon_{ijk}f_ju_k + \epsilon_{i3j} f_3 u_{g,j} - \frac{1}{\rho_0}\frac{\partial\pi^*}{\partial x_i} + \frac{\partial}{\partial x_j}\left( K_m\left(\frac{\partial u_i}{\partial x_j} + \frac{\partial u_j}{\partial x_i} \right)\right).
+  \frac{\partial u_i}{\partial t} = - \frac{\partial u_i u_j}{\partial x_j} -\epsilon_{ijk}f_ju_k + \epsilon_{i3j} f_3 u_{g,j} - \frac{1}{\rho_0}\frac{\partial\pi^*}{\partial x_i} + \frac{\partial}{\partial x_j}\left( K_m\left(\frac{\partial u_i}{\partial x_j} + \frac{\partial u_j}{\partial x_i} \right)\right)
 \end{align}
-In the limit of either well-resolved simulations or low $\RE$ (or both), $K_m \rightarrow \nu$, so the last term on the right hand side of eq.
-(\ref{gov_palm}) becomes $\nu\frac{\partial^2u_i}{\partial x_j^2}$, which complies with the Navier-Stokes equations for an incompressible fluid underlying our reference solution from DNS.
+for a neutrally stratified flow.
+In the limit of either very well resolved simulations or very low $\RE$ (or both), $K_m \rightarrow \nu$, so the last term on the right hand side of eq.
+(\ref{gov_palm}) becomes $\nu\frac{\partial^2u_i}{\partial x_j^2}$, which complies with the Navier-Stokes equations for an incompressible fluid.
 
@@
 
-In contrast to the interior closure of the LES where the direct effect of $\nu$ is of small relevance compared to the eddy viscosity---at least for high $\RE$---, this is not true for the viscid effects at the bottom boundary:  A constant flux layer is assumed in between the first grid point and the lowermost half point (on the Arakawa-C staggered grid) to estimate friction velocity and surface stresses by MOST:
+In contrast to the interior closure of the LES where the direct effect of $\nu$ is of small relevance compared to the eddy viscosity---at least for high $\RE$---, this is not true for the viscid effects at the bottom boundary.
+Here, a constant flux layer is usually assumed below the first grid point and the Monin--Obukhov similarity theory is used to compute the friction velocity and the stresses at the first half grid point (where the horizontal velocities are located on the Arakawa-C staggered grid):
 \begin{align}\label{most}
   u_\star = \frac{\kappa(U^2+V^2)^{0.5}}{\ln\left(z/z_0\right)},\\
-  -\overline{u''w''}_0 = \frac{\kappa Uu_\star}{\ln\left(z/z_0\right)}, 
+  -\overline{u''w''}_0 = \frac{\kappa Uu_\star}{\ln\left(z/z_0\right)}.
 \end{align}
-where double dashes refer to the unresolved, i.e.
-SGS-parameterized, fluctuations.
 In these expressions, viscosity enters indirectly by virtue of the roughness length $z_0$ when considering the law of the wall for a smooth surface:
@@
 \end{align}
-%This leads to the expression
-%\begin{align}
-%	z_0^+ = z_0\frac{u_\star}{\nu} = \exp\left(-\kappa C^+\right),
-%\end{align}
-For aerodynamically smooth flow, it is $z_0^+ = 0.1031$ based on Eq.~\ref{eqn:log-law}.
-(This is known to be the minimal roughness length of a turbulent boundary layer; see e.g.
-\citealp{kraus2008grundlagen}.) Hence,
+This leads to the expression
+\begin{align}
+	z_0^+ = z_0\frac{u_\star}{\nu} = \exp\left(-\kappa C^+\right),
+\end{align}
+and for an aerodynamically smooth flow, $z_0^+ = 0.1031$ (using the same values as in eq.
+\ref{log-law}).
+This is known to be the minimal roughness length of a turbulent boundary layer (see e.g.
+\cite{kraus2008grundlagen}).
+The roughness length in SI-units
 \begin{align}
@@
 \end{align}
-depends on the viscosity $\nu$ of the fluid. In other words, given fixed surface properties, a choice for $z_0$ implies a particular value for the viscosity $\nu$.
+hence depends on the viscosity of the fluid, which means that---given fixed surface properties---the choice of $z_0$ implies a particular value for the viscosity $\nu$.
 
@@
 In the case of very fine resolution, the first grid point might fall into the buffer layer or even the viscous layer of the flow, so the equations of MOST are no longer adequate to calculate the local stress.
-To avoid that, we follow the recommendation of \cite{kawai2012wall} and use the horizontal velocity from a higher layer $z_{sl}$ to compute the mean stress in the constant flux layer.
-Further, we adopt the boundary condition suggested by \cite{maronga2020improved} and use domain averaged velocities for the estimation $u_\star$:
+To avoid that, we follow the recommendation of \cite{kawai2012wall} to use the horizontal velocity from a higher layer $z_{sl}$ to compute the mean stress in the constant flux layer.
+Furthermore, we adopt the boundary condition suggested by \cite{maronga2020improved} and use the domain averaged velocities:
 \begin{align}\label{us_mean}
-  u_\star \approx u_{*,mean} = \frac{\kappa<u_h(z_{sl})>}{\ln(z_{sl}/z_0)},
+  u_{*,mean} = \frac{\kappa<u_h(z_{sl})>}{\ln(z_{sl}/z_0)},
 \end{align}
@@
 The mean stress is then used as a boundary condition at the first grid point ($z=z_1$).
-It is distributed locally to the unresolved stresses in x- and y-direction via
+It is distributed locally to the stresses in x- and y-direction via
 \begin{align}
-  \overline{u''w''}_0(x,y,z_1) = -u_{*,mean}^2 \frac{u(x,y,z_1)}{\sqrt{<u>^2(z_1)+<v>^2(z_1)}},
+  \overline{u'w'}_0(x,y,z_1) = -u_{*,mean}^2 \frac{u(x,y,z_1)}{\sqrt{<u>^2(z_1)+<v>^2(z_1)}},
 \end{align}
-and accordingly for $\overline{v''w''}_0$.
+and accordingly for $\overline{v'w'}_0$.
 This way, the domain average of the stress components yield the total stress of eq.
-\ref{us_mean} ($u_{*,mean}^2 = \sqrt{<\overline{u''w''}_0>^2+<\overline{v''w''}_0>^2}$).
-For the reference height we use the grid level closest to $z_{sl}^- \approx 0.1$.
-This alleviates two problems: First, $z_{sl}$ is within log-layer.
-Second, the flow at $z=z_{sl}$ is better resolved than closer to the surface.
+\ref{us_mean} ($u_{*,mean}^2 = \sqrt{<\overline{u'w'}_0>^2+<\overline{v'w'}_0>^2}$).
+As reference height we use $z_{sl}^- \approx 0.1$ (depending on the height of the closest grid level).
+By using this higher reference height for the boundary condition we solve two problems: First, the reference height is inside of the logarithmic layer, second, we use a velocity from a region where the flow is much better resolved than close to the surface.
 
@@
 	\caption{LES results: $\Delta^+$ is the grid cell size in wall units, $u_\star$ the friction veloctiy, $\alpha_\star$ the angle between geostrophic wind and negative surface shear stress, $\kappa_{LES}$ the Kármán-measure of the logarithmic layer (estimated by linear regression), $C^+$ the intercept (see eq.
-\ref{eqn:log-law}), and $\delta_{95}$ the height estimated by a linear interpolation of the shear stress reduction to 95\% of its surface value}
+\ref{log-law}), and $\delta_{95}$ the height estimated by a linear interpolation of the shear stress reduction to 95\% of its surface value}
 		\begin{tabular}{lcccccc}
@@
 % z0
-The geostrophic drag $Z=u_\star/G$ quantifies the conversion of mean-flow kinetic energy ($G^2$) to turbulence ($u_{\star}^2$).
-The ratio of $u_{\star}$ resulting from the simulations to the theoretical value $u_{\star,th}$ are shown in Tab.
+A key parameter governing the turbulent state of the flow is the geostrophic drag: it puts the surface friction in relation to the geostrophic wind and thus illustrates the conversion of mean-flow kinetic energy to turbulence.
+The ratio of $u_{\star}$ resulting from the simulations to the theoretical value $u_{\star,th}$ are shown in Tbl.
 \ref{simulation_results}.
@@
 From a top perspective, the horizontal velocity increases from its geostrophic value to the supergeostrophic maximum and then decreases with decreasing height.
-The horizontal mean velocity at the grid point closest to $z^-=0.1$ is used to calculate $u_\star$ according to Eq.~(\ref{us_mean}).
+The horizontal mean velocity at the grid point closest to $z^-=0.1$ is used to calculate $u_\star$ according to eq.
+\ref{us_mean}.
 It is remarkable that the choice of $z_0^+=0.1031$ leads to a value of $u_{\star}$ very close to the prediction by semi-empirical considerations (cf.
@@
 %alpha
-The surface veering $\alpha_\star$ is key to quantify the Ekman pumping and of central relevance for larger-scale applications.
-For the LES, we use the first derivatives of the velocity components calculated by 3$^\text{rd}$-order forward finite differences to estimate $\alpha_\star$:
-\begin{align}
-  \alpha_\star = \arctan\left(\left.\left\langle \frac{\partial v}{\partial z}\right\rangle_0\right/\left\langle \frac{\partial u}{\partial z}\right\rangle_0\right)-\arctan\left(\frac{V_G}{U_G}\right).
-\end{align}
-Nearly all LES yield an $\alpha_\star$ smaller than the theoretical value $\alpha_{\star,\text{th}}$ (cf.
-Tab.~\ref{simulation_results}), where we observe a slight increase of $\alpha_\star$ with resolution; The finest resolution of Re2 and Re3 generally align closely with the theoretical direction while for Re1\_200 the offset is 0.9°.
-The hodograph (Fig.~\ref{3Re_hodograph}) illustrates that this is not caused by poor quality of the simulation since all of the simulations of Re1 closely follow the theoretical hodograph; in fact, the turning angle at the respective height is actually slightly overestimated.
-Rather, the offset is due to the veering of the wind vector continues in the lower part of the boundary layer that is not present in the LES.
+Another key parameter of the Ekman flow is the surface veering angle $\alpha_\star$ between the directions of the negative surface shear stress and the geostrophic wind.
+For the LES, the direction of the negative stress is represented by the first derivatives of the velocity components calculated by 3rd order forward finite differences: $\alpha_\star = \tan^{-1}(\langle \partial v/\partial z\rangle_0/\langle \partial u/\partial z\rangle_0)-\tan^{-1}(V_G/U_G)$.
+Table \ref{simulation_results} shows that nearly all LES yield an $\alpha_\star$ smaller than the theoretical value $\alpha_{\star,th}$.
+In general, we observe an increase of $\alpha_\star$ with resolution.
+The finest resolution of $Re2$ and $Re3$ generally align closely with the theoretical direction while for $Re1\_200$ still 0.9° is missing.
+Fig.
+\ref{3Re_hodograph} illustrates that this is not caused by poor quality of the simulation since all of the simulations of $Re1$ follow the theoretical hodograph quite closely and even overestimate the angle at the respective height by a little.
+In contrast to the higher $\RE$, the veering of the wind vector continues in the lower parts of the boundary layer.
 Hence, for the low-$\RE$ case, the general observation of a higher $\alpha_\star$ with finer resolution is also caused by the approach of the final $\alpha_\star$ with decreasing height of the first grid point.
-In general, the finer resolved simulations show veering angles that are very close to the theoretical prediction.
+In general, the finer resolved simulations show veering angles that are very close to the theoretic predictions.
 
@@
 	}
-  \caption{Shear-aligned velocity profiles in outer scaling.
-Solid lines: u-component, dashed lines: v-component.
-The lowest grid point is marked by a cross ($\times$) colored according to the corresponding resolution.}
+  \caption{Shear-aligned velocity profiles in outer scaling. Solid lines: u-component, dashed lines: v-component. The lowest grid point is marked by $x$}
   \label{velocity_outer}
@@
 % profiles
-Velocity profiles from LES and our theoretical framework agree well for the different $\RE$ in Fig.\,\ref{velocity_outer}: Most prominently, the rotation across the Ekman layer agrees and a logarithmic layer for the U-component is present.
+Figure \ref{velocity_outer} shows the velocity profiles from the LES and the theoretical profiles for each of the $\RE$.
+In general, the velocity profiles of the LES agree well with the theoretical profiles: The course of the theoretical Ekman layer is matched and the simulations exhibit a logarithmic layer for the U-component.
 The lowest points of the LES of the higher $\RE$ fall into the logarithmic layer.
 For Re1, the lowest grid point falls into the buffer layer, which is visible as the curved course of the U-component below the logarithmic layer in the theoretical profile.
-The best resolved simulation with Deardorff-closure even seems to follow the course of the upper part of the buffer layer, but with a resolution of $\Delta^+=15$ this is a coincidence caused by a well-known S-shape of velocity profiles close to rigid wall \citep{brasseur2010designing}.
+The best resolved simulation with Deardorff-closure even seems to follow the course of the upper part of the buffer layer, but with a resolution of $\Delta^+=15$ this is a coincidence caused by a well-known S-shape of velocity profiles close to rigid walls \cite{brasseur2010designing}.
 This log-layer mismatch arises from a competition between the scales $u_\star$ and $z$ and other velocity and length scales introduced by the discretization of the dynamical system \citep{mason1992stochastic,brasseur2010designing}.
 In other words, at the lower boundary, the relevant eddies are too small to be resolved by the grid and their contribution to the flow has to be modeled.
-Also the vertical component is restricted by the wall's impermeability, known as blocking effect-- both effects illustrate a violation of isotropy near the wall and thus counter the SGS-closure assumptions.
+Also the vertical component is massively restricted by the non-permeability of the wall, known as blocking effect.
+On the contrary, the SGS-closure assumes isotropic turbulence, which is not the case on the grid scale near the wall.
 
@@
   \caption{K\'arm\'an measure $\kappa$ in the logarithmic region and above for different Reynolds numbers and resolutions}
-  \label{fig:3Re_kappa}
+  \label{3Re_kappa}
 \end{figure}
@@
 % kappa
-The goodness-of-fit in the logarithmic layer (Eq.\,\ref{eqn:log-law}) is assessed via the K\'arm\'an measure~(Fig.\ref{fig:3Re_kappa})
-\begin{align}
-  \kappa(z)=(z^+\partial U^+/\partial z^+)^{-1}\simeq const.
-\end{align}
-which is estimated here by linear regression of the mean velocity in x-direction against the logarithm of height between the seventh grid point and the $z^-=0.1$ for Re1 and $z^-=0.12$ for Re2 and Re3.
-The seventh grid point is chosen as lower boundary for the logarithmic layer as, in PALM, the mean velocity profile follows MOST aloft \citep{maronga2020improved}.
-The number of values for each regression is 6, 12, and 18 for ReX\_100, ReX\_150 and ReX\_200, respectively.
-In case Re1, the viscous sublayer occupies about $0.5\%$ of the boundary layer, but this share drops to $~10^{-5}-10^{-7}$ for cases Re2 and Re3 respectively.
+In the logarithmic region, the profile of the u-component should follow the logarithmic law of eq.
+\ref{log-law} with the K\'arm\'an measure $\kappa = (z^+\partial U^+/\partial z^+)^{-1}=const.$.
+In the low-Reynolds-number case, the viscous sublayer represents about $0.5\%$ of the boundary layer, while this layer is not visible for the high-Reynolds-number cases, where the share drops to $~10^{-5}-10^{-7}$.
 Above the viscous sublayer, the theoretical profile shows a near-constant value for $\kappa$ up to $z^-\approx0.1$ for the case Re1 and up to $z^-\approx0.12$ for the cases Re2 and Re3.
-We do not compute $\kappa$ for ReX\_50 since there are only six grid points below $z^-\approx 0.12$, and o constant K\'arm\'an measure $\kappa(z)$ can be observed.
-The resulting best estimates for $\kappa_{LES}$ are shown in Tab.\,\ref{simulation_results}, where the coefficient of determination ('r-squared') is above 0.99 for all cases.
-
-For finer resolution, $\kappa_{LES}$ decreases with increasing Re for the range of scale separation investigated here.
-ReX\_100 yields a value around $0.53$ (which is outside the range agreed on by atmospheric observation and laboratory measurements), the finer resolution yields values around $0.46$ for Re1\_200 and around $0.44$ for Re2\_200 and Re3\_200, which is in excellenet agreement with DNS data and laboratory measurements.
-The dynamic subgrid closure yields lower values for $\kappa_{LES}$: in Re1\_dyn we even see $\kappa_{LES}=0.39$, while Re2\_dyn and Re3\_dyn yield $0.42$ and $0.43$, respectively.
+In order to estimate $\kappa_{LES}$ from the simulations, we perform a linear regression of the mean velocity in x-direction and the logarithm of height between the seventh grid point and the grid point corresponding to the height $z^-=0.1$ for $Re1\_X$ and $z^-=0.12$ for $Re2\_X$ and $Re3\_X$.
+We consider only points above the 7th grid point following arguments of \cite{maronga2020improved}, that in PALM, the mean velocity profiles follow MOST above the seventh grid point.
+The number of values for each regression is 6, 12, and 18 for $ReX\_100$, $ReX\_150$ and $ReX\_200$, respectively.
+We do not compute $\kappa$ for $ReX\_50$ since there are only six grid points below $z^-\approx 0.12$.
+The resulting $\kappa_{LES}$ for the other simulations are shown in Tbl.
+\ref{simulation_results}.
+The coefficient of determination is above 0.99 for all fits.
+$\kappa_{LES}$ is decreasing with finer resolution for all $\RE$.
+While the values for $ReX\_100$ are rather high with a value around $0.53$, finer resolution leads to values closer to measurements around $0.46$ for $Re1\_200$ and around $0.44$ for $Re2\_200$ and $Re3\_200$.
+The dynamic subgrid closure yields lower values for $\kappa_{LES}$: in $Re1\_dyn$ we even see $\kappa_{LES}=0.39$, while $Re2\_dyn$ and $Re3\_dyn$ yield $0.42$ and $0.43$, respectively.
+Figure \ref{3Re_kappa} shows the K\'arm\'an measure in the logarithmic layer.
+For the simulations $ReX\_50$, no constant $\kappa$ can be observed.
+An increase in resolution leads to a profile of the K\'arm\'an measure closer to the theoretical curve for all cases.
 Close to the bottom, the K\'arm\'an measure is heavily influenced by the proximity of the wall and rather a function of the vertical index than of the physical distance from the wall.
-In accordance with \cite{maronga2014monin} and \cite{maronga2017formulation}, the kinks in the K\'arm\'an measure diminish around the seventh grid point for all simulations.
-Above, the curve smoothens and approaches the expected value of $\kappa$.
+In accordance with the observations of \cite{maronga2014monin} and \cite{maronga2017formulation}, the kinks in the K\'arm\'an measure diminish around the seventh grid point for all simulations.
+Above, the curve smoothens and approaches the expected value of $\kappa$ (at least for the finer resolutions).
 
@@
 
-The resolution near the surface is critical for an LES with isotropic grid spacing. \todo{Missing transition from resolution to two-point correlation!} The two-point correlation is defined as 
+As expected for an LES with isotropic grid spacing, the resolution near the surface is a critical part of the simulation.
+We can support our above observations by taking a look at the two-point correlation in the critical layers of the simulations.
+The two point correlation is defined as
 \begin{align}
-  B_{xx}\left(x^*\right) = \langle u\left(x-x^*\right)u\left(x\right)\rangle/\sigma_u^2,
+  B_{xx}\left(x^*\right) = \langle u\left(x-x^*\right)u\left(x\right)\rangle/\sigma_u^2.
 \end{align}
-and it supports the above findings \todo{which findings?}. 
-For an LES with isotropic grid and periodic horizontal boundaries, the turbulent structures of the velocity's w-component in the y-direction indicate how well the flow is resolved \citep{wurps2020grid}.
-Fig.\,\ref{2pc_008} shows the two-point correlation $B_{ww}(y^*)$ at a height of $z^-\approx 0.08$, which is right inside of the logarithmic layer.
+As suggested by \cite{wurps2020grid}, in an isotropic grid with periodic horizontal boundary conditions the turbulent structures of the velocity's w-component in y-direction can serve as an indicator how well the flow is resolved.
+Figure \ref{2pc_008} shows the two-point correlation $B_{ww}(y^*)$ at a height of $z^-\approx 0.08$, which is right inside of the logarithmic layer.
 We define the average size of a structure in the flow $\sigma$ as the distance where the two-point correlation drops to 0.3 (horizontal line in figure).
@@
 
-The theoretical formulation of mean velocity profiles within the turbulent Ekman layer can serve as benchmark for model frameworks such as LES.
+The theoretical formulation of mean velocity profiles within the turbulent Ekman layer can serve as comprehensive benchmark for model frameworks such as LES.
 These reference profiles offer a deeper analysis beyond the conventional assessment of the logarithmic increase in wind speed within the Prandtl layer.
@@
 \bibliographystyle{plainnat}
-\bibliography{papers.bib,ansorge.bib}
+\bibliography{papers.bib}
 

### 2024-07-18 write1/manuscript.tex
added data citations to part I

@@
 \usepackage{hyperref}
+\usepackage{caption}
 \setcitestyle{aysep={}}
@@
 scales ($u_\star, \nu$).
+%
+Another common measure of scale separation is the Reynolds number defined by the laminar Ekman layer
+thickness $D=\sqrt{2\nu/f}$, 
+\begin{equation}
+  \RE_D = \frac{G D} {\nu } = \sqrt{2 \RE_\Lambda}. 
+\end{equation}
 %
@@
   f_{V,\mathrm{log}}(z^+) = \frac{V_\mathrm{log}(z^+)}{G} \delta^+ = a_\mathrm{log} + b_\mathrm{log} \log z^+ + c_\mathrm{log} z^+. 
+  \label{eqn:abc_log}
 \end{equation} 
@@
 \begin{figure}
-  \includegraphics[width=0.5\textwidth]{../plot/c_log.pdf}
-  \caption{c-log.
-    \label{fig:clog}}
+  \centerline{\includegraphics[width=0.5\textwidth]{../plot/c_log.pdf}}
+  \caption{Coefficients $a_\text{log}$, $b_\text{log}$ and $c_\log$ (cf. Eq.~\ref{eqn:abc_log}) as a function of 
+    the viscous Reynolds number $Re_D$. The approximate range of scale separation relevant for atmospheric application 
+    is found in between the dotted lines, where $a_\text{log}\simeq 0$ and $c_\text{log})\simeq 0.4$ }
+  \label{fig:clog}
 \end{figure} 
@@
 the variation of $c_\mathrm{log}$ is, however, rather small.
+
+\subsection{Comparison with other theories}
+
+An alternative approach that considers viscous effects close to the surface is the van-Driest scaling \citep{vandriest:JOT1956}, 
+where an exponential damping of Prandtl's mixing length is considered near the wall to yield 
+\begin{align}
+  \frac{\p u^+}{\p y^+} = \frac{2}{ 1+ \sqrt{1+\left(2\kappa y^+\right)^2 (1-\exp\left[-y^+\right])}};
+\end{align}
+the spanwise component is zero as no rotational effects are considered. 
+%
+Comparing our proposed formulation for the stream-wise velocity in the inner layer to  \citeauthor{vandriest:JOT1956}'s formulation 
+yields later convergence of the velocity onto the logarithmic profile while, over all, it serves as an excellent model of the streamwise 
+velocity component:
+% 
+Notable deviations (on the order of few percent) only occur in the region $10 < z^+ < 30$, 
+where the velocity transitions from the linear to the logarithmic profile. 
+\begin{figure}
+  \centerline{\includegraphics[width=0.66\textwidth]{../plot/vanDriest_profile.png}}
+  \caption{Near-wall velocity profile according to the 
+    van-Driest scaling (blue, solid) in comparison with 
+    the present model (red, dashed), 
+    the viscous law of the wall (orange dashed), and 
+    the logarithmic law (green, solid)} 
+  \label{fig:vanDriest}
+\end{figure}
+
+\textbf{Compare} to Emeis (2002) and Gryning (2007); highlight explicit knowledge on veering-profile $\rightarrow$ directional sheer;
+
+\textbf{Remember} of interpretation in the context of eddy viscosity (Fig.~\ref{fig:stresses}) 
+Consider Townsend ``Turbulent Shear Flow''  Chapter 7.18: p. 319 
 %
@@
 
-\subsection{Comparison with other theories}
-
-\textbf{Compare} to Emeis (2002) and Gryning (2007); highlight explicit knowledge on veering-profile $\rightarrow$ directional sheer;
-
-\textbf{Compare} to van Driest scaling
-
-\textbf{Remember} of interpretation in the context of eddy viscosity (Fig.~\ref{fig:stresses}) 
-Consider Townsend ``Turbulent Shear Flow''  Chapter 7.18: p. 319 
 %

### 2024-07-18 plot/plot.py

@@ -23,14 +23,14 @@ plot_evisc=False
 plot_profiles=False
 plot_visc_outer=False
 print_table=False
-plot_outer_log=True
+plot_outer_log=False
 plot_convergence=False
 plot_re1600=False
 plot_les_comp=False
 plot_applications=False
 test_inner_streamwise=False
-plot_profile_comparison=True
-plot_vandriest=False
+plot_profile_comparison=False
+plot_vandriest=True
 plot_total_rot=False
 
 colors = {400 : 'gray',
@@ -2059,7 +2059,7 @@ if plot_profile_comparison:
 
 if plot_vandriest:
     KAPPA = sc.KAPPA
-    ASTAR=28.7
+    ASTAR=26.8
     fig=plt.figure(figsize=(5,3))
     ax=fig.add_axes([0.12,0.16,0.8,0.82])
 
@@ -2082,7 +2082,7 @@ if plot_vandriest:
     fig=plt.figure(figsize=(5,3))
     ax=fig.add_axes([0.12,0.16,0.8,0.82])
 
-    C0=5.52
+    C0=5.42
 
     u =np.zeros(len(zp)) 
     for i in range(len(zp)-1):
@@ -2103,7 +2103,7 @@ if plot_vandriest:
 
     ax.set_xscale('log') 
     ax.set_xlim(1,1e3)
-    ax.set_ylim(1,15)
+    ax.set_ylim(1,20)
     leg=plt.legend(loc='best')
     
     plt.savefig('vanDriest_profile.png',format='png',dpi=600)

### 2024-08-05 write2/manuscript_part2_2024.tex
minor edits

@@
-\documentclass[a4paper,11pt]{article}
+\documentclass[smallcondensed,draft]{svjour3}
 
-\usepackage[margin=2.5cm]{geometry}
+\usepackage{lmodern}
+%\usepackage[margin=2.5cm]{geometry}
 \usepackage{booktabs} 
@@
 \usepackage{graphicx}
-\usepackage{caption}
-\usepackage{natbib}
+%\usepackage{caption}
+\usepackage[authoryear]{natbib}
 \usepackage{xcolor}
 \usepackage{tikz}
+
+\usepackage{adjustbox}
 
@@
 \newcommand{\RE}{\mathbf{Re}}
+
 \hyphenation{Ol-den-burg}
-
+\hyphenation{non-sta-tio-na-ri-ty}
 \DeclareRobustCommand\sampleline[1]{%
@@
 
-\author{Hauke Wurps and Cedrick Ansorge}
+\author{Hauke Wurps $^{1}$ and Cedrick Ansorge $^{2}$}
+\authorrunning{H. Wurps and C. Ansorge}
 
-\date{Draft \today}
-\title{Towards a Universal Veering Profile for Turbulent Ekman Flow at arbitrary Reynolds number - Part 2\\
-  \normalsize LES and DNS of Turbulent Ekman Flow \\}
+\institute{$^{1}$ Uni Oldenburg \newline 
+$^{2}$ FU Berlin, Institut f\"ur Meteorologie, Carl-Heinrich-Becker-Weg 6--10, 12165 Berlin \email{cedrick@posteo.de}}
+
+\date{Submitted \today}
+\title{Wind veer in Turbulent Ekman Flow - Part 2: Extrapolation to atmospheric scale separation}
+\titlerunning{A Universal veering profile for Ekman flow - Part 2}
 
@@
   % Introduction % Theoretical profile
-  Turbulent Ekman flow is a canonical flow representation of the atmospheric boundary layer, characterized by a logarithmic increase of the wind speed near the surface and a turning of the wind vector aloft.
-In part I of this work, we derived a formulation of the mean wind velocity profiles of the stationary turbulent Ekman flow based on direct numerical simulation (DNS).
-Here, we explore the extrapolation of these profiles to atmospheric Reynolds numbers using large-eddy simulation (LES).
+  The atmospheric boundary layer is characterized by a logarithmic increase of the wind speed in the surface layer and a turning of the wind vector aloft.
+The simplest canonical configuration featuring both \emph{a logarithmic layer and the wind turning} is Ekman flow.
+A formulation of the mean velocity profile was derived from scaling considerations and calibrated using direct numerical simulation in part I of this work.
+Here, we explore the extrapolation to atmospheric Reynolds number (Re) using large-eddy simulation (LES).
   % Case desription and setup
-  We compare theoretical profiles of the wind vector for low Reynolds-number turbulent Ekman flow to results from DNS and LES.
-This necessitates the inclusion of viscous effects, usually neglected in LES, along with modifications to the standard bottom boundary condition.
+  Theoretical profiles of the wind vector are compared to the theoretical reference at intermediate Re, which necessitates consideration of viscous effects % [not required in abstract] (usually neglected in LES)
+  along with modifications to the standard bottom boundary condition.
   % Results and Discussion
-  Analysis of the grid-convergence for higher resolution unveils a convergence of LES data towards the theoretical profiles for both intermediate and high Reynolds numbers.
-Our LES thus confirm that--within the logarithmic layer (the lowest 10\,\% of the boundary layer)--- the turning of the wind vector rotates by approximately one-third of its total rotation across the entirety of the boundary layer.
+  Analysis of the grid convergence for higher resolution unveils a convergence of LES data towards the theoretical profiles for intermediate and high Re.
+The LES thus confirm that (i) the spanwise velocity scales as $\left.Re_\tau\right.^{-1}$, where $Re_\tau$ is the friction Reynolds number, and (ii) one third of the wind veer is confined to the surface layer.
+  % [this level of detail is not needed in the abstract] (the lowest 10\,\% of the boundary layer) 
   % Conclusions
-  Such agreement of our theoretical formulation and LES data raises our confidence in the underlying assumptions of the theory and thus reinforces the utility of the theoretical profiles as a reference for intermediate and a quasi-reference for higher Reynolds number simulations.
+  Such agreement of the theoretical formulation and LES data raises confidence in the underlying scaling assumptions and thus reinforces the utility of the theoretical profiles as a reference for intermediate and a quasi-reference for both idealized simulation and field observation.
 %
@@
 
-% \todo{start introduction rather with TEF than LES (rather subject than method)}
 The interaction of atmospheric boundary layer  (ABL) flow with its lower boundary condition, the surface, is a defining property of the ABL \citep{stull:1988}.
-Relevant aspects of the interaction with the boundary not only include its heterogeneity \citep{avissar:MWR1989,giorgi:RG1997,claussen:BM1991,garratt:BM1990} and roughness \citep{monin:ARF1970,brutsaert:WRR1975,raupach:AMR1991,kostelecky:JFM2024}, but also a wind veer in favor of the pressure gradient force when the surface is approaced \citep{ekman:AMA1905}.
-An important tool with capabilities to model the flow-surface interaction is Large-Eddy Simulation (LES).
-Indeed, LES is widely used to model turbulent flow in the atmospheric boundary layer (ABL) \citep{stoll2020large}.
-This includes various applications, such as simulating wind turbine and wind farm wakes, and interactions within wind farm clusters \citep{porte2011large,mehta2014large,breton2017survey}.
-Furthermore, LES is applied in complex environments like mountainous or urban areas \citep{stoll2020large,garcia2018predictive} and to investigate the dispersion of pollution \cite{han2018large}.
+Relevant aspects of the interaction with the boundary not only include its heterogeneity \citep{avissar:MWR1989,giorgi:RG1997,claussen:BM1991,garratt:BM1990} and roughness \citep{monin:ARF1970,brutsaert:WRR1975,raupach:AMR1991,kostelecky:JFM2024}, but also a wind veer in favor of the pressure gradient force when the surface is approached \citep{ekman:AMA1905}.
+
+Large-Eddy Simulation (LES) is widely used to study the turbulent flow in the atmospheric boundary layer (ABL) for a wide range of applications \citep{stoll2020large}.
+This includes the simulation of wind turbines and wind farm wakes, their interactions within wind farm clusters \citep{porte2011large,mehta2014large,breton2017survey}.
+Furthermore, LES is applied in complex environments like mountainous or urban areas \citep{stoll2020large,garcia2018predictive} and to investigate the dispersion of pollution \citep{han2018large}.
 These applications typically focus on the lower segment of the ABL on the rotating earth, known as the Prandtl or surface layer.
@@
 
-The simplest representation of the ABL taking into account rotation \emph{and} the quasi non-turbulent, free atmosphere aloft is the turbulent Ekman flow.
+The simplest representation of the ABL taking into account rotation and the quasi non-turbulent, free atmosphere aloft is Ekman flow.
 It is a horizontally homogeneous, statistically stationary boundary-layer flow over a rotating flat surface.
-Here we consider the problem under neutral stratification, where the potential temperature is constant across the whole domain.
-Turbulent Ekman flow shows the key characteristics shaping the real ABL: a logarithmic layer and the Ekman spiral.
-Near the surface, viscous forces dominate the flow.
-The stationary ABL-height is then defined by the interplay between turbulent growth due to the shear instability of the configuration \citep{lilly1966instability} and rotational suppression of turbulence due to the Coriolis effect.
-The system's statistical equilibrium is defined by one single parameter, the Reynolds number $\RE$, signifying the scale separation between the largest and smallest scale within the problem.
-The largest scales are given by shear-induced eddies constrained by the boundary layer's height, while the smallest scales reside within the dissipative range, where strong viscous forces suppress turbulence.
+Here we consider the problem under neutral stratification, where density effects on the flow due to temperature fluctuations are absent.
+This turbulent flow configuration features the key characteristics of the real ABL, a logarithmic layer and the Ekman spiral; in immediate vicinity of the surface, viscous forces dominate (\emph{viscous sublayer}).
+The stationary ABL-height is set by the interplay between turbulent growth due to the shear instability \citep{lilly1966instability} and rotational suppression of turbulence due to the Coriolis effect.
+The system's statistical equilibrium is characterized by a single parameter, the Reynolds number $\RE$, signifying the scale separation between the largest and smallest scales of the problem.
+The largest scales are given by shear-induced eddies constrained by the boundary-layer height, while the smallest scales reside within the dissipative range, where strong viscous forces suppress turbulence.
 The dimensional mean solution of the turbulent Ekman layer is influenced by three parameters: the geostrophic wind $G$, the Coriolis parameter $f$, and the kinematic viscosity $\nu$.
 These three parameters define the Reynolds number $Re_D = GD/\nu$.
-With $D=\sqrt{2\nu/f}$ the Ekman layer depth of a laminar flow we write
-
+Using $D=\sqrt{2\nu/f}$, the Ekman layer depth of a laminar flow, we write
+%
 \begin{align}\label{red}
@@
 \end{align}
-
+%
 The mean solutions of the velocity components are functions of only the Reynolds number and the height above the ground $z$: $u(\RE_D,z)$, $v(\RE_D,z)$ (see e.g. \cite{csanady1967resistance}).
@@
 However, it remains uncertain, if the extrapolation of the above framework towards atmospheric $\RE$ is valid.
-The substantial separation of scales in the ABL renders direct numerical simulation (DNS) of the entire turbulent flow at scale impossible.
+The substantial separation of scales in the ABL makes direct numerical simulation (DNS) of the entire turbulent flow at geophysical scale impossible.
 Within the framework of LES, only the larger eddies are resolved, that contain the main share of kinetic energy, while turbulent mixing below the subfilter-scale is parameterized.
@@
 As LES introduces the grid size as an additional parameter (often understood in relation to a filter scale, cf.
-\cite{pope2004ten}), we conduct a thorough analysis of its impact on simulation results.
-\cite{esau2004simulation} and \cite{jiang2018large} have executed similar LES analyses of neutral Ekman layers, however, they introduce an additional parameter by selecting an arbitrary roughness length $z_0$, whereas we derive the parameter $z_0$ corresponding to a fluid with a kinematic viscosity of $\nu = 1.5\cdot10^{-5}$ over flat surface.
+\cite{pope2004ten}), we analyse its impact on simulation results.
+\cite{esau2004simulation} and \cite{jiang2018large} have executed similar LES analyses of neutral Ekman layers, however, they introduce an additional parameter by selecting an arbitrary roughness length $z_0$, whereas we derive the parameter $z_0$ corresponding to a fluid with a kinematic viscosity of $\nu = 1.5\cdot10^{-5}$ over a flat surface.
 
@@
 part 1 of this study), $Re_D=1.5 \cdot 10^5$ and $10^6$ correspond to the scale separation found in typical atmospheric boundary layers.
-For $Re_D=1600$, we compare DNS and LES results using the exact same forcing and settings.
+For $Re_D=1600$, we compare DNS \citep{ansorge:2024b} and LES results using the exact same forcing and settings.
 This is the first comparison of this kind for the turbulent Ekman flow.
-The LES results converge to the DNS results with decreasing grid cell size.
-This approach raises our confidence in the LES results to a new level for the direct comparison with a resolving simulation.
-Furthermore, it allows us to extend the insight from DNS to atmospheric scale where an approach through DNS is clearly infeasible for computational constraints.
-We observe a clear $\RE$ dependency in the LES results, matching the predictions of the theoretical framework.
 
@@
 %Definitions
-We align the coordinate Ox with the surface stress, Oz points normal to the surface and normal to Ox, and Oy in the span-wise direction, normal to Oxz.
-%The mean velocity profiles for all $\RE$ exhibit a dual scaling nature: in proximity to the lower boundary, the velocity profiles collapse when scaled using
+We align the coordinate Ox with the surface stress (\cite{ansorge:BM2019}), Oz points normal to the surface and normal to Ox, and Oy in the span-wise direction, normal to Oxz.
 In proximity to the lower boundary, the boundary layer scales in viscous (inner) units, denoted by the index ``$+$'' ($z^+ = zu_\star/\nu$, $U^+=U/u_\star$, where $U$ is the velocity component in x-direction and $u_\star = \sqrt{\nu \p_z U}$ the friction velocity).
 From the top perspective, the profiles converge when scaled using outer units, denoted by the index ``$-$'' ($z^-=z/\delta$, $U^-=U/G$, where $\delta=u_\star/f$ is the boundary-layer depth scale, $f$ the Coriolis parameter and $G$ the geostrophic wind speed).
-At the lower boundary, the x-axis of the inner units aligns with the shear stress, while the x-axis of the outer units aligns with the geostrophic wind.
+At the lower boundary, the x-axis of the inner units aligns with the negative shear stress, while the x-axis of the outer units aligns with the geostrophic wind.
 The angle between both axes is denoted as $\alpha_\star$, the surface veer of the wind across the boundary layer.
@@
 \label{sec:ekmanlayer}
-The outer layer of the ABL is characterized by a triadic balance between turbulent flux, pressure gradient and the Coriolis; the vertical change of the Coriolis force causes a pronounced height dependence of the wind direction resulting in the Ekman spiral \citep{ekman:AMA1905}.
+The outer layer of the ABL is characterized by a triadic balance between turbulent flux, pressure gradient and the Coriolis force; the vertical change of the Coriolis force causes a pronounced height dependence of the wind direction resulting in the Ekman spiral \citep{ekman:AMA1905}.
 The classic solution employs Ekman dynamics down to the surface---and thus also in the surface layer, where this is a rather poor representation of turbulent mixing.
-Here, we model the surface layer by a lower boundary condition for the Ekman spiral determined from DNS:
+Here, we model the surface layer by a shifted lower boundary condition for the Ekman spiral determined from DNS:
 \begin{subequations}
@@
 
-In the viscous sublayer, the shear-aligned velocity $U^{\alpha_\star}$ is described by the law of the wall $U^{\alpha_\star+} = z^+$ and the span-wise velocity $V^{\alpha_\star}$ is zero by choice of the reference frame
-(the index $\alpha_\star$ indicates the alignment with the direction of the shear stress).
+In the viscous sublayer, the shear-aligned velocity $U^{\alpha_\star}$ is described by the law of the wall $U^{\alpha_\star+} = z^+$ and the span-wise velocity $V^{\alpha_\star}$ is zero by choice of the reference frame (the index $\alpha_\star$ indicates the alignment with the direction of the shear stress).
 Around $z^+=5$, the velocity begins to deviate from linearity, and the buffer layer forms the transition from the viscous to the  logarithmic layer.
@@
   \centerline{
-	\includegraphics[width=\textwidth]{figures_2024/alt_single_theor_Ekman_profiles_2022.pdf}}
+	\includegraphics[width=\textwidth]{figures/alt_single_theor_Ekman_profiles_2024.png}}
   \caption{a) Theoretical shear-aligned velocity profile ($U^+$,$-V^+$) of the turbulent Ekman flow for $Re_D = 2000$ (\sampleline{}) and $Re_D = 4000$ (\sampleline{dashed}) and the different layers (colors).
@@
 \label{setup}
-%
-%\subsection{Notation}
-%
-%\begin{table*}
-	%\centering
-%%	\caption{}
-	%\begin{tabular}{ll}
-    %\toprule
-		%$\Delta$ & grid cell size in LES \\
-		%$\overline{u''w''}_0$ &  \\
-		%$\overline{v''w''}_0$ &  \\
-		%$\delta_{95}$ & height where the turbulent stress decreases to 5\% of its surface value\\
-		%\bottomrule
-	%\end{tabular}
-%\end{table*}
 
@@
 $f$, $\nu$, and $G$ are input parameters and define the Reynolds numbers $Re_D$ and $Re_\tau$, while $u_\star$, $\alpha_\star$, and $\delta = u_\star/f$ are resulting properties of the flow}
-	\begin{tabular}{lcccccccc}
+	\begin{adjustbox}{width=\textwidth}
+    \begin{tabular}{l|ccccc|ccc}
           \toprule
@@
 	  \end{tabular}
+  \end{adjustbox}
 	\label{simulation_parameters}
@@
 An incompressible, turbulent Ekman flow over a hydrodynamically smooth surface is simulated using the Parallel Large Eddy Simulation model (PALM, \citealp{maronga2020overview}).
-We study three different Reynolds numbers Re$_D = 10^3;1.5\times 10^5;10^6$, hereafter Re1, Re2, and Re3, respectively (Tab.~\ref{simulation_parameters}).
+We study three different Reynolds numbers Re$_D = 1600;1.5\times 10^5;10^6$, hereafter Re1, Re2, and Re3, respectively (Tab.~\ref{simulation_parameters}).
 %
-The domain is rotating around the z-axis with an angular velocity corresponding to the Coriolis parameter $f=10^{-4}s^{-1}$.
+The domain is rotating around the z-axis with an angular velocity corresponding to the Coriolis parameter $f=10^{-4}s^{-1}$ (the y-component of the Coriolis acceleration was set to zero).
 The stratification of the flow is truly neutral, i.e., the potential temperature is constant across the whole domain.
@@
 Hence, a no-penetration boundary condition is used for the vertical velocity component while the horizontal components are prescribed according to geostrophic balance--a Dirichlet-type boundary condition.
-At the bottom, a constant-flux layer is assumed and Monin--Obukhov similarity theory (MOST) is used to calculate the surface momentum fluxes.
+Below the first grid point, a constant-flux layer is assumed and Monin--Obukhov similarity theory (MOST) is used to calculate the surface momentum fluxes.
 The Navier-Stokes equations are integrated using a 3$^\text{rd}$-order low-storage Runge--Kutta method.
@@
 \cite{spalart2008direct} used a horizontal domain of $L_x = 2\delta$, arguing that this length allows the resolution of the largest outer-layer eddies according to \cite{csanady1967resistance}.
-During a sensitivity test of the domain size we observed that simulations with domain sizes $L\geq 4\delta$ often tend to accumulate turbulence kinetic energy in the upper half of the the boundary layer.
-This TKE increases over several inertial oscillations with energy mostly on the scale of the domain size.
+During a sensitivity test of the domain size we observed that simulations with domain sizes $L\geq 4\delta$ often tend to accumulate turbulence kinetic energy in the upper half of the boundary layer.
+The TKE increases over several inertial oscillations with energy mostly on the scale of the domain size.
 Such a development was not observed in the DNS.
-We could successfully avoid such a behaviour by using a domain of size $L \approx 3\delta$ in combination with a shifted periodic boundary condition in y-direction, as described by \cite{munters2016shifted}.
+We could successfully avoid such a behavior by using a domain of size $L \approx 3\delta$ in combination with a shifted periodic boundary condition in y-direction, as described by \cite{munters2016shifted}.
 Although for the Ekman flow the direction of the mean flow is only aligned with the x-direction of the grid near the surface, a shift of the boundary condition by $L/3$ significantly helped to suppress the accumulation of TKE in the upper half of the boundary layer.
@@
 
-In LES at low $\RE$ or very high resolution the subgrid eddy viscosities may fall far below the molecular viscosity of air.
+In LES at low $\RE$ or very high resolution the subgrid eddy viscosity may fall far below the molecular viscosity of air.
 In the context of PALM's Deardorff closure, it is $K_m = c_0 \Delta \sqrt{e}$, where $c_0 = 0.1$ \citep{deardorff1980stratocumulus}, $\Delta$ is the grid size and $e$ is the SGS-TKE, calculated by a prognostic equation.
@@
 When this is the case, $\nu$ cannot be ignored anymore, and we let $K_m = c_0 \Delta \sqrt{e} + \nu$.
-Hence, (absent stratification) the governing equation of for momentum in our version of PALM is
+Hence, assuming absent stratification, the governing equation of for momentum in our version of PALM is
 \begin{align}\label{gov_palm}
-  \frac{\partial u_i}{\partial t} = - \frac{\partial u_i u_j}{\partial x_j} -\epsilon_{ijk}f_ju_k + \epsilon_{i3j} f_3 u_{g,j} - \frac{1}{\rho_0}\frac{\partial\pi^*}{\partial x_i} + \frac{\partial}{\partial x_j}\left( K_m\left(\frac{\partial u_i}{\partial x_j} + \frac{\partial u_j}{\partial x_i} \right)\right).
+  \frac{\partial u_i}{\partial t} = - \frac{\partial u_i u_j}{\partial x_j} -\epsilon_{ijk}f_ju_k + \epsilon_{i3j} f_3 u_{g,j} - \frac{1}{\rho_0}\frac{\partial\pi^*}{\partial x_i} + \frac{\partial}{\partial x_j}\left( K_m\left(\frac{\partial u_i}{\partial x_j} + \frac{\partial u_j}{\partial x_i} \right)\right),
 \end{align}
+where $u_i$ is the filtered velocity, $\epsilon_{ijk}$ the Levi-Civita symbol, $f_i$ the Coriolis acceleration, $u_{g,i}$ the geostrophic wind, $\rho_0$ the density of dry air, and $\pi^*$ the modified perturbation pressure.
 In the limit of either well-resolved simulations or low $\RE$ (or both), $K_m \rightarrow \nu$, so the last term on the right hand side of eq.
@@
 
-In contrast to the interior closure of the LES where the direct effect of $\nu$ is of small relevance compared to the eddy viscosity---at least for high $\RE$---, this is not true for the viscid effects at the bottom boundary:  A constant flux layer is assumed in between the first grid point and the lowermost half point (on the Arakawa-C staggered grid) to estimate friction velocity and surface stresses by MOST:
+In contrast to the interior closure of the LES, where the direct effect of $\nu$ is of small relevance compared to the eddy viscosity---at least for high $\RE$--- this is not true for the viscid effects at the bottom boundary:  The assumption of a constant flux layer between the first grid point and the bottom (on the Arakawa-C staggered grid) leads to the following estimations of friction velocity and surface stresses:
 \begin{align}\label{most}
@@
 	\centering
-	\includegraphics[width=0.7\textwidth]{figures_2024/d3y_3Re_hodograph}
+	\includegraphics[width=0.7\textwidth]{figures/d3y_3Re_hodograph}
   \caption{Geostrophy aligned hodographs of the LES in comparison to the theoretical and laminar hodographs.
@@
 The hodograph (Fig.~\ref{3Re_hodograph}) illustrates that this is not caused by poor quality of the simulation since all of the simulations of Re1 closely follow the theoretical hodograph; in fact, the turning angle at the respective height is actually slightly overestimated.
-Rather, the offset is due to the veering of the wind vector continues in the lower part of the boundary layer that is not present in the LES.
+Rather, the offset is due to the veering of the wind vector continuing in the lower part of the boundary layer which is not present in the LES.
 Hence, for the low-$\RE$ case, the general observation of a higher $\alpha_\star$ with finer resolution is also caused by the approach of the final $\alpha_\star$ with decreasing height of the first grid point.
@@
   \centerline{
-	  \includegraphics[width=\textwidth]{figures_2024/d3y_3Re_profiles_outer.png}
+	  \includegraphics[width=\textwidth]{figures/d3y_3Re_profiles_outer.png}
 	}
@@
   \centerline{
-	  \includegraphics[width=\textwidth]{figures_2024/d3y_3Re_kappa}
+	  \includegraphics[width=\textwidth]{figures/d3y_3Re_kappa}
   }
@@
 Above the viscous sublayer, the theoretical profile shows a near-constant value for $\kappa$ up to $z^-\approx0.1$ for the case Re1 and up to $z^-\approx0.12$ for the cases Re2 and Re3.
-We do not compute $\kappa$ for ReX\_50 since there are only six grid points below $z^-\approx 0.12$, and o constant K\'arm\'an measure $\kappa(z)$ can be observed.
+We do not compute $\kappa$ for ReX\_50 since there are only six grid points below $z^-\approx 0.12$, and no constant K\'arm\'an measure $\kappa(z)$ can be observed.
 The resulting best estimates for $\kappa_{LES}$ are shown in Tab.\,\ref{simulation_results}, where the coefficient of determination ('r-squared') is above 0.99 for all cases.
@@
 For finer resolution, $\kappa_{LES}$ decreases with increasing Re for the range of scale separation investigated here.
-ReX\_100 yields a value around $0.53$ (which is outside the range agreed on by atmospheric observation and laboratory measurements), the finer resolution yields values around $0.46$ for Re1\_200 and around $0.44$ for Re2\_200 and Re3\_200, which is in excellenet agreement with DNS data and laboratory measurements.
+ReX\_100 yields a value around $0.53$ (which is outside the range agreed on by atmospheric observation and laboratory measurements), the finer resolution yields values around $0.46$ for Re1\_200 and around $0.44$ for Re2\_200 and Re3\_200, which is in excellent agreement with DNS data and laboratory measurements.
 The dynamic subgrid closure yields lower values for $\kappa_{LES}$: in Re1\_dyn we even see $\kappa_{LES}=0.39$, while Re2\_dyn and Re3\_dyn yield $0.42$ and $0.43$, respectively.
@@
 \begin{figure}[ht]
-  \centerline{\includegraphics[width=\textwidth]{figures_2024/d3y_2pc_ww_3Re_0.079.png}}
+  \centerline{\includegraphics[width=\textwidth]{figures/d3y_2pc_ww_3Re_0.079.png}}
   \caption{Two-point correlation of w-component in y-direction at $z^-\approx 0.08$. The horizontal line at $B_{ww}=0.3$ is indicative of the average size of turbulent structures in the flow.}
@@
 
-The resolution near the surface is critical for an LES with isotropic grid spacing. \todo{Missing transition from resolution to two-point correlation!} The two-point correlation is defined as 
+In order to estimate the quality of the flow resolution, the horizontally averaged two-point correlation can be utilized.
+For an LES with an isotropic grid and periodic horizontal boundaries, the span-wise two-point correlation of the vertical velocity component, $B_{ww}(y^*)$, indicates how well the flow is resolved \citep{wurps2020grid}.
+The two-point correlation is defined as
 \begin{align}
@@
 \end{align}
-and it supports the above findings \todo{which findings?}. 
-For an LES with isotropic grid and periodic horizontal boundaries, the the span-wise two-point correlation of the vertical velocity component $B_{ww}(y^*)$ indicates how well the flow is resolved \citep{wurps2020grid}.
-We define $\sigma$, the average size of a structure in the flow, as the separation at which the two-point correlation at a height $z^-\approx 0.08$, right inside the logarithmic layer, drops to 0.3, i.e.
-$B_{ww}(\sigma^\star) = 0.3$ (Fig.\,\ref{fig:2pc_008}).
+(other components and directions accordingly).
+We define $\sigma$, the average size of a structure in the flow, as the separation at which the two-point correlation drops to 0.3, i.e.
+$B_{ww}(\sigma^\star) = 0.3$.
+Fig.\,\ref{fig:2pc_008} shows the two-point correlations at a height $z^-\approx 0.08$, right inside the logarithmic layer.
 The number of cells by which the corresponding structure is resolved is then $\sigma/\Delta$.
@@
 This corresponds to the resolution from which on a logarithmic layer with a constant $\kappa$ can be seen.
-It also fits to the rule of thumb given in \cite{wurps2020grid}, that in a sufficiently resolved part of an LES the average structures of the w-component in y-direction should be resolved by at least 4 cells.
+It also fits to the rule of thumb given in \cite{wurps2020grid}, that in a sufficiently resolved part of an LES, the average structures of the w-component in y-direction should be resolved by at least 4 cells.
 The dynamic closure shows structures that are slightly smaller than the structures of $ReX\_200$, which is caused by a tendency to lower eddy-viscosities and, hence, a weaker coupling of neighboring grid cells.
@@
 \begin{figure}[ht]
-  \centerline{\includegraphics[width=\textwidth]{figures_2024/d3y_2pc_ww_3Re_4.png}}
+  \centerline{\includegraphics[width=\textwidth]{figures/d3y_2pc_ww_3Re_4.png}}
   \caption{Two-point correlation of w-component in y-direction at the 8th grid point above the surface}
@@
 
-In vicinity of the surface, the two-point statistics crucially depend on how the surface-interaction is modelled \citep{maronga2014monin}.
-which may cause strong sensitivity on the resolution.
+In vicinity of the surface, the two-point statistics crucially depend on how the surface-interaction is modelled \citep{maronga2014monin}, which may cause strong sensitivity on the resolution.
 Indeed, the two-point correlations $B_{ww}(y^*)$ evaluated at the 8th grid point above the bottom (Fig.~\ref{fig:2pc_4}) appear rather universal.
-This is in spite of different heights the simulations ($z^-=0.16, 0.08, 0.053, 0.04, 0.04$ for $ReX\_50$, $ReX\_100$, $ReX\_150$, $ReX\_200$, $ReX\_dyn$, respectively).
+This is in spite of different physical heights corresponding to this grid point in the simulations ($z^-=0.16, 0.08, 0.053, 0.04, 0.04$ for $ReX\_50$, $ReX\_100$, $ReX\_150$, $ReX\_200$, $ReX\_dyn$, respectively).
 For $Re1$, the curves almost collapse perfectly while for $Re2$ and $Re3$ we see slightly larger structures for $ReX\_50$ and slightly smaller structures for $ReX\_dyn$.
@@
 \begin{figure}[ht]
-  \centerline{\includegraphics[width=\textwidth]{figures_2024/d3y_3Re_v_outer_lin.png}}
+  \centerline{\includegraphics[width=\textwidth]{figures/d3y_3Re_v_outer_lin.png}}
   \caption{Span-wise velocity component in outer scaling}
@@
 
-he span-wise velocity component in the lower part of the boundary layer is shown in Fog.~\ref{fig:3Re_v_lin}.
-\todo{if we keep $V(z=\Delta)$ in the conclusion, add this here}.
-In the lowest part of $Re1$ ($z^-\lessapprox 0.05$), a part of the viscous layer is visible, where the velocity is slightly underestimated by the the LES.
-Above, the simulations Re1\_200 and Re1\_dyn are very close to the theoretical curve.
-Lower resolution leads to an overestimation of the velocity in the logarithmic layer, which is consistent with a lack of large eddies and thus too little mixing, reduced turbulence intensity and increased velocity gradients.
-Around $z^-=0.25$, while blending into the Ekman layer, the curves cross each other and aloft the coarse simulations underestimate the velocity.
-For the higher $Re$ all simulations overestimate the velocity in the logarithmic layer while finer resolution loeads to lower velocity and better agreement with the theoretical curve.
-Again, we see a crossing of the curves, between $z^-=0.2$ and $z^-=0.25$.
-A possible explanation of the fast increase and the steeper course of the coarser simulations is that the layers are coupled via less cells.
+The span-wise velocity component in the lower part of the boundary layer is shown in Fig.~\ref{fig:3Re_v_lin}.
+The orientation of the grid is chosen so that the stress at the bottom points into the x-direction; therefore, the v-component should be very small at the first grid point ($V(z=\Delta/2)\ll u_\star$).
+Particularly for the higher $\RE$, $-V^+$ is in the vicinity of $u_\star$ in the coarse simulations but approaches zero with higher resolution, indicating insufficient turning of the wind vector and a lower-than-expected $\alpha$.
+
+In the lowest part of $Re1$ ($z^-\lesssim 0.05$), a portion of the viscous layer is visible, where the velocity is slightly underestimated by the LES.
+Above, the simulations Re1\_200 and Re1\_dyn align closely with the theoretical curve.
+Lower resolution results in an overestimation of velocity in the logarithmic layer, which is consistent with a lack of large eddies and thus too little mixing, reduced turbulence intensity and increased velocity gradients.
+Around $z^-=0.25$, while blending into the Ekman layer, the curves cross each other, and aloft this point, the coarse simulations underestimate the velocity.
+
+For the higher $Re$, all simulations overestimate the velocity in the logarithmic layer, while finer resolutions lead to lower velocity and better agreement with the theoretical curve.
+Again, we observe a crossing of the curves between $z^-=0.2$ and $z^-=0.25$.
+The rapid increase and steeper profile in the coarser simulations may be due to the layers being coupled via fewer cells, resulting in less effective vertical momentum redistribution.
 
@@
 
-Above the logarithmic layer, the Ekman layer follows, characterized by a change of wind direction.
-The course of the wind velocity vector is visualized by hodographs, as plotted in Fig.
+Above the logarithmic layer the Ekman layer follows, characterized by a change of wind direction.
+While some studies, such as \cite{jiang2018large}, speculate about the existence of a second logarithmic layer, we did not find evidence of it, neither in the theoretical formulation nor in our simulations.
+The trajectory of the wind velocity vector is depicted by hodographs, as shown in Fig.
 \ref{3Re_hodograph}.
@@
 
-\todo{translational statement to hodograph missing} In Fig.
-\ref{3Re_hodograph}, the cross indicates the lowest grid point and the plus indicates the height where the velocity for the boundary condition is taken from.
-To avoid taking a velocity from the first layer of the simulation, where the turbulent flow is poorly resolved, we took the horizontal velocity near $z^-=0.1$, where the mean velocity already veered away from the direction of the surface stress by around one third of $\alpha$.
-However, the veering does not seem to influence the resulting bulk stress $u_\star$ at the bottom: all but the coarsest resolutions yield a $u_\star$ very close to the theoretic reference.
+In Fig.
+\ref{3Re_hodograph}, the cross indicates the lowest grid point, and the plus sign indicates the height from which the velocity for the boundary condition is taken from.
+To avoid using a velocity from the first layer of the simulation, where the turbulent flow is poorly resolved, we selected the horizontal velocity near $z^-=0.1$.
+At this height, the mean velocity has already veered away from the direction of the surface stress by around one-third of $\alpha$.
+However, the veering does not appear to affect the resulting bulk stress $u_\star$ at the bottom: all but the coarsest resolutions yield a $u_\star$ very close to the theoretical reference.
 
@@
   \centerline{
-	\includegraphics[width=\textwidth]{figures_2024/d3y_3Re_Ekman_lin.png}
+	\includegraphics[width=\textwidth]{figures/d3y_3Re_Ekman_lin.png}
 }
@@
 
-The shear-aligned velocity deficit $U^--U^-_G$ (v-component accordingly) agrres well with the expected DNS-based profiles.
-The largest deviations occur for the coarsest LES near the super-geostrophic maximum of the U-component.
-Finer resolution leads to excellent agreement, also holds for the higher $\RE$.
+The shear-aligned velocity deficit $U^--U^-_G$ (v-component accordingly) agrees well with the expected DNS-based profiles (see Fig.
+\ref{fig:3Re_Ekman}).
+The largest deviations occur in the coarsest LES near the super-geostrophic maximum of the U-component.
+Finer resolution leads to excellent agreement, which also holds for the higher $\RE$.
 The super-geostrophic maximum is a consequence of vertical redistribution of momentum across the ABL; less redistribution leads to a less pronounced maximum.
 At the same time, of course, one would expect an increased level of turbulence to cause a flatter hodograph.
-But, the surface veering is set rather by the interplay of surface-layer turbulence and the prescribed roughness while the profile in the Ekman layer is resolved.
-However, for higher Reynolds numbers $Re2$ and $Re3$, the maximum velocity in x-direction is still not reached by the LES and the approach of the geostrophic wind in the upper part of the boundary layer takes place at lower heights than in theory.
-This is even more obvious for the V-component, where the maximum value is at a significantly lower height than in the theoretical curve.
-In general, a finer resolution leads to a lower location of the maximum.
-The maximum at lower heights might also explain the too high velocities in the logarithmic layer for the higher $Re$.
+However, surface veering is primarily influenced by the interplay of surface-layer turbulence and the prescribed roughness while the profile in the Ekman layer is resolved.
+
+For higher Reynolds numbers ($Re2$ and $Re3$), the LES does not reach the maximum velocity in x-direction, and the approach of the geostrophic wind in the upper boundary layer occurs at lower heights than predicted by theory.
+This discrepancy is even more pronounced for the V-component, where the maximum value is at a significantly lower height than in the theoretical curve.
+In general, finer resolution leads to a lower location of the maximum.
+This lower maximum might also explain the too high velocities observed in the logarithmic layer for the higher $Re$.
 
@@
 
-We studied truly neutral Ekman boundary layers---absent of capping inversions, non-stationarity, and surface heterogeneity.
+The truly neutral Ekman boundary layer---absent of capping inversions, non-statio\-narity, and surface heterogeneity---is studied as a prototype representation of the ABL.
 For the first time, an LES of the Ekman layer is \emph{a posteriori} validated against a DNS for an identical configuration.
-Despite the common assumption that viscosity as a parameter drops out in LES as a consequence of the turbulence closure, we considered a range of scale separations (Reynolds numbers) and found that for an exact representation of the surface turning and wall friction as well as the wind-turning profile, viscous effects need consideration.
+Despite the common assumption that viscosity as a parameter drops out in LES as a consequence of the turbulence closure, we considered a range of scale separations ($\RE$) and found that for an exact representation of the surface turning and wall friction as well as the wind-turning profile, viscous effects need consideration.
 We derive an analytical relation demonstrating that --- given fixed surface properties --- the specification of $z_0$ implicitly defines a viscosity.
@@
 
-With the wind-profile formulation developed in Part 1 of this work, we have a reference based on first principles for intermediate-Re simulations, and a quasi-reference for the higher-Re simulations.
-While we acknowledge that both the LES and the wind-profile model suffer from assumptions for high Reynolds number, their quantitative agreement across a wide range of scale separations hints towards the consistency.
-In particular, we understand that the grid convergence of LES towards the theoretical profiles underpins the inviscid scaling hypotheses in the development of the theory underlying part 1 of this paper.
+With the wind-profile formulation developed in Part 1 of this work, we establish a reference based on first principles for intermediate $\RE$, and a quasi-reference for geophysical $\RE$.
+While we acknowledge that both the LES and the wind-profile model suffer from assumptions for high $\RE$, their quantitative agreement across a wide range of scale separations hints towards consistency.
+The grid convergence of LES towards the theoretical profiles underpins the inviscid scaling hypotheses in the development of the theory underlying part 1 of this paper.
+This approach raises our confidence in the LES to a new level for the direct comparison with a resolving simulation.
+Furthermore, it allows us to extend the insight from DNS to atmospheric scale where an approach through DNS is clearly infeasible for computational constraints.
 
-We simulated three different Reynolds numbers while having DNS results for the lowest $\RE$.
-For the low $\RE$, viscous forces have a significant contribution to the balance of forces on the grid scale of the LES, hence we adapted the LES code to consider the fluid's viscosity in addition to the modeled eddy-viscosity.
+We simulated three different Reynolds numbers while DNS data is available for the lowest $\RE$ only.
+For this lowest $\RE$, viscous forces significantly contribute to the balance of forces on the LES grid, and we adapted the LES code to consider the molecular fluid viscosity in addition to the modeled eddy-viscosity.
 We did not introduce a roughness length $z_0$ at the lower boundary as additional parameter but deduced $z_0$ according to the law of the wall and the shear velocity expected according to the semi-empirical law by \cite{spalart1989theoretical}.
@@
 \begin{itemize}
-\item The roughness parameter ($z_0$) essentially defines a Reynolds number of the LES Problem
-\item $V|_{z=\Delta} < u_\star$ for an accurate matching of the hodograph and proper quantification of $\alpha_\star$. \todo{where is this in the results?}
+\item The roughness parameter ($z_0$) essentially defines a Reynolds number of the LES Problem.
+\item $V|_{z=\Delta/2} \ll u_\star$ for an accurate matching of the hodograph and proper quantification of $\alpha_\star$.
 \item There should exist a logarithmic layer, or explicit consideration of viscous interaction with the surface; otherwise the assumptions of the surface closure (MOST / dynamic wall model) will fail.
@@
 
-When these considerations are taken into account, LES becomes possible at uncommonly low Reynolds number and resolution, which allows a quantitative comparison to state-of-the art DNS.
+When these considerations are taken into account, LES becomes possible at uncommonly low Reynolds number and  high resolution, which allows a quantitative comparison to state-of-the art DNS.
+Most importantly, we observe a clear $\RE$-dependency in the LES results, matching the predictions of the theoretical framework, illustrating that, also in  LES, the choice of $z_0$ implicitly defines a $\RE$.
 
@@
 
-We did not see a second logarithmic layer as \cite{jiang2018large} neither in the theoretical formulation nor in our simulations.
-A reason might be their introduction of the additional parameter $z_0$.
-\todo{keep this paragraph?
-After our discussion about the change of Re when introducing a roughness length I don't believe this explanation anymore...}
-
 The theoretical formulation of mean velocity profiles within the turbulent Ekman layer can serve as benchmark for model frameworks such as LES.
@@
 The simulations were performed at the HPC Cluster EDDY, located at the University of Oldenburg (Germany) and funded by the Federal Ministry for Economic Affairs and Energy (Bundesministeriums für Wirtschaft und Energie) under grant number 0324005.
+The DNS were generated on the HPC systems \texttt{juwels} at J\"ulich Supercomputing Centre under the (grants \texttt{stadit}, \texttt{hku24)} and \texttt{hawk} at H\"ochstleistungs\-rechen\-zentrum Stuttgart (Bundesprojekt \texttt{trainABL}, \#44187).
+CA acknowledges funding by the European Commission through ERC Starting Grant \#851374.
+We are thankful to the refubium team of FU Berlin for making available the DNS data used in this publication through their repository.
 
-\bibliographystyle{plainnat}
+\bibliographystyle{spbasic_updated}
 \bibliography{papers.bib,ansorge.bib}

### 2024-08-05 write1/manuscript.tex
diff for part2; edits in part1

@@
-\RequirePackage{fix-cm} 
 \documentclass[smallcondensed,final]{svjour3}
@@
 \usepackage{longtable}
-\usepackage{natbib}
+\usepackage[authoryear]{natbib}
 
@@
 \usepackage{hyperref}
-\usepackage{caption}
 \setcitestyle{aysep={}}
@@
 
-\date{\scriptsize Draft - Version \today}
-\title{Profiles of Wind Veer and Wind Speed in Turbulent Ekman Flow}
-\author{Cedrick Ansorge, Hauke Wurps}
+\date{\scriptsize Submitted \today}
+\title{Wind Veer and Speed in Turbulent Ekman Flow - Part 1: Scaling analysis and universal profile model}
+\titlerunning{}
+\author{Cedrick Ansorge$^1$ and Hauke Wurps$^2$}
+\authorrunning{C Ansorge and H Wurps}
+\institute{$^{1}$\,FU Berlin, Institut f\"ur Meteorologie, Carl-Heinrich-Becker-Weg 6--10, 12165 Berlin {cedrick@posteo.de}\\
+  $^{2}$\,Uni Oldenburg 
+  }
 
 \begin{document} 
-
+%
 \maketitle
-
+%
 \begin{abstract}
@@
 Knowledge about $u_\star$ and $\alpha$ is thus a prerequisite for any quantitative theory of the surface layer, 
-and \citet{rossby:PPO1935} constrained the two parameters based on integral relations in the ABL.
+and \citet{rossby:PIP1935} constrained the two parameters based on integral relations in the ABL.
 %
@@
 The turbulence scale separation in a wall-bounded flow is commonly characterized by the friction Reynolds number
-\citep{jimenez:ARF2014}:
+\citep{jimenez:ARF2012}:
 \begin{equation}
@@
 \end{subequations}
-where $u_i=U_i/G$ are the non-dimensional components velocity, $\pi$ is non-dimensional pressure, $g_j=G_j/G$ are non-dimensionalized components geostrophic wind (with $g_{1}^2+g_{2}^2=1$ by construction), and $\epsilon$ is the Levi--Civita tensor.
+where $u_i=U_i/G$ are the non-dimensional components velocity, $\pi$ is non-dimen\-sio\-nal pressure, $g_j=G_j/G$ are non-dimensionalized components geostrophic wind (with $g_{1}^2+g_{2}^2=1$ by construction), and $\epsilon$ is the Levi--Civita tensor.
 %
@@
 %
-No collapse in inner units is found for the profiles of spanwise velocity $V^{\alpha_\star+}(z^+)$. 
+No collapse is found for the profiles of spanwise velocity when considered in inner units, $V^{\alpha_\star+}(z^+)$. 
 %
@@
   \centerline{\includegraphics[width=0.5\textwidth]{../plot/c_log.pdf}}
-  \caption{Coefficients $a_\text{log}$, $b_\text{log}$ and $c_\log$ (cf. Eq.~\ref{eqn:abc_log}) as a function of 
+  \caption{Coefficients $a_\text{log}$, $b_\text{log}$ and $c_\text{log}$ 
+    (cf. Eq.~\ref{eqn:abc_log}) 
+    as a function of 
     the viscous Reynolds number $Re_D$. The approximate range of scale separation relevant for atmospheric application 
-    is found in between the dotted lines, where $a_\text{log}\simeq 0$ and $c_\text{log})\simeq 0.4$ }
+    is found in between the dotted lines, where $a_\text{log}\simeq 0$ and $c_\text{log}\simeq 0.4$ }
   \label{fig:clog}

### 2024-08-30 write1/manuscript.tex
continued work on comparison section in part 1

@@
 %
-Comparing our proposed formulation for the stream-wise velocity in the inner layer to  \citeauthor{vandriest:JOT1956}'s formulation 
+Comparing our proposed formulation (Fig.~\ref{fig:vanDriest}) for the stream-wise velocity in the inner layer to  \citeauthor{vandriest:JOT1956}'s formulation 
 yields later convergence of the velocity onto the logarithmic profile while, over all, it serves as an excellent model of the streamwise 
@@
 
+For the higher layers of the ABL, the Ekman spiral is the simplest model available.
+When employed across the entirety of the ABL, the spiral is flattened, corresponding to a reduction of the veering angle both at the surface and throughout the ABL.
+We, however, find that a modified version of the Ekman spiral explicitly taking into account the surface boundary condition, is a consistent model and yields excellent
+agreement with the velocity profiles from DNS (Sec.~\ref{sec:ekman}).
+
+Given a surface veering, a formulation for the velocity profile in the logartihmic and Ekman layer can be obtained following the arguments by \cite{etling:2008}, cf.
+\cite{emeis:2018}.
+This simple model is compared to our DNS data and theoretical profiles in Fig.~\ref{fig:etling}.
+
+\begin{figure}
+  \includegraphics[width=0.5\textwidth]{../plot/etling_direction_1.pdf} 
+  \includegraphics[width=0.5\textwidth]{../plot/etling_magnitude_1.pdf}
+  \caption{}
+  \label{fig:etling} 
+\end{figure}
+outlined in 
 \textbf{Compare} to Emeis (2002) and Gryning (2007); highlight explicit knowledge on veering-profile $\rightarrow$ directional sheer;

### 2024-10-01 write1/manuscript.tex
update

@@
 \subsection{Comparison with other theories}
-
+%%% VAN DRIEST SCALING
 An alternative approach that considers viscous effects close to the surface is the van-Driest scaling \citep{vandriest:JOT1956}, 
@@
 
+%EKMAN SPIRAL 
 For the higher layers of the ABL, the Ekman spiral is the simplest model available.
-When employed across the entirety of the ABL, the spiral is flattened, corresponding to a reduction of the veering angle both at the surface and throughout the ABL.
-We, however, find that a modified version of the Ekman spiral explicitly taking into account the surface boundary condition, is a consistent model and yields excellent
-agreement with the velocity profiles from DNS (Sec.~\ref{sec:ekman}).
-
+When employed across the entirety of the ABL, the spiral of a turbulent Ekman layer is flattened with respect to Ekman's lamniar solution, which  corresponds to a reduction of the veering angle both at the surface and throughout the ABL.
+We, however, find that a modified version of the Ekman spiral explicitly taking into account the surface boundary condition, is a consistent model and yields excellent agreement with the velocity profiles from DNS (Sec.~\ref{sec:ekman}).
+
+%ETLING MODEL 
 Given a surface veering, a formulation for the velocity profile in the logartihmic and Ekman layer can be obtained following the arguments by \cite{etling:2008}, cf.

### 2024-10-01 plot/plot.py

@@ -23,14 +23,14 @@ plot_evisc=False
 plot_profiles=False
 plot_visc_outer=False
 print_table=False
-plot_outer_log=False
+plot_outer_log=True
 plot_convergence=False
 plot_re1600=False
 plot_les_comp=False
 plot_applications=False
 test_inner_streamwise=False
 plot_profile_comparison=False
-plot_vandriest=True
+plot_vandriest=False
 plot_total_rot=False
 
 colors = {400 : 'gray',
@@ -1800,11 +1800,11 @@ if plot_profile_comparison:
         sc_dct[0] = 0. 
         ax.plot(ym[1:],et_dct/et_dct[0] ,ls='-',label=r'$Re_D={},\alpha_0={}^\circ$'.format(re,int(al*10)/10.),c=ReD_col[re]) 
         ax.plot(ym,(-sc_dct+sc_dct[-1])/sc_dct[-1] ,ls='--',c=ReD_col[re])
-        axm.plot(ym,us*np.sqrt(up*up+vp*vp),ls='--',c=ReD_col[re])
+        axm.plot(yp,np.sqrt(up*up+vp*vp),ls='--',c=ReD_col[re])
         eek_u[:i_gamma] = 1/sc.KAPPA *np.log(yp[:i_gamma]*np.exp(sc.KAPPA*sc.C))*np.cos(al*D2R)
         eek_v[:i_gamma] = 1/sc.KAPPA *np.log(yp[:i_gamma]*np.exp(sc.KAPPA*sc.C))*np.sin(al*D2R)
         m=np.sqrt(eek_u**2+eek_v**2)
-        axm.plot(ym[1:],m/m[-1],ls='-',c=ReD_col[re],label=r'$Re_D={}; u_\star={}G$'.format(re,int(us*1000)/1000.))
+        axm.plot(yp[1:],m,ls='-',c=ReD_col[re],label=r'$Re_D={}; u_\star={}G$'.format(re,int(us*1000)/1000.))
     lg=plt.legend()
     lg.get_frame().set_linewidth(0.0)
     ax.set_xlabel(r'$z^-$')
@@ -1855,12 +1855,12 @@ if plot_profile_comparison:
         sc_dct[0] = 0. 
[…]

### 2024-10-10 write1/manuscript.tex
finished paragraph on Etling model

@@
 %ETLING MODEL 
-Given a surface veering, a formulation for the velocity profile in the logartihmic and Ekman layer can be obtained following the arguments by \cite{etling:2008}, cf.
+A two-layer model consolidating both the logartihmic and Ekman layer can be obtained following the arguments by \cite{etling:2008}, cf.
 \cite{emeis:2018}.
-This simple model is compared to our DNS data and theoretical profiles in Fig.~\ref{fig:etling}.
+Given a surface veering and a matching height (extent of the logarithmic layer), a formulation for the velocity profile across both the logarithmic and the outer layer is obtained.
+A comparison using the surface veering based on our model and a matching height of $z_\text{prandtl}=0.05\delta$, which gives better results than the matching height of $0.1\delta$ suggested by \cite{etling:2008}, is shown in Fig.~\ref{fig:etling}.
+% 
+The overall shape of velocity magnitude is matched apart from the viscous and buffer layer (cf.
+Fig.~\ref{fig:etling}b) that is neglected by the two-layer model.
+However, quantitative departures on the order of 10-20\% occur across the inner layer: it turns out that the non-rotating profile, of the two-layer model in the logarithmic region yields too low overall velocity as the spanwise component contributes to the velocity across the inner layer.
+% 
+Deviations also occur with respect to wind direction; despite the rather low matching height, a substantial fraction of the rotation occurs within the lower part of the ABL and about 20\% of the wind veer is not captured by the two-layer model.
+% 
+As the overall veer is given, the non-captured veer close to the surface is then compensated across the logarithmic layer. In the upper part of the boundary layer, both profiles match well. 
+%
 
@@
   \includegraphics[width=0.5\textwidth]{../plot/etling_magnitude_1.pdf}
-  \caption{}
+  \caption{Comparison of the DNS data (dashed lines) with the two-layer model proposed by Etling (solid lines) for different Reynolds number.
+Left panel shows the relative wind veer, right panel shows the velocity magnitude.
+The etling model is calibrated by the surface veer from the DNS and roughness parameter for the Etling model is chosen according to the correction factor $\exp{\kappa A}$ such that the total veer and velocity magnitude agree.
+DNS profile in the buffer layer, defined here as $z^+<30$ are shown as opaque dashed lines as the two-layer model does not consider viscous effects.}
   \label{fig:etling} 
@@
 outlined in 
-\textbf{Compare} to Emeis (2002) and Gryning (2007); highlight explicit knowledge on veering-profile $\rightarrow$ directional sheer;
+\textbf{Compare} Gryning (2007); highlight explicit knowledge on veering-profile $\rightarrow$ directional sheer;
 

### 2024-10-10 write2/manuscript_part2_2024.tex
finished paragraph on Etling model

@@
   % Case desription and setup
-  Theoretical profiles of the wind vector are compared to the theoretical reference at intermediate Re, which necessitates consideration of viscous effects % [not required in abstract] (usually neglected in LES)
+  Theoretical profiles of the wind vector are compared to the theoretical reference at intermediate Re, which necessitates consideration of viscous effects % [not required in abstract] (usually neglected in LES) a
   along with modifications to the standard bottom boundary condition.

### 2024-10-10 write2/manuscript_part2_20240801_hw.tex
finished paragraph on Etling model

--- 2024-08-05 write2/manuscript_part2_2024.tex
@@
-\documentclass[smallcondensed,draft]{svjour3}
+\documentclass[smallcondensed,final]{svjour3}
 
 \usepackage{lmodern}
-%\usepackage[margin=2.5cm]{geometry}
+\usepackage[margin=2.5cm]{geometry}
 \usepackage{booktabs} 
@@
 %\usepackage{caption}
-\usepackage[authoryear]{natbib}
+\usepackage{natbib}
 \usepackage{xcolor}
 \usepackage{tikz}
-
-\usepackage{adjustbox}
 
@@
 \newcommand{\RE}{\mathbf{Re}}
+\hyphenation{Ol-den-burg}
 
-\hyphenation{Ol-den-burg}
-\hyphenation{non-sta-tio-na-ri-ty}
 \DeclareRobustCommand\sampleline[1]{%
@@
 
-\author{Hauke Wurps $^{1}$ and Cedrick Ansorge $^{2}$}
-\authorrunning{H. Wurps and C. Ansorge}
+\author{Hauke Wurps and Cedrick Ansorge}
 
-\institute{$^{1}$ Uni Oldenburg \newline 
-$^{2}$ FU Berlin, Institut f\"ur Meteorologie, Carl-Heinrich-Becker-Weg 6--10, 12165 Berlin \email{cedrick@posteo.de}}
-
-\date{Submitted \today}
-\title{Wind veer in Turbulent Ekman Flow - Part 2: Extrapolation to atmospheric scale separation}
-\titlerunning{A Universal veering profile for Ekman flow - Part 2}
+\date{Draft \today}
+\title{Towards a Universal Veering Profile for Turbulent Ekman Flow at arbitrary Reynolds number - Part 2\\
+  \normalsize LES and DNS of Turbulent Ekman Flow \\}
 
@@
   % Introduction % Theoretical profile
-  The atmospheric boundary layer is characterized by a logarithmic increase of the wind speed in the surface layer and a turning of the wind vector aloft.
-The simplest canonical configuration featuring both \emph{a logarithmic layer and the wind turning} is Ekman flow.
-A formulation of the mean velocity profile was derived from scaling considerations and calibrated using direct numerical simulation in part I of this work.
-Here, we explore the extrapolation to atmospheric Reynolds number (Re) using large-eddy simulation (LES).
+  Turbulent Ekman flow is a canonical flow representation of the atmospheric boundary layer, characterized by a logarithmic increase of the wind speed near the surface and a turning of the wind vector aloft.
+In part I of this work, we derived a formulation of the mean wind velocity profiles of the stationary turbulent Ekman flow based on direct numerical simulation (DNS).
+Here, we explore the extrapolation of these profiles to atmospheric Reynolds numbers using large-eddy simulation (LES).
   % Case desription and setup
-  Theoretical profiles of the wind vector are compared to the theoretical reference at intermediate Re, which necessitates consideration of viscous effects % [not required in abstract] (usually neglected in LES)
-  along with modifications to the standard bottom boundary condition.
+  We compare theoretical profiles of the wind vector for low Reynolds-number turbulent Ekman flow to results from DNS and LES.
+This necessitates the inclusion of viscous effects, usually neglected in LES, along with modifications to the standard bottom boundary condition.
   % Results and Discussion
-  Analysis of the grid convergence for higher resolution unveils a convergence of LES data towards the theoretical profiles for intermediate and high Re.
-The LES thus confirm that (i) the spanwise velocity scales as $\left.Re_\tau\right.^{-1}$, where $Re_\tau$ is the friction Reynolds number, and (ii) one third of the wind veer is confined to the surface layer.
-  % [this level of detail is not needed in the abstract] (the lowest 10\,\% of the boundary layer) 
+  Analysis of the grid-convergence for higher resolution unveils a convergence of LES data towards the theoretical profiles for both intermediate and high Reynolds numbers.
+Our LES thus confirm that---within the logarithmic layer (the lowest 10\,\% of the boundary layer)--- the direction of the wind vector turns by approximately one-third of its total rotation across the entirety of the boundary layer.
   % Conclusions
-  Such agreement of the theoretical formulation and LES data raises confidence in the underlying scaling assumptions and thus reinforces the utility of the theoretical profiles as a reference for intermediate and a quasi-reference for both idealized simulation and field observation.
+  Such agreement of our theoretical formulation and LES data raises our confidence in the underlying assumptions of the theory and thus reinforces the utility of the theoretical profiles as a reference for intermediate and a quasi-reference for higher Reynolds number simulations.
 %
@@
 Relevant aspects of the interaction with the boundary not only include its heterogeneity \citep{avissar:MWR1989,giorgi:RG1997,claussen:BM1991,garratt:BM1990} and roughness \citep{monin:ARF1970,brutsaert:WRR1975,raupach:AMR1991,kostelecky:JFM2024}, but also a wind veer in favor of the pressure gradient force when the surface is approached \citep{ekman:AMA1905}.
-
-Large-Eddy Simulation (LES) is widely used to study the turbulent flow in the atmospheric boundary layer (ABL) for a wide range of applications \citep{stoll2020large}.
-This includes the simulation of wind turbines and wind farm wakes, their interactions within wind farm clusters \citep{porte2011large,mehta2014large,breton2017survey}.
+An important tool with capabilities to model the ABL flow is Large-Eddy Simulation (LES).
+Indeed, LES is widely used to model turbulent flow in the atmospheric boundary layer (ABL) \citep{stoll2020large}.
+This includes various applications, such as simulating wind turbine and wind farm wakes, and interactions within wind farm clusters \citep{porte2011large,mehta2014large,breton2017survey}.
 Furthermore, LES is applied in complex environments like mountainous or urban areas \citep{stoll2020large,garcia2018predictive} and to investigate the dispersion of pollution \citep{han2018large}.
@@
 
-The simplest representation of the ABL taking into account rotation and the quasi non-turbulent, free atmosphere aloft is Ekman flow.
+The simplest representation of the ABL taking into account rotation \emph{and} the quasi non-turbulent, free atmosphere aloft is the turbulent Ekman flow.
 It is a horizontally homogeneous, statistically stationary boundary-layer flow over a rotating flat surface.
-Here we consider the problem under neutral stratification, where density effects on the flow due to temperature fluctuations are absent.
-This turbulent flow configuration features the key characteristics of the real ABL, a logarithmic layer and the Ekman spiral; in immediate vicinity of the surface, viscous forces dominate (\emph{viscous sublayer}).
-The stationary ABL-height is set by the interplay between turbulent growth due to the shear instability \citep{lilly1966instability} and rotational suppression of turbulence due to the Coriolis effect.
-The system's statistical equilibrium is characterized by a single parameter, the Reynolds number $\RE$, signifying the scale separation between the largest and smallest scales of the problem.
-The largest scales are given by shear-induced eddies constrained by the boundary-layer height, while the smallest scales reside within the dissipative range, where strong viscous forces suppress turbulence.
+Here we consider the problem under neutral stratification, where the potential temperature is constant across the whole domain.
+Turbulent Ekman flow shows the key characteristics shaping the real ABL: a logarithmic layer and the Ekman spiral.
+Near the surface, viscous forces dominate the flow.
+The stationary ABL-height is then defined by the interplay between turbulent growth due to the shear instability of the configuration \citep{lilly1966instability} and rotational suppression of turbulence due to the Coriolis effect.
+The system's statistical equilibrium is defined by one single parameter, the Reynolds number $\RE$, signifying the scale separation between the largest and smallest scale within the problem.
+The largest scales are given by shear-induced eddies constrained by the boundary layer's height, while the smallest scales reside within the dissipative range, where strong viscous forces suppress turbulence.
 The dimensional mean solution of the turbulent Ekman layer is influenced by three parameters: the geostrophic wind $G$, the Coriolis parameter $f$, and the kinematic viscosity $\nu$.
 These three parameters define the Reynolds number $Re_D = GD/\nu$.
-Using $D=\sqrt{2\nu/f}$, the Ekman layer depth of a laminar flow, we write
-%
+With $D=\sqrt{2\nu/f}$ the Ekman layer depth of a laminar flow we write
+
 \begin{align}\label{red}
@@
 \end{align}
-%
+
 The mean solutions of the velocity components are functions of only the Reynolds number and the height above the ground $z$: $u(\RE_D,z)$, $v(\RE_D,z)$ (see e.g. \cite{csanady1967resistance}).
@@
 However, it remains uncertain, if the extrapolation of the above framework towards atmospheric $\RE$ is valid.
-The substantial separation of scales in the ABL makes direct numerical simulation (DNS) of the entire turbulent flow at geophysical scale impossible.
+The substantial separation of scales in the ABL renders direct numerical simulation (DNS) of the entire turbulent flow at scale impossible.
 Within the framework of LES, only the larger eddies are resolved, that contain the main share of kinetic energy, while turbulent mixing below the subfilter-scale is parameterized.
@@
 As LES introduces the grid size as an additional parameter (often understood in relation to a filter scale, cf.
-\cite{pope2004ten}), we analyse its impact on simulation results.
+\cite{pope2004ten}), we conduct a thorough analysis of its impact on simulation results.
 \cite{esau2004simulation} and \cite{jiang2018large} have executed similar LES analyses of neutral Ekman layers, however, they introduce an additional parameter by selecting an arbitrary roughness length $z_0$, whereas we derive the parameter $z_0$ corresponding to a fluid with a kinematic viscosity of $\nu = 1.5\cdot10^{-5}$ over a flat surface.
@@
 part 1 of this study), $Re_D=1.5 \cdot 10^5$ and $10^6$ correspond to the scale separation found in typical atmospheric boundary layers.
-For $Re_D=1600$, we compare DNS \citep{ansorge:2024b} and LES results using the exact same forcing and settings.
+For $Re_D=1600$, we compare DNS and LES results using the exact same forcing and settings.
 This is the first comparison of this kind for the turbulent Ekman flow.
+\todo{keep rest of the paragraph in the introdution or move to results/conclusions?} The LES results converge to the DNS results with decreasing grid cell size.
+This approach raises our confidence in the LES results to a new level for the direct comparison with a resolving simulation.
+Furthermore, it allows us to extend the insight from DNS to atmospheric scale where an approach through DNS is clearly infeasible for computational constraints.
+We observe a clear $\RE$ dependency in the LES results, matching the predictions of the theoretical framework.
 
@@
 %Definitions
-We align the coordinate Ox with the surface stress (\cite{ansorge:BM2019}), Oz points normal to the surface and normal to Ox, and Oy in the span-wise direction, normal to Oxz.
+We align the coordinate Ox with the surface stress, Oz points normal to the surface and normal to Ox, and Oy in the span-wise direction, normal to Oxz.
 In proximity to the lower boundary, the boundary layer scales in viscous (inner) units, denoted by the index ``$+$'' ($z^+ = zu_\star/\nu$, $U^+=U/u_\star$, where $U$ is the velocity component in x-direction and $u_\star = \sqrt{\nu \p_z U}$ the friction velocity).
@@
 
-In the viscous sublayer, the shear-aligned velocity $U^{\alpha_\star}$ is described by the law of the wall $U^{\alpha_\star+} = z^+$ and the span-wise velocity $V^{\alpha_\star}$ is zero by choice of the reference frame (the index $\alpha_\star$ indicates the alignment with the direction of the shear stress).
+In the viscous sublayer, the shear-aligned velocity $U^{\alpha_\star}$ is described by the law of the wall $U^{\alpha_\star+} = z^+$ and the span-wise velocity $V^{\alpha_\star}$ is zero by choice of the reference frame
+(the index $\alpha_\star$ indicates the alignment with the direction of the shear stress).
 Around $z^+=5$, the velocity begins to deviate from linearity, and the buffer layer forms the transition from the viscous to the  logarithmic layer.
@@
 $f$, $\nu$, and $G$ are input parameters and define the Reynolds numbers $Re_D$ and $Re_\tau$, while $u_\star$, $\alpha_\star$, and $\delta = u_\star/f$ are resulting properties of the flow}
-	\begin{adjustbox}{width=\textwidth}
 	\begin{tabular}{l|ccccc|ccc}
@@
 	\end{tabular}
-  \end{adjustbox}
 	\label{simulation_parameters}
@@
 
-The truly neutral Ekman boundary layer---absent of capping inversions, non-statio\-narity, and surface heterogeneity---is studied as a prototype representation of the ABL.
+We studied truly neutral Ekman boundary layers---absent of capping inversions, non-stationarity, and surface heterogeneity.
 For the first time, an LES of the Ekman layer is \emph{a posteriori} validated against a DNS for an identical configuration.
-Despite the common assumption that viscosity as a parameter drops out in LES as a consequence of the turbulence closure, we considered a range of scale separations ($\RE$) and found that for an exact representation of the surface turning and wall friction as well as the wind-turning profile, viscous effects need consideration.
+Despite the common assumption that viscosity as a parameter drops out in LES as a consequence of the turbulence closure, we considered a range of scale separations (Reynolds numbers) and found that for an exact representation of the surface turning and wall friction as well as the wind-turning profile, viscous effects need consideration.
 We derive an analytical relation demonstrating that --- given fixed surface properties --- the specification of $z_0$ implicitly defines a viscosity.
@@
 
-With the wind-profile formulation developed in Part 1 of this work, we establish a reference based on first principles for intermediate $\RE$, and a quasi-reference for geophysical $\RE$.
-While we acknowledge that both the LES and the wind-profile model suffer from assumptions for high $\RE$, their quantitative agreement across a wide range of scale separations hints towards consistency.
-The grid convergence of LES towards the theoretical profiles underpins the inviscid scaling hypotheses in the development of the theory underlying part 1 of this paper.
-This approach raises our confidence in the LES to a new level for the direct comparison with a resolving simulation.
-Furthermore, it allows us to extend the insight from DNS to atmospheric scale where an approach through DNS is clearly infeasible for computational constraints.
+With the wind-profile formulation developed in Part 1 of this work, we have a reference based on first principles for intermediate-Re simulations, and a quasi-reference for the higher-Re simulations.
+While we acknowledge that both the LES and the wind-profile model suffer from assumptions for high Reynolds number, their quantitative agreement across a wide range of scale separations hints towards the consistency.
+In particular, we understand that the grid convergence of LES towards the theoretical profiles underpins the inviscid scaling hypotheses in the development of the theory underlying part 1 of this paper.
 
-We simulated three different Reynolds numbers while DNS data is available for the lowest $\RE$ only.
-For this lowest $\RE$, viscous forces significantly contribute to the balance of forces on the LES grid, and we adapted the LES code to consider the molecular fluid viscosity in addition to the modeled eddy-viscosity.
+We simulated three different Reynolds numbers while having DNS results for the lowest $\RE$.
+For the low $\RE$, viscous forces have a significant contribution to the balance of forces on the grid scale of the LES, hence we adapted the LES code to consider the fluid's viscosity in addition to the modeled eddy-viscosity.
 We did not introduce a roughness length $z_0$ at the lower boundary as additional parameter but deduced $z_0$ according to the law of the wall and the shear velocity expected according to the semi-empirical law by \cite{spalart1989theoretical}.
@@
 
-When these considerations are taken into account, LES becomes possible at uncommonly low Reynolds number and  high resolution, which allows a quantitative comparison to state-of-the art DNS.
-Most importantly, we observe a clear $\RE$-dependency in the LES results, matching the predictions of the theoretical framework, illustrating that, also in  LES, the choice of $z_0$ implicitly defines a $\RE$.
+When these considerations are taken into account, LES becomes possible at uncommonly low Reynolds number and resolution, which allows a quantitative comparison to state-of-the art DNS.
 
@@
 The simulations were performed at the HPC Cluster EDDY, located at the University of Oldenburg (Germany) and funded by the Federal Ministry for Economic Affairs and Energy (Bundesministeriums für Wirtschaft und Energie) under grant number 0324005.
-The DNS were generated on the HPC systems \texttt{juwels} at J\"ulich Supercomputing Centre under the (grants \texttt{stadit}, \texttt{hku24)} and \texttt{hawk} at H\"ochstleistungs\-rechen\-zentrum Stuttgart (Bundesprojekt \texttt{trainABL}, \#44187).
-CA acknowledges funding by the European Commission through ERC Starting Grant \#851374.
-We are thankful to the refubium team of FU Berlin for making available the DNS data used in this publication through their repository.
 
-\bibliographystyle{spbasic_updated}
+\bibliographystyle{plainnat}
 \bibliography{papers.bib,ansorge.bib}

### 2024-10-10 plot/plot.py

@@ -23,13 +23,13 @@ plot_evisc=False
 plot_profiles=False
 plot_visc_outer=False
 print_table=False
-plot_outer_log=True
+plot_outer_log=False
 plot_convergence=False
 plot_re1600=False
 plot_les_comp=False
 plot_applications=False
 test_inner_streamwise=False
-plot_profile_comparison=False
+plot_profile_comparison=True
 plot_vandriest=False
 plot_total_rot=False
 
@@ -77,27 +77,49 @@ def etling_profile(yp,ret,et_al,gamma_p=0.1):
     kappa=sc.KAPPA
     off_A=sc.C 
 
-    et_al = et_al * D2R 
+    et_al = et_al *D2R 
     
     ny,yp,ym=build_grid(ret,thick=3)
     i_01 =mp.geti(yp,gamma_p*ret)
-    zp = gamma_p*ret
+    yprandtl = gamma_p*ret
     surface_factor=np.exp(kappa*off_A) 
-    z0 = zp / (surface_factor*gamma_p*ret)
-    et_us = etling_us(z0,zp,et_al) 
+    z0 = yprandtl / (surface_factor*gamma_p*ret)
+    et_us = etling_us(z0,yprandtl,et_al)
+    iprandtl = mp.geti(yp,yprandtl)
+    print(iprandtl)
+
+    etling_pr_up=np.zeros(ny) 
+    etling_pr_vp=np.zeros(ny)
+    etling_ek_up=np.zeros(ny)
+    etling_ek_vp=np.zeros(ny)
+
+    print('ETLING PROFILE FOR RE_tau={}: us={}, al={}'.format(ret,et_us,et_al*D2R))
 
     d_ek_m = np.sqrt(2*kappa*gamma_p)
     d_ek_p = d_ek_m * ret
 
     # profiles
-    z_tilde = (yp - gamma_p*reTau ) / d_ek_p
+    z_tilde = (yp - gamma_p*ret ) / d_ek_p 
     z_arg=z_tilde[1:]+np.pi/4-et_al
     damp = np.exp(-z_tilde[1:])
-    etling_pr_up=1./sc.KAPPA*np.log(yp[1:]*surface_factor)*np.cos(et_al) 
[…]

### 2024-10-10 write1/manuscript.tex
continued work on Section 5.2 (Profile comparison)

@@
 Fig.~\ref{fig:etling}b) that is neglected by the two-layer model.
-However, quantitative departures on the order of 10-20\% occur across the inner layer: it turns out that the non-rotating profile, of the two-layer model in the logarithmic region yields too low overall velocity as the spanwise component contributes to the velocity across the inner layer.
+However, quantitative departures on the order of 10\% occur across the inner layer: it turns out that the non-rotating profile, of the two-layer model in the logarithmic region yields too low overall velocity as the spanwise component contributes to the velocity across the inner layer.
 % 
@@
 % 
-
 \begin{figure}
+  \textbf{(a)\hspace{0.48\textwidth}(b)}\\
   \includegraphics[width=0.5\textwidth]{../plot/etling_direction_1.pdf} 
@@
   \label{fig:etling} 
+\end{figure}
+
+\begin{figure}
+  \textbf{(a)\hspace{0.48\textwidth}(b)}\\
+  \includegraphics[width=0.5\textwidth]{../plot/stress_rotation.pdf}
+  \includegraphics[width=0.5\textwidth]{../plot/misalignment.pdf} 
+  \caption{} 
+  \label{fig:alignment}
 \end{figure}

### 2024-10-10 plot/plot.py

@@ -29,9 +29,10 @@ plot_re1600=False
 plot_les_comp=False
 plot_applications=False
 test_inner_streamwise=False
-plot_profile_comparison=True
+plot_profile_comparison=False
 plot_vandriest=False
 plot_total_rot=False
+plot_shear_vs_rotation=True
 
 colors = {400 : 'gray',
           500 : 'pink',
@@ -1782,17 +1783,107 @@ if test_inner_streamwise:
     plt.savefig('innertest.pdf',format='pdf')
     plt.close('all')
 
+ReD_col={500: 'blue',
+         1000:'orange',
+         1600:'red',
+         10000:'gray',
+         40000:'black' }
 
-    
+if plot_shear_vs_rotation: 
+
+    ReD_arr = [ 500,1000,1600, 10000, 40000]
+    fig=plt.figure(figsize=(5,3))
+    fig2=plt.figure(figsize=(5,3))
+    ax=fig.add_axes([0.15,0.15,0.83,0.80]) 
+    ax2=fig2.add_axes([0.15,0.15,0.83,0.80]) 
+
+
+    for re in ReD_arr: 
+
+        us,al=sc.ustar_alpha(re)
+        al=al/np.pi*180
+        reTau = (re*us)**2/2. 
+        ny,yp,ym=build_grid(reTau,thick=3)
+
+        up,vp=sc.profile_plus(yp,re)
+        um,vm=up/us,vp/us
+        up_z=np.gradient(up,yp,edge_order=2)
+        vp_z=np.gradient(vp,yp,edge_order=2)
+
+        alpha = np.arctan2(vp,up)
+        alpha_shear = np.arctan2(-vp_z,-up_z)
+        if re == 500:
+            ax.plot(ym,alpha,label='Re={} (DD)'.format(re),c=ReD_col[re],ls='--',alpha=0.5,lw=3)
+            ax.plot(ym,alpha_shear-alpha,label='stress vs. DD',c=ReD_col[re],ls='-')
+        else:
+            ax.plot(ym,alpha,label='Re={}'.format(re),c=ReD_col[re],ls='--',alpha=0.5,lw=3)
+            ax.plot(ym,alpha_shear-alpha,c=ReD_col[re],ls='-')
+
+        if re in files.keys():
+            f_nc = Dataset(files[re],'r')
+            u_geo=f_nc.variables['rU'][-1,-1]
[…]

### 2024-10-10 write1/manuscript.tex
added figure Caption for comparison with K-Theorty

@@
 %%% VAN DRIEST SCALING
-An alternative approach that considers viscous effects close to the surface is the van-Driest scaling \citep{vandriest:JOT1956}, 
-where an exponential damping of Prandtl's mixing length is considered near the wall to yield 
-\begin{align}
-  \frac{\p u^+}{\p y^+} = \frac{2}{ 1+ \sqrt{1+\left(2\kappa y^+\right)^2 (1-\exp\left[-y^+\right])}};
-\end{align}
-the spanwise component is zero as no rotational effects are considered. 
-%
-Comparing our proposed formulation (Fig.~\ref{fig:vanDriest}) for the stream-wise velocity in the inner layer to  \citeauthor{vandriest:JOT1956}'s formulation 
-yields later convergence of the velocity onto the logarithmic profile while, over all, it serves as an excellent model of the streamwise 
-velocity component:
-% 
-Notable deviations (on the order of few percent) only occur in the region $10 < z^+ < 30$, 
-where the velocity transitions from the linear to the logarithmic profile. 
 \begin{figure}
@@
   \label{fig:vanDriest}
-\end{figure}
+\end{figure}  An alternative approach that considers viscous effects close to the surface is the van-Driest scaling \citep{vandriest:JOT1956}, 
+where an exponential damping of Prandtl's mixing length is considered near the wall to yield 
+\begin{align}
+  \frac{\p u^+}{\p y^+} = \frac{2}{ 1+ \sqrt{1+\left(2\kappa y^+\right)^2 (1-\exp\left[-y^+\right])}};
+\end{align}
+the spanwise component is zero as no rotational effects are considered. 
+%
+Comparing our proposed formulation (Fig.~\ref{fig:vanDriest}) for the stream-wise velocity in the inner layer to  \citeauthor{vandriest:JOT1956}'s formulation 
+yields later convergence of the velocity onto the logarithmic profile while, over all, it serves as an excellent model of the streamwise 
+velocity component:
+% 
+Notable deviations (on the order of few percent) only occur in the region $10 < z^+ < 30$, 
+where the velocity transitions from the linear to the logarithmic profile.
 
@@
 %ETLING MODEL 
-A two-layer model consolidating both the logartihmic and Ekman layer can be obtained following the arguments by \cite{etling:2008}, cf.
-\cite{emeis:2018}.
-Given a surface veering and a matching height (extent of the logarithmic layer), a formulation for the velocity profile across both the logarithmic and the outer layer is obtained.
-A comparison using the surface veering based on our model and a matching height of $z_\text{prandtl}=0.05\delta$, which gives better results than the matching height of $0.1\delta$ suggested by \cite{etling:2008}, is shown in Fig.~\ref{fig:etling}.
-% 
-The overall shape of velocity magnitude is matched apart from the viscous and buffer layer (cf.
-Fig.~\ref{fig:etling}b) that is neglected by the two-layer model.
-However, quantitative departures on the order of 10\% occur across the inner layer: it turns out that the non-rotating profile, of the two-layer model in the logarithmic region yields too low overall velocity as the spanwise component contributes to the velocity across the inner layer.
-% 
-Deviations also occur with respect to wind direction; despite the rather low matching height, a substantial fraction of the rotation occurs within the lower part of the ABL and about 20\% of the wind veer is not captured by the two-layer model.
-% 
-As the overall veer is given, the non-captured veer close to the surface is then compensated across the logarithmic layer. In the upper part of the boundary layer, both profiles match well. 
-% 
 \begin{figure}
@@
   \label{fig:etling} 
-\end{figure}
+\end{figure}A two-layer model consolidating both the logartihmic and Ekman layer can be obtained following the arguments by \cite{etling:2008}, cf.
+\cite{emeis:2018}.
+Given a surface veering and a matching height (extent of the logarithmic layer), a formulation for the velocity profile across both the logarithmic and the outer layer is obtained.
+A comparison using the surface veering based on our model and a matching height of $z_\text{prandtl}=0.05\delta$, which gives better results than the matching height of $0.1\delta$ suggested by \cite{etling:2008}, is shown in Fig.~\ref{fig:etling}.
+% 
+The overall shape of velocity magnitude is matched apart from the viscous and buffer layer (cf.
+Fig.~\ref{fig:etling}b) that is neglected by the two-layer model.
+However, quantitative departures on the order of 10\% occur across the inner layer: it turns out that the non-rotating profile, of the two-layer model in the logarithmic region yields too low overall velocity as the spanwise component contributes to the velocity across the inner layer.
+% 
+Deviations also occur with respect to wind direction; despite the rather low matching height, a substantial fraction of the rotation occurs within the lower part of the ABL and about 20\% of the wind veer is not captured by the two-layer model.
+% 
+As the overall veer is given, the non-captured veer close to the surface is then compensated across the logarithmic layer. In the upper part of the boundary layer, both profiles match well. 
+%
 
@@
   \includegraphics[width=0.5\textwidth]{../plot/misalignment.pdf} 
-  \caption{} 
+  \caption{Panel \textbf{(a)} shows the wind direction (DD) as thick dashed lines and the direction of stress relative to DD as thin solid lines, both according to our profile model.
+The direction of the negative flux vector $(\overline{u'w'},\overline{v'w'})$ based on DNS data is shown as a dotted line.
+In panel \textbf{(b)}, we plot the misalignment, i.e.
+the angle, between the flux vector and the shear vector, where the shear vector is defined accordingly as $(\p_z U, \p_z V)$.}
   \label{fig:alignment}

### 2024-10-14 write1/manuscript.tex
edits
started work on conclusions

@@
 \title{Wind Veer and Speed in Turbulent Ekman Flow - Part 1: Scaling analysis and universal profile model}
-\titlerunning{}
+\titlerunning{A Universal Velocity Profile for Ekman Flow}
 \author{Cedrick Ansorge$^1$ and Hauke Wurps$^2$}
@@
 % 
-
+\par
+% 
+The interpretation of flux and gradient profiles in terms of the K-theory (cf.
+Sec.~\ref{sec:scaling}, Fig.~\ref{fig:stresses}) suggests a certain universality in the inner layer, while a global collapse, i.e.
+across the boundary layer, is not
+obtained. 
+% 
+While the K-Profiles concern the total stress, a consistent formulation of the turning would also require its partitioning to the individual components, i.e. 
+the orientation of the stress vector $(\overline{u'w'}, \overline{v'w'})$ in the horizontal plane.
+If K-theory shall be used, this stress vector needs to be anti-parallel to the corresponding stress vector $(\partial_z U, \partial_z V)$.
+% 
+Fig.~\ref{fig:alignment} shows the direction of the velocity, gradient and stress 
+vectors across the boundary layer. It turns out that the negative stress vector with respect to the wind direction and the flux vectors (absolute) have an approximately similar direction. 
+%
+It appears that both rotate by about $270^\circ$ ($3\pi/2$) across the boundary layer. 
+%
+Apart from the lower part of the Ekman layer ($0.2<z^-<0.7$), where there is a slight dependence on $Re$, the direction of stress appears to be universal, which is a consequence of the Ekman profile introduced in Sec.~\ref{sec:ekman}.
+%
+However, this implies the a misalignment between the flux and the stress on the order of the wind turning, and indeed Fig.~\ref{fig:alignment}b shows a misalignment up to $90^\circ$.
+This is in accordance with the expectation by \cite[][\S 7.18]{Townsend:1976} and prevents the transfer of Energy from the mean flow to turbulence at these heights, thus preventing a boundary-layer growth.
+% 
+While the behavior in the inner layer seems to depend on $Re$, there emerges universality in the misalignment across the outer layer, suggesting that a universal consideration of misalignment in the context of K-theory is possible for formulations of higher-order quantities.
+%
 \begin{figure}
@@
 \end{figure}
-outlined in 
-\textbf{Compare} Gryning (2007); highlight explicit knowledge on veering-profile $\rightarrow$ directional sheer;
-
-\textbf{Remember} of interpretation in the context of eddy viscosity (Fig.~\ref{fig:stresses}) 
-Consider Townsend ``Turbulent Shear Flow''  Chapter 7.18: p. 319 
-%
+
+% \textbf{Compare} Gryning (2007); highlight explicit knowledge on veering-profile % $\rightarrow$ directional sheer;
+%% \textbf{Remember} of interpretation in the context of eddy viscosity (Fig.~\ref{fig:stresses}) 
+% Consider Townsend ``Turbulent Shear Flow''  Chapter 7.18: p. 319 
+% Implications for \textbf{K-theory} (we now can consider that shear and stress are not necessarily perfectly aligned).
+% $\rightarrow$ can we do something to infer a K-profile from these theoretical considerations? 
 \begin{figure}
@@
     Thick, solid lines show theory, dashed lines data from DNS. }
-  \label{fig:hodograph} 
+  \label{fig:summary} 
 \end{figure}
@@
 %
-Implications for \textbf{K-theory} (we now can consider that shear and stress are not necessarily perfectly aligned).
-$\rightarrow$ can we do something to infer a K-profile from these theoretical considerations? 
-%
-\begin{figure}
-  \includegraphics[width=\textwidth]{../plot/mag_dir.pdf} 
-  \caption{Total velocity and veering (in degrees) vs inner and outer height.
-    Dashed lines show DNS data, thick lines are from semi-empirical theory.} 
-\end{figure} 
+
+%
+%\begin{figure}
+%  \includegraphics[width=\textwidth]{../plot/mag_dir.pdf} 
+%  \caption{Total velocity and veering (in degrees) vs inner and outer height.
+%    Dashed lines show DNS data, thick lines are from semi-empirical theory.} 
+%\end{figure} 
 %
@@
 %
-Applications:
+\subsection{Summary}
+We investigate the wind veer in Ekman flow.
+Based on scaling considerations and direct numerical simulation spanning on decade in external separation from $Re_\Lambda=1.25 \times 10^5$ to $Re_\Lambda=1.28\times 10^6$, we derive a universal formulation for the both horizontal components of the velocity profile in a stationary, smooth Ekman layer.
+% 
+This formulation is consistent with the DNS data and also yields reasonable results 
+at geophysical scale separation; the logarithmic law of the mean velocity is recovered with the well-known limits and deviations towards the surface and Ekman layer. 
+%
+The classic formulation of the Ekman layer, employing the surface boundary condition, is replaced by a modified solution that can be obtained by Ekman's system of governing equations, but using different boundary conditions that are more appropriate of the actual situation encountered in the planetary boundary layer.
+% 
+The three parameters that characterize this boundary condition are estimated based on DNS data. 
+% 
+\par
+% 
+To quantify the spanwise velocity component consistently across the boundary layer, 
+we derive a universal scaling of the spanwise velocity component in a shear-aligned reference frame. For the \emph{inner layer}, we find the mixed scaling  
+\[V^{\alpha_\star}/G = \frac{1}{\delta^+} f(z^+),\]
+that is, the spanwise velocity normalized by the outer velocity scale is a universal function of the inner height and Friction Reynolds number $Re_\tau = \delta^+$. 
+% 
+This scaling is derived here based on scaling considerations, and it is in excellent agreement with the DNS data available.   
+% 
+In the outer layer, the spanwise velocity follows outer scaling, consistently with the Ekman model mentioned above (Sec.~\ref{sec:ekman}).
+% 
+\par
+%
+The velocity profiles are shown in inner and outer scaling as well as in ho\-do\-graph- view in Fig.~\ref{fig:summary}.
+The shear-aligned velocity deficit is shown in outer scaling highlighting the universality in the outer layer.
+% 
+In the inner layer, the logarithmic profile is (blue, panel \textbf{(b)}) is ecnountered across a widening region for increasing Reynolds number. 
+% 
+In this simple inner scaling, the spanwise velocity (which follows a mixed scaling) 
+does not collapse but seems to depend on $Re$ (the collapse is seen in Fig.~\ref{fig:inner_w}).
+
+\subsection{Applications \& Outlook} 
 \begin{itemize}

### 2024-10-14 plot/plot.py

@@ -18,7 +18,7 @@ D2R = np.pi / 180.
 use_data=False   # if true -- the netcdf files from DNS are needed 
 plot_ustar_alpha=False
 plot_ekman_ideal=False 
-plot_summary=False
+plot_summary=True
 plot_evisc=False
 plot_profiles=False
 plot_visc_outer=False
@@ -32,7 +32,7 @@ test_inner_streamwise=False
 plot_profile_comparison=False
 plot_vandriest=False
 plot_total_rot=False
-plot_shear_vs_rotation=True
+plot_shear_vs_rotation=False
 
 colors = {400 : 'gray',
           500 : 'pink',
@@ -1204,8 +1204,8 @@ if plot_summary :
         ax3.plot(umr_loc,wmr_loc,alpha=alpha_loc,c='black') #c=c_loc)
     
         i_loc = mp.geti(wmr_loc,np.amax(wmr_loc))
-        if re_loc < 10001:
-            ax3.text(umr_loc[i_loc],wmr_loc[i_loc],'Re='+str(re_loc),c='black')#c_loc)
+        if re_loc in [500, 1600,5000,30000,1000000]:
+            ax3.text(umr_loc[i_loc]*0+0.8,wmr_loc[i_loc],'Re='+str(re_loc),c='black',alpha=alpha_loc)#c_loc)
             yp01 = 0.08 *us_loc**2 / nu_loc
 
         ax2.plot(yp_loc[1:]*sqdp, up_loc[1:], c='blue',lw=2,ls='-',alpha=alpha_loc)

### 2024-10-17 write1/manuscript.tex
first complete draft

@@
 %
-\subsection{Summary}
 We investigate the wind veer in Ekman flow.
@@
 % 
-In the inner layer, the logarithmic profile is (blue, panel \textbf{(b)}) is ecnountered across a widening region for increasing Reynolds number. 
+In the inner layer, the logarithmic profile (blue, panel \textbf{(b)}) is ecnountered across a widening region for increasing Reynolds number. 
 % 
@@
 does not collapse but seems to depend on $Re$ (the collapse is seen in Fig.~\ref{fig:inner_w}).  
-
-\subsection{Applications \& Outlook} 
-\begin{itemize}
-\item reference-shear for neutral profile approaches(systematic!) $\rightarrow$ wind engineering! 
-\item initial condition for LES/DNS to eliminate/minimize inertial oscillation in Benchmark simulations
-\item
-\end{itemize}
-
-\bibliographystyle{spbasic_updated} 
-\bibliography{ansorge}
+% 
+While, at a fixed height, the spanwise velocity scales approximately as $Re_\tau^{-1}$, the fraction 
+of turning that is encountered within the inner layer of the flow amounts to about 1/3 of the total wind veer 
+(Fig.~\ref{fig:etling}a).
+% 
+The hodographs show the well-known Ekman shape with the laminar profile as an outer limit and 'flatter' 
+hodographs, corresponding to less turning, for increasing Reynolds number. 
+% 
+Importantly, our results suggest that there is no lower limit of the turning. 
+% 
+This implies that, despite its very large scale separation and huge Reynolds number, the ABL cannot be 
+in the 'limit' of high Reynolds number from the perspective of wind veer.
+
+%\subsection{Applications \& Outlook} 
+%\begin{itemize}
+%\item reference-shear for neutral profile approaches(systematic!) $\rightarrow$ wind engineering! 
+%\item initial condition for LES/DNS to eliminate/minimize inertial oscillation in Benchmark simulations
+%\item
+%\end{itemize}
 
@@
 
+\vspace{1em}
+\noindent\textbf{Acknowledgements} This study is funded by the European Commission's 7$^{th}$ Framework Programme through 
+the ERC 2019 Starting Grant trainABL (ID 851374). Simulations were carried out on the German supercomputing systems 
+\texttt{hawk} at H\"ocstleistungsrechenzentrum Stuttgart (Bundesprojekt 44187) and \texttt{juwels} at Gauss Centre for supercomputing, 
+For\-schungs\-zen\-trum J\"ulich (Projects \texttt{hku24} and \texttt{stadit}). 
+We extend our thanks to Sally Issa for valuable feedback on the manuscript.
+
+\section*{Declarations}
+
+\noindent\textbf{Author Contributions} CA conceived the study, acquired funding, performed the simulation, analysed and plotted the data for this work.
+CA and HW interpreted the data and compiled the manuscript.
+\vspace{1em}
+
+\noindent\textbf{Funding} CA is funded by ERC Starting Grant \emph{trainABL} (ID 851374). Open Access funding for this publication through the DEAL agreement. 
+\vspace{1em}
+
+\noindent\textbf{Data Availability} The data is available in long-term repositories and indexed by DOIs as cited in the list of references.
+Further analysis tools and processed data can be made available upon direct request.
+\vspace{1em}
+
+\noindent\textbf{Conflict of Interest} The Authors declare no competing interest.
+\vspace{1em}
+
+\noindent\textbf{Open Access} This article is licensed under a Creative Commons Attribution 4.0 International License, which permits use, sharing, adaptation, distribution and reproduction in any medium or format, as long as you give appropriate credit to the original author(s) and the source, provide a link to the Creative Commons licence, and indicate if changes were made.
+To view a copy of this licence, visit \href{http://creativecommons.org/licenses/by/4.0/}{https://creativecommons.org/licenses/by/4.0/}.
+
+\bibliographystyle{spbasic_updated}  
+\bibliography{ansorge}
+
+\end{document}
 \appendix

### 2024-10-31 write1/manuscript.tex
worked on comments from Hauke

@@
 In any quantitative description of the surface layer, the friction velocity $u_\star$ is the dynamic scale
-and $\alpha$ defines the alignment of the frame of reference.
+and $\alpha$ defines the horizontal alignment of the frame of reference, i.e. the rotation of the surface 
+friction with respect to the outer wind direction. 
 %
@@
 Attempts were also undertaken to obtain profiles of the wind speed: 
-One approach is to match the inner and outer layer at a reference height;
-\cite{etling:2002, emeis:2018}  (Sec. 21.10; Eq. 21.48) choose the Prandtl-layer height $z_\text{Prandtl}$ to match
+One approach is to match the inner and outer layer at a reference height as usggested by
+\cite{etling:2002, emeis:2018}  (Sec. 21.10; Eq. 21.48); they choose the Prandtl-layer height $z_\text{Prandtl}$ to match
 the wind speed profiles, which, however, requires external prescription of $\alpha(z_\text{prandtl})$, the veering at that height. 
@@
 Ekman-layer models are roughly based on Ekman's seminal \citeyear{ekman:AMA1905} paper in combinations with
-additional assumptions, such as a prescribed profile shape for eddy viscosity \cite{ellison:QJR1955} and two-layer models of the ABL take into account rotational effects at higher altitudes,
+additional assumptions.
+One option is a prescribed profile shape for eddy viscosity (\cite{ellison:QJR1955}), another are two-layer models of the ABL that take into account rotational effects at higher altitudes,
 for instance when the wind speed needs to be evaluated at heights on the order of $100-200\,\mathrm{m}$,
-a particular concern when it comes to wind-power forecasting \citep{optis:BM2014}.
+a particular concern when it comes to wind-power forecasting \citep{etling:2002, emeis:2018, optis:BM2014}.
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Problem Formulation and Numerical Approach}
+\section{Problem formulation and numerical approach}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Notation and Governing Equations}
+\subsection{Notation and governing equations}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-The dimensional velocity vector of the numerical simulations is $\underline{U} = (U_1,U_2,U_3) = (U,V,W)$ over the coordinate system
-$Oxyz$, where an approximate alignment (plus/ minus few degrees) of the direction $Ox$ with the surface shear stress is achieved.
+The dimensional velocity vector of the numerical simulations is $\underline{U} = (U_1,U_2,U_3) = (U,V,W)$ over the coordinate system $Oxyz$, where an approximate alignment (plus/ minus few degrees) of the direction $Ox$ with the surface shear stress is achieved.
+%
+We consider velocity profiles only, i.e. all velocities are averaged over horizontal planes and in time, that is, they 
+correspond to an Ensemble average. 
+%
 The coordinate $Oz$ points away from the wall, and $Oy$ points in the span-wise direction normal to $Oxz$.
@@
 %
-We denote the modulus of the surface shear, the surface friction, by
+We denote the square root of the modulus of surface shear, the surface friction, by
 \begin{equation}
@@
 %
-Another common measure of scale separation is the Reynolds number defined by the laminar Ekman layer
-thickness $D=\sqrt{2\nu/f}$, 
-\begin{equation}
-  \RE_D = \frac{G D} {\nu } = \sqrt{2 \RE_\Lambda}. 
-\end{equation}
-%
-Another common measure of scale separation is the Reynolds number defined by the laminar Ekman layer
-thickness $D=\sqrt{2\nu/f}$, 
-\begin{equation}
-  \RE_D = \frac{G D} {\nu } = \sqrt{2 \RE_\Lambda}. 
-\end{equation}
+Another common measure of scale separation is the Reynolds number $\RE_D = G D /\nu $  defined by the laminar Ekman layer thickness $D=\sqrt{2\nu/f}$.
 %
@@
 \begin{eqnarray}
-  \frac{\partial u_i}{\partial t} &=& \frac{\p \pi}{\p x_i} - u_j \frac{\p u_i}{\p x_j} + \epsilon_{i2j} (u_j -G_j)   + \frac{1}{\RE} \frac{\p^2 u_i}{\p x_{j}^2} \\ 
+  \frac{\partial u_i}{\partial t} &=& \frac{\p \pi}{\p x_i} - u_j \frac{\p u_i}{\p x_j} + \epsilon_{i2j} (u_j -g_j)   + \frac{1}{\RE_\Lambda} \frac{\p^2 u_i}{\p x_{j}^2} \\ 
   \frac{\partial u_j}{\partial x_j} &=& 0,   
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Numerical Simulations}
+\subsection{Numerical simulations}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 and discussed in more detail as the drag law below in Sec.~\ref{sec:drag-law}
-(we find the expected slight decrease of $u_\star$ and $\alpha_\star$ with increasing $\RE$).
+(we find the expected slight decrease of $u_\star/G$ and $\alpha_\star$ with increasing $\RE$).
 %
@@
       \toprule
-      Re & $\delta^+=\RE_\tau$& $u_\star$ & $\alpha_\star$ & $\delta_{95}/\delta$ & $f u_\star^{-3}\int_{0}^{\delta} e \mathrm{d} z $ & $ G^{-3}\int_{0}^{\delta} \epsilon \mathrm{d}z$\\
+      Re & $\delta^+=\RE_\tau$& $u_\star/G$ & $\alpha_\star$ & $\delta_{95}/\delta$ & $f u_\star^{-3}\int_{0}^{\delta} e \mathrm{d} z $ & $ G^{-3}\int_{0}^{\delta} \epsilon \mathrm{d}z$\\
       \midrule 
@@
 %
-Velocity profiles in inner units ($U^+(z^+)$, Fig.~\ref{fig:profiles}a)and outer units ($U^-(z^-)$, Fig.~\ref{fig:profiles}b)
+Velocity profiles in inner units ($U^{\alpha+}(z^+)$, Fig.~\ref{fig:profiles}a)and outer units ($U^{\alpha-}(z^-)$, Fig.~\ref{fig:profiles}b)
 are in accordance with previous work \citep{coleman:JFM1992,spalart:PF2008,spalart:PF2009,ansorge:BM2014,ansorge:BM2019}:
@@
     \caption{
-      \textbf{a)} Shear-aligned velocity profiles in inner units;
-      \textbf{b)} Shear-aligned velocity deficit in outer units; circles mark the
-      height $z^-=0.15$; dashed line gives the best logarithmic fit to the data. 
+      \textbf{a)} Shear-aligned velocity profiles in inner units; circles mark the
+      height $z^-=0.15$;
+      \textbf{b)} Shear-aligned velocity deficit in outer units; diamonds mark the grid point closest to the height $z^+=50$
       \label{fig:profiles}
@@
     \textbf{c-d)} Normalized eddy viscosity $\nu_E$ (solid) plotted versus
-    inner \textbf{(c)} and outer height \textbf{(d)}.
+    inner height \textbf{(c)} and outer height \textbf{(d)}. \textbf{(c)} uses inner normalization; 
+    \textbf{(d)} uses normalization by $\nu \delta^+$ which approximately collapses data in the 
+    outer layer. 
     Different colors are for different Reynolds numbers (cf. Tab.~\ref{tab:sim-setup}).
@@
     $z^-=0.15$, diamonds in \textbf{(b)} and \textbf{(d)} are for $z^+=50$
-    as in Fig.~\ref{fig:profiles}.
+    as in Fig.~\ref{fig:profiles}
     \label{fig:stresses}} 
@@
 \begin{equation}
-  S_\mathrm{turb} = \sqrt{\overline{uw}^2 + \overline{vw}^2}. 
-\end{equation} 
+  S_\mathrm{turb} = \sqrt{\overline{u' w'}^2 + \overline{v'w'}^2}, 
+\end{equation} 
+where dashed quantities $u'$, $v'$, $w'$ indicate deviations from the mean and the overbar denotes 
+horizontal and time averaging. 
+%
 Indeed, $S_\mathrm{turb}$ only follows inner normalization below $z^+\lesssim 20$
@@
     \includegraphics[trim=1152 0 0 0, clip, width=\textwidth]{../plot/PLANES/tke_xy_1.pdf}}
-  \caption{horizontal slices of turbulence kinetic energy in the
+  \caption{Horizontal slices of turbulence kinetic energy in the
     Buffer layer (i=10),
     logarithmic Layer (i=100), and
-    outer layer (i=400);
+    outer layer (i=400) of the case with $Re_\tau=2978$;
     coloring between percentiles 4 and 96 of the respective image. Lower panel:streamwise--vertical intersect through the domain 
@@
 \end{figure} 
-The organization of the flow is depicted in terms of the turbulence kinetic energy in Fig.~\ref{fig:slices}.
+The organization of the flow with $Re_\tau=2978$ is depicted in terms of the turbulence kinetic energy in Fig.~\ref{fig:slices}.
 %
@@
 % 
-\subsection{Drag-Law}
+\subsection{Drag-law}
 \label{sec:drag-law}
@@
 \end{subequations}
-with  $z_\mathrm{ek} = \delta_\mathrm{ek} (y^- -s_\mathrm{ek})$.
+with  $z_\mathrm{ek} = \delta_\mathrm{ek} (z^- -s_\mathrm{ek})$.
 % 
@@
 \end{equation}
-Thus, the reduction of the area under the Ekman hodograph is directly linked to the variation
-in the friction velocity, which quantifies the well-known qualitative observation that a more turbulent boundary
-layer leads to a flatter hodograph in Ekman flow.
+%Thus, the reduction of the area under the Ekman hodograph is directly linked to the variation
+%in the friction velocity, which quantifies the well-known qualitative observation that a more turbulent boundary
+%layer leads to a flatter hodograph in Ekman flow.
 %
@@
     \label{fig:profiles_outer}
-    Shear-aligned velocity deficit for the wall-streamwise (panel \textbf{(a)}) and wall-spanwise (panel \textbf{(b)})
+    Shear-aligned velocity deficit for the streamwise (panel \textbf{(a)}) and spanwise (panel \textbf{(b)})
     components of the mean velocity $U^\alpha$ and $V^\alpha$.
     Solid lines show DNS data, dashed lines the Ekman profiles $U_\mathrm{ek}$ and $V_\mathrm{ek}$ as defined in
-    Eq.~\ref{eqn:profile_ekman}. Region below $z^-=0.3$ is shown opaque. Variations in $U_\mathrm{ek}$ and $V_\mathrm{ek}$ are
-    a consequence of the normalization and related to chagnes in $u_\star$ and $\alpha_\star$ among the different $\RE_D$. 
+    Eq.~\ref{eqn:profile_ekman}. 
+    %Region below $z^-=0.3$ is shown opaque. 
+    Variations in $U_\mathrm{ek}$ and $V_\mathrm{ek}$ are
+    a consequence of the normalization and related to changes in $u_\star$ and $\alpha_\star$ among the different $\RE_D$. 
   }
@@
 
-\subsection{Streamwise Velocity Component}
+\subsection{Streamwise velocity component}
 \label{sec:streamwise} 
-For the streamwise velocity profile (that in non-rotating flows due to the geometry is always aligned with the surface shear stress),
-well-established theories exist for various regimes according to their distance from the wall and the relative role of
-viscosity, turbulence and interaction with the outer region of the flow with the logarithmic law for the mean velocity as a central anchor point.
+Well-established theories exist for the streamwise velocity profile, which in non-rotating flows is aligned with the surface shear stress due to the geometry.
+These theories cover various regimes based on their distance from the wall and the relative influence of viscosity, turbulence, and interaction with the outer flow region, with the logarithmic law for the mean velocity serving as a central reference point.
 %
@@
 %
-We therefore introduce a two-step correction procedure, accounting for the smaller-than linear growth beyond $y^+\approx 5$, and assuring smooth matching with the logarithmic law at $y^+=40$:
-\begin{equation}
-  U_\text{inner}^{{\alpha_\star}^+} = \frac{z^+}{1+c_1 (y^+)^2} +  (c_2 y^+ - a_\mathrm{match} )\frac{1+tanh\left[0.2(y^+-22)\right]}{2} + c_3 e^{-c_4(y^+-22)^2}. \label{eqn:u_inner} 
+We therefore introduce a two-step correction procedure, accounting for the smaller-than linear growth beyond $z^+\approx 5$, and assuring smooth matching with the logarithmic law at $z^+=40$:
+\begin{equation}
+  U_\text{inner}^{{\alpha_\star}^+} = \frac{z^+}{1+c_1 (y^+)^2} +  (c_2 z^+ - a_\mathrm{match} )\frac{1+tanh\left[0.2(z^+-22)\right]}{2} + c_3 e^{-c_4(z^+-22)^2}. \label{eqn:u_inner} 
 \end{equation} 
@@
 %
-The second and third terms on the right hand side vanish for $y^+\ll 22$, and $c_1=0.00185$
-implies an approximately 5\% correction at $y^+=5$ and an 18.5\% correction at $y^+=10$.
+The second and third terms on the right hand side vanish for $z^+\ll 22$, and $c_1=0.00185$
+implies an approximately 5\% correction at $z^+=5$ and an 18.5\% correction at $z^+=10$.
 %
@@
 \begin{equation}
-  {U_\text{log}^{\alpha_\star}}^+ = \frac{1}{\kappa} \log y^+ + C 
+  {U_\text{log}^{\alpha_\star}}^+ = \frac{1}{\kappa} \log z^+ + C 
 \end{equation}
 with the von-K\'arm\'an constant $\kappa=0.416$ and the boundary condition $C=5.4605$. %  log-layer (K\'arm\'an, Prandtl, Jimenez, Marusic)
-For this logarithmic law, $a_\mathrm{match}=3.569861$ for a matching at $y^+=40$. 
+For this logarithmic law, $a_\mathrm{match}=3.569861$ for a matching at $z^+=40$. 
 %
@@
 On the other hand,--provided the outer velocity deficit is $\RE$ independent--the $\RE$-dependence
-of $\alpha_\star$ and $u_\star$ implies that this difference cannot be constant as a function of $\RE$
-%
-\par 
+of $\alpha_\star$ and $u_\star$ implies that this difference cannot be constant as a function of $\RE$.
+%
+%\par 
 %
@@
   \begin{flushleft}
-    \textbf{\ \ a}\hspace{0.45\textwidth}\textbf{\ \ b}\\
+    \textbf{\ \ a)}\hspace{0.45\textwidth}\textbf{\ \ b)}\\
   \end{flushleft} 
   \includegraphics[width=1.0\textwidth]{../plot/w_profile.pdf}
-  \caption{Profiles of shear-aligned span-wise velocity $(W^\alpha)^+$ versus inner height. 
+  \caption{Profiles of shear-aligned span-wise velocity $(V^\alpha)^+$ versus inner height. 
     Dashed lines show DNS data, thick, opaque lines are from the semi-emprical theory developed above.
-    Left panel show standard inner normalization. Right panel shows the inviscid normalization yielding
+    Panel \textbf{(a)} shows standard inner normalization, panel \textbf{(b)} the inviscid normalization yielding
     a universal profile for the spanwise component of velocity in the inner layer. }
@@
 %
+\begin{subequations} 
+  \label{eqn:match}
 \begin{align}
   f_{V,\mathrm{log}}(z^+=10) &=& f_{V,\mathrm{visc}}(z^+=10) &=:& v_{10} \simeq 27.3\\
-  \frac{\p}{\p z^+} \left[ f_{V,\mathrm{log}}\right]_{z^+=10} &=& \frac{\p}{\p z^+}\left[ F_{V,\mathrm{visc}}\right]_{z^+=10} &=:&d_{10}\simeq 4.01 \\
+  \frac{\p}{\p z^+} \left[ f_{V,\mathrm{log}}\right]_{z^+=10} &=& \frac{\p}{\p z^+}\left[ f_{V,\mathrm{visc}}\right]_{z^+=10} &=:&d_{10}\simeq 4.01 \\
   f_{V,\mathrm{log}}(z^+=0.3\delta^+) &=& V^{\alpha_\star}_\mathrm{ek}(z^-=0.3)\delta^+&=:& v_{03} 
 \end{align}
+\end{subequations}
 where $v_{03}$ is determined by $V_\mathrm{ek}(0.3)$ and $U_\mathrm{ek}(0.3)$ and depends on $\RE$.
@@
 %
+\par
+% 
+\subsection{Matching of the inner and outer layer profiles}
+% 
+The formulations introduced above are continuous across the transition from the inner to 
+the outer layer. 
+% 
+However, the requirement of smooth derivatives would over-constrain the velocity profiles and is hence not applied. 
+% 
+This can, in particular for low or extremely high $\RE$ cause discontinuity in the derivatives of the velocity profiles around the transition from the inner to the outer layer. 
+% 
+To avoid such artificial discontinuity, the profiles are blended by an error-function transition using a blending 
+height $z_{blend}^-=0.28-2.25\sqrt{1./Re_\tau}$ and a transition thickness 
+$z_\text{trans}=2$, 
+such that the weighting function $\omega(z^-)$ becomes 
+\[
+  \omega(z^-) = \frac12 \left[\text{erf}\left(z_\text{trans}\log\frac{z^-}{z_{\text{blend}}^{-}}\right) +1\right].
+\]
+hence 
+ $u_\text{total} = (1-\omega) \times u_\text{inner} + \omega \times u_\text{outer}$ and similar for $v$.
+
+%
+\begin{figure}
+  \phantom{AAA}\textbf{(a)\hspace{0.3\textwidth}(b)\hspace{0.3\textwidth}(c)}\\
+  \includegraphics[width=\textwidth]{../plot/ekman_profiles.pdf}
+  \caption{ \textbf{(a)} Velocity deficit, \textbf{(b)} velocity profile in shear-aligned hodographs and \textbf{(c)} hodograph in geostrophy-aligned coordinates.
+   In panels \textbf{(a)} and \textbf{(b)}, blue lines correspond to the streamwise component and red to the spanwise. }
+  \label{fig:summary} 
+\end{figure} 
+The resulting velocity profiles across the entire boundary layer are shown in inner and outer scaling as well as in ho\-do\-graph- view in Fig.~\ref{fig:summary}.
+The shear-aligned velocity deficit is shown in outer scaling highlighting the universality in the outer layer.
+% 
+The logarithmic scaling of the streamwise component is encountered as straight blue lines in panels \textbf{(a)} and \textbf{(b)} where 
+the extent of the logarithmic range increases with $\RE$ towards lower values of $z^-$ and higher values of $z^+$ depending on the scaling. 
+Importantly, the logarithmic region is widening for increasing Reynolds number -- irespective of the scaling. 
+% 
+In this simple inner scaling, the spanwise velocity (which follows a mixed scaling) 
+Does not collapse but seems to depend on $Re$ (the collapse is seen in Fig.~\ref{fig:inner_w}).
+% 
+However, the velocity deficit in outer units becomes approximately universal, also across the inner layer; 
+this reflects the compensation of reduced turning ($\alpha$) by increased drag ($u_\star$), 
+and is consistent with the theoretical considerations discussed in \ref{sec:drag-law}. 
+% 
+\par
+%
+The spanwise velocity at a fixed height scales approximately as $Re_\tau^{-1}$ (Sec.~\ref{sec:spanwise}).
+%
+However, the fraction  of turning that is encountered within the inner layer of the flow amounts to about 1/3 of the total wind veer  (Fig.~\ref{fig:etling}a).  
+%
+This is because, in inner units, the inner layer grows as $Re_\tau$  which exactly compensates the reduced gradient of spanwise velocity. 
+% 
+The hodographs show the well-known Ekman shape with the laminar profile as an outer limit and 'flatter'  hodographs, corresponding to less turning, for increasing Reynolds number.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Implications for surface-layer scaling} 
-Eq.~(\ref{eqn:scaling_spanwise}) establishes a universal mixed scaling for the spanwise velocity
-in the viscous layer:
-%
-While it requires the vertical coordinate to be expressed in inner units,
-the velocity itself is normalized by the geostrophic wind, and becomes
-inversely proportional to the friction Reynolds number $\RE_\tau=\delta^+$
-when considered at a fixed height.
-%
-In vicinity of the surface, such mixed scaling has already been identified for higher-order statistics in convective 
-flows \citep{mellado:BM2016,li:JAS2018}, where large scales leave their signatures in vicinity of the surface.
-%
-It is important to note here that, while $V$ is a first-order statistic from a statistical perspective,
-the spanwise velocity is a higher-order correction term from the perspective of similarity theory and
-from the viewpoint of the channel-flow analogy that is routinely employed in the surface layer.
-%
-Further, this is consistent with the scaling for the velocity hodograph found in
-Eq.~(\ref{eqn:hodograph_scaling}) where the friction velocity also drops out.
-%
-\par
-%
-In the surface layer, there is not only a mixed scaling--as we had already identified
-in the viscous layer--, but we cannot find a universal function onto which the profiles
-of spanwise velocity collapse. 
-%
-This additional degree of freedom reflects the inner--outer matching problem for the
-spanwise velocity, and rather than giving a universal profile for this region,
-as is usually done, we resort here to a parametric description of the problem,
-namely in terms of the function $f_{V,\mathrm{log}}$ determined by the
-parameters $a_\mathrm{log}$, $b_\mathrm{log}$, $c_\mathrm{log}$ which can be estimated
-based on the above scaling considerations for any Reynolds number. 
-%
-We note that, once the parameter $a_\mathrm{log}$ is known, the parameters $b_\mathrm{log}$
-and $c_\mathrm{log}$ can be estimated solely based on $f_{V,\mathrm{visc}}$, i.e. using the
-value $v_{10}$ and $d_{10}$ found for the viscous region of the flow.
-%
-For the range of Reynolds number relevant to geophysical problems ($10^4\lesssim \RE_D \lesssim 10^6$),
-the variation of $c_\mathrm{log}$ is, however, rather small.
+Eq.~(\ref{eqn:scaling_spanwise}) establishes a universal mixed scaling for the spanwise velocity in the viscous layer:
+%
+While it requires the vertical coordinate to be expressed in inner units, the velocity itself is normalized by the geostrophic wind, and becomes inversely proportional to the friction Reynolds number $\RE_\tau=\delta^+$ when considered at a fixed height.
+%
+In vicinity of the surface, such mixed scaling has already been identified for higher-order statistics in convective flows \citep{mellado:BM2016,li:JAS2018}, where large scales leave their signatures in vicinity of the surface.
+%
+It is important to note here that, while $V$ is a first-order statistic from a statistical perspective, the spanwise velocity is a higher-order correction term from the perspective of similarity theory and from the viewpoint of the channel-flow analogy that is routinely employed in the surface layer.
+%
+Further, this is consistent with the scaling for the velocity hodograph found in Eq.~(\ref{eqn:hodograph_scaling}) where the friction velocity also drops out.
+%
+\par
+%
+In the surface layer, there is not only a mixed scaling--as we had already identified in the viscous layer--, but we cannot find a universal function onto which the profiles of spanwise velocity collapse.
+%
+This additional degree of freedom reflects the inner--outer matching problem for the spanwise velocity. 
+%
+Rather than giving a universal profile for this region, we resort here to a parametric description of the problem in terms of the function $f_{V,\mathrm{log}}$ determined by the parameters $a_\mathrm{log}$, $b_\mathrm{log}$, $c_\mathrm{log}$ which can be estimated based on the above scaling considerations for any Reynolds number.
+%
+We note that, once the parameter $a_\mathrm{log}$ is known, the parameters $b_\mathrm{log}$ and $c_\mathrm{log}$ can be estimated solely based on $f_{V,\mathrm{visc}}$, i.e.
+using the value $v_{10}$ and $d_{10}$ found for the viscous region of the flow.
+%
+For the range of Reynolds number relevant to geophysical problems ($10^4\lesssim \RE_D \lesssim 10^6$), the variation of $c_\mathrm{log}$ is, however, rather small.
 
@@
 \begin{align}
-  \frac{\p u^+}{\p y^+} = \frac{2}{ 1+ \sqrt{1+\left(2\kappa y^+\right)^2 (1-\exp\left[-y^+\right])}};
+  \frac{\p u^+}{\p z^+} = \frac{2}{ 1+ \sqrt{1+\left(2\kappa z^+\right)^2 (1-\exp\left[-z^+\right])}};
 \end{align}
@@
 %
-Comparing our proposed formulation (Fig.~\ref{fig:vanDriest}) for the stream-wise velocity in the inner layer to  \citeauthor{vandriest:JOT1956}'s formulation 
-yields later convergence of the velocity onto the logarithmic profile while, over all, it serves as an excellent model of the streamwise 
-velocity component:
-% 
-Notable deviations (on the order of few percent) only occur in the region $10 < z^+ < 30$, 
-where the velocity transitions from the linear to the logarithmic profile.
-
+Comparing our proposed formulation for the stream-wise velocity in the inner layer to  \citeauthor{vandriest:JOT1956}'s formulation yields later convergence of the velocity onto the logarithmic profile while, over all, it serves as an excellent model of the streamwise velocity cmponent (Fig.~\ref{fig:vanDriest}):
+% 
+Notable deviations (on the order of few percent) only occur in the region $10 < z^+ < 30$,  where the velocity transitions from the linear to the logarithmic profile. 
+%
+\par
+%
 %EKMAN SPIRAL 
 For the higher layers of the ABL, the Ekman spiral is the simplest model available. 
-When employed across the entirety of the ABL, the spiral of a turbulent Ekman layer is flattened with respect to Ekman's lamniar solution, which  corresponds to a reduction of the veering angle both at the surface and throughout the ABL.
+%
+When employed across the entirety of the ABL, the spiral of a turbulent Ekman layer is flattened with respect to Ekman's laminar solution, which  corresponds to a reduction of the veering angle both at the surface and throughout the ABL.
+%
 We, however, find that a modified version of the Ekman spiral explicitly taking into account the surface boundary condition, is a consistent model and yields excellent agreement with the velocity profiles from DNS (Sec.~\ref{sec:ekman}).
@@
 \begin{figure}
-  \textbf{(a)\hspace{0.48\textwidth}(b)}\\
-  \includegraphics[width=0.5\textwidth]{../plot/etling_direction_1.pdf} 
-  \includegraphics[width=0.5\textwidth]{../plot/etling_magnitude_1.pdf}
+  \textbf{(a)\hspace{0.48\textwidth}(b)}\\[-0.15em]
+  \includegraphics[width=0.5\textwidth]{../plot/etling_direction_0.png} 
+  \includegraphics[width=0.5\textwidth]{../plot/etling_magnitude_0.png}
   \caption{Comparison of the DNS data (dashed lines) with the two-layer model proposed by Etling (solid lines) for different Reynolds number.
 Left panel shows the relative wind veer, right panel shows the velocity magnitude.
-The etling model is calibrated by the surface veer from the DNS and roughness parameter for the Etling model is chosen according to the correction factor $\exp{\kappa A}$ such that the total veer and velocity magnitude agree.
-DNS profile in the buffer layer, defined here as $z^+<30$ are shown as opaque dashed lines as the two-layer model does not consider viscous effects.}
+The Etling model is calibrated by the surface veer from the DNS and the roughness parameter is chosen according to the correction factor $\exp{\kappa A}$ such that the total veer and velocity magnitude agree.
+Profiles in the buffer layer, defined here as $z^+<30$ are shown as opaque dashed lines as the two-layer model does not consider viscous effects.}
   \label{fig:etling} 
-\end{figure}A two-layer model consolidating both the logartihmic and Ekman layer can be obtained following the arguments by \cite{etling:2008}, cf.
+\end{figure}
+A two-layer model consolidating both the logartihmic and Ekman layer can be obtained following the arguments by \cite{etling:2008}, cf.
 \cite{emeis:2018}.
@@
 % 
-Fig.~\ref{fig:alignment} shows the direction of the velocity, gradient and stress 
-vectors across the boundary layer. It turns out that the negative stress vector with respect to the wind direction and the flux vectors (absolute) have an approximately similar direction. 
+Fig.~\ref{fig:alignment} shows the direction of the velocity, gradient and stress  vectors across the boundary layer.
+It turns out that the negative stress vector with respect to the wind direction and the flux vectors (absolute) have an approximately similar direction.
 %
@@
 %
-However, this implies the a misalignment between the flux and the stress on the order of the wind turning, and indeed Fig.~\ref{fig:alignment}b shows a misalignment up to $90^\circ$.
-This is in accordance with the expectation by \cite[][\S 7.18]{Townsend:1976} and prevents the transfer of Energy from the mean flow to turbulence at these heights, thus preventing a boundary-layer growth.
-% 
-While the behavior in the inner layer seems to depend on $Re$, there emerges universality in the misalignment across the outer layer, suggesting that a universal consideration of misalignment in the context of K-theory is possible for formulations of higher-order quantities.
+However, this implies a misalignment between the flux and the stress on the order of the wind turning, and indeed Fig.~\ref{fig:alignment}b shows a misalignment up to $\pi/8$.
+This is in accordance with the expectation by \cite[][Chap.~7.18]{Townsend:1976} and prevents the transfer of energy from the mean flow to turbulence at these heights, thus preventing a boundary-layer growth.
+% 
+While the behavior in the inner layer seems to depend on $Re$, there emerges universality in the misalignment across the outer layer, suggesting that a consideration of misalignment in the context of K-theory is possible when developing formulations for higher-order quantities.
 %
 \begin{figure}
-  \textbf{(a)\hspace{0.48\textwidth}(b)}\\
+  \textbf{(a)\hspace{0.48\textwidth}(b)}\\[-0.15em]
   \includegraphics[width=0.5\textwidth]{../plot/stress_rotation.pdf}
@@
 % $\rightarrow$ can we do something to infer a K-profile from these theoretical considerations? 
-\begin{figure}
-  \phantom{AAA}\textbf{(a)\hspace{0.3\textwidth}(b)\hspace{0.3\textwidth}(c)}\\
-  \includegraphics[width=\textwidth]{../plot/ekman_profiles.pdf}
-  \caption{ \textbf{(a)} Velocity deficity, \textbf{(b)} velocity profile in shear-aligned hodographs and \textbf{(c)} hodograph in geostrophy-aligned coordinates.
-    Thick, solid lines show theory, dashed lines data from DNS. }
-  \label{fig:summary} 
-\end{figure}
-
 %
@@
 %
-We investigate the wind veer in Ekman flow.
+We investigate the wind veer and fundamental scaling properties of the velocity profiles in Ekman flow. 
+%
 Based on scaling considerations and direct numerical simulation spanning on decade in external separation from $Re_\Lambda=1.25 \times 10^5$ to $Re_\Lambda=1.28\times 10^6$, we derive a universal formulation for the both horizontal components of the velocity profile in a stationary, smooth Ekman layer.
 % 
-This formulation is consistent with the DNS data and also yields reasonable results 
-at geophysical scale separation; the logarithmic law of the mean velocity is recovered with the well-known limits and deviations towards the surface and Ekman layer. 
+This formulation is consistent with the DNS data and also yields reasonable results  at geophysical scale separation; the logarithmic law of the mean velocity is recovered with the well-known limits and deviations towards the surface and Ekman layer.
 %
@@
 % 
-To quantify the spanwise velocity component consistently across the boundary layer, 
-we derive a universal scaling of the spanwise velocity component in a shear-aligned reference frame. For the \emph{inner layer}, we find the mixed scaling  
-\[V^{\alpha_\star}/G = \frac{1}{\delta^+} f(z^+),\]
+To quantify the spanwise velocity component consistently across the boundary layer, we derive a universal scaling of the spanwise velocity component in a shear-aligned reference frame.
+For the \emph{inner layer}, we find the mixed scaling
+%
+\begin{equation} 
+  V^{\alpha_\star}/G = \frac{1}{\delta^+} f(z^+)
+\end{equation}
+%
 that is, the spanwise velocity normalized by the outer velocity scale is a universal function of the inner height and Friction Reynolds number $Re_\tau = \delta^+$. 
@@
 % 
-In the outer layer, the spanwise velocity follows outer scaling, consistently with the Ekman model mentioned above (Sec.~\ref{sec:ekman}).
-% 
-\par
-%
-The velocity profiles are shown in inner and outer scaling as well as in ho\-do\-graph- view in Fig.~\ref{fig:summary}.
-The shear-aligned velocity deficit is shown in outer scaling highlighting the universality in the outer layer.
-% 
-In the inner layer, the logarithmic profile (blue, panel \textbf{(b)}) is ecnountered across a widening region for increasing Reynolds number. 
-% 
-In this simple inner scaling, the spanwise velocity (which follows a mixed scaling) 
-does not collapse but seems to depend on $Re$ (the collapse is seen in Fig.~\ref{fig:inner_w}).  
-% 
-While, at a fixed height, the spanwise velocity scales approximately as $Re_\tau^{-1}$, the fraction 
-of turning that is encountered within the inner layer of the flow amounts to about 1/3 of the total wind veer 
-(Fig.~\ref{fig:etling}a).
-% 
-The hodographs show the well-known Ekman shape with the laminar profile as an outer limit and 'flatter' 
-hodographs, corresponding to less turning, for increasing Reynolds number. 
-% 
-Importantly, our results suggest that there is no lower limit of the turning. 
-% 
-This implies that, despite its very large scale separation and huge Reynolds number, the ABL cannot be 
-in the 'limit' of high Reynolds number from the perspective of wind veer.
+In the outer layer, the spanwise velocity follows outer scaling, consistently with the Ekman model discussed above (Sec.~\ref{sec:ekman}).
+% 
+\par
+% 
+Our results suggest that there is no lower limit of the turning. 
+% 
+Hence---despite its very large scale separation / huge Reynolds number---the ABL is not 
+in the 'limit' of high Reynolds number from the perspective of wind veer, but always in a regime where 
+changes in $\RE$ impact on the vertical partitioning of rotation.
 
@@
 \label{app:ekman_solution}
-
+The following Ekman system is obtained by averaging the Navier--Stokes equations horizontally and over time and neglecting the turbulent transport terms (turbulence can then be considered via the eddy-viscosity concept through variations in the viscosity $\nu$):
 \begin{subequations}
@@
     \p_t U\\
-    \p_t W  
+    \p_t V  
   \end{matrix}\right)&=&\left(\begin{matrix}
-     fW &+& \nu \p_z^2 U\\ 
-    -f(U-G) &+& \nu \p_z^2 W
+     fV &+& \nu \p_z^2 U\\ 
+    -f(U-G) &+& \nu \p_z^2 V
   \end{matrix}\right)\\ 
-  \Rightarrow \partial_t (U+iW) &=& f(W-i(U-G))) + \nu \p_z^2(U+iW)
+  \Rightarrow \partial_t (U+iV) &=& f(V-i(U-G))) + \nu \p_z^2(U+iV)
 \end{eqnarray}
@@
   \hat{u}(z) &=& U_{\infty} + e^{-\gamma z} \left[A \cos\gamma z + B \sin\gamma z\right]   \\
-  \hat{w}(z) &=& W_{\infty} + e^{-\gamma z} \left[-A \sin\gamma z + B \cos\gamma z\right]
+  \hat{v}(z) &=& V_{\infty} + e^{-\gamma z} \left[-A \sin\gamma z + B \cos\gamma z\right]
 \end{eqnarray} 
 \end{subequations}
-where the constants $U_\infty$, $W_\infty$ set the top boundary condition and $A$ and $B$ set the bottom boundary condition. 
-%
-The most common boundary condition for a surface Ekman layer is $A=U_{\infty}=G$, $B=0$, and $W_{\infty}=0$.
+where the constants $U_\infty$, $V_\infty$ set the top boundary condition and $A$ and $B$ set the bottom boundary condition. 
+%
+The most common boundary condition for a surface Ekman layer is $A=U_{\infty}=G$, $B=0$, and $V_{\infty}=0$.
 %
@@
 %
+\begin{subequations}
 \begin{eqnarray}
@@
     u(z_\text{match}) \equiv u_\text{match} &=& U_{\infty} + e^{-\gzm} \left [ A \cos\gzm + B\sin\gzm \right] \\
-    w(z_\text{match}) \equiv w_\text{match} &=& W_{\infty} + e^{-\gzm} \left [-A \sin\gzm + B\cos\gzm \right]
+    v(z_\text{match}) \equiv v_\text{match} &=& V_{\infty} + e^{-\gzm} \left [-A \sin\gzm + B\cos\gzm \right]
   \end{matrix} 
@@
     u_\text{match}-U_{\infty} \\ 
-    w_\text{match}-W_{\infty}
+    v_\text{match}-V_{\infty}
   \end{matrix}\right) = e^{-\gzm}\left(\begin{matrix}A \\ B \end{matrix}\right)\left(\begin{matrix}
@@
 \end{eqnarray}
+\end{subequations}
 Matching the profile at $\xi=0$, one obtains
@@
 %
-As, however, in our context, the thickness and position of the spiral can already be controlled by
-the eddy viscosity and an offset in $\zeta$, here we let $B=0$.
+However, in our context, the thickness and position of the spiral can already be controlled by
+the eddy viscosity and an offset in $\xi$, here we let $B=0$.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2024-10-31 plot/EkmanProfileClass.py

@@ -267,6 +267,7 @@ class EkmanUniversalClass:
         u[i_mtc:]  = log_law[i_mtc:] 
             
         # outer-layer deficit
+        trans_scale=2.0
         ctr=0.28- 2.25*np.sqrt(1./re)
         wgt_o=self.erf_transition(trans_scale,ctr,ym[1:]) #(sp.special.erf(trans_scale*np.log(ym[1:]/ctr))+1)/2 #starts at index 1! 
         u[1:] -= ( wgt_o*(log_law[1:]-ous[1:]) )

### 2024-10-31 plot/plot.py

@@ -18,7 +18,7 @@ D2R = np.pi / 180.
 use_data=False   # if true -- the netcdf files from DNS are needed 
 plot_ustar_alpha=False
 plot_ekman_ideal=False 
-plot_summary=True
+plot_summary=False
 plot_evisc=False
 plot_profiles=False
 plot_visc_outer=False
@@ -32,7 +32,7 @@ test_inner_streamwise=False
 plot_profile_comparison=False
 plot_vandriest=False
 plot_total_rot=False
-plot_shear_vs_rotation=False
+plot_shear_vs_rotation=True
 
 colors = {400 : 'gray',
           500 : 'pink',
@@ -641,9 +641,9 @@ if plot_profiles:
     plt.show()
     plt.close('all') 
 
-    fig1=plt.figure(figsize=(12,4.5))
-    ax1=fig1.add_axes([0.06,0.11,0.43,0.87])
-    ax2=fig1.add_axes([0.56,0.11,0.43,0.87])
+    fig1=plt.figure(figsize=(9,3.25))
+    ax1=fig1.add_axes([0.08,0.13,0.41,0.85])
+    ax2=fig1.add_axes([0.56,0.13,0.43,0.85])
 
     fig2=plt.figure(figsize=(5,4))
     ax3=fig2.add_axes([0.1,0.1,0.85,0.85]) 
@@ -656,7 +656,7 @@ if plot_profiles:
     vlog=sc.profile_log(ylog) 
     ax1.plot(ylog,sc.profile_log(ylog),lw=3,ls='-',c='black',alpha=0.5,
              label=r'$\kappa^{-1} log z^+ + C$')
-    ax1.plot([50,50],[0,21],ls='--',lw=0.5,c='black') 
+    # ax1.plot([50,50],[0,21],ls='--',lw=0.5,c='black') 
     for re in [500,750,1000,1300,1600]:
         c=colors[re]
         nu=2./re**2
@@ -680,8 +680,8 @@ if plot_profiles:
         ax2.plot(ym[1:],um[1:]-um[-1],ls='-',c=c,lw=2)
         ax2.plot(ym[1:],vm[1:]-vm[-1],ls='--',c=c,lw=2)
         ax2.plot([0,1.5],[0,0],ls='--',lw=0.5,c='black')
-        ax2.plot([0.3,0.3],[-7,1.1],ls='--',lw=0.5,c='black')
[…]

### 2024-10-31 write1/manuscript.tex
added affiliation for Hauke

@@
 \date{\scriptsize Submitted \today}
-\title{Wind Veer and Speed in Turbulent Ekman Flow - Part 1: Scaling analysis and universal profile model}
+\title{Wind Veer and Speed in Turbulent Ekman Flow Part I: Scaling analysis and universal profile model}
 \titlerunning{A Universal Velocity Profile for Ekman Flow}
-\author{Cedrick Ansorge$^1$ and Hauke Wurps$^2$}
+\author{Cedrick Ansorge$^1$ and Hauke Wurps$^{2,3}$}
 \authorrunning{C Ansorge and H Wurps}
-\institute{$^{1}$\,FU Berlin, Institut f\"ur Meteorologie, Carl-Heinrich-Becker-Weg 6--10, 12165 Berlin {cedrick@posteo.de}\\
-  $^{2}$\,Uni Oldenburg 
+\institute{$^{1}$\,FU Berlin, Institut f\"ur Meteorologie, Carl-Heinrich-Becker-Weg 6--10, 12165 Berlin, Germany {cedrick@posteo.de}\\
+  $^{2}$\,Carl von Ossietzky Universit\"at Oldenburg, School of Mathematics and Science, Institute of Physics \\ 
+  $^{3}$\,ForWind - Center for Wind Energy Research, K\"upkersweg 70, 26129 Oldenburg, Germany 
   }

### 2024-11-07 write1/manuscript.tex
Minor edits by Hauke

@@
 %
-Another common measure of scale separation is the Reynolds number $\RE_D = G D /\nu $  defined by the laminar Ekman layer thickness $D=\sqrt{2\nu/f}$.
+Another common measure of scale separation is the Reynolds number 
+\begin{equation}
+  \RE_D = \frac{G D}{\nu} 
+\end{equation}  defined by the laminar Ekman layer thickness $D=\sqrt{2\nu/f}$.
 %
@@
 
-\section{Laminar Ekman solution with consideration of inner layer}
+\section*{Appendix A: Laminar Ekman solution with consideration of inner layer}
 \label{app:ekman_solution}
@@
 The lower boundary condition, however, neglects the existence of the surface layer, and it appears reasonable to define
-$A=c G$ where $c<1$ is a constant that icorporates the increased shear in the surface layer.
+$A=c G$ where $c<1$ is a constant that incorporates the increased shear in the surface layer.
 %
@@
 
-\section{Matching the spanwise velocity profiles in the inner layer}
+\section*{Appendix B: Matching the spanwise velocity profiles in the inner layer}
+
 \label{app:log-coefficients}

### 2024-11-07 write1/manuscript.tex
Fixed labels and numbering of appendices

@@
 \appendix
-
-\section*{Appendix A: Laminar Ekman solution with consideration of inner layer}
+\section{Appendices}
+\subsection{Laminar Ekman solution with consideration of inner layer}
 \label{app:ekman_solution}
@@
 
-\section*{Appendix B: Matching the spanwise velocity profiles in the inner layer}
+\subsection{Matching the spanwise velocity profiles in the inner layer}
 

### 2024-11-13 cover_letter/20241114_coverletter.tex
minor edits based on suggestion from Hauk; added cover letter

\documentclass[brief]{fub}  
\usepackage[utf8]{inputenc}   % ggf. das inputencoding aendern, z.B. "latin1"
\usepackage[english]{babel}
\usepackage{ansorge}         % hier steht der "basename" Ihrer .sty-Datei

\renewcommand{\adressvermerk}{} % ggf. Eintrag vornehmen

\renewcommand{\empfaengeradresse}{
  Boundary-Layer Meteorology \\ 
  To the editor\\
  Prof. William Anderson \\[0.5em]
  -- via E-Mail --

}

\begin{document}

\betreff{Submission of our manuscripts ``Wind veer and speed in turbulent Ekman flow Part I and II'' }

\anrede{
  Dear Professor Anderson, 
}

We are pleased to submit two connected research papers on a topic of crucial importance for Boundary-Layer Meteorology,
namely the profiles on wind speed and direction in the lower part of the boundary layer.
%
The research papers are entitled
\begin{itemize}
\item[I] Wind veer and speed in turbulent Ekman flow Part I: Scaling Analysis and universal profile model
\item[II] Wind veer and speed in turbulent Ekman flow Part II: Extrapolation to Atmospheric Scale Separation
\end{itemize}

%
This work tackles the problem of a consistent scaling framework across the different layers of the boundary layer,
including the viscous sub-layer, buffer layer, logarithmic layer, Ekman layer, and outer layer.
%
\par
%
In Part I, we approach the problem through DNS at unseen scale separation and domain size, we obtain high-quality data across all
these regimes that allow use to identify some robust features and scaling properties and yield valuable insight
on the scaling of stress, turbulent viscosity and the velocity profiles.
%
As we put a particular focus on the wind veer, the span-wise velocity component (which also accomplishes the Ekman transport)
is of central importance and we identify a novel mixed scaling of the component in the surface layer which allows us
to project the turning of the wind also within the lower part of the surface layer for arbitrary Reynolds number.
% 
\par
%
Part II builds upon these results using large-eddy simulation (LES). We being with an a posteriori validation of the LES where
the DNS set-up is matched as exactly as possible in terms of Reynolds number and the z-nough parameter.
This yields valuable insight regarding the physical convergence of the LES to the DNS solution that serves as a reference here. 
%
Studies in typical, idealized amtmospheric-boundary-layer settings illustrate that the LES results are consistent with
the profile model that has been developed and part I and was calibrated using DNS data, provided the LES uses about 150 grid points to resolve the boundary layer.
%
The LES and the profile model developed are independent paradigms to study the velocity profile. Hence, the perfect agreement of LES and DNS data at low Reynolds number together with the agreement between LES and the profile model evaluated for geophysical Reynolds number suggests that both the LES and the profile model behave in a physical way for high Reynolds number. 
%
\par
%
We believe that our research makes a relevant contribution to the field.
%
The findings of this study, particularly of the surface layer similarity, are consistent and
complementary with atmospheric measurements.
%
The model developed in Part I is of practical relevance for applications in wind-power projection, numerical modelling
and can be used for a-priori testing of turbulence parameterizations in idealized settings. 
%
\par
%
We hope that the manuscript raises your interest and complies with the high standard of Boun\-da\-ry-Layer Meteorology and  would be delighted if the manuscript were considered for publication in the journal.

\grussformelsigniert{With Kind Regards \vspace{-0.25em}}

\end{document}

### 2024-11-13 write1/manuscript.tex
minor edits based on suggestion from Hauk; added cover letter

@@
 
-The Coriolis force bends the path of motion on a rotating sphere and establishes geostrophic equilibrium when in balance with a pressure gradient force.
+The Coriolis force bends the apparent path of motion on a rotating sphere and establishes geostrophic equilibrium when in balance with a pressure gradient force.
 Wind veer away from the wind direction in geostrophic equilibrium is
@@
 Attempts were also undertaken to obtain profiles of the wind speed: 
-One approach is to match the inner and outer layer at a reference height as usggested by
+One approach is to match the inner and outer layer at a reference height as suggested by
 \cite{etling:2002, emeis:2018}  (Sec. 21.10; Eq. 21.48); they choose the Prandtl-layer height $z_\text{Prandtl}$ to match
@@
 %
-We therefore resort to the physical manifestation of the eddy diffuysivity in an Ekman layer, and use the
+We therefore resort to the physical manifestation of the eddy diffusivity in an Ekman layer, and use the
 boundary layer depth $\delta_\mathrm{ek} = 0.66 \delta \times 2\pi$.
@@
 % 
-This can, in particular for low or extremely high $\RE$ cause discontinuity in the derivatives of the velocity profiles around the transition from the inner to the outer layer. 
+This can, in particular for low or extremely high $\RE$, cause discontinuity in the derivatives of the velocity profiles around the transition from the inner to the outer layer. 
 % 
 To avoid such artificial discontinuity, the profiles are blended by an error-function transition using a blending 
-height $z_{blend}^-=0.28-2.25\sqrt{1./Re_\tau}$ and a transition thickness 
+height $z_{blend}^-=0.28-2.25\sqrt{1/Re_\tau}$ and a transition thickness 
 $z_\text{trans}=2$, 
 such that the weighting function $\omega(z^-)$ becomes 
-\[
-  \omega(z^-) = \frac12 \left[\text{erf}\left(z_\text{trans}\log\frac{z^-}{z_{\text{blend}}^{-}}\right) +1\right].
-\]
+\begin{align}
+  \omega(z^-) = \frac12 \left[\text{erf}\left(z_\text{trans}\log\frac{z^-}{z_{\text{blend}}^{-}}\right) +1\right],
+\end{align}
 hence 
@@
 the extent of the logarithmic range increases with $\RE$ towards lower values of $z^-$ and higher values of $z^+$ depending on the scaling. 
-Importantly, the logarithmic region is widening for increasing Reynolds number -- irespective of the scaling. 
+Importantly, the logarithmic region is widening for increasing Reynolds number -- irrespective of the scaling. 
 % 
 In this simple inner scaling, the spanwise velocity (which follows a mixed scaling) 
-Does not collapse but seems to depend on $Re$ (the collapse is seen in Fig.~\ref{fig:inner_w}).
+does not collapse but seems to depend on $Re$ (the collapse is seen in Fig.~\ref{fig:inner_w}).
 % 
@@
 %
-Comparing our proposed formulation for the stream-wise velocity in the inner layer to  \citeauthor{vandriest:JOT1956}'s formulation yields later convergence of the velocity onto the logarithmic profile while, over all, it serves as an excellent model of the streamwise velocity cmponent (Fig.~\ref{fig:vanDriest}):
+Comparing our proposed formulation for the stream-wise velocity in the inner layer to  \citeauthor{vandriest:JOT1956}'s formulation yields later convergence of the velocity onto the logarithmic profile while, over all, it serves as an excellent model of the streamwise velocity component (Fig.~\ref{fig:vanDriest}):
 % 
@@
 \end{figure}
-A two-layer model consolidating both the logartihmic and Ekman layer can be obtained following the arguments by \cite{etling:2008}, cf.
+A two-layer model consolidating both the logarithmic and Ekman layer can be obtained following the arguments by \cite{etling:2008}, cf.
 \cite{emeis:2018}.

### 2024-11-20 cover_letter/20241114_coverletter.tex
updated cover letter; consistent lower-case headings in part1

@@
 This work tackles the problem of a consistent scaling framework across the different layers of the boundary layer,
-including the viscous sub-layer, buffer layer, logarithmic layer, Ekman layer, and outer layer.
+including the viscous sub-layer, buffer layer, logarithmic layer, Ekman layer, and outer layer. Both papers 
+are concerned with the study of idealized, stationary and truly neutral Ekman flow over a smooth surface. 
 %
@@
 %
-Part II builds upon these results using large-eddy simulation (LES). We being with an a posteriori validation of the LES where
-the DNS set-up is matched as exactly as possible in terms of Reynolds number and the z-nough parameter.
+Part II builds upon these results using large-eddy simulation (LES). We beging with an a posteriori validation of the LES where
+the DNS set-up is matched as exactly as possible in terms of Reynolds number and the z-nought parameter.
 This yields valuable insight regarding the physical convergence of the LES to the DNS solution that serves as a reference here. 
@@
 Studies in typical, idealized amtmospheric-boundary-layer settings illustrate that the LES results are consistent with
-the profile model that has been developed and part I and was calibrated using DNS data, provided the LES uses about 150 grid points to resolve the boundary layer.
+the profile model that has been developed in part I and was calibrated using DNS data, provided the LES uses about 150 grid points to resolve the boundary layer.
 %
 The LES and the profile model developed are independent paradigms to study the velocity profile.
-Hence, the perfect agreement of LES and DNS data at low Reynolds number together with the agreement between LES and the profile model evaluated for geophysical Reynolds number suggests that both the LES and the profile model behave in a physical way for high Reynolds number.
+Hence, the excellent agreement of LES and DNS data at low Reynolds number together with the agreement between LES and the profile model evaluated for geophysical Reynolds number suggests that both the LES and the profile model behave in a physical way for high Reynolds number.
 %
@@
 
-\grussformelsigniert{With Kind Regards \vspace{-0.25em}}
+With Kind Regards, \vspace{-1em}
+
+\includegraphics[height=1.3cm]{signature_hauke.jpg}~~\includegraphics[height=1.5cm]{signature_transparent.png}\vspace{-2em}
+
+Hauke Wurps \ \ \ \ \ \ \ \ \  Cedrick Ansorge
+
+%\grussformelsigniert{With Kind Regards \vspace{-0.25em}}
 

### 2024-11-20 write1/manuscript.tex
updated cover letter; consistent lower-case headings in part1

@@
 \date{\scriptsize Submitted \today}
-\title{Wind Veer and Speed in Turbulent Ekman Flow Part I: Scaling analysis and universal profile model}
+\title{Wind veer and speed in turbulent Ekman flow part I: scaling analysis and universal profile model}
 \titlerunning{A Universal Velocity Profile for Ekman Flow}
@@
 \end{abstract}
-%
+~\\[1em] 
+\textbf{Keywords} Direct Numerical Simulation $\cdot$ Scale Separation $\cdot$ Ekman Layer $\cdot$ Surface Layer $\cdot$ Hodograph
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\noindent\textbf{Author Contributions} CA conceived the study, acquired funding, performed the simulation, analysed and plotted the data for this work.
+\noindent\textbf{Author contributions} CA conceived the study, acquired funding, performed the simulation, analysed and plotted the data for this work.
 CA and HW interpreted the data and compiled the manuscript.
@@
 
-\noindent\textbf{Data Availability} The data is available in long-term repositories and indexed by DOIs as cited in the list of references.
+\noindent\textbf{Data availability} The data is available in long-term repositories and indexed by DOIs as cited in the list of references.
 Further analysis tools and processed data can be made available upon direct request.
@@
 
-\noindent\textbf{Conflict of Interest} The Authors declare no competing interest.
+\noindent\textbf{Conflict of interest} The Authors declare no competing interest.
 \vspace{1em}
 
-\noindent\textbf{Open Access} This article is licensed under a Creative Commons Attribution 4.0 International License, which permits use, sharing, adaptation, distribution and reproduction in any medium or format, as long as you give appropriate credit to the original author(s) and the source, provide a link to the Creative Commons licence, and indicate if changes were made.
+\noindent\textbf{Open access} This article is licensed under a Creative Commons Attribution 4.0 International License, which permits use, sharing, adaptation, distribution and reproduction in any medium or format, as long as you give appropriate credit to the original author(s) and the source, provide a link to the Creative Commons licence, and indicate if changes were made.
 To view a copy of this licence, visit \href{http://creativecommons.org/licenses/by/4.0/}{https://creativecommons.org/licenses/by/4.0/}.

### 2024-11-21 cover_letter/20241114_coverletter.tex
ready to submit (?)

@@
 
-\betreff{Submission of our manuscripts ``Wind veer and speed in turbulent Ekman flow Part I and II'' }
+\betreff{Submission of our two-part manuscript ``Wind veer and speed in turbulent Ekman flow" }
 
@@
 
-We are pleased to submit two connected research papers on a topic of crucial importance for Boundary-Layer Meteorology,
+We are pleased to submit two connected research papers on a central topic in Boundary-Layer Meteorology,
 namely the profiles on wind speed and direction in the lower part of the boundary layer.
@@
 \begin{itemize}
-\item[I] Wind veer and speed in turbulent Ekman flow Part I: Scaling Analysis and universal profile model
-\item[II] Wind veer and speed in turbulent Ekman flow Part II: Extrapolation to Atmospheric Scale Separation
+\item[I] Wind veer and speed in turbulent Ekman flow part I: scaling analysis and universal profile model
+\item[II] Wind veer and speed in turbulent Ekman flow part II: extrapolation to atmospheric scale separation
 \end{itemize}

### 2024-11-21 write1/manuscript.tex
ready to submit (?)

@@
   for increasing scale separation vide Reynolds numbers. 
+  % 
+  The extrapolation to these scaling to atmospheric scale separation is confirmed via large-eddy simulation 
+  in part II of this manuscript. 
 \end{abstract}
@@
     \textbf{a)}\hspace{0.47\textwidth} \textbf{b)}\\
-    \includegraphics[width=\textwidth]{../plot/uv_innerouter.pdf}\\
+    \includegraphics[width=\textwidth]{Fig1.eps}\\ %uv_innerouter.eps}\\
     \caption{
@@
   \textbf{a)}\hspace{0.47\textwidth} \textbf{b)}\\
-  \includegraphics[width=\textwidth]{../plot/stresses.pdf}
+  \includegraphics[width=\textwidth]{Fig2ab.eps}\\  %stresses.eps}
   \textbf{c)}\hspace{0.47\textwidth} \textbf{d)}\\    
-  \includegraphics[width=\textwidth]{../plot/eddy_viscosity.pdf}\\
+  \includegraphics[width=\textwidth]{Fig2cd.eps}\\ %-eddy_viscosity.eps}\\
   \caption{
@@
   \centerline{
-    \includegraphics[trim=0 0 1152 0, clip, width=0.33\textwidth]{../plot/PLANES/tke_xz_10.pdf}
-    \includegraphics[trim=1152 0 0 0, clip, width=0.33\textwidth]{../plot/PLANES/tke_xz_50.pdf}
-    \includegraphics[trim=0 0 1152 0, clip, width=0.33\textwidth]{../plot/PLANES/tke_xz_400.pdf}}
+    \includegraphics[trim=0 0 1152 0, clip, width=0.33\textwidth]{Fig3a.png}
+    \includegraphics[trim=1152 0 0 0, clip, width=0.33\textwidth]{Fig3b.png}
+    \includegraphics[trim=0 0 1152 0, clip, width=0.33\textwidth]{Fig3c.png}}
   \centerline{
-    \includegraphics[trim=1152 0 0 0, clip, width=\textwidth]{../plot/PLANES/tke_xy_1.pdf}}
+    \includegraphics[trim=1152 0 0 0, clip, width=\textwidth]{Fig3d.png}}
   \caption{Horizontal slices of turbulence kinetic energy in the
@@
 \begin{figure}
-  \centerline{\includegraphics[width=0.7\textwidth]{../plot/ustar_alpha.pdf}}
+  \centerline{\includegraphics[width=0.7\textwidth]{Fig4.eps}} % ustar_alpha.eps}}
   \caption{Variation of geostrophic drag, $Z_\star$, and surface veering, $\alpha_\star$, with Reynolds number according to
@@
   \end{flushleft} 
-  \includegraphics[width=\textwidth]{../plot/ekman_ideal.pdf}
+  \includegraphics[width=\textwidth]{Fig5.eps} %ekman_ideal.eps}
   \caption{\label{fig:ekman_ideal}
@@
   \end{flushleft} 
-  \includegraphics[width=0.5\textwidth]{../plot/outer_layer_u.pdf}\includegraphics[width=0.5\textwidth]{../plot/outer_layer_w.pdf}
+  \includegraphics[width=0.5\textwidth]{Fig6a.eps} %outer_layer_u.eps}
+  \includegraphics[width=0.5\textwidth]{Fig6b.eps} %outer_layer_w.eps}
   \caption{
@@
   \end{flushleft} 
-  \includegraphics[width=\textwidth]{../plot/u_profile.pdf}
+  \includegraphics[width=\textwidth]{Fig7.eps} % u_profile.eps}
   \caption{Shear-aligned profiles of velocity components $U^{\alpha_\star+}$ in inner (left) and outer (right) units.
@@
   \end{flushleft} 
-  \includegraphics[width=1.0\textwidth]{../plot/w_profile.pdf}
+  \includegraphics[width=1.0\textwidth]{Fig8.eps} %w_profile.eps}
   \caption{Profiles of shear-aligned span-wise velocity $(V^\alpha)^+$ versus inner height. 
@@
 \begin{figure}
-  \centerline{\includegraphics[width=0.5\textwidth]{../plot/c_log.pdf}}
+  \centerline{\includegraphics[width=0.5\textwidth]{Fig9.eps}} %c_log.eps}}
   \caption{Coefficients $a_\text{log}$, $b_\text{log}$ and $c_\text{log}$ 
@@
   \phantom{AAA}\textbf{(a)\hspace{0.3\textwidth}(b)\hspace{0.3\textwidth}(c)}\\
-  \includegraphics[width=\textwidth]{../plot/ekman_profiles.pdf}
+  \includegraphics[width=\textwidth]{Fig10.eps}  %ekman_profiles.eps 
   \caption{ \textbf{(a)} Velocity deficit, \textbf{(b)} velocity profile in shear-aligned hodographs and \textbf{(c)} hodograph in geostrophy-aligned coordinates.
@@
 \begin{figure}
-  \centerline{\includegraphics[width=0.66\textwidth]{../plot/vanDriest_profile.png}}
+  \centerline{\includegraphics[width=0.66\textwidth]{Fig11.eps}} %vanDriest_profile.eps}}
   \caption{Near-wall velocity profile according to the 
@@
   \textbf{(a)\hspace{0.48\textwidth}(b)}\\[-0.15em]
-  \includegraphics[width=0.5\textwidth]{../plot/etling_direction_0.png} 
-  \includegraphics[width=0.5\textwidth]{../plot/etling_magnitude_0.png}
+  \includegraphics[width=0.5\textwidth]{Fig12a.eps} %etling_direction.eps} 
+  \includegraphics[width=0.5\textwidth]{Fig12b.eps} %etling_magnitude.eps}
   \caption{Comparison of the DNS data (dashed lines) with the two-layer model proposed by Etling (solid lines) for different Reynolds number.
@@
   \textbf{(a)\hspace{0.48\textwidth}(b)}\\[-0.15em]
-  \includegraphics[width=0.5\textwidth]{../plot/stress_rotation.pdf}
-  \includegraphics[width=0.5\textwidth]{../plot/misalignment.pdf} 
+  \includegraphics[width=0.5\textwidth]{Fig13a.eps} %stress_rotation.eps}
+  \includegraphics[width=0.5\textwidth]{Fig13b.eps} %misalignment.eps} 
   \caption{Panel \textbf{(a)} shows the wind direction (DD) as thick dashed lines and the direction of stress relative to DD as thin solid lines, both according to our profile model.
@@
 %\begin{figure}
-%  \includegraphics[width=\textwidth]{../plot/mag_dir.pdf} 
+%  \includegraphics[width=\textwidth]{mag_dir.eps} 
 %  \caption{Total velocity and veering (in degrees) vs inner and outer height.
@@
 \vspace{1em}
-\noindent\textbf{Acknowledgements} This study is funded by the European Commission's 7$^{th}$ Framework Programme through 
-the ERC 2019 Starting Grant trainABL (ID 851374). Simulations were carried out on the German supercomputing systems 
+\noindent\textbf{Acknowledgements} Simulations were carried out on the German supercomputing systems 
 \texttt{hawk} at H\"ocstleistungsrechenzentrum Stuttgart (Bundesprojekt 44187) and \texttt{juwels} at Gauss Centre for supercomputing, 
@@
 
-\noindent\textbf{Funding} CA is funded by ERC Starting Grant \emph{trainABL} (ID 851374). Open Access funding for this publication through the DEAL agreement. 
+\noindent\textbf{Funding} CA is funded by the European Commission's 7$^{th}$ Framework Programme through 
+the ERC 2019 Starting Grant trainABL (ID 851374). HW is funded by the Ministry for Science and Culture of 
+Lower Saxony (project \emph{Zukunftskonzept Windenergieforschung}). Open Access funding for this publication through the DEAL agreement. 
 \vspace{1em}
@@
 \bibliographystyle{spbasic_updated}  
-\bibliography{ansorge}
+% \bibliography{ansorge}
+\begin{thebibliography}{57}
+\providecommand{\natexlab}[1]{#1}
+\providecommand{\url}[1]{{#1}}
+\providecommand{\urlprefix}{URL }
+\expandafter\ifx\csname urlstyle\endcsname\relax
+  \providecommand{\doi}[1]{DOI~\discretionary{}{}{}#1}\else
+  \providecommand{\doi}{DOI~\discretionary{}{}{}\begingroup
+  \urlstyle{rm}\Url}\fi
+\providecommand{\eprint}[2][]{\url{#2}}
+
+\bibitem[{Ansorge(2019)}]{ansorge:BM2019}
+Ansorge C (2019) Scale {{Dependence}} of {{Atmosphere}}--{{Surface Coupling
+  Through Similarity Theory}}. Boundary-Layer Meteorol 170(1):1--27,
+  \doi{10.1007/s10546-018-0386-y}
+
+\bibitem[{Ansorge(2024{\natexlab{a}})}]{ansorge:2024c}
+Ansorge C (2024{\natexlab{a}}) Direct numerical simulation of turbulent
+  {{Ekman}} flow ({{Re}}=0500): {{DOI}} 10.17169/refubium-42505.
+  \doi{10.17169/refubium-42505}
+
+\bibitem[{Ansorge(2024{\natexlab{b}})}]{ansorge:2024a}
+Ansorge C (2024{\natexlab{b}}) Direct numerical simulation of turbulent
+  {{Ekman}} flow ({{Re}}=1000): {{DOI}} 10.17169/refubium-42507.
+  \doi{10.17169/refubium-42507}
+
+\bibitem[{Ansorge(2024{\natexlab{c}})}]{ansorge:2024}
+Ansorge C (2024{\natexlab{c}}) Direct numerical simulation of turbulent
+  {{Ekman}} flow ({{Re}}=1300): {{DOI}} 10.17169/refubium-42508.
+  \doi{10.17169/refubium-42508}
+
+\bibitem[{Ansorge(2024{\natexlab{d}})}]{ansorge:2024b}
+Ansorge C (2024{\natexlab{d}}) Direct numerical simulation of turbulent
+  {{Ekman}} flow ({{Re}}=1600); {{DOI}} 10.17169/refubium-42509.
+  \doi{10.17169/refubium-42509}
+
+\bibitem[{Ansorge and Mellado(2014)}]{ansorge:BM2014}
+Ansorge C, Mellado JP (2014) Global {{Intermittency}} and {{Collapsing
+  Turbulence}} in the {{Stratified Planetary Boundary Layer}}. Boundary-Layer
+  Meteorol 153(1):89--116, \doi{10.1007/s10546-014-9941-3}
+
+\bibitem[{Ansorge and Mellado(2016)}]{ansorge:JFM2016}
+Ansorge C, Mellado JP (2016) Analyses of external and global intermittency in
+  the logarithmic layer of {{Ekman}} flow. J Fluid Mech 805:611--635,
+  \doi{10.1017/jfm.2016.534}
+
+\bibitem[{Baars and Marusic(2020)}]{baars:JFM2020a}
+Baars WJ, Marusic I (2020) Data-driven decomposition of the streamwise
+  turbulence kinetic energy in boundary layers. {{Part}} 2. {{Integrated}}
+  energy and. J Fluid Mech 882:A26, \doi{10.1017/jfm.2019.835}
+
+\bibitem[{Barenblatt(1993)}]{barenblatt:JFM1993}
+Barenblatt GI (1993) Scaling laws for fully developed turbulent shear flows.
+  {{Part}} 1. {{Basic}} hypotheses and analysis. J Fluid Mech 248:513--520,
+  \doi{10.1017/S0022112093000874}
+
+\bibitem[{Barenblatt and Goldenfeld(1995)}]{barenblatt:PF1995}
+Barenblatt GI, Goldenfeld N (1995) Does fully developed turbulence exist?
+  {{Reynolds}} number independence versus asymptotic covariance. Phys Fluids
+  7(12):3078--3082, \doi{10/bw2sqs}
+
+\bibitem[{Blackadar and Tennekes(1968)}]{blackadar:JAS1968}
+Blackadar AK, Tennekes H (1968) Asymptotic {{Similarity}} in {{Neutral
+  Barotropic Planetary Boundary Layers}}. Journal of the Atmospheric Sciences
+  25:1015--1020, \doi{10.1175/1520-0469(1968)025<1015:ASINBP>2.0.CO;2}
+
+\bibitem[{Brown et~al.(2005)Brown, Beljaars, Hersbach, Hollingsworth, Miller,
+  and Vasiljevic}]{brown:QJR2005}
+Brown AR, Beljaars ACM, Hersbach H, Hollingsworth A, Miller M, Vasiljevic D
+  (2005) Wind turning across the marine atmospheric boundary layer. Quarterly
+  Journal of the Royal Meteorological Society 131(607):1233--1250,
+  \doi{10/dw2bxz}
+
+\bibitem[{Calaf et~al.(2010)Calaf, Meneveau, and Meyers}]{calaf:PF2010}
+Calaf M, Meneveau C, Meyers J (2010) Large eddy simulation study of fully
+  developed wind-turbine array boundary layers. Phys Fluids 22(1):015,110,
+  \doi{10/b7gc6v}
+
+\bibitem[{Coleman et~al.(1992)Coleman, Ferziger, and Spalart}]{coleman:JFM1992}
+Coleman GN, Ferziger JH, Spalart PR (1992) Direct {{Simulation}} of the
+  {{Stably Stratified Turbulent Ekman Layer}}. Journal of Fluid Mechanics
+  244:677--712, \doi{10.1017/S0022112092003264}
+
+\bibitem[{{da Silva} et~al.(2014){da Silva}, Hunt, Eames, and
+  Westerweel}]{dasilva:ARF2014}
+{da Silva} CB, Hunt JC, Eames I, Westerweel J (2014) Interfacial {{Layers
+  Between Regions}} of {{Different Turbulence Intensity}}. Annu Rev Fluid Mech
+  46(1):567--590, \doi{10.1146/annurev-fluid-010313-141357}
+
+\bibitem[{Dimotakis(2005)}]{dimotakis:ARF2005}
+Dimotakis PE (2005) {{TURBULENT MIXING}}. Annual Review of Fluid Mechanics
+  37(1):329--356, \doi{10.1146/annurev.fluid.36.050802.122015}
+
+\bibitem[{Ekman(1905)}]{ekman:AMA1905}
+Ekman VW (1905) On the influence of the earth's rotation on ocean currents. Ark
+  Mat Astron Fys, Vol 2 (1905), pp 1-53 2:1--53
+
+\bibitem[{Ellison(1955)}]{ellison:QJR1955}
+Ellison TH (1955) The {{Ekman}} spiral. Q J Roy Met Soc 81(350):637--638,
+  \doi{10.1002/qj.49708135025}
+
+\bibitem[{Emeis(2018)}]{emeis:2018}
+Emeis S (2018) Wind {{Energy Meteorology}}, 2nd edn. Atmospheric {{Physics}}
+  for {{Wind Power Generation}}, Springer, Heidelberg
+
+\bibitem[{Emeis et~al.(2007)Emeis, {Baumann-Stanzer}, Piringer, Kallistratova,
+  Kouznetsov, and Yushkov}]{emeis:m2007}
+Emeis S, {Baumann-Stanzer} K, Piringer M, Kallistratova M, Kouznetsov R,
+  Yushkov V (2007) Wind and turbulence in the urban boundary layer analysis
+  from acoustic remote sensing data and fit to analytical relations. metz
+  16(4):393--406, \doi{10.1127/0941-2948/2007/0217}
+
+\bibitem[{Esau(2004)}]{esau:EFM2004}
+Esau I (2004) Simulation of {{Ekman Boundary Layers}} by {{Large Eddy Model}}
+  with {{Dynamic Mixed Subfilter Closure}}. Environmental Fluid Mechanics
+  4(3):273--303, \doi{10/b8f3kh}
+
+\bibitem[{Etling(2002)}]{etling:2002}
+Etling D (2002) Theoretische {{Meteorologie}}, 2nd edn. Eine
+  {{Einf{\"u}hrung}}, Springer-Verlag, Berlin, Heidelberg
+
+\bibitem[{Etling(2008)}]{etling:2008}
+Etling D (2008) {Theoretische Meteorologie: eine Einf{\"u}hrung}, 3rd edn.
+  Springer, Berlin Heidelberg
+
+\bibitem[{Foken(2002)}]{foken:m2002}
+Foken T (2002) Some aspects of the viscous sublayer. metz 11(4):267--272,
+  \doi{10.1127/0941-2948/2002/0011-0267}
+
+\bibitem[{Foken et~al.(1978)Foken, Kitajgorodskij, and Kuznecov}]{foken:BM1978}
+Foken {\relax Th}, Kitajgorodskij SA, Kuznecov OA (1978) On the dynamics of the
+  molecular temperature boundary layer above the sea. Boundary-Layer Meteorol
+  15(3):289--300, \doi{10.1007/BF02652602}
+
+\bibitem[{Ghannam and {Bou-Zeid}(2021)}]{ghannam:QJR2021}
+Ghannam K, {Bou-Zeid} E (2021) Baroclinicity and directional shear explain
+  departures from the logarithmic wind profile. Quarterly Journal of the Royal
+  Meteorological Society 147:434--464, \doi{10/gnj6z2}
+
+\bibitem[{Gryning et~al.(2007)Gryning, Batchvarova, Br{\"u}mmer, J{\o}rgensen,
+  and Larsen}]{gryning:BM2007}
+Gryning SE, Batchvarova E, Br{\"u}mmer B, J{\o}rgensen H, Larsen S (2007) On
+  the extension of the wind profile over homogeneous terrain beyond the surface
+  boundary layer. Boundary-Layer Meteorol 124(2):251--268,
+  \doi{10.1007/s10546-007-9166-9}
+
+\bibitem[{H{\"o}gstr{\"o}m(1988)}]{hogstrom:BM1988}
+H{\"o}gstr{\"o}m U (1988) Non-dimensional wind and temperature profiles in the
+  atmospheric surface layer: {{A}} re-evaluation. Boundary-Layer Meteorology
+  42:55--78, \doi{10.1007/BF00119875}
+
+\bibitem[{H{\"o}gstr{\"o}m(1996)}]{hogstrom:BM1996}
+H{\"o}gstr{\"o}m U (1996) Review of some basic characteristics of the
+  atmospheric surface layer. Boundary-Layer Meteorology 78(3-4):215--246,
+  \doi{10.1007/BF00120937}
+
+\bibitem[{Jacobs and Van~Boxel(1988)}]{jacobs:AFM1988}
+Jacobs AF, Van~Boxel JH (1988) Changes of the displacement height and roughness
+  length of maize during a growing season. Agricultural Forest Meteorol
+  42(1):53--62, \doi{10.1016/0168-1923(88)90066-4}
+
+\bibitem[{Jiang et~al.(2018)Jiang, Wang, and Sullivan}]{jiang:JAS2018}
+Jiang Q, Wang S, Sullivan P (2018) Large-{{Eddy Simulation Study}} of {{Log
+  Laws}} in a {{Neutral Ekman Boundary Layer}}. Journal of the Atmospheric
+  Sciences 75(6):1873--1889, \doi{10.1175/JAS-D-17-0153.1}
+
+\bibitem[{Jim{\'e}nez(2012)}]{jimenez:ARF2012}
+Jim{\'e}nez J (2012) Cascades in {{Wall-Bounded Turbulence}}. Ann Rev Fluid
+  Mech 44(Volume 44, 2012):27--45, \doi{10.1146/annurev-fluid-120710-101039}
+
+\bibitem[{Kelly and Gryning(2010)}]{kelly:BM2010}
+Kelly M, Gryning SE (2010) Long-{{Term Mean Wind Profiles Based}} on
+  {{Similarity Theory}}. Boundary-Layer Meteorol 136(3):377--390,
+  \doi{10.1007/s10546-010-9509-9}
+
+\bibitem[{Kelly and Troen(2016)}]{kelly:WE2016}
+Kelly M, Troen I (2016) Probabilistic stability and `tall' wind profiles:
+  Theory and method for use in wind resource assessment. Wind Energy
+  19(2):227--241, \doi{10.1002/we.1829}
+
+\bibitem[{Klein et~al.(2021)Klein, Maier, and Schmidt}]{klein:PAM2021a}
+Klein M, Maier RE, Schmidt H (2021) Stochastic modeling of transient neutral
+  and stably-stratified {{Ekman}} boundary layers. P A M M 21(1):e202100,146,
+  \doi{10.1002/pamm.202100146}
+
+\bibitem[{Li et~al.(2018)Li, Gentine, Mellado, and McColl}]{li:JAS2018}
+Li Q, Gentine P, Mellado JP, McColl KA (2018) Implications of {{Nonlocal
+  Transport}} and {{Conditionally Averaged Statistics}} on {{Monin}}--{{Obukhov
+  Similarity Theory}} and {{Townsend}}'s {{Attached Eddy Hypothesis}}. J Atmos
+  Sci 75(10):3403--3431, \doi{10.1175/JAS-D-17-0301.1}
+
+\bibitem[{Lindvall and Svensson(2019)}]{lindvall:QJR2019}
+Lindvall J, Svensson G (2019) Wind turning in the atmospheric boundary layer
+  over land. Q J Roy Met Soc 145(724):3074--3088, \doi{10.1002/qj.3605}
+
+\bibitem[{Mellado and Ansorge(2012)}]{mellado:ZaM2012}
+Mellado J, Ansorge C (2012) Factorization of the {{Fourier}} transform of the
+  pressure-{{Poisson}} equation using finite differences in colocated grids. Z
+  angew Math Mech 92(5):380--392, \doi{10.1002/zamm.201100078}
+
+\bibitem[{Mellado et~al.(2016)Mellado, {van Heerwaarden}, and
+  Garcia}]{mellado:BM2016}
+Mellado JP, {van Heerwaarden} CC, Garcia JR (2016) Near-{{Surface Effects}} of
+  {{Free Atmosphere Stratification}} in {{Free Convection}}. Boundary-Layer
+  Meteorol 159(1):69--95, \doi{10.1007/s10546-015-0105-x}
+
+\bibitem[{Mirocha et~al.(2018)Mirocha, Churchfield, {Mu{\~n}oz-Esparza}, Rai,
+  Feng, Kosovi{\'c}, Haupt, Brown, Ennis, Draxl, Sanz~Rodrigo, Shaw, Berg,
+  Moriarty, Linn, Kotamarthi, Balakrishnan, Cline, Robinson, and
+  Ananthan}]{mirocha:WES2018}
+Mirocha JD, Churchfield MJ, {Mu{\~n}oz-Esparza} D, Rai RK, Feng Y, Kosovi{\'c}
+  B, Haupt SE, Brown B, Ennis BL, Draxl C, Sanz~Rodrigo J, Shaw WJ, Berg LK,
+  Moriarty PJ, Linn RR, Kotamarthi VR, Balakrishnan R, Cline JW, Robinson MC,
+  Ananthan S (2018) Large-eddy simulation sensitivities to variations of
+  configuration and forcing parameters in canonical boundary-layer flows for
+  wind energy applications. Wind Energy Science 3(2):589--613, \doi{10/gn3nh4}
+
+\bibitem[{Moin and Mahesh(1998)}]{moin:ARF1998}
+Moin P, Mahesh K (1998) Direct numerical simulation: {{A}} tool in turbulence
+  research. Annual Review of Fluid Mechanics 30:539--578,
+  \doi{10.1146/annurev.fluid.30.1.539}
+
+\bibitem[{Momen and {Bou-Zeid}(2016)}]{momen:JAS2016}
+Momen M, {Bou-Zeid} E (2016) Large-{{Eddy Simulations}} and {{Damped-Oscillator
+  Models}} of the {{Unsteady Ekman Boundary Layer}}*. Journal of the
+  Atmospheric Sciences 73(1):25--40, \doi{10.1175/JAS-D-15-0038.1}
+
+\bibitem[{Momen et~al.(2018)Momen, {Bou-Zeid}, Parlange, and
+  Giometto}]{momen:JAS2018}
+Momen M, {Bou-Zeid} E, Parlange MB, Giometto M (2018) Modulation of {{Mean
+  Wind}} and {{Turbulence}} in the {{Atmospheric Boundary Layer}} by
+  {{Baroclinicity}}. Journal of the Atmospheric Sciences 75(11):3797--3821,
+  \doi{10/gfmn67}
+
+\bibitem[{Monin(1970)}]{monin:ARF1970}
+Monin AS (1970) The {{Atmospheric Boundary Layer}}. Annual Review of Fluid
+  Mechanics 2:225--250, \doi{10.1146/annurev.fl.02.010170.001301}
+
+\bibitem[{Monin and Yaglom(1975)}]{monin:1975}
+Monin AS, Yaglom AM (1975) Statistical {{Fluid Mechanics}}, {{Vol}}. {{II}},
+  Dover {{Publications}} on {{Physics}}, vol~II. Dover Publications, Inc.,
+  Mineola
+
+\bibitem[{Optis et~al.(2014)Optis, Monahan, and Bosveld}]{optis:BM2014}
+Optis M, Monahan A, Bosveld FC (2014) Moving {{Beyond Monin-Obukhov Similarity
+  Theory}} in {{Modelling Wind-Speed Profiles}} in the {{Lower Atmospheric
+  Boundary Layer}} under {{Stable Stratification}}. Boundary-Layer Meteorology
+  153(3):497--514, \doi{10.1007/s10546-014-9953-z}
+
+\bibitem[{Rossby and Montgomery(1935)}]{rossby:PIP1935}
+Rossby CG, Montgomery RB (1935) The layer of frictional influence in wind and
+  ocean currents. Papers in Physical Oceanography and Meteorology III(3):1--101
+
+\bibitem[{Sakagami et~al.(2020)Sakagami, Haas, and Passos}]{sakagami:BM2020}
+Sakagami Y, Haas R, Passos JC (2020) Generalized {{Non-dimensional Wind}} and
+  {{Temperature Gradients}} in the {{Surface Layer}}. Boundary-Layer Meteorol
+  175(3):441--451, \doi{10.1007/s10546-020-00510-3}
+
+\bibitem[{Spalart(1989)}]{spalart:JFM1989}
+Spalart PR (1989) Theoretical and numerical study of a three-dimensional
+  turbulent boundary layer. J Fluid Mech 205(-1):319,
+  \doi{10.1017/S0022112089002053}
+
+\bibitem[{Spalart et~al.(2008)Spalart, Coleman, and Johnstone}]{spalart:PF2008}
+Spalart PR, Coleman GN, Johnstone R (2008) Direct numerical simulation of the
+  {{Ekman}} layer: {{A}} step in {{Reynolds}} number, and cautious support for
+  a log law with a shifted origin. Phys Fluids 20(10):101,507,
+  \doi{10.1063/1.3005858}
+
+\bibitem[{Spalart et~al.(2009)Spalart, Coleman, and Johnstone}]{spalart:PF2009}
+Spalart PR, Coleman GN, Johnstone R (2009) Retraction: ``{{Direct}} numerical
+  simulation of the {{Ekman}} layer: {{A}} step in {{Reynolds}} number, and
+  cautious support for a log law with a shifted origin'' [{{Phys}}. {{Fluids}}
+  20, 101507 (2008)]. Phys Fluids 21(10):109,901, \doi{10.1063/1.3247176}
+
+\bibitem[{Stoll et~al.(2020)Stoll, Gibbs, Salesky, Anderson, and
+  Calaf}]{stoll:BM2020}
+Stoll R, Gibbs JA, Salesky ST, Anderson W, Calaf M (2020) Large-{{Eddy
+  Simulation}} of the {{Atmospheric Boundary Layer}}. Boundary-Layer Meteorol
+  177(2-3):541--581, \doi{10/gmbmzw}
+
+\bibitem[{Svensson and Holtslag(2009)}]{svensson:BM2009}
+Svensson G, Holtslag AAM (2009) Analysis of {{Model Results}} for the
+  {{Turning}} of the {{Wind}} and {{Related Momentum Fluxes}} in the {{Stable
+  Boundary Layer}}. Boundary-Layer Meteorol 132(2):261--277, \doi{10/bwknmt}
+
+\bibitem[{Tennekes(1973)}]{tennekes:JAS1973}
+Tennekes H (1973) A {{Model}} for the {{Dynamics}} of the {{Inversion Above}} a
+  {{Convective Boundary Layer}}. Journal of the Atmospheric Sciences
+
+\bibitem[{Townsend(1976)}]{Townsend:1976}
+Townsend AA (1976) The Structure of Turbulent Shear Flow, 2nd edn. Cambridge
+  University Press
+
+\bibitem[{Van~Driest(1956)}]{vandriest:JOT1956}
+Van~Driest ER (1956) On {{Turbulent Flow Near}} a {{Wall}}. Journal of the
+  Aeronautical Sciences 23(11):1007--1011, \doi{10.2514/8.3713}
+
+\bibitem[{Zikanov et~al.(2003)Zikanov, Slinn, and Dhanak}]{zikanov:JFM2003}
+Zikanov O, Slinn DN, Dhanak MR (2003) Large-eddy simulations of the
+  wind-induced turbulent {{Ekman}} layer. J Fluid Mech 495:343--368,
+  \doi{10/ccbpbw}
+
+\end{thebibliography}
 

### 2024-11-21 cover_letter/20241114_coverletter.tex
fixed typos in cover letter

@@
 %
-In Part I, we approach the problem through DNS at unseen scale separation and domain size, we obtain high-quality data across all
+In part I, we approach the problem through DNS at unseen scale separation and domain size, we obtain high-quality data across all
 these regimes that allow use to identify some robust features and scaling properties and yield valuable insight
@@
 %
-Part II builds upon these results using large-eddy simulation (LES). We beging with an a posteriori validation of the LES where
+Part II builds upon these results using large-eddy simulation (LES). We begin with an a posteriori validation of the LES where
 the DNS set-up is matched as exactly as possible in terms of Reynolds number and the z-nought parameter.
@@
 %
-The model developed in Part I is of practical relevance for applications in wind-power projection, numerical modelling
+The model developed in part I is of practical relevance for applications in wind-power projection, numerical modelling
 and can be used for a-priori testing of turbulence parameterizations in idealized settings. 

### 2024-11-21 cover_letter/20241114_coverletter.tex
explained z-nought in cover letter

@@
 Part II builds upon these results using large-eddy simulation (LES). We begin with an a posteriori validation of the LES where
-the DNS set-up is matched as exactly as possible in terms of Reynolds number and the z-nought parameter.
+the DNS set-up is matched as exactly as possible in terms of Reynolds number and the roughness parameter z-nought ($z_0$).
 This yields valuable insight regarding the physical convergence of the LES to the DNS solution that serves as a reference here. 

### 2025-03-12 write1/REVISION1/responses.tex
finished work on Reviewer 1 responses

\documentclass[a4paper,10pt]{article} 
\usepackage[margin=3.2cm]{geometry}
\usepackage{booktabs} 
\usepackage{graphicx} 
\usepackage{amssymb}
\usepackage{amsmath} 
\usepackage{xcolor}
\usepackage{setspace} 
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{doi}
\usepackage{pdfpages}

\onehalfspacing

\newcommand{\revpoint}[3] 
{ \hrule ~\\[0.5em]
\noindent\textcolor{gray}{\noindent#1} \\[0.5em]  
\noindent\textcolor{blue}{\textit{\noindent#2}} \\[0.5em]
\noindent\textcolor{black}{\noindent{#3}} \\[0.5em]
}

\newcommand{\com}[1]{\textcolor{red}{\noindent#1}}

\title{Response to Reviewers} 
\author{Cedrick Ansorge and Hauke Wurps }
\begin{document}
\maketitle

%-------------------------------------------------------------------------------------%

%######################################################################################%
% not all comments are included, only where action is needed.
%######################################################################################%

\section*{Response to Reviewer 1}

Dear Reviewer 1, \\[1em]

\noindent Many thanks for the positive feedback on our manuscript. We have addressed the minor points raised in your review and hope that you agree with the edits. Pleas find the detailed responses below.  \\[1em]

\noindent Kind regards, \\[0.5em]

\noindent Cedrick Ansorge and Hauke Wurps  \\[0.5em]
%-------------------------------------------------------------------------------------%

\revpoint{The authors analyze wind speed and direction profiles in turbulent Ekman flow using asymptotic theory and DNS data. They found that the streamwise velocity follows classical scaling laws, while outer-layer profiles align with a modified Ekman spiral. They showed that the spanwise component exhibits mixed scaling near the surface and outer scaling farther out. \newline  Overall, the analysis and DNS are conducted rigorously, providing valuable insights and contributing meaningfully to advancements in the field.}{We sincerely thank you for the positive assessment of our manuscript and the appreciation of our work.}{}

\revpoint{In line 47, friction velocity is denoted as $u_* = \sqrt{|\tau_{sfc}/\rho}$. However, in Eq. (1a) (below line 179), $\tau_{sfc}$ is defined as the kinematic shear stress. Please make them consistent. }{Manz thanks for spotting this inconsistency in nomenclature. We have aligned $\tau_\text{sfc}$ with the conventional definition of dynamic shear stress and now include the division bz density in Eq. (1b).}{\begin{subequations}\begin{equation}
    \underline{\tau}_\mathrm{sfc} = \left(\begin{array}{c} \tau_x\\ \tau_y\\\tau_z\end{array}\right) = - \rho\nu \left( \frac{\partial U}{\partial z} \hat{e}_x + \frac{\partial V}{\partial z} \hat{e}_y\right)
  \end{equation}
  and labelled by an upper index $\alpha$ as in $\underline{U}^\alpha$ for the velocity vector, and
  (ii) the coordinate system aligned with the free-atmosphere geostrophic wind labelled by an upper index $G$ as in $\underline{U}^G$.
  %
  We denote the square root of the modulus of surface shear, the surface friction, by
  \begin{equation}
    u_\star = \sqrt{ \frac{|| - \underline{\tau}_\mathrm{sfc} ||}{\rho} } 
    \label{eqn:ustar_def}
  \end{equation} and let $Z_\star=G/u_\star$;
\end{subequations}}

\revpoint{Location for the definition of $z^-$. The first appearance of $z^-$ is at line 262, and figure 1b, but it was defined on line 307. It would be better to define it close to where it appears. }{Thank you - we now introduce the outer length scale much earlier in Sec. 2.1 ("Notation and governing equations"). We also think, this is more appropriate.}{The outer normalization (with respect to the boundary-layer depth $\delta$ and velocity $u_\star$) is denoted 
by a superscript $'-'$, i.e. $z^-=z/\delta$.  }

\revpoint{Figure 4: Should the legend be $Z_* = 4 log(Re_D) -8$, instead of $u_\star* = ...$"?}{Indeed, there was a typo, please find the revised Figure below.}{
\includegraphics[width=\textwidth]{Fig4.eps}
}
\newpage
\section*{Response to Reviewer 2}
\revpoint{The authors perform direct numerical simulations of the Ekman layer at moderate Reynolds number, and use their simulations to examine the vertical mean velocity profiles in rotating, neutrally-stratified flows that serve as a proxy for the neutral atmospheric boundary layer. They compare their DNS results to a model that matches the Ekman model (in the outer layer) with the classical results for the inner layer (viscous layer plus log layer) in wall-bounded turbulent flows. Specifically, they examine the scaling of the spanwise velocity profile, and find that a matching procedure—combining both inner- and outer-scaling—is necessary. The article is generally clear and well-written, and addresses the important problem of how velocity profiles behave in the Ekman layer. Despite being an old problem, it has not been resolved, and there are many subtleties that are often neglected in the literature. I am overall supportive of this manuscript, but below I provide some specific comments which I believe will help improve the clarity of communication in the article. I am recommending that the manuscript is accepted in BLM, pending minor revisions. }{}{}

\revpoint{L. 12, 76: discussion of span-wise velocity component. Can you explain how you are defining streamwise and spanwise here? Are you assuming the streamwise component is aligned with the geostrophic wind direction far above the surface? }{}{}
\revpoint{L. 16 outer height - do you mean outer length scale? }{}{}
\revpoint{L. 19 “vide Reynolds numbers” - I’m not sure this makes sense to me. }{}{}
\revpoint{L. 63 Prandtl-layer height is undefined here}{}{}
\revpoint{L. 124 - “require one to consider” or “require consideration of” }{}{}
\revpoint{ll. 168-170 - This seems to be contradictory. The authors state that rotation only acts on the horizontal components of velocity, but go on to state that they neglect rotational effects on the horizontal components of velocity. }{}{}
\revpoint{l. 176 - I do not think that Ensemble should be capitalized }{}{}
\revpoint{L. 193 “this state is defined by a Reynolds number, the only non-dimensional parameter that governs the system.” Shouldn’t the Rossby number also enter as a non-dimensional parameter? Given a domain height, geostrophic wind speed G, and Coriolis parameter f, one can also define the Rossby number. The problem would depend on Reynolds number alone in the absence of rotation, but if there is rotation, the relevant dimensionless parameters should be both Re and Ro. }{}{}
\revpoint{L. 221 - It might be better to use scientific notation here. }{}{}
\revpoint{L. 277, 281 - the notation is a bit awkward here, e.g. $V^{\alpha_* +}(z^+)$. Is there a way to make it clearer? }{}{}
\revpoint{Fig. 1a - Does the solid black line correspond to the Re=1600 case? If so, why is there i a discontinuity just above $10^1$? Or are you also using a solid black line to denote the inner layer velocity profile, $u^+ = z^+$? }{}{}
\revpoint{L. 308 Eddy should not be capitalized}{}{}
\revpoint{L. 311 - The assumptions of horizontal homogeneity and W=0 are not exactly the same thing as symmetries in the flow. }{}{}
\revpoint{Fig. 3: It would be better to present heights in terms of z+ values rather than the index of the vertical coordinate. }{}{}

\newpage
\section*{Response to Reviewer 3}
\revpoint{This is an excellent paper that should be published in BLM. However, I request the authors to tone
down their narrative, including the title. It is an empirical study that tries to fit DNS data using various
formulations. Some of these formulations have a physical basis, while others are ad-hoc. Please refrain
from calling the results “universal”. } {}{}

\revpoint{At the heart of this paper is Eq.~10a. Following Etling's work and others, this equation has been derived in Appendix A1. \newline 
  \[
    \frac{1}{G} \left(\begin{array}{c}U_\text{Ek} \\ V_\text{Ek} \end{array}\right) = \left(\begin{array}{c} 1 \\0 \end{array}\right) + e^{-z_\text{Ek}} \left[ a_\text{ek} \left(\begin{array}{c} -\cos z_\text{ek} \\ \sin z_\text{ek} \end{array}\right)+ b_\text{ek}\left(\begin{array}{c}\sin z_\text{ek} \\ \cos z_\text{ek}\end{array}\right)   \right]
  \] \newline  
  The problem with this equation is the assumption of constant eddy viscosity. As the authors showed in
  Figure 2c, the eddy viscosity profile from the DNS is far from constant. It has a clear maximum and a
  specific shape, which depends on Re.
  So, the authors took an empirical (engineering) approach on page 14 (line 403). They estimated a bulk
  eddy viscosity indirectly. This is a major weakness of this paper and could be avoided. Please see the
  following paper where we analytically derived the eddy viscosity profile from the Ekman equations.
  Since this is a relatively new paper, the authors might not have come across it. That is why I am revealing
  my identity.
  \begin{quotation}Basu, Sukanta, and Albert AM Holtslag. “A novel approach for deriving the stable
  boundary layer height and eddy viscosity profiles from the Ekman
  equations.” Boundary-Layer Meteorology 187.1 (2023): 105-115.
  \end{quotation}
  The authors might want to revise their derivations based on our analytical eddy viscosity profile. If that
  is difficult to do, I suggest highlighting the limitations of their current bulk approach. Also, emphasize
  the limitations of Eq. (10a) as it assumes constant eddy viscosity. }{}{}

\revpoint{Could the authors comment on Fig. 6a in the revised manuscript? The discrepancy between DNS and
  their empirical prediction is large and not explained.
}{}{}
\end{document}

### 2025-03-12 plot/plot.py

@@ -16,7 +16,7 @@ from EkmanProfileClass import Complex, EkmanUniversalClass, build_grid
 D2R = np.pi / 180. 
 
 use_data=False   # if true -- the netcdf files from DNS are needed 
-plot_ustar_alpha=False
+plot_ustar_alpha=True
 plot_ekman_ideal=False 
 plot_summary=False
 plot_evisc=False
@@ -1286,9 +1286,9 @@ if plot_ustar_alpha:
     print('TOTAL ERROR (us,alpha): {},{}'.format(err_us,err_al))
     fig=plt.figure(figsize=(6,3))
     ax=fig.add_axes([0.12,0.15,0.85,0.83])
-    ax.plot(re_arr,1./us_arr,label=r'$G/u_\star$',c='blue')
+    ax.plot(re_arr,1./us_arr,label=r'$Z_\star=G/u_\star$',c='blue')
     us_est=3.9*np.log(re_arr)-8
-    ax.plot(re_arr,us_est,ls=':',lw=1,c='black',label=r'$u_\star=4\log(Re)-8$') 
+    ax.plot(re_arr,us_est,ls=':',lw=1,c='black',label=r'$Z_\star=4\log(Re)-8$') 
     ax.plot(re_arr,al_arr/D2R,label=r'$\alpha$',c='red')
     lg=plt.legend(loc='center right')
     ax.set_xlim(4e2,1.5e6)

### 2025-03-27 write1/REVISION1/manuscript.tex
finished work on Reviewer 1 responses
finished work on Review two and most part of no. 3

--- 2024-11-21 write1/manuscript.tex
@@
 \usepackage{hyperref}
+\journalname{Bounday-Layer Meteorology}
 \setcitestyle{aysep={}}
@@
 \date{\scriptsize Submitted \today}
-\title{Wind veer and speed in turbulent Ekman flow part I: scaling analysis and universal profile model}
-\titlerunning{A Universal Velocity Profile for Ekman Flow}
+\title{Wind veer and speed in turbulent Ekman flow part I: scaling analysis and velocity profile model}
+\titlerunning{Scaling Analysis and Velocity Profile for Ekman Flow}
 \author{Cedrick Ansorge$^1$ and Hauke Wurps$^{2,3}$}
@@
   %
-  The profile of the streamwise component follows the classical viscous, logarithmic and wake scaling.
+  The profile of the streamwise component, considered in a wall-stress-aligned reference frame, follows the classical viscous, logarithmic and wake scaling.
   %
@@
   Viscous stress scales universally across the boundary layer in inner units while the total stress
-  becomes universal as a function of outer height.
+  becomes universal as a function of the outer height, commonly denoted as $z^-$.
   %
   This implies a mixed scaling for the turbulent stress and eddy viscosity across the inner layer
-  and convergence to a universal scaling as function of the outer height across the outer layer
-  for increasing scale separation vide Reynolds numbers. 
+  and convergence to a scaling as function of the outer height across the outer layer
+  for increasing scale separation, i.e. for increasing Reynolds number. 
   % 
@@
 Not only does the veering set the frame of reference for surface layer theory, it also has
-effects at small and large scales from large-scale dispersion via plume spreading to
-cyclone spin-down \citep{svensson:BM2009}
-and on the capabilities of data assimilation and accuracy of surface flux estimates \citep{brown:QJR2005}. 
+effects at small and large scales from large-scale dispersion via plume spreading to cyclone 
+spin-down \citep{svensson:BM2009} and on the capabilities of data assimilation 
+and accuracy of surface flux estimates \citep{brown:QJR2005}. 
 %
@@
 One approach is to match the inner and outer layer at a reference height as suggested by
-\cite{etling:2002, emeis:2018}  (Sec. 21.10; Eq. 21.48); they choose the Prandtl-layer height $z_\text{Prandtl}$ to match
-the wind speed profiles, which, however, requires external prescription of $\alpha(z_\text{prandtl})$, the veering at that height. 
+\cite{etling:2002, emeis:2018}  (Sec. 21.10; Eq. 21.48); they choose the Prandtl-layer depth, 
+a measure for the depth of the surface layer, to match the wind speed profiles. 
+However, this requires external prescription of the veering at that height. 
 A one-dimensional profile with constant veering is given by \citet[Sec. 3; Eq. 3.1-3.19]{emeis:m2007}. 
@@
 Turbulent Ekman flow is considered here as a conceptual model of the
-homogeneous, stationary ABL over a flat surface under neutral stratification. 
-%
-Universal profiles of the wind vector for turbulent Ekman flow not only are a well-described limit
-for theoretical exploration or higher-order approaches taking into account possible effects of stratification,
+homogeneous, stationary ABL over a flat surface and under neutral stratification. 
+%
+The description of velocity profiles for this strongly simplified problem is not only a canonical fluid-mechanical problem, but it 
+also constitutes a well-described limit for theoretical exploration or higher-order approaches taking into account possible effects of stratification,
 roughness or other physical complications encountered in the real geophysical system.
@@
 %
-(iv) Barotropic and baroclinic effects on the velocity profile require to consider 
+(iv) Barotropic and baroclinic effects on the velocity profile require consideration of
 the height-dependence of the veer and stress misalignment \citep{momen:JAS2018, ghannam:QJR2021}.
@@
 %
-The f-plane approximation is applied such that rotation only acts on horizontal velocity components;
-we thus neglect rotational effects on the horizontal components of velocity and dynamical effects
-due to latitudinal variation of the rate of rotation.
+The f-plane approximation is applied, i.e. rotation acts on the horizontal components of velocity alone;
+rotational effects on the vertical component of velocity and dynamical effects
+due to latitudinal variation of the rate of rotation are neglected.
 %
@@
 We consider velocity profiles only, i.e. all velocities are averaged over horizontal planes and in time, that is, they 
-correspond to an Ensemble average. 
+correspond to an ensemble average. 
 %
@@
 \begin{equation}
-  \underline{\tau}_\mathrm{sfc} = \left(\begin{array}{c} \tau_x\\ \tau_y\\\tau_z\end{array}\right) = - \nu \left( \frac{\partial U}{\partial z} \hat{e}_x + \frac{\partial V}{\partial z} \hat{e}_y\right)
+  \underline{\tau}_\mathrm{sfc} = \left(\begin{array}{c} \tau_x\\ \tau_y\\\tau_z\end{array}\right) = - \rho\nu \left( \frac{\partial U}{\partial z} \hat{e}_x + \frac{\partial V}{\partial z} \hat{e}_y\right)
 \end{equation}
@@
 \begin{equation}
-  u_\star = \sqrt{ || - \underline{\tau}_\mathrm{sfc} || } 
+  u_\star = \sqrt{ \frac{|| - \underline{\tau}_\mathrm{sfc} ||}{\rho} } 
   \label{eqn:ustar_def}
@@
 statistics do not depend on time; and this state is defined by a Reynolds number,
-the only non-dimensional parameter that governs the system.
+the only non-dimensional parameter that governs the system (in Ekman flow, the ratio of Rossby and Reynolds number 
+are not independent as $Re_D=2\Lambda_\text{Ro} / D$).
 %
@@
 scales ($u_\star, \nu$). 
+%
+The outer normalization (with respect to the boundary-layer depth $\delta$ and velocity $u_\star$) is denoted 
+by a superscript $'-'$, i.e. $z^-=z/\delta$.  
 %
@@
 %
-The Eddy viscosity plays a crucial part when modelling profiles and the vertical
+The eddy viscosity plays a crucial part when modelling profiles and the vertical
 transport in turbulent flow. 
@@
 %
-For the symmetries in the flow (horizontal homogeneity, and $W=0$), it is
+For the assumptions of horizontal homogeneity and incompressibility (which implies  $W=0$, i.e. no mean wall-normal velocity), it is
 %
@@
   \caption{Horizontal slices of turbulence kinetic energy in the
-    Buffer layer (i=10),
-    logarithmic Layer (i=100), and
-    outer layer (i=400) of the case with $Re_\tau=2978$;
+    Buffer layer (i=10, $z^+\approx 10.5$),
+    logarithmic layer (i=100, $z^+ \approx 150$), and
+    outer layer (i=400, $z^+ \approx 1200$) of the case with $Re_\tau=2978$;
     coloring between percentiles 4 and 96 of the respective image. Lower panel:streamwise--vertical intersect through the domain 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{A universal velocity profile for the turbulent Ekman layer}
+\section{A Reynolds-number-independent velocity profile for the turbulent Ekman layer}
 \label{sec:profiles} 
@@
 In view of this situation, we use an adapted Ekman spiral that does not enforce the
-boundary conditions at the surface but at a different height. 
+boundary conditions at the surface but at a different height while maintaining the restriction 
+to constant eddy viscosity. 
 %
@@
 \subsection{Implications for surface-layer scaling} 
-Eq.~(\ref{eqn:scaling_spanwise}) establishes a universal mixed scaling for the spanwise velocity in the viscous layer:
+Eq.~(\ref{eqn:scaling_spanwise}) establishes a mixed scaling for the spanwise velocity in the viscous layer:
 %
@@
 %
-Rather than giving a universal profile for this region, we resort here to a parametric description of the problem in terms of the function $f_{V,\mathrm{log}}$ determined by the parameters $a_\mathrm{log}$, $b_\mathrm{log}$, $c_\mathrm{log}$ which can be estimated based on the above scaling considerations for any Reynolds number.
+Rather than giving a profile for this region, we resort here to a parametric description of the problem in terms of the function $f_{V,\mathrm{log}}$ determined by the parameters $a_\mathrm{log}$, $b_\mathrm{log}$, $c_\mathrm{log}$ which can be estimated based on the above scaling considerations for any Reynolds number.
 %
@@
 %
-Based on scaling considerations and direct numerical simulation spanning on decade in external separation from $Re_\Lambda=1.25 \times 10^5$ to $Re_\Lambda=1.28\times 10^6$, we derive a universal formulation for the both horizontal components of the velocity profile in a stationary, smooth Ekman layer.
+Based on scaling considerations and direct numerical simulation spanning on decade in external separation from $Re_\Lambda=1.25 \times 10^5$ to $Re_\Lambda=1.28\times 10^6$, we derive a closed formulation for both horizontal components of the velocity profile in a stationary, smooth Ekman layer.
 % 
@@
 \label{app:ekman_solution}
-The following Ekman system is obtained by averaging the Navier--Stokes equations horizontally and over time and neglecting the turbulent transport terms (turbulence can then be considered via the eddy-viscosity concept through variations in the viscosity $\nu$):
+The following Ekman system is obtained by averaging the Navier--Stokes equations horizontally and over time and neglecting the turbulent transport terms (turbulence can then be considered via variation of the vertically constant eddy viscosity $\nu$; for a height-dependent eddy-viscosity model, the reader is referred to \cite{basu:BM2023}):
 \begin{subequations}
@@
 \vspace{1em}
-\noindent\textbf{Acknowledgements} Simulations were carried out on the German supercomputing systems 
+\noindent\textbf{Acknowledgements} We thank Sukanta Basu and two anonymous reviewers for their valuable reviews of this manuscript. Simulations were carried out on the German supercomputing systems 
 \texttt{hawk} at H\"ocstleistungsrechenzentrum Stuttgart (Bundesprojekt 44187) and \texttt{juwels} at Gauss Centre for supercomputing, 
@@
   7(12):3078--3082, \doi{10/bw2sqs}
+
+\bibitem[{Basu and Holtslag(2023)}]{basu:BM2023}
+Basu S, Holtslag AAM (2023) A novel approach for deriving the stable boundary layer height and eddy viscosity concepts from the Ekman equations. Boundary-Layer Meteorol (187):105--115, \doi{}
 

### 2025-03-27 write1/REVISION1/responses.tex
finished work on Review two and most part of no. 3

@@
 I am recommending that the manuscript is accepted in BLM, pending minor revisions.
-}{}{}
+}{We appreciate your summary of our work and thank you for the time spent on our manuscript.
+Below, please find the detailed responses to your comments.}{}
 
@@
 Are you assuming the streamwise component is aligned with the geostrophic wind direction far above the surface?
-}{}{}
-\revpoint{L. 16 outer height - do you mean outer length scale? }{}{}
-\revpoint{L. 19 “vide Reynolds numbers” - I’m not sure this makes sense to me. }{}{}
-\revpoint{L. 63 Prandtl-layer height is undefined here}{}{}
-\revpoint{L. 124 - “require one to consider” or “require consideration of” }{}{}
+}{This is indeed important and we inserted an apposition phrase for clarification}{[\ldots], considered in a wall-stress-aligned reference frame, [\ldots]}
+
+\revpoint{L.
+16 outer height - do you mean outer length scale?
+}{We mean 'outer height' here, but we realize the phrase may be uncommon to some BLM readers and have clarified as follows}{[\ldots] commonly denoted as $z^-$.}
+
+\revpoint{L. 19 “vide Reynolds numbers” - I’m not sure this makes sense to me. }{What we mean here is that an 
+increase in scale separation is the same as an increase in Reynolds number. We rephrased}{[\ldots], i.e. an increase in Reynolds number.}
+
+\revpoint{L.
+63 Prandtl-layer height is undefined here}{The symbol was superfluous and is not needed.
+We now describe what the Prandtl-layer height is and trust that the interested reader will be able to find details on the exact definition in the reference that is mentioned here.
+}{[\ldots] they choose the Prandtl-layer depth, a measure for the depth of the surface layer, to match the wind speed profiles.
+However, this requires external prescription of the veering at that height. }
+
+\revpoint{L.
+124 - “require one to consider” or “require consideration of” }{Thank you!
+We chose the latter.}{[\ldots] require consideration of the height-dependence of the veer and stress misalignment [\ldots].}
+
 \revpoint{ll.
@@
 The authors state that rotation only acts on the horizontal components of velocity, but go on to state that they neglect rotational effects on the horizontal components of velocity.
-}{}{}
-\revpoint{l. 176 - I do not think that Ensemble should be capitalized }{}{}
+}{Thank you for spotting this inconsistency, of course rotational effects on the \textbf{horizontal} components of velocity are considered, they are only neglected for the \textbf{vertical} components.
+}{The f-plane approximation is applied, i.e.
+rotation acts on the horizontal components of velocity alone; rotational effects on the vertical component of velocity and dynamical effects
+due to latitudinal variation of the rate of rotation are neglected.}
+
+\revpoint{l. 176 - I do not think that Ensemble should be capitalized }{adopted.}{}
+
 \revpoint{L.
@@
 The problem would depend on Reynolds number alone in the absence of rotation, but if there is rotation, the relevant dimensionless parameters should be both Re and Ro.
-}{}{}
-\revpoint{L. 221 - It might be better to use scientific notation here. }{}{}
-\revpoint{L. 277, 281 - the notation is a bit awkward here, e.g. $V^{\alpha_* +}(z^+)$. Is there a way to make it clearer? }{}{}
+}{In general, this is true.
+However, when Ekman flow is considered, the ratio of Rossby and Reynolds number (which is the Ekman number) is fixed.
+To clarify the impact of this constraint: it also implies that - given the Coriolis parameter $f$, i.e.
+a latitude - the geostrophic wind defines the Rossby radius.
+That is, Ekman flow is  a rather strong constraint form the perspective of large-scale dynamics.
+Nonetheless, it is a widely adopted simplification in the context of boundary-layer dynamics.
+We add a phrase clarifying the uniqueness of the problem given $Re$.
+For clarification, one may see that the Ekman-based Reynolds number $Re_D$ can be written as Ratio of the Rossby Radius $\Lambda_\text{Ro}$ and laminar Ekman-layer scale $D$, which is another illustration of the link between large-scale and boundary-layer dynamics: \begin{equation} Re_D = \frac{GD}{\nu} = \frac{G \sqrt{2\nu}}{\sqrt{f}\nu}= \sqrt{\frac{2 G^2}{f\nu}}=\sqrt{\frac{G^2}{f^2}\frac{2f}{\nu}}=\sqrt{4 \Lambda_\text{Ro}^{2} \frac{f}{2\nu}}=2\Lambda_\text{Ro}/D.\label{eqn:re-relation}\end{equation}
+Analogously, one may see that 
+$Ro = \frac{G}{\Lambda_\text{Ro}f} = 1$ for the definition of the Rossby number.
+Alternatively -- when replacing $\Lambda_\text{Ro}$ by $D$, which also yields a valid definition of Rossby number from the perspective of boundary-layer dynamics-- it is (cf.
+Eq.~\ref{eqn:re-relation})
+\begin{equation}
+\text{Ro}_D=\frac{G}{Df} = \text{Ro} \frac{ D}{\Lambda_\text{Ro}} = \frac{2}{Re_D}
+\end{equation} 
+}{}
+
+\revpoint{L.
+221 - It might be better to use scientific notation here.
+}{As we do need five significant digits for the middle number, we do not see the benefit of moving to scientific notation}{ // no change incurred //}
+
+\revpoint{L.
+277, 281 - the notation is a bit awkward here, e.g.
+$V^{\alpha_\star+}(z^+)$.
+Is there a way to make it clearer?
+}{We have thought a lot about this notation and checked many different approaches.
+It is very important to note that the index $a_\star$ cannot simply ommited (in fact, above it has been demanded to clarify the alignment, which is the sole purpose of the index $\alpha_\star$).
+At the moment, we do not see a better option - if there may be more aesthetic options, all of them which come to our mind, would sacrifize clarity.}{// no change incurred // }
 \revpoint{Fig.
@@
 }{}{}
-\revpoint{L. 308 Eddy should not be capitalized}{}{}
-\revpoint{L. 311 - The assumptions of horizontal homogeneity and W=0 are not exactly the same thing as symmetries in the flow. }{}{}
-\revpoint{Fig. 3: It would be better to present heights in terms of z+ values rather than the index of the vertical coordinate. }{}{}
+\revpoint{L. 308 Eddy should not be capitalized}{adopted}{The eddy viscosity [\ldots]. }
+\revpoint{L.
+311 - The assumptions of horizontal homogeneity and W=0 are not exactly the same thing as symmetries in the flow.
+}{We realize that this was an imprecise phrasing and clarified.}{For the assumptions of horizontal homogeneity and incompressibility (which implies  $W=0$, i.e.
+no mean wall-normal velocity), it is [\ldots].
+}
+\revpoint{Fig.
+3: It would be better to present heights in terms of $z^+$ values rather than the index of the vertical coordinate.
+}{We believe that both ways of conveying this information have their value and added the requested distance from the wall in inner units.}{\textbf{Fig.\,3\ }Horizontal slices of turbulence kinetic energy in the
+Buffer layer (i=10, $z^+\approx 10.5$),
+logarithmic layer (i=100, $z^+ \approx 150$), and
+outer layer (i=400, $z^+ \approx 1200$) of the case with $Re_\tau=2978$;
+coloring between percentiles 4 and 96 of the respective image. Lower panel:streamwise--vertical intersect through the domain 
+\label{fig:slices}}
 
 \newpage
-\section*{Response to Reviewer 3}
+\section*{Response to Reviewer 3 (Prof. Sukanta Basu)}
+
+Dear Professor Basu, \\[1em]
+
+\noindent we like to extend our sincere thanks for your assessment of our manuscript. We understand your criticism of our somewhat bold claim of universality and have amended the manuscript.\\[1em]
+
+\noindent We ask for your understanding that, as you already anticipated, a complete reconsideration of our results in the framework of your 2023 paper is not carried out at this point.
+We, however, would be happy to provide processed and unprocessed data if a colleague or potentially member of your working group is interested in carrying these ideas further.\\[1em]
+
+\noindent As you unveiled your identity and you gave your consent to be mentioned, we also thank for your contribution in the acknowledgments of the revised manuscript. \\[1em]
+
+\noindent Cedrick Ansorge and Hauke Wurps \\[2em]
+\hrule
+
 \revpoint{This is an excellent paper that should be published in BLM. However, I request the authors to tone
@@
 formulations. Some of these formulations have a physical basis, while others are ad-hoc. Please refrain
-from calling the results “universal”. } {}{}
+from calling the results “universal”.
+}{We critically reassessed our use of the work 'universal' in the manuscript and indeed agree that in some respect, in particular with respect to the derived velocity-profile formulations the use was somewhat overstating.
+We removed these occurrences in the introduction, conclusion and some parts of the results.
+Most importantly, we removed the word from the running title and from the section title.
+However, in the context of some the classic universal scalings *(logarithmic scaling, viscous scaling), we keep the word -- also in the context of our results.}{A number of changes was carried out throughout the manuscript in response to this point, we kindly ask you to refer to the annotated pdf file for deatils.
+Primarily we undertook the following modifications \\
+  (1) changed title to \textbf{Wind veer and speed in turbulent Ekman flow part I: scaling analysis and velocity profile model}\\
+  (2) changed heading of Section 4.1 to \textbf{A Reynolds-number-independent velocity profile for the turbulent Ekman layer}\\
+  (3) Conclusions: universal formulation $\rightarrow$ \textbf{closed formulation} }
 

### 2025-03-27 plot/plot.py

@@ -16,9 +16,9 @@ from EkmanProfileClass import Complex, EkmanUniversalClass, build_grid
 D2R = np.pi / 180. 
 
 use_data=False   # if true -- the netcdf files from DNS are needed 
-plot_ustar_alpha=True
+plot_ustar_alpha=False
 plot_ekman_ideal=False 
-plot_summary=False
+plot_summary=True
 plot_evisc=False
 plot_profiles=False
 plot_visc_outer=False
@@ -32,7 +32,7 @@ test_inner_streamwise=False
 plot_profile_comparison=False
 plot_vandriest=False
 plot_total_rot=False
-plot_shear_vs_rotation=True
+plot_shear_vs_rotation=False
 
 colors = {400 : 'gray',
           500 : 'pink',
@@ -58,6 +58,8 @@ files = { 400: base+'avg_flw_ri0.0_re400.0_1024x384x1024.nc',
           1301:base+'avg_flw_ri0.0_re1300.0_co0_2560x640x5120.nc',
           1600:base+'avg_flw_ri0.0_re1600.0_co0_3840x960x7680.nc' } 
 
+
+print(Dataset(files[1600]).variables['y'][[10,100,400]])
 
 def etling_us(z0,zp,al):
     # universal parameter lz = lambda_z * 1/sqrt(2)

### 2025-04-21 write1/REVISION1/responses.tex
continued reviewer response

@@
 
+\renewcommand{\familydefault}{\sfdefault}
 \onehalfspacing
@@
 
-\title{Response to Reviewers} 
+\newcommand{\explainformat}{\noindent In our response below, we repeat your original comment \textcolor{gray}{in gray font}.
+Next comes \textcolor{blue}{\textit{our response in blue font}}, and we close by the changes to the manuscript in regular font.
+Comments are separated by horizontal lines for readability.\\[1em]}
+
+\title{Response to the Decision Letter of the Editor and Reviewers\\
+\normalsize -- Wind veer and speed in turbulent Ekman flow part I: scaling analysis and velocity profile model -- }
 \author{Cedrick Ansorge and Hauke Wurps }
@@
 \maketitle
-
+\tableofcontents 
+\newpage 
 %-------------------------------------------------------------------------------------%
@@
 %######################################################################################%
+\begin{flushright}{Berlin, \today}\\ \end{flushright}
+\section{Response to the Editor} 
+\noindent Dear Professor Anderson, \\[1em]
 
-\section*{Response to Reviewer 1}
+\noindent we thank you very much for the assessment of both parts of this combined manuscript and the commitment of BLM to consider this body of work as a whole. \\[1em]
+
+\noindent Below, we detail our responses including changes incurred to the manuscript.
+We also provide an annotated manuscript (manuscript\_diff.pdf) to ease the assessment of changes and expedit the further processing of our manuscript.
+\\[1em]
+
+\noindent As it happens, some changes to this part of the manuscript are required due to Reviewer comments on the second part of the paper.
+We detail these changes below -- before the responses to the individual reviewers.
+\\[1em]
+
+\explainformat{}
+
+\noindent Kind Regards, \\
+\noindent Cedrick Ansorge and Hauke Wurps
+
+\newpage 
+\begin{flushright}{Berlin, \today}\\ \end{flushright}
+\section{Response to Reviewer 1}
 
@@
 \\[1em]
+
+\explainformat{}
 
@@
 Please make them consistent.
-}{Manz thanks for spotting this inconsistency in nomenclature.
+}{Many thanks for spotting this inconsistency in nomenclature.
 We have aligned $\tau_\text{sfc}$ with the conventional definition of dynamic shear stress and now include the division bz density in Eq.
@@
 \newpage
-\section*{Response to Reviewer 2}
+\begin{flushright}{Berlin, \today}\\ \end{flushright}
+\section{Response to Reviewer 2}
+Dear Reviewer 2, \\[1em]
+
+\noindent Many thanks for the positive feedback on our manuscript.
+We have addressed the minor points raised in your review and hope that you agree with the edits.
+Pleas find the detailed responses below.
+\\[1em]
+
+\explainformat{}
+
+\noindent Kind regards, \\[0.5em]
+
+\noindent Cedrick Ansorge and Hauke Wurps  \\[0.5em]
+
 \revpoint{The authors perform direct numerical simulations of the Ekman layer at moderate Reynolds number, and use their simulations to examine the vertical mean velocity profiles in rotating, neutrally-stratified flows that serve as a proxy for the neutral atmospheric boundary layer.
@@
 \newpage
-\section*{Response to Reviewer 3 (Prof. Sukanta Basu)}
+\section{Response to Reviewer 3 (Prof. Sukanta Basu)}
 
@@
 
-\noindent We ask for your understanding that, as you already anticipated, a complete reconsideration of our results in the framework of your 2023 paper is not carried out at this point.
+\noindent We ask for your understanding that, as you already anticipated, a complete reconsideration of our results in the framework of your 2023 paper is not carried out at this point (the detailed reasoning is given below).
 We, however, would be happy to provide processed and unprocessed data if a colleague or potentially member of your working group is interested in carrying these ideas further.\\[1em]
@@
   is difficult to do, I suggest highlighting the limitations of their current bulk approach. Also, emphasize
-  the limitations of Eq. (10a) as it assumes constant eddy viscosity. }{}{}
+  the limitations of Eq. (10a) as it assumes constant eddy viscosity.}
+  {We agree that eddy viscosity is not constant as a function of height.
+In fact, we even discuss the height dependency of eddy viscosity in Fig.
+2c,d / Sec.
+2.
+At the same time, the revised Ekman model / Eq.
+(10) makes use of Ekman's assumption of height-constant eddy viscosity.
+The important point here is that this is only used in the upper part of the boundary layer, where variation of the eddy viscosity is not the dominant mechanism for the wind profile.
+We find excellent agreement above $z^-\approx 0.5$ (cf.
+Fig.
+6a/b in the manuscript).
+We are aware that deviations from the idealized Ekman profile below $z^-\approx 0.5$ are due to variations of eddy viscosity.
+  The key novelty in our approach regarding the Ekman spiral is, however, to apply it only in parts of the outer layer and to acknowledge its inappropriateness for most of what is commonly considered the turbulent boundary layer (about 2/3 of $\delta$).
+Across this range, where we argue to apply the Ekman model the eddy viscosity is not constant, but its variation appears limited, also in view of our own results presented in Sec.
+2.
+We do hence not see an advantage to incorporate the model using variable eddy viscosity suggested by Basu et al.
+(2023) but rather understand this as an alternative approach, whichs is now mentioned in the revised manuscript.}{}
 

### 2025-04-23 write1/REVISION1/manuscript.tex
finished responses on Part I

@@
 %
+While alternative approaches exist that incorporate a variation of the eddy viscosity \citep{basu:BM2023}, 
+we stick here to \citeauthor{ekman:AMA1905}'s classic assumption of constant eddy viscosity. 
+As the Ekman spiral is applied only for $z>z_{blend}\approx 0.28$ (cf. Sec.~\ref{sec:matching}
+below), the assumption of constant eddy viscosity is also weakened in the sense that it is only required for 
+$z>z_{blend}$. 
+% 
 Corresponding profiles are shown in comparison with data from three DNS runs in Fig.~\ref{fig:profiles_outer}.
-The idealized profiles capture the secondary minimum and convergence to the geostrophic equlibrium in the
+The idealized profiles capture the secondary minimum and convergence to the geostrophic equilibrium in the
 non-turbulent flow very well.
 %
+Deviations from the Ekman spiral are found towards the lower end of the corresponding height range, 
+marked by a dot in the figure; this is consistent with an increasing variation of eddy viscosity when 
+the more turbulent part of the boundary layer is entered from aloft.
 
@@
 \subsection{Matching of the inner and outer layer profiles}
+\label{sec:matching}
 % 

### 2025-04-23 write1/REVISION1/responses.tex
finished responses on Part I

@@
 \maketitle
-\tableofcontents 
-\newpage 
 %-------------------------------------------------------------------------------------%
@@
 \begin{flushright}{Berlin, \today}\\ \end{flushright}
-\section{Response to the Editor} 
+\section*{Response to the Editor} 
 \noindent Dear Professor Anderson, \\[1em]
@@
 
-\revpoint{The authors analyze wind speed and direction profiles in turbulent Ekman flow using asymptotic theory and DNS data.
+\revpoint{1.
+The authors analyze wind speed and direction profiles in turbulent Ekman flow using asymptotic theory and DNS data.
 They found that the streamwise velocity follows classical scaling laws, while outer-layer profiles align with a modified Ekman spiral.
@@
 
-\revpoint{In line 47, friction velocity is denoted as $u_* = \sqrt{|\tau_{sfc}/\rho}$.
+\revpoint{2.
+In line 47, friction velocity is denoted as $u_* = \sqrt{|\tau_{sfc}/\rho}$.
 However, in Eq.
@@
 
-\revpoint{Location for the definition of $z^-$.
+\revpoint{3.
+Location for the definition of $z^-$.
 The first appearance of $z^-$ is at line 262, and figure 1b, but it was defined on line 307.
@@
 
-\revpoint{Figure 4: Should the legend be $Z_* = 4 log(Re_D) -8$, instead of $u_\star* = ...$"?}{Indeed, there was a typo, please find the revised Figure below.}{
+\revpoint{4. Figure 4: Should the legend be $Z_* = 4 log(Re_D) -8$, instead of $u_\star* = ...$"?}{Indeed, there was a typo, please find the revised Figure below.}{
 \includegraphics[width=\textwidth]{Fig4.eps}
@@
 
-\revpoint{This is an excellent paper that should be published in BLM. However, I request the authors to tone
+\revpoint{1. This is an excellent paper that should be published in BLM. However, I request the authors to tone
 down their narrative, including the title. It is an empirical study that tries to fit DNS data using various
@@
 
-\revpoint{At the heart of this paper is Eq.~10a. Following Etling's work and others, this equation has been derived in Appendix A1. \newline 
+\revpoint{2. At the heart of this paper is Eq.~10a. Following Etling's work and others, this equation has been derived in Appendix A1. \newline 
   \[
@@
   is difficult to do, I suggest highlighting the limitations of their current bulk approach. Also, emphasize
-  the limitations of Eq. (10a) as it assumes constant eddy viscosity.}
-  {We agree that eddy viscosity is not constant as a function of height.
+  the limitations of Eq.
+(10a) as it assumes constant eddy viscosity.}{We agree that eddy viscosity is not constant as a function of height.
 In fact, we even discuss the height dependency of eddy viscosity in Fig.
@@
 We do hence not see an advantage to incorporate the model using variable eddy viscosity suggested by Basu et al.
-(2023) but rather understand this as an alternative approach, whichs is now mentioned in the revised manuscript.}{}
+(2023) but rather understand this as an alternative approach, whichs is now mentioned in the revised manuscript.
+The corresponding paragraph has been amended accordingly.}{In summary, the outer layer of Ekman flow is characterized by a turning of the wind velocity and the super-geostrophic maximum that is sustained by momentum convergence at the inflection point of the velocity profile.
+  %
+  The super-geostrophic maximum of streamwise velocity and a secondary minimum aloft the bulk-turbulent part of
+  the boundary layer are well-described by a classic Ekman spiral with adapted boundary conditions and a shift
+  in reference height.
+  %
+  While alternative approaches exist that incorporate a variation of the eddy viscosity (Basu, 1923), 
+  we stick here to Ekman's classic assumption of constant eddy viscosity. 
+  As the Ekman spiral is applied only for $z>z_{blend}\approx 0.28$ (cf. Sec.~4.5 below), the assumption of constant eddy viscosity is also weakened in the sense that it is only required for 
+  $z>z_{blend}$. 
+  % 
+  Corresponding profiles are shown in comparison with data from three DNS runs in Fig.~6.
+  The idealized profiles capture the secondary minimum and convergence to the geostrophic equilibrium in the
+  non-turbulent flow very well.
+  %
+  Deviations from the Ekman spiral are found towards the lower end of the corresponding height range, 
+  marked by a dot in the figure; this is consistent with an increasing variation of eddy viscosity when 
+  the more turbulent part of the boundary layer is entered from aloft.}
 
-\revpoint{Could the authors comment on Fig. 6a in the revised manuscript? The discrepancy between DNS and
+\revpoint{3. Could the authors comment on Fig. 6a in the revised manuscript? The discrepancy between DNS and
   their empirical prediction is large and not explained.
-}{}{}
+}{We now discuss the discrepancy in Fig. 6a, see response to previous point.}{}
 \end{document}
